### 2017-11-11 manuscript-md/02_methods.Rmd
Add first revision draft

# Materials and Methods

```{r options, echo=FALSE, message=FALSE}
library(knitr)
opts_chunk$set(
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  dpi = 300,
  cache = TRUE
)
options(knitr.table.format = "pandoc")
suppressPackageStartupMessages({
  library(tidyverse)
})
source("scripts/pft_abbreviations.R")
try_data <- readRDS("extdata/traits_analysis.rds")

try_sub <- try_data %>%
  select(pft = clm45, Jmax_area:Vcmax_mass,
         -Latitude, -Longitude, -LMA)

sample_size <- try_sub %>%
  select(pft, one_of(both_params)) %>%
  mutate(
    pft = factor(pft, abbr2pft) %>% forcats::lvls_revalue(pft2abbr)
  ) %>%
  group_by(pft) %>%
  summarize_all(~sum(!is.na(.))) %>%
  gather("param", "sample_size", -pft) %>%
  mutate(param = factor(param, both_params))
```

## Trait data

Foliar trait data for this analysis comes from the TRY global traits database
(@kattge_try_2011; see Supplementary Table 1 for a complete table of component datasets and references.).
We focused our research on seven foliar traits:
Leaf longevity (months),
specific leaf area (SLA, mm$^2$ mg$^{-1}$),
leaf nitrogen content ($N_{mass}$, mg N g$^{-1}$ or $N_{area}$, g m$^{-2}$),
leaf phosphorus content ($P_{mass}$, mg P g$^{-1}$ or $P_{area}$, g m$^{-2}$),
leaf dark respiration at leaf temperature ($R_{d,mass}$, µmol g$^{-1}$ s$^{-1}$, or $R_{d,area}$, µmol m$^{-2}$ s$^{-1}$),
maximum Rubisco carboxylation rate at 25°C ($V_{c,max,mass}$, µmol g$^{-1}$ s$^{-1}$, or $V_{c,max,area}$, µmol m$^{-2}$ s$^{-1}$),
and maximum electron transport rate ($J_{max,mass}$, µmol g$^{-1}$ s$^{-1}$, or $J_{max,area}$, µmol m$^{-2}$ s$^{-1}$).
We performed analyses separately for both mass- and area-normalized traits to avoid artifacts caused by different trait normalization [@osnas_global_2013; @lloyd_les].
We restricted our analysis to TRY data that have been quality-controlled and for which adequate species information was provided for functional type classification (see @kattge_try_2011).

Although the light- and CO2-saturated photosynthetic rate ($A_{max}$) was an important trait in previous studies, we did not include it in our study for two reasons.
First of all, data on raw photosynthetic rates are highly sensitive to measurement methodology and environmental conditions, which were generally inconsistent or unavailable in TRY.
Second, $A_{max}$ is not a good measure of photosynthetic capacity because it integrates over variability in many physiologically independent traits such as $V_{c,max}$, $J_{max}$, and stomatal conductance, and is therefore not used in vegetation models as a photosynthetic parameter [@Ali_2015].

Following past studies [e.g. @wright_worldwide_2004; @wright_assessing_2005; @diaz_global_2016], we log-transformed all trait values to correct for their strong right-skewness.

## Plant functional types

```{r "pfts"}
n_species <- try_data %>%
  distinct(clm45, AccSpeciesID) %>%
  select(pft = clm45, species = AccSpeciesID) %>%
  count(pft) %>%
  mutate(
    pft = as.character(pft),
    label = pft2abbr[pft],
    pft = tools::toTitleCase(gsub("_", " ", pft))
  ) %>%
  select(`Label` = label, `PFT` = pft, `Number of species` = n)
species_caption <- paste(
  "Names, labels, and species counts for plant functional types (PFTs)",
  "used in this analysis."
)
kable(n_species, caption = species_caption, align = rep("c", ncol(n_species)))
```

(ref:ssfigcap) Sample sizes for each trait-PFT pair. $y$ axis is scaled logarithmically.

```{r "samplesize", fig.cap='(ref:ssfigcap)', fig.width=7, fig.height=7}
ss_plot <- sample_size %>%
  mutate(param = forcats::lvls_revalue(param, param_simple[both_params])) %>%
  ggplot() + 
  aes(x = pft, y = sample_size, fill = pft) + 
  geom_col() +
  scale_y_log10(breaks = c(10, 100, 1000, 10000)) +
  facet_wrap(~param, ncol = 2, labeller = label_parsed) +
  scale_fill_manual(values = pft_colors) +
  xlab("Plant functional type") +
  ylab("Sample size") +
  guides(fill = guide_legend(title = "PFT")) +
  theme_bw() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )
ss_plot
```

We assigned each species to a unique plant functional type (PFT) following the scheme in the Community Land Model (CLM4.5, @clm45_note; Tab. \@ref(tab:pfts), Fig. \@ref(fig:samplesize)).
We obtained categorical data on growth form, leaf type, phenology, and photosynthetic pathway from the TRY database.
Where the same species was reported to have different attributes in different datasets, we assigned the attribute that was observed most frequently, and, in cases of ties, selected the attribute that was more likely based on phylogenetically similar species.
For biome specification, we collected all latitude and longitude data for each species,
matched these data to 30 second ($\sim1 km^{2}$ ) mean annual temperature ($AMT$, averaged 1970-2000) data from WorldClim-2 [@worldclim],
calculated the mean temperature for each species,
and then binned these species based on the following cutoffs: boreal/arctic ($AMT \leq 5^\circ C$), temperate ($AMT \leq 20^\circ C$), and tropical ($AMT > 20^\circ C$).

## Multivariate analysis

### Basic model description

In this study, we compared three different models representing different levels of complexity.

The simplest model was the ‘univariate’ model, in which each trait was modeled independently.
For an observation $x_{i,t}$ of trait $t$ and sample $i$...

$$x_{i,t} \sim N(\mu_t, \sigma_t)$$

...where $N$ is the univariate normal (Gaussian) distribution with mean $\mu_t$ and standard deviation $\sigma_t$ for trait $t$.

The second-simplest model was the ‘multivariate’ model, in which traits were modeled as samples from a multivariate distribution with a single mean vector and covariance matrix.
For the observed vector of traits ${\mathbf{x_i}}$ for sample $i$...

$$\mathbf{x_i} \sim mvN(\mathbf{\mu}, \mathbf{\Sigma})$$

...where $mvN$ is the multivariate normal (Gaussian) distribution with mean vector $\mathbf{\mu}$ and variance-covariance matrix $\mathbf{\Sigma}$.
We ran both of these models independently for each PFT as well as for the entire dataset (as if every observation belonged to the same PFT).

The most complex model was the ‘hierarchical’ model, in which observed trait values were drawn from a PFT-specific multivariate normal distribution describing within-PFT variation and whose parameters were themselves sampled from a global multivariate distribution describing the variation among PFTs.
For the observed vector of traits $\mathbf{x}_{i,p}$ for sample $i$ belonging to PFT $p$...

$$\mathbf{x}_{i,p} \sim mvN(\mathbf{\mu}_p, \mathbf{\Sigma}_p)$$
$$\mathbf{\mu}_p \sim mvN(\mathbf{\mu}_g, \mathbf{\Sigma}_g)$$

…where $\mathbf{\mu}_p$ and $\mathbf{\Sigma}_p$ are the mean vector and variance-covariance matrix describing variation within PFT $p$, and $\mathbf{\mu}_g$ and $\mathbf{\Sigma}_g$ are the mean vector and variance-covariance matrix describing among-PFT (global) variation.

### Model implementation

We fit the above models using a Gibbs sampling algorithm that leveraged known conjugate prior relationships for efficient exploration of the sampling space.
For priors on all multivariate mean vectors ($\mathbf{\mu}$), we used normal distributions…

$$P(\mathbf{\mu}) \sim mvN(\mathbf{\mu}_0, {\mathbf{\Sigma}}_0)$$

...which gives rise to the following expression for the posterior...

$$P(\mathbf{\mu} \mid 
    \mathbf{x}, \mathbf{\Sigma}, 
    \mathbf{\mu}_0, \mathbf{\Sigma}_0)
  \sim
  mvN(\mathbf{\mu^*}, \mathbf{\Sigma^*})$$

...where...

$$\mathbf{\Sigma^*} = (\mathbf{\Sigma}_0^{-1} + n \mathbf{\Sigma}^{-1})^{-1}$$
$$\mathbf{\mu^*} = \mathbf{\mu}_0 \mathbf{\Sigma}_0^{-1} + \bar{\mathbf{x}} n \mathbf{\Sigma}^{-1}$$

...where ${\bar{{\mathbf{x}}}}$ are the sample means of the data and $n$ is the number of rows in the data.

For priors on all multivariate variance-covariance matrices, we used the Wishart distribution ($W$)...

$$P(\mathbf{\Sigma}) \sim W(\nu_0, \mathbf{S}_0)$$

...which gives rise to the following expression for the posterior...

$$P(\mathbf{\Sigma} \mid
  \mathbf{x}, \mathbf{\mu},
  \nu_0, \mathbf{\Sigma}_0)
  \sim
  (W(\nu^*, S^*))^{-1}$$

...where...

$$\nu^* = 1 + \nu_0 + n + m$$
$$\mathbf{x^*} = \mathbf{x} - \bar{\mu}$$
$$\mathbf{SS} = \mathbf{x^*}^{T} \mathbf{x^*}$$
$$\mathbf{S^*} = (\mathbf{S}_0 + \mathbf{SS})^{-1}$$

...where $n$ is the number of rows and $m$ is the number of columns in data matrix $x$. For further details on this derivation, see @gelman_bayesian.

The fundamentally multivariate nature of the sampling procedure described above makes it incapable of accommodating missing values in the data ($x$).
Therefore, our algorithm also included imputation of partially missing data, which proceeded as follows: For a block of data $\mathbf{x'}$ containing missing observations in the columns $\mathbf{m}$ and present observations in columns $\mathbf{p}$...

$$\mathbf{x'}[m] \sim mvN(\mathbf{\mu}', \mathbf{\Sigma}')$$

...where...

$$\mathbf{\mu'} = 
  (\mathbf{x'}[p] - \mathbf{\mu'}[p]) 
  (\mathbf{\Sigma}[p,p]^{-1} \mathbf{\Sigma}[p,m])$$
$$\mathbf{\Sigma'} = \mathbf{\Sigma}[m,m] - 
  \mathbf{\Sigma}[m,p]
  (\mathbf{\Sigma}[p,p]^{-1} \mathbf{\Sigma}[p,m])$$

For each model fit, we ran five parallel MCMC chains, automatically increasing the sample size until the final result achieved convergence as determined by a Gelman-Rubin potential scale reduction statistic less than 1.05 (Gelman and Rubin 1992).

We implemented this sampling algorithm in R (version 3.4.2; @rstats).
The code for fitting these models to data is available as an open source R package (`mvtraits`) at <http://github.com/ashiklom/mvtraits>.

### Analysis of results

To assess the impact of multivariate and hierarchical constraint on trait estimates,
we compared the mean and size of the 95% confidence intervals of mean estimates of each trait for each PFT (Fig. \@ref(fig:meancompare)).
In addition, we report the mean and 95% confidence intervals of mean estimates for each trait and PFT from the hierarchical model in tabular form (Tab. 2 and 3). 
To test the hypothesis that the multivariate and hierarchical models offer more value in terms of uncertainty constraint at smaller sample sizes, we calculated the relative uncertainty ($\alpha$) as a function of the mean ($\mu$) and upper ($q_{0.975}$) and lower ($q_{0.025}$) confidence limits of trait estimates as...

$$ \alpha = \frac{q_{0.975} - q_{0.025}}{\mu} $$

We then fit a generalized linear model relating relative uncertainty to sample size ($n$) for each of the model types (univariate, multivariate, and hierarchical; Fig. \@ref(fig:civssamplesize)) as...

$$ \log{\alpha} = b_0 + b_1 \log{n} $$

To assess the consistency of among- and with-PFT trait trade-offs, we calculated the eigenvalues from pairwise variance-covariance matrix estimates for each trait pair and, where the covariance was significantly different from zero ($p < 0.05$) plotted the dominant eigenvectors centered on the mean estimates (Fig. \@ref(fig:stickpairs)).
Due to the small number of points used to estimate across-PFT covariance in the hierarchical model, none of the global correlations estimated in the hierarchical model were significantly different from zero ($p < 0.05$).
For this reason, we compared within-PFT covariances as estimated by the hierarchical model with the covariances estimated by fitting a multivariate model to all of the data.

The R code for all analyses is publicly available online at <http://github.com/ashiklom/np_trait_analysis>.
The TRY data used for this analysis can be requested at <http://try-db.org>.

### 2017-11-11 manuscript-md/03_results.Rmd
Add first revision draft

# Results

```{r prepresults}
cachefile <- ".cache/mvtraits_results.rds"
results_all <- readRDS(cachefile)

max_ss <- max(sample_size$sample_size)

multi_summary <- results_all %>%
  filter(model_type %in% c("uni", "multi")) %>%
  mutate(
    mu_mean = map(data, c("mu", "Mean")) %>%
      map(~as_tibble(as.list(.))),
    mu_lo = map(data, c("mu", "2.5%")) %>%
      map(~as_tibble(as.list(.))),
    mu_hi = map(data, c("mu", "97.5%")) %>%
      map(~as_tibble(as.list(.)))
      ) %>%
  select(model_type:pft, mu_mean:mu_hi) %>%
  unnest(.sep = ".") %>%
  gather("variable", "value", -(model_type:pft), na.rm = TRUE) %>%
  separate(variable, c("stat", "param"), sep = "\\.") %>%
  spread(stat, value) %>%
  mutate(
    pft = factor(pft, abbr2pft) %>% forcats::lvls_revalue(pft2abbr),
    param = factor(param, both_params)
  )

hier_prep <- results_all %>%
  filter(model_type == "hier") %>%
  select(model_type:pft_type, data) %>%
  mutate(
    mu_vals = map(data, "mu_group"),
    pft = map(mu_vals, names),
    mu_df = map2(pft, mu_vals, ~tibble(pft = ..1, dat = ..2))
  ) %>%
  select(-data, -mu_vals, -pft) %>%
  unnest()

hier_summary <- hier_prep %>%
  mutate(
    mu_mean = map(dat, c("Mean")) %>%
      map(~as_tibble(as.list(.))),
    mu_lo = map(dat, c("2.5%")) %>%
      map(~as_tibble(as.list(.))),
    mu_hi = map(dat, c("97.5%")) %>%
      map(~as_tibble(as.list(.)))
      ) %>%
  select(model_type:pft, mu_mean:mu_hi) %>%
  unnest(.sep = ".") %>%
  gather("variable", "value", -(model_type:pft), na.rm = TRUE) %>%
  separate(variable, c("stat", "param"), sep = "\\.") %>%
  spread(stat, value) %>%
  mutate(
    pft = if_else(is.na(pft), "global", pft),
    pft = factor(pft, abbr2pft) %>% forcats::lvls_revalue(pft2abbr),
    param = factor(param, both_params)
  )

# Absolute comparison of means
means_dat <- multi_summary %>%
  filter(pft != "GLOB", !is.na(pft)) %>%
  bind_rows(hier_summary) %>%
  left_join(sample_size) %>%
  mutate(
    model_type = factor(model_type, c("uni", "multi", "hier")) %>%
      forcats::lvls_revalue(c("univariate", "multivariate", "hierarchical")),
    mass_area = factor(mass_area, c("mass", "area")),
  ) %>%
  mutate_at(
    c("mu_mean", "mu_lo", "mu_hi"),
    ~10 ^ .
  ) %>%
  filter(
    !(param %in% c("leaf_lifespan", "SLA") & mass_area == "area")
  )
```

## Estimates of PFT-level means

```{r meanvaluetable, results = "asis"}
## Write summary table
table_cap_mass <- paste(
  "Mean and 95\\% confidence interval of trait estimates",
  "for mass-normalized traits from the hierarchical model."
)
table_cap_area <- gsub("mass", "area", table_cap_mass)

table_dat <- means_dat %>%
  filter(
    model_type == "hierarchical",
    !(param %in% c("SLA", "leaf_lifespan") & mass_area == "area")
  ) %>%
  mutate(
    mu_lo = if_else(param == "Rdmass", mu_lo * 1000, mu_lo),
    mu_mean = if_else(param == "Rdmass", mu_mean * 1000, mu_mean),
    mu_hi = if_else(param == "Rdmass", mu_hi * 1000, mu_hi),
    lo_f = formatC(mu_lo, digits = 3),
    mid_f = formatC(mu_mean, digits = 3),
    hi_f = formatC(mu_hi, digits = 3),
    value = sprintf("%s (%s,%s)", mid_f, lo_f, hi_f),
    value = if_else(grepl("Vcmax|Jmax", param) & pft == "C4G", "---", value)
  ) %>%
  select(pft, param, value) %>%
  mutate(
    param = factor(param, both_params) %>% 
      forcats::lvls_revalue(param_markdown_nounit[both_params]),
  )%>%
  spread(param, value) %>%
  arrange(pft)

table_dat %>%
  select(-matches("area")) %>%
  kable(caption = table_cap_mass, format = "latex",
        escape = FALSE, booktabs = TRUE) %>%
  kableExtra::kable_styling(latex_options = c("scale_down"))
table_dat %>%
  select(-matches("mass")) %>%
  kable(caption = table_cap_area, format = "latex",
        escape = FALSE, booktabs = TRUE) %>%
  kableExtra::kable_styling(latex_options = c("scale_down"))
```

(ref:mccap) Mean and 95% confidence interval on best estimates of traits for each plant functional type from the univariate, multivariate, and hierarchical models.
For leaf lifespan and SLA, results were not significantly different between the mass- and area-based models, so only results from the mass-based model are shown.
For some PFT-trait combinations, where large error bars resulting from the relatively uninformative priors are substantially larger than the variability among means, the $y$ axes are constrained to facilitate comparison.

```{r meancompare, fig.width=7, fig.height=7, fig.cap='(ref:mccap)'}
pclip <- function(p, pname, value, lim, hi = TRUE) {
  f <- ifelse(hi, `>`, `<`)
  if_else(p == pname & f(value, lim), lim, value)
}

plot_dat <- means_dat %>%
  mutate(
    #too_few = sample_size < 2,
    #mu_hi = if_else(too_few, NA_real_, mu_hi),
    #mu_lo = if_else(too_few, NA_real_, mu_lo),
    #mu_mean = if_else(too_few, NA_real_, mu_mean),
    irrelevant = grepl("Vcmax|Jmax", param) & pft == "C4G",
    mu_hi = if_else(irrelevant, NA_real_, mu_hi),
    mu_lo = if_else(irrelevant, NA_real_, mu_lo),
    mu_mean = if_else(irrelevant, NA_real_, mu_mean),
    ## TODO: Fix these in the prior and re-run
    mu_hi = pclip(param, "Vcmax_mass", mu_hi, 1.5),
    mu_hi = pclip(param, "Vcmax_area", mu_hi, 100),
    mu_hi = pclip(param, "Jmax_mass", mu_hi, 1.8),
    mu_hi = pclip(param, "Jmax_area", mu_hi, 150),
    mu_mean = pclip(param, "Jmax_area", mu_mean, 150),
    mu_hi = pclip(param, "Rdmass", mu_hi, 0.03),
    mu_hi = pclip(param, "Rdarea", mu_hi, 1.7),
    mu_mean = pclip(param, "Rdarea", mu_mean, 1.7),
    mu_lo = pclip(param, "Rdarea", mu_lo, 0.5, FALSE),
    mu_hi = pclip(param, "Parea", mu_hi, 0.285),
    #mu_lo = pclip(param, "Jmax_area", mu_lo, 40, FALSE),
    param = forcats::lvls_revalue(param, param_fancy_chr[both_params])
  )

p_means <- ggplot(plot_dat) +
  aes(x = interaction(model_type, pft),
      y = mu_mean, ymin = mu_lo, ymax = mu_hi,
      color = pft, shape = model_type) +
  geom_pointrange(size = 0.3) +
  facet_wrap(~param, scales = "free", ncol = 2,
              labeller = label_parsed) +
  scale_y_continuous() +
  scale_color_manual(values = pft_colors) +
  guides(
    shape = guide_legend(title = "Model type"),
    color = guide_legend(title = "PFT")
  ) +
  ylab("Trait estimate mean and 95% CI") +
  xlab("Model type and PFT") +
  theme_bw() +
  theme(
    text = element_text(size = 12),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid = element_blank()
  )
p_means
```

In general, leaf trait estimates from the univariate, multivariate, and hierarchical models were similar (Fig. \@ref(fig:meancompare)).
Where we observed differences between modles, the largest were between the univariate and multivariate models, while the additional constraint from the hierarchical model tended to have a minimal effect on trait estimates.
Significant differences in trait estimates between univariate and multivariate models occurred even for traits with relatively large sample sizes, such as leaf nitrogen content.

Evergreen PFTs had by far the largest leaf lifespan, with the longest lifespan observed for temperate and boreal needleleaf evergreen species.
Meanwhile, all of deciduous species had lifespans shorter than 7 months.
Among deciduous species, lifespan was generally longer in warmer biomes than colder ones.

Across-PFT patterns in SLA and $N_{mass}$, $P_{mass}$, and $R_{d,mass}$ were similar.
Temperate broadleaved deciduous trees and shrubs generally had among the highest values of these traits, while temperate evergreen trees and shrubs had generally among the lowest.
However, none of these patterns were universal to all four traits.
For example, tropical evergreen trees had relatively high $N_{mass}$ and average SLA and $R_{d,mass}$, but among the lowest $P_{mass}$.
Similarly, temperate and boreal shrubs had higher $N_{mass}$ and $P_{mass}$ than any of the grasses, but comparable SLA.

Across-PFT patterns in $N_{area}$, $P_{area}$, and $R_{d,area}$ were different from their mass-normalized counterparts.
For example, tropical broadleaved evergreen and needleleaf evergreen trees had among the lowest $N_{mass}$ and $P_{mass}$ basis but among the highest $N_{area}$ and $P_{area}$, while the opposite was true of deciduous temperate trees and shrubs.
Species with N contents near the middle of the observed range did not shift as dramatically depending on type of normalization.

C3 grasses had both the highest $V_{c,max,mass}$ and $V_{c,max,area}$.
Compared to broadleaved trees, temperate needleleaved evergreen trees had lower $V_{c,max,mass}$ but higher $V_{c,max,area}$.
Among broadleaved trees, deciduous trees had higher $V_{c,max,mass}$ and slightly higher $V_{c,max,area}$ than evergreen trees.

C3 grasses also had the highest $J_{max,area}$, but temperate broadleaved deciduous trees had the highest $J_{max,mass}$.
All of the shrub PFTs had the lowest $J_{max,mass}$ but average or above-average $J_{max,area}$, while the opposite was true of broadleaved tropical PFTs.
Of the tree PFTs, needleleaved evergreen trees had among the highest $J_{max,area}$ but the lowest $J_{max,mass}$.

We observed clear differences in the relative uncertainties of mean estimates with respect to sample size.
All of the high-latitude PFTs consistently had among the largest error bars around their mean estimates relative to other PFTs, while the traits with the largest uncertainties were dark respiration, $V_{c,max}$, and $J_{max}$.
For some of these trait-PFT combinations, such as estimates $J_{max,mass}$ for shrub PFTs, the additional constraint from trait covariance substantially reduced error bars, making it possible to compare estimates against those of other PFTs.
Our analysis of the relationship between model type, sample size, and estimate relative uncertainty found that this covariance-based constraint from the multivariate model becomes increasingly important at smaller sample sizes (\@ref(fig:civssamplesize)).
However, this analysis revealed no consistent significant benefit from the hierarchical model for our specific analysis.

(ref:sscap) Relative uncertainty in PFT-level trait estimates as a function of sample size for each model type.
Lines represent linear models ($\log(y) = b_0 + b_1 \log(x)$) fit independently for each model type.

```{r "civssamplesize", fig.width=4.4, fig.height=3, fig.cap='(ref:sscap)'}
## Relative uncertainty
cv_dat <- plot_dat %>%
  mutate(
    cv = abs((mu_hi - mu_lo) / mu_mean)
  ) %>%
  filter(sample_size > 3)

cv_mod <- cv_dat %>%
  group_by(model_type) %>%
  nest() %>%
  mutate(
    cvfit = map(data, ~lm(log10(cv) ~ log10(sample_size), data = .))
  )

cv_pred <- cv_mod %>%
  mutate(
    grid = map(data, modelr::data_grid, sample_size = seq(5, max_ss)),
    cvpred = map2(grid, cvfit, modelr::add_predictions)
  ) %>%
  unnest(cvpred) %>%
  mutate(cv = 10 ^ pred)

x_lims <- range(cv_pred$sample_size)
y_lims <- range(cv_pred$cv)

relative_ci <- ggplot(cv_dat) +
  aes(x = sample_size, y = cv, color = model_type) +
  geom_point(size = 0.5) +
  geom_line(data = cv_pred) +
  scale_x_continuous(trans = "log10") +
  coord_cartesian(xlim = x_lims, ylim = y_lims) +
  xlab("Sample size") +
  ylab("Relative width of 95% CI") +
  scale_color_brewer(type = "qual", palette = 2) +
  guides(color = guide_legend(title = "Model type")) +
  theme_bw() +
  theme(
    legend.position = c(0.85, 0.75),
    panel.grid = element_blank()
  )
relative_ci
```

## Trait correlation patterns among- and within-PFTs

(ref:stickcap) Pairwise trait mean and covariance estimates for all data pooled globally (black) and for each PFT (colored).
Correlations not significantly different from zero are indicated by x symbols at the mean estimate.
$x$ and $y$ axes vary on a log scale, reflecting the fact that the model was fit using the base 10 log of all traits.

```{r stickpairs, fig.width = 5, fig.height = 5, fig.cap = '(ref:stickcap)'}
suppressPackageStartupMessages({
  library(mvtraits)
})
results_sub <- results_all %>%
  filter((model_type == "multi" & pft == "global") |
         (model_type == "hier"))
dat_list <- results_sub$data
names(dat_list) <- with(results_sub, paste(model_type, mass_area, sep = "_"))

stickplot_pairs(
  #mu_global_lower = dat_list$hier_mass$mu_global,
  #Sigma_global_lower = dat_list$hier_mass$Sigma_global,
  mu_global_lower = dat_list$multi_mass$mu,
  Sigma_global_lower = dat_list$multi_mass$Sigma,
  mu_group_lower = dat_list$hier_mass$mu_group,
  Sigma_group_lower = dat_list$hier_mass$Sigma_group,
  vars_lower = mass_params,
  group_names = pft2abbr[-1],
  vars_label = param_labels,
  unlog_axes = TRUE,
  #mu_global_upper = dat_list$hier_area$mu_global,
  #Sigma_global_upper = dat_list$hier_area$Sigma_global,
  mu_global_upper = dat_list$multi_area$mu,
  Sigma_global_upper = dat_list$multi_area$Sigma,
  mu_group_upper = dat_list$hier_area$mu_group,
  Sigma_group_upper = dat_list$hier_area$Sigma_group,
  vars_upper = area_params,
  col_group = pft_colors,
  cex_g = 1.2,
  lwd_g = 1,
  lty_g = 1,
  lwd_s = 1.3,
  par_plot = list(oma = c(0, 5, 3, 5), cex = 0.4),
  par_legplot = list(mar = c(0, 0, 0, 0)),
  par_legend = list(ncol = 5, x = "center", cex = 0.6)
)
mtext("Mass-normalized", side = 2, cex = 0.7, line = 3)
mtext("Area-normalized", side = 4, cex = 0.7, line = 1)
```

Where pairwise trait correlations were statistically significant, these correlations were generally consistent in direction both globally and within each PFT (Fig. \@ref(fig:stickpairs)).
In particular, mass- and area-normalized traits were all positively correlated with each other and, respectively, positively and negatively correlated with SLA, both globally and within each PFT.
On the other hand, correlations between leaf lifespan and area-normalized traits were more variable.
The correlation between $N_{area}$ and leaf lifespan was positive both globally and for evergreen shrubs, but was negative for broadleaved deciduous trees and not significant for any other PFTs.
Similarly, the correlation between $P_{area}$ and leaf lifespan was positive for temperate broadleaved evergreen trees and needleleaved evergreen trees but negative for evergreen shrubs.

A large number of pairwise trait correlations were not significant.
In some cases, this was driven by sample size (Fig. \@ref(fig:samplesize)).
For instance, needleleaved deciduous trees were often the only PFT for which a correlation was not statistically significant.
In other cases, though, PFTs with smaller sample sizes had significant pairwise correlations while PFTs with much larger sample sizes had none.
For example, tropical broadleaved evergreen trees were relatively well-sampled for all traits, but none of their traits were well correlated with leaf lifespan.
Meanwhile, temperate broadleaved evergreen trees had significant correlations between leaf lifespan and SLA, $P_{area}$, and $R_{d,area}$.
In general, we observed fewer significant trait correlations among area-normalized traits than mass-normalized traits.

```{r boxprep}
hier_mass_file <- "../output/hier.mass.clm45.NA.2017-11-09-1749.rds"
hier_area_file <- "../output/hier.area.clm45.NA.2017-11-10-0937.rds"
multi_mass_file <- "../output/multi.mass.clm45.NA.2017-11-09-1733.rds"
multi_area_file <- "../output/multi.area.clm45.NA.2017-11-09-1741.rds"

hier_summary_mass <- readRDS(hier_mass_file)[["summary_table"]] %>%
  mutate(mass_area = "mass")
hier_summary_area <- readRDS(hier_area_file)[["summary_table"]] %>%
  mutate(mass_area = "area")
multi_summary_mass <- readRDS(multi_mass_file)[["summary_table"]] %>%
  mutate(mass_area = "mass")
multi_summary_area <- readRDS(multi_area_file)[["summary_table"]] %>%
  mutate(mass_area = "area")

proc <- function(dat, hier = TRUE) {
  if (hier) {
    dat <- filter(dat, group != "global")
  } else {
    dat <- mutate(dat, group = "global")
  }
  dat <- dat %>%
    filter(variable == "Corr") %>%
    separate(index, c("xvar", "yvar"), sep = "\\.\\.") %>%
    mutate(
      pft = factor(group, abbr2pft) %>% forcats::lvls_revalue(pft2abbr)
    ) %>%
    select(mass_area, pft, xvar, yvar, Mean, `2.5%`, `97.5%`)
}

hsm <- proc(hier_summary_mass)
hsa <- proc(hier_summary_area)
msm <- proc(multi_summary_mass, FALSE)
msa <- proc(multi_summary_area, FALSE)
hs <- bind_rows(hsm, hsa, msm, msa)
```

(ref:corrbarscap) Mean and 95% CI on estimates of pairwise correlation coefficients
for all data pooled globally (dark grey) and for each PFT (colored).

```{r corrbars, fig.width=7, fig.height=7, fig.cap='(ref:corrbarscap)'}
.zzz <- close.screen(all.screens = TRUE)
corr_bar <- function(dat, xvar, yvar, mass_area) {
  dp <- dat %>%
    filter(
      mass_area == !!mass_area,
      (xvar == !!xvar & yvar == !!yvar) | 
        (xvar == !!yvar & yvar == !!xvar),
    ) %>%
    arrange(pft) %>%
    mutate(color = c("grey25", pft_colors))
  x <- barplot(dp$Mean, col = dp$color, border = dp$color,
               ylim = c(-1, 1), axes = FALSE)
  arrows(x, dp[["2.5%"]], x, dp[["97.5%"]],
         length = 0.01, angle = 90, code = 3)
  box()
}

nv <- length(mass_params)
main_screens <- split.screen(rbind(c(0, 1, 0.2, 1), c(0, 1, 0, 0.2)))
par(mar = rep(0.1, 4), oma = c(0, 8, 1, 5), cex = 0.4)
screens <- split.screen(c(nv, nv), screen = main_screens[1])
for (i in seq_len(nv)) {
  for (j in seq_len(nv)) {
    k <- j + nv * (i - 1)
    screen(screens[k])
    if (i == j) {
      # Label
      plot(0:1, 0:1, type = "n", ann = FALSE, axes = FALSE)
      box()
      text(x = 0.5, y = 0.5, labels = param_labels[i], cex = 2)
    } else {
      if (j > i) {
        # Upper triangle
        mass_area <- "area"
        xvar <- area_params[j]
        yvar <- area_params[i]
      } else {
        # Lower triangle
        mass_area <- "mass"
        xvar <- mass_params[j]
        yvar <- mass_params[i]
      }
      corr_bar(hs, xvar, yvar, mass_area)
    }
    if (i == 2 && j == 1) {
      axis(side = 2)
    }
    if (i == 1 && j == nv) {
      axis(side = 4)
    }
  }
}
## Draw legend
screen(main_screens[2])
par(mar = c(0, 0, 0, 0), oma = c(0, 0, 0, 0))
plot(0, 0, type = "n", axes = FALSE, ann = FALSE)
legend("center", legend = levels(hs$pft), fill = c("grey25", pft_colors), 
       ncol = 5, bty = "n")
close.screen(all.screens = TRUE)
mtext("Mass-normalized", side = 2, cex = 0.7, line = 1)
mtext("Mean and 95% CI of pairwise correlation", side = 2, cex = 1, line = 2)
mtext("Area-normalized", side = 4, cex = 0.7, line = 0)
```

The strength of pairwise trait correlations varied substantially depending on scale, PFT, and trait (Fig. \@ref(fig:corrbars)).
The two pairwise trait correlations that exhibited the most consistent strength globally and within each PFT were the correlation between SLA and $N_{area}$, and between $N_{mass}$ and $P_{mass}$.
Correlation strength was often related to sample size, with well-sampled PFTs exhibiting stronger correlations and undersampled PFTs exhibiting weaker correlations.

### 2017-11-11 manuscript-md/index.Rmd
Add first revision draft

---
title: "Covariance as constraint: Leveraging leaf economic relationships to constrain plant functional type trait estimates"
author: >
  Alexey N. Shiklomanov$^1$,
  Elizabeth M. Cowdery$^1$,
  Michael Bahn$^2$,
  Sabina Burrascano$^3$,
  Chaeho Byun$^4$,
  Joseph Craine$^5$,
  Andrés Gonzalez-Melo$^6$,
  Alvaro G. Gutiérrez$^7$,
  Steven Jansen$^8$,
  Nathan Kraft$^9$,
  Koen Kramer$^{10}$,
  Vanessa Minden$^{11,12}$
  Ülo Niinemets$^{13}$,
  Yusuke Onoda$^{14}$,
  Enio Egon Sosinski$^{15}$,
  Nadejda A. Soudzilovskaia$^{16}$,
  Michael C. Dietze$^1$
bibliography:
  - library.bib
  - extra.bib
csl: new-phytologist.csl
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    toc: no
---
$^1$Department of Earth & Environment, Boston University, Boston, MA 02215, USA;
$^2$Institute of Ecology, University of Innsbruck, 6020 Innsbruck, Austria;
$^3$Department of Environmental Biology, Sapienza University of Rome, Piazzale Aldo Moro 5, 00185, Rome, Italy;
$^4$School of Biological Sciences, Seoul National University, Gwanak-ro 1, 08826, Seoul, Seoul, South Korea;
$^5$Jonah Ventures, Manhattan, KS 66502, USA;
$^6$Universidad del Rosario. Facultad de Ciencias Naturales y Matemáticas. Carrera 26 #63B-48. Bogotá, Colombia;
$^7$Departamento de Ciencias Ambientales y Recursos Naturales, Facultad de Ciencias Agronomicas, Universidad de Chile, Av. Santa Rosa 11315, La Pintana 8820808, Santiago, Chile;
$^8$Institute of Systematic Botany and Ecology, Ulm University, Ulm, Germany;
$^9$Department of Ecology and Evolutionary Biology, University of California Los Angeles, Los Angeles, CA 90095, USA;
$^{10}$Vegetation, Forest, and Landscape Ecology, Wageningen Environmental Research and Wageningen University, Wageningen, Gelderland, The Netherlands;
$^{11}$Institute for Biology and Environmental Sciences, Carl von Ossietzky-University of Oldenburg, Carl von Ossietzky Str. 9-11, 26129 Oldenburg, Germany;
$^{12}$Department of Biology, Ecology and Evolution, Vrije Universiteit Brussel, Pleinlaan 2, 1050 Brussels;
$^{13}$Institute of Agricultural and Environmental Sciences, Estonian University of Life Sciences, Kreutzwaldi 1, 51014 Tartu, Estonia;
$^{14}$Graduate School of Agriculture, Kyoto University, Kyoto, 606-8502, Japan;
$^{15}$Embrapa Temperate Agriculture, Rodovia BR 392, Km 78, CEP 96010-971 Pelotas, RS, Brazil;
$^{16}$Conservation Biology Department, Institute of Environmental Sciences, Leiden University, Leiden, The Netherlands

Corresponding author:
Alexey Shiklomanov,
telephone: +1 (302) 897-3348,
email: ashiklom\@bu.edu

Key words: Functional trade-off; hierarchical modeling; trait variation; ecological modeling; leaf morphology; leaf biochemistry

```{r wordcount, echo = FALSE}
fnames <- c(
  introduction = "01_introduction.md",
  methods = "02_methods.Rmd",
  results = "03_results.Rmd",
  discussion = "04_discussion.md",
  conclusions = "05_conclusions.md"
)
files <- file.path(
  "~/Projects/new-phytologist-traits/np-trait-analysis/manuscript-md", 
  fnames
)
names(files) <- names(fnames)

getwc <- function(f) {
  ct <- withr::with_options(
    list(knitr.table.format = "pandoc"),
    wordcountaddin::text_stats(f)
  )
  wline <- ct[3]
  wlist <- strsplit(wline, split = "[[:space:]]+")
  as.numeric(wlist[[1]][3])
}

wc_table <- data.frame(wordcount = vapply(files, getwc, numeric(1)))
wc_total <- rbind(wc_table, sum(wc_table$wordcount))
wc_total$percent <- 100 * wc_total$wordcount / 
  sum(wc_total$wordcount[-nrow(wc_total)])
rownames(wc_total)[nrow(wc_total)] <- "total"
knitr::kable(wc_total, digits = 1)
```

# Summary {-}

* We investigated whether global leaf economic relationships across plant functional types (PFTs) can be reconciled with within-PFT relationships, and the extent to which this hierarchical structure can be used to constrain trait estimates.

* We developed a hierarchical multivariate Bayesian model that assumes separate means and covariance structures within and across PFTs and fit this model to seven leaf traits from the TRY database: longevity, specific leaf area, nitrogen and phosphorus contents, and rates of dark respiration, maximum Rubisco carboxylation, and maximum electron transport. 
We evaluated the ability of this model to estimate trait means and reproduce leaf economic trade-offs.

* Leveraging covariance in multivariate models reduced uncertainties in mean trait estimates, particularly for undersampled trait-PFT combinations.
Sometimes, the multivariate model significantly shifted trait means, resulting in estimates better aligned with patterns predicted from global ecology and biogeography.
However, additional constraint from the across-PFT hierarchy was limited compared to the simple multivariate models.

* Trait correlations were generally consistent in direction within and across PFTs, and consistent with predictions of the leaf economic spectrum.
However, correlation strength varied substantially across PFTs indicating that leaf economic relationships within PFTs are often confounded by other biotic and abiotic factors.

### 2017-11-11 scripts/base_compare_mean_variance.R
Update scripts from draft

@@ -1,8 +1,6 @@
 library(mvtraits)
 library(tidyverse)
 
-# TODO: Redraw both of these as single figures, faceted as param ~ mass_area.
-
 source("scripts/pft_abbreviations.R")
 cachefile <- ".cache/mvtraits_results.rds"
 results_all <- readRDS(cachefile)
@@ -10,6 +8,27 @@ try_data <- readRDS("extdata/traits_analysis.rds")
 try_sub <- try_data %>%
   select(pft = clm45, Jmax_area:Vcmax_mass,
          -Latitude, -Longitude, -LMA)
+
+n_species <- try_data %>%
+  distinct(clm45, AccSpeciesID) %>%
+  select(pft = clm45, species = AccSpeciesID) %>%
+  count(pft) %>%
+  mutate(
+    pft = as.character(pft),
+    label = pft2abbr[pft],
+    pft = tools::toTitleCase(gsub("_", " ", pft))
+  ) %>%
+  select(`Label` = label, `PFT` = pft, `Number of species` = n)
+species_table <- "manuscript-md/pftdefinitions.md"
+species_caption <- paste("Names, labels, and species counts",
+                         "for plant functional types (PFTs)",
+                         "used in this analysis.",
+                         "{#tbl:pfts}")
+if (FALSE) {
+  sink(file = species_table)
+  pander::pandoc.table(n_species, caption = species_caption, split.tables = Inf)
+  sink(NULL)
+}
 
 sample_size <- try_sub %>%
   select(pft, one_of(both_params)) %>%
@@ -20,6 +39,25 @@ sample_size <- try_sub %>%
   summarize_all(~sum(!is.na(.))) %>%
   gather("param", "sample_size", -pft) %>%
   mutate(param = factor(param, both_params))
+
+ss_plot <- sample_size %>%
+  mutate(param = forcats::lvls_revalue(param, param_simple[both_params])) %>%
+  ggplot() + 
+  aes(x = pft, y = sample_size, fill = pft) + 
+  geom_col() +
+  scale_y_log10(breaks = c(10, 100, 1000, 10000)) +
+  facet_wrap(~param, ncol = 2, labeller = label_parsed) +
+  scale_fill_manual(values = pft_colors) +
+  xlab("Plant functional type") +
+  ylab("Sample size") +
[…]

### 2017-11-11 scripts/clean_output.R

@@ -1,0 +1,22 @@
+library(mvtraits)
+
+try_dat <- readRDS("extdata/traits_analysis.rds")
+pfts <- levels(try_dat$clm45)
+group_names <- sprintf("group%02d", seq_along(pfts))
+
+fname <- "output/hier.area.clm45.NA.2017-10-25-1632.rds"
+
+out <- readRDS(fname)
+
+summary_table <- out$summary_table
+
+### Fix names
+group_vars <- c("mu_group", "Sigma_group", "Corr_group")
+clean_out$means[group_vars] <- lapply(
+  clean_out$means[group_vars],
+  "names<-",
+  pfts
+)
+names(out)[2] <- "stats"
+saveRDS(out, fname)
+

### 2017-11-11 scripts/data_map.R

@@ -1,0 +1,28 @@
+library(tidyverse)
+
+try_dat <- readRDS("extdata/traits_analysis.rds")
+try_clim <- readRDS("extdata/traits_with_climate.rds")
+
+miss <- function(x) {
+  as.integer(is.na(x))
+}
+
+map_dat <- try_dat %>%
+  mutate(
+    nonmissing_mass = miss(leaf_lifespan) + miss(SLA) + miss(Nmass) +
+      miss(Pmass) + miss(Rdmass) + miss(Vcmax_mass) + miss(Jmax_mass),
+    nonmissing_area = miss(leaf_lifespan) + miss(SLA) + miss(Narea) +
+      miss(Parea) + miss(Rdarea) + miss(Vcmax_area) + miss(Jmax_area)
+  ) %>%
+  group_by(Latitude, Longitude) %>%
+  summarize_at(vars(one_of("nonmissing_mass", "nonmissing_area")), sum) %>%
+  filter(nonmissing_mass > 0, nonmissing_area > 0)
+
+ggplot(map_dat) +
+  aes(x = Longitude, y = Latitude) +
+  borders("world", colour = "black") +
+  geom_point(aes(color = nonmissing_mass)) +
+  scale_color_gradient(low = "azure1", high = "red4", trans = "log10") +
+  guides(color = guide_colorbar(title = "Sample size")) +
+  theme_bw()
+ggsave("figures/data_map.png", width = 10, height = 6, units = "in", dpi = 300)

### 2017-11-11 scripts/exploratory.R

@@ -1,0 +1,13 @@
+library(tidyverse)
+d <- readRDS("extdata/traits_analysis.rds")
+
+d %>%
+  filter(
+    #clm45 == "broadleaf_deciduous_temperate",
+    !is.na(Narea), !is.na(Vcmax_area)
+  ) %>%
+  ggplot() +
+  aes(x = Narea, y = Vcmax_area) +
+  geom_point() +
+  geom_smooth(method = "lm")
+

### 2017-11-11 scripts/explore_prior.R

@@ -22,9 +22,10 @@ try_data <- readRDS('extdata/traits_analysis.rds')
 #   spread(vartype, value)
 
 try_long <- try_data %>%
+  select(-Latitude, -Longitude, -LMA) %>%
   select_if(is_double) %>%
   gather() %>%
-  filter(!is.na(value), key != 'LMA') %>%
+  filter(!is.na(value)) %>%
   mutate(value = log10(value))
 
 source('scripts/informative_prior.R')
@@ -39,4 +40,5 @@ plot_df <- bind_rows(try_long %>% mutate(type = 'data'),
 
 ggplot(plot_df) +
   aes(x = key, y = value, fill = type) +
-  geom_violin()
+  geom_violin() +
+  facet_wrap(~key, scales = "free")

### 2017-11-11 scripts/gg_compare_mean_variance.R

@@ -1,0 +1,24 @@
+library(nptraits)
+
+all_summaries <- readRDS('results/summaries_processed.rds')
+figures_dir <- 'figures'
+
+plot_data <- all_summaries %>%
+    distinct(area_mass, pft_scheme) %>%
+    mutate(plots = map2(area_mass, pft_scheme, compare_mv, summary_data = all_summaries)) %>%
+    mutate_if(is.factor, as.character) %>%
+    unnest()
+
+plot_data %$%
+    pwalk(list(area_mass, pft_scheme, mean_plot),
+          ~ ggsave(filename = file.path(
+              figures_dir, paste('compare_mean', ..1, ..2, 'png', sep = '.')
+          ),
+          plot = ..3))
+
+plot_data %$%
+    pwalk(list(area_mass, pft_scheme, var_plot),
+          ~ ggsave(filename = file.path(
+              figures_dir, paste('compare_variance', ..1, ..2, 'png', sep = '.')
+          ),
+          plot = ..3))

### 2017-11-11 scripts/informative_prior.R

@@ -1,17 +1,17 @@
 prior_df <- tribble(
   ~param, ~mean, ~stdev,
-  'Jmax_area', 2, 0.5,
-  'Jmax_mass', 0, 0.4,
-  'leaf_lifespan', 0.75, 0.5,
-  'Narea', 0.25, 0.5,
-  'Nmass', 1.25, 0.5,
-  'Parea', -1, 0.5,
+  'Jmax_area', 2, 0.26,
+  'Jmax_mass', 0, 0.26,
+  'leaf_lifespan', 0.75, 0.4,
+  'Narea', 0.25, 0.3,
+  'Nmass', 1.25, 0.3,
+  'Parea', -1, 0.3,
   'Pmass', 0, 0.3,
-  'Rdarea', 0, 0.5,
-  'Rdmass', -2, 0.5,
-  'SLA', 1, 0.75,
-  'Vcmax_area', -0.5, 0.5,
-  'Vcmax_mass', 1.7, 0.5
+  'Rdarea', 0, 0.27,
+  'Rdmass', -2, 0.26,
+  'SLA', 1, 0.6,
+  'Vcmax_mass', -0.4, 0.26,
+  'Vcmax_area', 1.7, 0.19
 ) %>%
   mutate(variance = stdev ^ 2)
 

### 2017-11-11 scripts/pft_abbreviations.R

@@ -4,7 +4,8 @@ abbr2pft <- c(
   BlETe = "broadleaf_evergreen_temperate",
   BlDTr = "broadleaf_deciduous_tropical",
   BlDTe = "broadleaf_deciduous_temperate",
-  NlE = "needleleaf_evergreen",
+  NlETe = "needleleaf_evergreen_temperate",
+  NlEBo = "needleleaf_evergreen_boreal",
   NlD = "needleleaf_deciduous",
   ShE = "shrub_evergreen",
   ShDTe = "shrub_deciduous_temperate",
@@ -19,39 +20,88 @@ names(pft2abbr) <- abbr2pft
 
 mass_params <- c("leaf_lifespan", "SLA", "Nmass", "Pmass", "Rdmass", "Vcmax_mass", "Jmax_mass")
 area_params <- gsub("mass$", "area", mass_params)
+both_params <- c("leaf_lifespan", "SLA", "Nmass", "Narea", "Pmass", "Parea",
+                 "Rdmass", "Rdarea", "Vcmax_mass", "Vcmax_area",
+                 "Jmax_mass", "Jmax_area")
 param_labels <- c("Leaf lifespan", "SLA", "N", "P", "Rd", "Vcmax", "Jmax")
+
+param_simple <- c(
+  leaf_lifespan = "Leaf~lifespan",
+  SLA = "SLA",
+  Nmass = "N[mass]",
+  Narea = "N[area]",
+  Pmass = "P[mass]",
+  Parea = "P[area]",
+  Rdmass = "R[list(d, mass)]",
+  Rdarea = "R[list(d, area)]",
+  Vcmax_mass = "V[list(c, max, mass)]",
+  Vcmax_area = "V[list(c, max, area)]",
+  Jmax_mass = "J[list(max, mass)]",
+  Jmax_area = "J[list(max, area)]"
+)
+
+param_markdown_nounit <- c(
+  leaf_lifespan = "Leaf lifespan",
+  SLA = "SLA",
+  Nmass = "$N_{mass}$",
+  Narea = "$N_{area}$",
+  Pmass = "$P_{mass}$",
+  Parea = "$P_{area}$",
[…]

### 2017-11-11 scripts/run_model.R

@@ -118,6 +118,7 @@ if (model_type == 'hier') {
                           parallel = parallel,
                           autofit = autofit,
                           priors = prior,
+                          threshold = threshold,
                           max_attempts = max_attempts,
                           keep_samples = keep_samples,
                           save_progress = save_progress)

### 2017-11-11 scripts/species_trait_comparison.R

@@ -1,0 +1,68 @@
+library(tidyverse)
+
+spp_file <- "../preprocess-try/pfts_species/tps_species.rds"
+data_file <- "extdata/trait_data.rds"
+
+try_spp <- readRDS(spp_file)
+try_data <- readRDS(data_file)
+
+if (FALSE) {
+
+  try_spp %>%
+    filter(AccSpeciesName == "Acer saccharum")
+
+}
+
+early_hardwood <- c(
+  "Acer barbatum",
+  "Acer grandidentatum",
+  "Acer leucoderme",
+  "Acer negundo",
+  "Acer nigrum",
+  "Acer pensylvanicum",
+  "Acer platanoides",
+  "Acer rubrum",
+  "Acer saccharinum",
+  "Acer saccharum",
+  "Acer spicatum",
+  "Aesculus flava",
+  "Alnus glutinosa",
+  "Alnus rhombifolia",
+  "Calycanthus floridus",
+  "Carpinus caroliniana",
+  "Cornus florida",
+  "Diospyros texana",
+  "Diospyros virginiana",
+  "Fagus grandifolia",
+  "Halesia carolina",
+  "Halesia parviflora",
+  "Ilex montana",
+  "Oxydendrum arboreum",
+  "Platanus occidentalis",
+  "Symplocos tinctoria",
+  "Tilia americana",
+  "Fagus sylvatica"
+)
+sp_ids <- try_spp %>%
+  filter(AccSpeciesName %in% early_hardwood) %>%
+  select(species = AccSpeciesName, AccSpeciesID)
+
+sp_compare <- sp_ids %>%
+  inner_join(try_data) %>%
+  mutate(
+    species = factor(species) %>%
+      forcats::fct_reorder(SLA, median, na.rm = TRUE)
+  )
+
+sp_means <- sp_compare %>%
+  group_by(species) %>%
+  summarize(SLA = median(SLA, na.rm = TRUE))
+
+ggplot(sp_means) +
+  aes(x = species, y = SLA) +
+  geom_col()
+
+ggplot(sp_compare) +
+  aes(x = species) +
+  geom_boxplot(aes(y = SLA))
+

### 2017-11-11 scripts/stick_pairs.R

@@ -42,8 +42,10 @@ stickplot_pairs(
   lwd_g = 1,
   lty_g = 1,
   lwd_s = 1.3,
-  par_plot = list(oma = c(0, 3, 3, 3), cex = 0.4),
+  par_plot = list(oma = c(0, 5, 3, 5), cex = 0.4),
   par_legplot = list(mar = c(0, 0, 0, 0)),
   par_legend = list(ncol = 5, x = "center", cex = 0.6)
 )
+mtext("Mass normalization", side = 2, cex = 0.7, line = 3)
+mtext("Area normalization", side = 4, cex = 0.7, line = 1)
 dev.off()

### 2017-11-11 scripts/summary_table.R

@@ -1,0 +1,48 @@
+library(mvtraits)
+library(tidyverse)
+
+source("scripts/pft_abbreviations.R")
+cachefile <- ".cache/mvtraits_results.rds"
+results_all <- readRDS(cachefile)
+
+results_hier <- results_all %>%
+  filter(model_type == "hier") %>%
+  select(mass_area, data) %>%
+  mutate(
+    mu_group = map(data, "mu_group"),
+    Sigma_group = map(data, "Sigma_group"),
+    pft_names = map(mu_group, names),
+    pft_dat = pmap(
+      list(mu_group, Sigma_group, pft_names),
+      ~tibble(pft = ..3, mu = ..1, Sigma = ..2)
+    )
+  ) %>%
+  unnest(pft_dat) %>%
+  mutate(
+    mu_mean = map(mu, "Mean"),
+    Sigma_mean = map(Sigma, "Mean") %>% map(diag),
+    param = map(mu_mean, names)
+  ) %>%
+  unnest(mu_mean, Sigma_mean, param)
+
+results_table <- results_hier %>%
+  mutate(
+    lo = 10 ^ (mu_mean - 1.96 * Sigma_mean),
+    mid = 10 ^ mu_mean,
+    hi = 10 ^ (mu_mean + 1.96 * Sigma_mean),
+    lo_f = formatC(lo, digits = 3),
+    mid_f = formatC(mid, digits = 3),
+    hi_f = formatC(hi, digits = 3),
+    value = sprintf("%s (%s - %s)", mid_f, lo_f, hi_f)
+  )
+results_table %>%
+  filter(!(param %in% c("SLA", "leaf_lifespan") & mass_area == "mass")) %>%
+  select(pft, param, value) %>%
+  mutate(
+    param = factor(param, both_params) %>% 
+      forcats::lvls_revalue(param_markdown[both_params]),
+    pft = factor(pft, abbr2pft) %>% forcats::lvls_revalue(pft2abbr)
+  ) %>%
+  spread(param, value) %>%
+  arrange(pft) %>%
+  pander::pandoc.table("")

### 2017-11-13 manuscript-md/03_results.Rmd
Fix results plots
Add equation annotation to sample size plot.

Use unaltered means (instead of adjusted plotting data) for sample size
model.

Fix limits on Rdarea in means plot.

@@
     mu_hi = pclip(param, "Rdmass", mu_hi, 0.03),
-    mu_hi = pclip(param, "Rdarea", mu_hi, 1.7),
-    mu_mean = pclip(param, "Rdarea", mu_mean, 1.7),
-    mu_lo = pclip(param, "Rdarea", mu_lo, 0.5, FALSE),
+    mu_hi = pclip(param, "Rdarea", mu_hi, 0.0035),
+    mu_mean = pclip(param, "Rdarea", mu_mean, 0.0035),
+    mu_lo = pclip(param, "Rdarea", mu_lo, 0.0035),
     mu_hi = pclip(param, "Parea", mu_hi, 0.285),
@@
 ## Relative uncertainty
-cv_dat <- plot_dat %>%
+cv_dat <- means_dat %>%
   mutate(
@@
   ) %>%
-  filter(sample_size > 3)
+  filter(sample_size > 1)
 
@@
     cvfit = map(data, ~lm(log10(cv) ~ log10(sample_size), data = .))
+  )
+
+cv_coef <- cv_mod %>%
+  mutate(
+    coef = map(cvfit, broom::tidy)
+  ) %>%
+  unnest(coef) %>%
+  select(model_type, term, estimate) %>%
+  spread(term, estimate) %>%
+  "names<-"(c("model_type", "intercept", "slope")) %>%
+  mutate(
+    eqn = sprintf("'%s: '~log[10](y) == %0.2f - %0.2f~log[10](x)", model_type, intercept, abs(slope))
   )
@@
   geom_line(data = cv_pred) +
+  annotate("text", x = rep(2500, 3), y = c(1.4, 1.25, 1.1), 
+           label = cv_coef$eqn, parse = TRUE, 
+           size = 2, vjust = "right") +
   scale_x_continuous(trans = "log10") +
@@
   par_legplot = list(mar = c(0, 0, 0, 0)),
-  par_legend = list(ncol = 5, x = "center", cex = 0.6)
+  par_legend = list(ncol = 5, x = "center", cex = 0.6),
+  reorder_legend = TRUE
 )
@@
 ```{r boxprep}
-hier_mass_file <- "../output/hier.mass.clm45.NA.2017-11-09-1749.rds"
-hier_area_file <- "../output/hier.area.clm45.NA.2017-11-10-0937.rds"
-multi_mass_file <- "../output/multi.mass.clm45.NA.2017-11-09-1733.rds"
-multi_area_file <- "../output/multi.area.clm45.NA.2017-11-09-1741.rds"
+hier_mass_file <- "../output/hier.mass.clm45.NA.2017-11-13-1250.rds"
+hier_area_file <- "../output/hier.area.clm45.NA.2017-11-13-1312.rds"
+multi_mass_file <- "../output/multi.mass.clm45.NA.2017-11-13-1231.rds"
+multi_area_file <- "../output/multi.area.clm45.NA.2017-11-13-1245.rds"
 

### 2017-11-13 scripts/pft_abbreviations.R
Add broadleaf deciduous boreal to PFT list

@@ -4,6 +4,7 @@ abbr2pft <- c(
   BlETe = "broadleaf_evergreen_temperate",
   BlDTr = "broadleaf_deciduous_tropical",
   BlDTe = "broadleaf_deciduous_temperate",
+  BlDBo = "broadleaf_deciduous_boreal",
   NlETe = "needleleaf_evergreen_temperate",
   NlEBo = "needleleaf_evergreen_boreal",
   NlD = "needleleaf_deciduous",
@@ -103,5 +104,21 @@ param_fancy <- sapply(param_fancy_chr, formula)
 #)
 
 source("scripts/paultolcolors.R")
-pft_colors <- tol21rainbow[c(1, 3, 4, 6, 7, 8, 9, 13:15, 16:18)]
+pft_inds <- c(
+  BlETr = 1,
+  BlETe = 3,
+  BlDTr = 4,
+  BlDTe = 5,
+  BlDBo = 6,
+  NlETe = 7,
+  NlEBo = 8,
+  NlD = 9,
+  ShE = 13,
+  ShDTe = 14,
+  ShDBo = 15,
+  C3GAr = 16,
+  C3GTe = 17,
+  C4G = 18
+)
+pft_colors <- tol21rainbow[pft_inds]
 #pft_colors <- tol14rainbow[-c(10, 9)]

### 2017-11-13 scripts/explore_prior.R
Update priors

@@ -29,16 +29,13 @@ try_long <- try_data %>%
   mutate(value = log10(value))
 
 source('scripts/informative_prior.R')
-
 prior_long <- prior_df %>%
   mutate(samples = map2(mean, stdev, ~rnorm(5000, .x, .y))) %>%
   select(key = param, value = samples) %>%
   unnest()
-
 plot_df <- bind_rows(try_long %>% mutate(type = 'data'),
                      prior_long %>% mutate(type = 'prior'))
-
 ggplot(plot_df) +
-  aes(x = key, y = value, fill = type) +
-  geom_violin() +
+  aes(x = key, y = 10^value, fill = type) +
+  geom_boxplot() +
   facet_wrap(~key, scales = "free")

### 2017-11-13 scripts/informative_prior.R

@@ -1,17 +1,17 @@
 prior_df <- tribble(
   ~param, ~mean, ~stdev,
-  'Jmax_area', 2, 0.26,
-  'Jmax_mass', 0, 0.26,
-  'leaf_lifespan', 0.75, 0.4,
-  'Narea', 0.25, 0.3,
-  'Nmass', 1.25, 0.3,
-  'Parea', -1, 0.3,
-  'Pmass', 0, 0.3,
-  'Rdarea', 0, 0.27,
-  'Rdmass', -2, 0.26,
-  'SLA', 1, 0.6,
-  'Vcmax_mass', -0.4, 0.26,
-  'Vcmax_area', 1.7, 0.19
+  "Jmax_area", 2, 0.19,
+  "Jmax_mass", 0, 0.26,
+  "leaf_lifespan", 0.8, 0.4,
+  "Narea", 0.25, 0.3,
+  "Nmass", 1.25, 0.3,
+  "Parea", -1, 0.3,
+  "Pmass", 0, 0.3,
+  "Rdarea", log10(0.0011), 0.15,
+  "Rdmass", log10(0.8), 0.20,
+  "SLA", 1, 0.6,
+  "Vcmax_mass", log10(0.45), 0.2,
+  "Vcmax_area", log10(50), 0.18
 ) %>%
   mutate(variance = stdev ^ 2)
 

### 2017-11-14 manuscript-md/03_results.Rmd
Cleanup latest round of results

@@
 
+censor <- function(rxp, pft, pfts, param, value) {
+  if_else(grepl(rxp, param) & pfts %in% pft, "---", value)
+}
+
 table_dat <- means_dat %>%
@@
     mu_hi = if_else(param == "Rdmass", mu_hi * 1000, mu_hi),
+    mu_lo = if_else(param == "Rdarea", mu_lo * 1000, mu_lo),
+    mu_mean = if_else(param == "Rdarea", mu_mean * 1000, mu_mean),
+    mu_hi = if_else(param == "Rdarea", mu_hi * 1000, mu_hi),
     lo_f = formatC(mu_lo, digits = 3),
@@
     value = sprintf("%s (%s,%s)", mid_f, lo_f, hi_f),
-    value = if_else(grepl("Vcmax|Jmax", param) & pft == "C4G", "---", value)
+    value = censor("Vcmax|Jmax", "C4G", pft, param, value),
+    value = censor("Rdmass", "ShDBo", pft, param, value),
+    value = censor("Jmax_mass", c("NlEBo", "NlD"), pft, param, value),
+    value = censor("Jmax_area", c("NlD"), pft, param, value)
   ) %>%
@@
     mu_hi = pclip(param, "Rdmass", mu_hi, 0.03),
-    mu_hi = pclip(param, "Rdarea", mu_hi, 0.0035),
-    mu_mean = pclip(param, "Rdarea", mu_mean, 0.0035),
-    mu_lo = pclip(param, "Rdarea", mu_lo, 0.0035),
+    mu_mean = pclip(param, "Rdmass", mu_mean, 0.03),
+    mu_lo = pclip(param, "Rdmass", mu_lo, 0.03),
+    mu_hi = pclip(param, "Rdarea", mu_hi, 0.0025),
+    mu_mean = pclip(param, "Rdarea", mu_mean, 0.0025),
+    mu_lo = pclip(param, "Rdarea", mu_lo, 0.0025),
     mu_hi = pclip(param, "Parea", mu_hi, 0.285),
@@
 ```{r boxprep}
-hier_mass_file <- "../output/hier.mass.clm45.NA.2017-11-13-1250.rds"
-hier_area_file <- "../output/hier.area.clm45.NA.2017-11-13-1312.rds"
-multi_mass_file <- "../output/multi.mass.clm45.NA.2017-11-13-1231.rds"
-multi_area_file <- "../output/multi.area.clm45.NA.2017-11-13-1245.rds"
+outdir <- "../output"
+hier_mass_file <- tail(dir(outdir, "hier.mass.clm45", full.names = TRUE), 1)
+hier_area_file <- tail(dir(outdir, "hier.area.clm45", full.names = TRUE), 1)
+multi_mass_file <- tail(dir(outdir, "multi.mass.clm45", full.names = TRUE), 1)
+multi_area_file <- tail(dir(outdir, "multi.area.clm45", full.names = TRUE), 1)
 

### 2017-11-14 manuscript-md/02_methods.Rmd
Fix trait selection in methods
Add interpretation to figure captions
Also, minor stylistic and grammatical revisions to address specific
reviewer comments.

@@
 try_sub <- try_data %>%
-  select(pft = clm45, Jmax_area:Vcmax_mass,
+  select(pft = clm45, one_of(both_params),
          -Latitude, -Longitude, -LMA)
@@
 
-To assess the consistency of among- and with-PFT trait trade-offs, we calculated the eigenvalues from pairwise variance-covariance matrix estimates for each trait pair and, where the covariance was significantly different from zero ($p < 0.05$) plotted the dominant eigenvectors centered on the mean estimates (Fig.
+To assess the consistency of among- and within-PFT trait trade-offs, we calculated the eigenvalues from pairwise variance-covariance matrix estimates for each trait pair and, where the covariance was significantly different from zero ($p < 0.05$) plotted the dominant eigenvectors centered on the mean estimates (Fig.
 \@ref(fig:stickpairs)).

### 2017-11-14 manuscript-md/03_results.Rmd
Add interpretation to figure captions
Also, minor stylistic and grammatical revisions to address specific
reviewer comments.

@@
 Lines represent linear models ($\log(y) = b_0 + b_1 \log(x)$) fit independently for each model type.
+In general, differences in estimate uncertainty between the univariate and multivariate models were minimal at large sample sizes but increasingly important at low sample sizes.
+However, differences in estimate uncertainty between the multivariate and hierarchical models were consistently negligible.
 
@@
 (ref:stickcap) Pairwise trait mean and covariance estimates for all data pooled globally (black) and for each PFT (colored).
-Correlations not significantly different from zero are indicated by x symbols at the mean estimate.
+Covariance estimates not significantly different from zero ($p < 0.05$) are indicated by x symbols at the mean estimate.
 $x$ and $y$ axes vary on a log scale, reflecting the fact that the model was fit using the base 10 log of all traits.
+With the exception of leaf lifespan, pairwise covariances are consistent in direction but vary somewhat in magnitude between PFTs, and when comparing PFT-level and global estimates.
+However, many pairwise covariances are not statistically significant, particularly (but not always) for undersampled traits and PFTs.
 
@@
 for all data pooled globally (dark grey) and for each PFT (colored).
+For most PFT-trait pairs, correlations are mutually consistent in magnitude but vary in strength.
 

### 2017-11-15 manuscript-md/02_methods.Rmd
Complete first pass at reviewer response letter
Also, some associated revisions to to the manuscript.

@@
 
-For each model fit, we ran five parallel MCMC chains, automatically increasing the sample size until the final result achieved convergence as determined by a Gelman-Rubin potential scale reduction statistic less than 1.05 (Gelman and Rubin 1992).
+For each model fit, we ran five parallel MCMC chains, automatically increasing the sample size until the final result achieved convergence as determined by a Gelman-Rubin potential scale reduction statistic less than 1.1 [@gelman_1992_inference].
 
-We implemented this sampling algorithm in R (version 3.4.2; @rstats).
-The code for fitting these models to data is available as an open source R package (`mvtraits`) at <http://github.com/ashiklom/mvtraits>.
+We implemented this sampling algorithm in R [version 3.4.2; @rstats].
+The code for fitting these models to arbitrary data is available as an open source R package (`mvtraits`) at <http://github.com/ashiklom/mvtraits>.
 
@@
 To assess the impact of multivariate and hierarchical constraint on trait estimates,
-we compared the mean and size of the 95% confidence intervals of mean estimates of each trait for each PFT (Fig. \@ref(fig:meancompare)).
-In addition, we report the mean and 95% confidence intervals of mean estimates for each trait and PFT from the hierarchical model in tabular form (Tab. 2 and 3). 
-To test the hypothesis that the multivariate and hierarchical models offer more value in terms of uncertainty constraint at smaller sample sizes, we calculated the relative uncertainty ($\alpha$) as a function of the mean ($\mu$) and upper ($q_{0.975}$) and lower ($q_{0.025}$) confidence limits of trait estimates as...
+we compared the mean and size of the 95% confidence intervals of mean estimates of each trait for each PFT from each model (Fig.~\@ref(fig:meancompare)).
+In addition, we report the mean and 95% confidence intervals of mean estimates for each trait and PFT from the hierarchical model in tabular form (Tab.~2 and 3).
+
+To test the hypothesis that the multivariate and hierarchical models offer more value in terms of uncertainty constraint at smaller sample sizes, we calculated the relative uncertainty ($\alpha$) as a function of the mean ($\mu$) and upper ($q_{0.975}$) and lower ($q_{0.025}$) confidence limits of trait estimates.
 
@@
 
-We then fit a generalized linear model relating relative uncertainty to sample size ($n$) for each of the model types (univariate, multivariate, and hierarchical; Fig. \@ref(fig:civssamplesize)) as...
+We then fit a generalized linear model relating relative uncertainty to sample size ($n$) for each of the model types (univariate, multivariate, and hierarchical; Fig. \@ref(fig:civssamplesize)).
 
@@
 
-To assess the consistency of among- and within-PFT trait trade-offs, we calculated the eigenvalues from pairwise variance-covariance matrix estimates for each trait pair and, where the covariance was significantly different from zero ($p < 0.05$) plotted the dominant eigenvectors centered on the mean estimates (Fig.
-\@ref(fig:stickpairs)).
-Due to the small number of points used to estimate across-PFT covariance in the hierarchical model, none of the global correlations estimated in the hierarchical model were significantly different from zero ($p < 0.05$).
+If all three models perform equally well at all sample sizes, their respective slope coefficients would be statistically indistinguishable.
+On the other hand, models that perform better should have lower slope ($b_1$) coefficients, indicating a slower increase in the uncertainty ($\alpha$) with decreasing sample size ($n$).
+
+To assess the consistency of among- and within-PFT trait trade-offs, we calculated the eigenvalues from pairwise variance-covariance matrix estimates for each trait pair and, where the covariance was significantly different from zero ($p < 0.05$), plotted the dominant eigenvectors centered on the mean estimates (Fig.~\@ref(fig:stickpairs)).
+This figure provides a visual representation of relative positions of PFTs in trait space and both the direction and extent of within-PFT trait covariance, and is directly analogous to conceptual figures describing hierarchical trait variability across environmental gradients as presented in, for instance, @cornwell_community_2009 and @albert_intraspecific_2010.
+Due to the small number of points used to estimate across-PFT covariance in the hierarchical model, none of the across-PFT covariances estimated in the hierarchical model were significantly different from zero ($p < 0.05$).
 For this reason, we compared within-PFT covariances as estimated by the hierarchical model with the covariances estimated by fitting a multivariate model to all of the data.
 
-The R code for all analyses is publicly available online at <http://github.com/ashiklom/np_trait_analysis>.
+Besides the consistency in the direction of trait covariance globally and between different PFTs, we also investigated the strength and predictive power of these covariances, which is represented by correlation coefficients (i.e.
+the pairwise covariance normalized to the variances of the component variables).
+To do this, we plotted the mean and 95% confidence interval of the pairwise trait correlation coefficients for the global estimate from the pooled multivariate model and PFT-level estimates from the hierarchical model (Fig.~\@ref(fig:corrbars)).
+
+The R code for running these analyses is publicly available online at <http://github.com/ashiklom/np_trait_analysis>.
 The TRY data used for this analysis can be requested at <http://try-db.org>.

### 2017-11-16 scripts/getbibtex.sh
Minor revisions to introduction

@@ -1,0 +1,5 @@
+#!/bin/bash
+
+while read line; do
+    doi2bibtex $line
+done < scripts/ref_dois

### 2017-11-16 scripts/process_references.R

@@ -1,0 +1,55 @@
+library(tidyverse)
+library(rcrossref)
+library(stringi)
+
+try_refs <- src_sqlite("../preprocess-try/try.sqlite") %>%
+  tbl("orig_citations") %>%
+  collect() %>%
+  right_join(read_csv("../preprocess-try/traits/references.csv"))
+
+try_refs_proc <- try_refs %>%
+  mutate(
+    reference_proc = stri_enc_toascii(Reference)
+  ) %>%
+  select(reference_proc, n, ReferenceID) %>%
+  filter(!grepl("unpub.|Unpublished literature compilation",
+                reference_proc, ignore.case = TRUE))
+
+crossref_results <- list()
+for (i in seq_len(nrow(try_refs_proc))) {
+  message(i, " of ", nrow(try_refs_proc))
+  qry <- try_refs_proc$reference_proc[i]
+  result <- cr_works(query = try_refs_proc$reference_proc[i], limit = 1)
+  print(result$data$DOI)
+  crossref_results[[i]] <- result
+}
+
+custom_dois <- c(
+  `39` = "10.3334/ORNLDAAC/1106",
+  `40` = "10.1890/ES10-00175.1",
+  `43` = "10.3334/ORNLDAAC/1104",
+  `113` = "10.1139/cjb-2012-0162",
+  `116` = "10.3334/ORNLDAAC/1097"
+)
+
+crossref_dois <- map_chr(crossref_results, c("data", "DOI"), .default = NA)
+crossref_dois[as.numeric(names(custom_dois))] <- unname(custom_dois)
+keys <- sprintf("try_data_%02d", seq_along(crossref_dois))
+keyrefs <- paste0("@", keys)
+
+crossref_table_raw <- tibble(DOI = crossref_dois, key = keys)
+
+write.table(
+  crossref_table_raw,
+  file = "scripts/ref_dois",
+  quote = FALSE,
+  sep = " -k ",
+  row.names = FALSE,
+  col.names = FALSE
+)
+
+crossref_table <- try_refs_proc %>%
+  mutate(Reference = keyrefs) %>%
+  select(Reference, Observations = n)
+
+saveRDS(crossref_table, file = "extdata/crossref_table.rds")

### 2017-11-16 manuscript-md/02_methods.Rmd
Finished manuscript draft

@@
 
-Foliar trait data for this analysis comes from the TRY global traits database
-(@kattge_try_2011; see Supplementary Table 1 for a complete table of component datasets and references.).
+Foliar trait data for this analysis comes from the TRY global traits database (@kattge_try_2011; Tab.\ \@ref(tab:trydata)).
 We focused our research on seven foliar traits:
@@
 and maximum electron transport rate ($J_{max,mass}$, µmol g$^{-1}$ s$^{-1}$, or $J_{max,area}$, µmol m$^{-2}$ s$^{-1}$).
-We performed analyses separately for both mass- and area-normalized traits to avoid artifacts caused by different trait normalization [@osnas_global_2013; @lloyd_les].
-We restricted our analysis to TRY data that have been quality-controlled and for which adequate species information was provided for functional type classification (see @kattge_try_2011).
+We performed analyses separately for both mass- and area-normalized traits to avoid potential artifacts caused by different trait normalization [@osnas_global_2013; @lloyd_les].
+We restricted our analysis to TRY data that have been quality-controlled and for which adequate species information was provided for functional type classification [see @kattge_try_2011].
 
@@
 
-Following past studies [e.g. @wright_worldwide_2004; @wright_assessing_2005; @diaz_global_2016], we log-transformed all trait values to correct for their strong right-skewness.
+Following past studies [e.g. @wright_worldwide_2004; @wright_assessing_2005; @onoda_2011_global; @diaz_global_2016], we log-transformed all trait values to correct for their strong right-skewness.
+
+```{r "trydata"}
+crtable <- readRDS("extdata/crossref_table.rds") %>%
+  mutate(
+    Reference = case_when(
+      Reference == "@try_data_03" ~ "@baraloto_decoupled_2010",
+      Reference == "@try_data_09" ~ "@wright_worldwide_2004",
+      Reference == "@try_data_14" ~ "@kattge_2009_vcmax",
+      Reference == "@try_data_15" ~ "@laughlin_multi-trait_2010",
+      Reference == "@try_data_26" ~ "@onoda_2011_global",
+      Reference == "@try_data_30" ~ "@messier_how_2010",
+      Reference == "@try_data_33" ~ "@pierce_allocating_2013",
+      Reference == "@try_data_41" ~ "@poorter_causes_2009",
+      Reference == "@try_data_44" ~ "@choat_global_2012",
+      Reference == "@try_data_59" ~ "@laughlin_climatic_2011",
+      Reference == "@try_data_67" ~ "@wright_does_2012",
+      Reference == "@try_data_80" ~ "@choat_global_2012",
+      Reference == "@try_data_114" ~ "@kichenin_2013_contrasting",
+      Reference == "@try_data_134" ~ "@diaz_plant_2004",
+      TRUE ~ Reference
+    )
+  )
+cap <- "Data sources and observation counts for foliar trait data"
+kable(crtable, caption = cap)
+```
 
@@
 
-We assigned each species to a unique plant functional type (PFT) following the scheme in the Community Land Model (CLM4.5, @clm45_note; Tab. \@ref(tab:pfts), Fig. \@ref(fig:samplesize)).
+We assigned each species to a unique plant functional type (PFT) following the scheme in the Community Land Model [CLM4.5, @clm45_note; Tab.\ \@ref(tab:pfts), Fig.\ \@ref(fig:samplesize)].
 We obtained categorical data on growth form, leaf type, phenology, and photosynthetic pathway from the TRY database.
-Where the same species was reported to have different attributes in different datasets, we assigned the attribute that was observed most frequently, and, in cases of ties, selected the attribute that was more likely based on phylogenetically similar species.
+Where the same species was reported to have different attributes in different datasets, we assigned the attribute that was observed most frequently, and, in cases of ties, selected the attribute that was more likely based on phylogenetically similar species or more likely to occur in nature.
 For biome specification, we collected all latitude and longitude data for each species,
-matched these data to 30 second ($\sim1 km^{2}$ ) mean annual temperature ($AMT$, averaged 1970-2000) data from WorldClim-2 [@worldclim],
-calculated the mean temperature for each species,
+matched these data to 30 second ($\sim1 km^{2}$) mean annual temperature ($AMT$, averaged 1970-2000) data from WorldClim-2 [@worldclim],
+calculated the mean AMT for all sites at which that species was observed,
 and then binned these species based on the following cutoffs: boreal/arctic ($AMT \leq 5^\circ C$), temperate ($AMT \leq 20^\circ C$), and tropical ($AMT > 20^\circ C$).
@@
 
-The fundamentally multivariate nature of the sampling procedure described above makes it incapable of accommodating missing values in the data ($x$).
-Therefore, our algorithm also included imputation of partially missing data, which proceeded as follows: For a block of data $\mathbf{x'}$ containing missing observations in the columns $\mathbf{m}$ and present observations in columns $\mathbf{p}$...
+The fundamentally multivariate nature of the sampling procedure described above makes it incapable of accommodating partially missing observations.
+Therefore, our algorithm also included imputation of partially missing data, which proceeded as follows:
+For a block of data $\mathbf{x'}$ containing missing observations in columns $\mathbf{m}$ and present observations in columns $\mathbf{p}$,
+the missing values $\mathbf{x'}[m]$ are drawn randomly from a marginal multivariate normal distribution at each iteration of the sampling algorithm...
 
@@
 For each model fit, we ran five parallel MCMC chains, automatically increasing the sample size until the final result achieved convergence as determined by a Gelman-Rubin potential scale reduction statistic less than 1.1 [@gelman_1992_inference].
-
-We implemented this sampling algorithm in R [version 3.4.2; @rstats].
-The code for fitting these models to arbitrary data is available as an open source R package (`mvtraits`) at <http://github.com/ashiklom/mvtraits>.
+We implemented this sampling algorithm in an open source, publicly available R [version 3.4.2; @rstats] package (<http://github.com/ashiklom/mvtraits>).
 
@@
 To assess the impact of multivariate and hierarchical constraint on trait estimates,
-we compared the mean and size of the 95% confidence intervals of mean estimates of each trait for each PFT from each model (Fig.~\@ref(fig:meancompare)).
-In addition, we report the mean and 95% confidence intervals of mean estimates for each trait and PFT from the hierarchical model in tabular form (Tab.~2 and 3).
+we compared the mean and size of the 95% confidence intervals of mean estimates of each trait for each PFT from each model (Fig.\ \@ref(fig:meancompare)).
+In addition, we report the mean and 95% confidence intervals of mean estimates for each trait and PFT from the hierarchical model in tabular form (Tab.\ 3 and 4).
 
@@
 
-If all three models perform equally well at all sample sizes, their respective slope coefficients would be statistically indistinguishable.
+If all three models performed equally well at all sample sizes, their respective slope coefficients would be statistically indistinguishable.
 On the other hand, models that perform better should have lower slope ($b_1$) coefficients, indicating a slower increase in the uncertainty ($\alpha$) with decreasing sample size ($n$).
 
-To assess the consistency of among- and within-PFT trait trade-offs, we calculated the eigenvalues from pairwise variance-covariance matrix estimates for each trait pair and, where the covariance was significantly different from zero ($p < 0.05$), plotted the dominant eigenvectors centered on the mean estimates (Fig.~\@ref(fig:stickpairs)).
+To assess the consistency of among- and within-PFT trait trade-offs, we calculated the eigenvalues from pairwise variance-covariance matrix estimates for each trait pair and, where the covariance was significantly different from zero ($p < 0.05$), plotted the dominant eigenvectors centered on the mean estimates (Fig.\ \@ref(fig:stickpairs)).
 This figure provides a visual representation of relative positions of PFTs in trait space and both the direction and extent of within-PFT trait covariance, and is directly analogous to conceptual figures describing hierarchical trait variability across environmental gradients as presented in, for instance, @cornwell_community_2009 and @albert_intraspecific_2010.
@@
 the pairwise covariance normalized to the variances of the component variables).
-To do this, we plotted the mean and 95% confidence interval of the pairwise trait correlation coefficients for the global estimate from the pooled multivariate model and PFT-level estimates from the hierarchical model (Fig.~\@ref(fig:corrbars)).
+To do this, we plotted the mean and 95% confidence interval of the pairwise trait correlation coefficients for the global estimate from the pooled multivariate model and PFT-level estimates from the hierarchical model (Fig.\ \@ref(fig:corrbars)).
 

### 2017-11-16 manuscript-md/03_results.Rmd
Finished manuscript draft

@@
 
-In general, leaf trait estimates from the univariate, multivariate, and hierarchical models were similar (Fig. \@ref(fig:meancompare)).
-Where we observed differences between modles, the largest were between the univariate and multivariate models, while the additional constraint from the hierarchical model tended to have a minimal effect on trait estimates.
+In general, leaf trait estimates from the univariate, multivariate, and hierarchical models were similar (Fig.\ \@ref(fig:meancompare)).
+Where we observed differences between models, the largest were between the univariate and multivariate models, while the additional constraint from the hierarchical model tended to have a minimal effect on trait estimates.
 Significant differences in trait estimates between univariate and multivariate models occurred even for traits with relatively large sample sizes, such as leaf nitrogen content.
@@
 Among broadleaved trees, deciduous trees had higher $V_{c,max,mass}$ and slightly higher $V_{c,max,area}$ than evergreen trees.
+Between the deciduous and evergreen tree PFTs, we observed no significant trend by climate zone.
 
-C3 grasses also had the highest $J_{max,area}$, but temperate broadleaved deciduous trees had the highest $J_{max,mass}$.
+C3 grasses and temperate needleleaved evergreen trees had the highest $J_{max,area}$, but temperate broadleaved deciduous trees had the highest $J_{max,mass}$.
 All of the shrub PFTs had the lowest $J_{max,mass}$ but average or above-average $J_{max,area}$, while the opposite was true of broadleaved tropical PFTs.
-Of the tree PFTs, needleleaved evergreen trees had among the highest $J_{max,area}$ but the lowest $J_{max,mass}$.
+Of the tree PFTs, needleleaved evergreen trees had the highest $J_{max,area}$ but the lowest $J_{max,mass}$.
 
@@
 All of the high-latitude PFTs consistently had among the largest error bars around their mean estimates relative to other PFTs, while the traits with the largest uncertainties were dark respiration, $V_{c,max}$, and $J_{max}$.
-For some of these trait-PFT combinations, such as estimates $J_{max,mass}$ for shrub PFTs, the additional constraint from trait covariance substantially reduced error bars, making it possible to compare estimates against those of other PFTs.
+For many of these trait-PFT combinations, the additional constraint from trait covariance provided by the multivariate and hierarchical models substantially reduced error bars, making it possible to compare estimates against those of other PFTs.
 Our analysis of the relationship between model type, sample size, and estimate relative uncertainty found that this covariance-based constraint from the multivariate model becomes increasingly important at smaller sample sizes (\@ref(fig:civssamplesize)).
-However, this analysis revealed no consistent significant benefit from the hierarchical model for our specific analysis.
+However, this analysis revealed no consistent significant benefit from the hierarchical model.
 
@@
 
-Where pairwise trait correlations were statistically significant, these correlations were generally consistent in direction both globally and within each PFT (Fig. \@ref(fig:stickpairs)).
+Where pairwise trait correlations were statistically significant, these correlations were generally consistent in direction both globally and within each PFT (Fig.\ \@ref(fig:stickpairs)).
 In particular, mass- and area-normalized traits were all positively correlated with each other and, respectively, positively and negatively correlated with SLA, both globally and within each PFT.
-On the other hand, correlations between leaf lifespan and area-normalized traits were more variable.
-The correlation between $N_{area}$ and leaf lifespan was positive both globally and for evergreen shrubs, but was negative for broadleaved deciduous trees and not significant for any other PFTs.
+The same was generally true of correlations of mass-based traits with leaf lifespan, but correlations of leaf lifespan with area-normalized traits were more variable.
+The correlation between $N_{area}$ and leaf lifespan was positive globally and for evergreen shrubs but negative for broadleaved deciduous trees and not significant for any other PFTs.
 Similarly, the correlation between $P_{area}$ and leaf lifespan was positive for temperate broadleaved evergreen trees and needleleaved evergreen trees but negative for evergreen shrubs.
+The correlation between leaf lifespan and $R_{d,area}$ significant and positive globally, but was not significant within any PFTs.
+The correlations of leaf lifespan with $V_{c,max,area}$ and $J_{max,area}$ were negative for temperate broadleaved deciduous trees, but not significant for any other PFTs or globally.
 
 A large number of pairwise trait correlations were not significant.
-In some cases, this was driven by sample size (Fig. \@ref(fig:samplesize)).
+In some cases, this was driven by sample size (Fig.\ \@ref(fig:samplesize)).
 For instance, needleleaved deciduous trees were often the only PFT for which a correlation was not statistically significant.
@@
 
-The strength of pairwise trait correlations varied substantially depending on scale, PFT, and trait (Fig. \@ref(fig:corrbars)).
+The strength of pairwise trait correlations varied substantially depending on scale, PFT, and trait (Fig.\ \@ref(fig:corrbars)).
 The two pairwise trait correlations that exhibited the most consistent strength globally and within each PFT were the correlation between SLA and $N_{area}$, and between $N_{mass}$ and $P_{mass}$.
-Correlation strength was often related to sample size, with well-sampled PFTs exhibiting stronger correlations and undersampled PFTs exhibiting weaker correlations.
+Correlation strength was often, but not always, related to sample size, with well-sampled PFTs exhibiting stronger correlations and undersampled PFTs exhibiting weaker correlations.

### 2017-11-16 manuscript-md/index.Rmd
Finished manuscript draft

@@
   - extra.bib
+  - tryrefs.bib
 csl: new-phytologist.csl
@@
 
-* We investigated whether global leaf economic relationships across plant functional types (PFTs) can be reconciled with within-PFT relationships, and the extent to which this hierarchical structure can be used to constrain trait estimates.
+* We investigated whether global leaf economic relationships across plant functional types (PFTs) area also present within-PFTs, and the extent to which this hierarchical structure can be used to constrain trait estimates.
 
@@
 * Leveraging covariance in multivariate models reduced uncertainties in mean trait estimates, particularly for undersampled trait-PFT combinations.
-Sometimes, the multivariate model significantly shifted trait means, resulting in estimates better aligned with patterns predicted from global ecology and biogeography.
 However, additional constraint from the across-PFT hierarchy was limited compared to the simple multivariate models.

### 2017-11-17 manuscript-md/index.Rmd
Add line numbers to PDF output

@@
     toc: no
+    includes:
+      in_header: preamble.tex
 ---

### 2017-11-17 manuscript-md/03_results.Rmd
Clean up Makefile. Minor fig ref fix in results

@@
 For many of these trait-PFT combinations, the additional constraint from trait covariance provided by the multivariate and hierarchical models substantially reduced error bars, making it possible to compare estimates against those of other PFTs.
-Our analysis of the relationship between model type, sample size, and estimate relative uncertainty found that this covariance-based constraint from the multivariate model becomes increasingly important at smaller sample sizes (\@ref(fig:civssamplesize)).
+Our analysis of the relationship between model type, sample size, and estimate relative uncertainty found that this covariance-based constraint from the multivariate model becomes increasingly important at smaller sample sizes (Fig.\ \@ref(fig:civssamplesize)).
 However, this analysis revealed no consistent significant benefit from the hierarchical model.

### 2018-01-03 manuscript-md/index.Rmd
Remove Sabina from author list

--- 2017-11-16 manuscript-md/index.Rmd
@@
   Michael Bahn$^2$,
-  Sabina Burrascano$^3$,
-  Chaeho Byun$^4$,
-  Joseph Craine$^5$,
-  Andrés Gonzalez-Melo$^6$,
-  Alvaro G. Gutiérrez$^7$,
-  Steven Jansen$^8$,
-  Nathan Kraft$^9$,
-  Koen Kramer$^{10}$,
-  Vanessa Minden$^{11,12}$
-  Ülo Niinemets$^{13}$,
-  Yusuke Onoda$^{14}$,
-  Enio Egon Sosinski$^{15}$,
-  Nadejda A. Soudzilovskaia$^{16}$,
+  Chaeho Byun$^3$,
+  Joseph Craine$^4$,
+  Andrés Gonzalez-Melo$^5$,
+  Alvaro G. Gutiérrez$^6$,
+  Steven Jansen$^7$,
+  Nathan Kraft$^8$,
+  Koen Kramer$^{9}$,
+  Vanessa Minden$^{10,11}$
+  Ülo Niinemets$^{12}$,
+  Yusuke Onoda$^{13}$,
+  Enio Egon Sosinski$^{14}$,
+  Nadejda A. Soudzilovskaia$^{15}$,
   Michael C. Dietze$^1$
@@
 $^2$Institute of Ecology, University of Innsbruck, 6020 Innsbruck, Austria;
-$^3$Department of Environmental Biology, Sapienza University of Rome, Piazzale Aldo Moro 5, 00185, Rome, Italy;
-$^4$School of Biological Sciences, Seoul National University, Gwanak-ro 1, 08826, Seoul, Seoul, South Korea;
-$^5$Jonah Ventures, Manhattan, KS 66502, USA;
-$^6$Universidad del Rosario. Facultad de Ciencias Naturales y Matemáticas. Carrera 26 #63B-48. Bogotá, Colombia;
-$^7$Departamento de Ciencias Ambientales y Recursos Naturales, Facultad de Ciencias Agronomicas, Universidad de Chile, Av. Santa Rosa 11315, La Pintana 8820808, Santiago, Chile;
-$^8$Institute of Systematic Botany and Ecology, Ulm University, Ulm, Germany;
-$^9$Department of Ecology and Evolutionary Biology, University of California Los Angeles, Los Angeles, CA 90095, USA;
-$^{10}$Vegetation, Forest, and Landscape Ecology, Wageningen Environmental Research and Wageningen University, Wageningen, Gelderland, The Netherlands;
-$^{11}$Institute for Biology and Environmental Sciences, Carl von Ossietzky-University of Oldenburg, Carl von Ossietzky Str. 9-11, 26129 Oldenburg, Germany;
-$^{12}$Department of Biology, Ecology and Evolution, Vrije Universiteit Brussel, Pleinlaan 2, 1050 Brussels;
-$^{13}$Institute of Agricultural and Environmental Sciences, Estonian University of Life Sciences, Kreutzwaldi 1, 51014 Tartu, Estonia;
-$^{14}$Graduate School of Agriculture, Kyoto University, Kyoto, 606-8502, Japan;
-$^{15}$Embrapa Temperate Agriculture, Rodovia BR 392, Km 78, CEP 96010-971 Pelotas, RS, Brazil;
-$^{16}$Conservation Biology Department, Institute of Environmental Sciences, Leiden University, Leiden, The Netherlands
+$^3$School of Biological Sciences, Seoul National University, Gwanak-ro 1, 08826, Seoul, Seoul, South Korea;
+$^4$Jonah Ventures, Manhattan, KS 66502, USA;
+$^5$Universidad del Rosario. Facultad de Ciencias Naturales y Matemáticas. Carrera 26 #63B-48. Bogotá, Colombia;
+$^6$Departamento de Ciencias Ambientales y Recursos Naturales, Facultad de Ciencias Agronomicas, Universidad de Chile, Av. Santa Rosa 11315, La Pintana 8820808, Santiago, Chile;
+$^7$Institute of Systematic Botany and Ecology, Ulm University, Ulm, Germany;
+$^8$Department of Ecology and Evolutionary Biology, University of California Los Angeles, Los Angeles, CA 90095, USA;
+$^{9}$Vegetation, Forest, and Landscape Ecology, Wageningen Environmental Research and Wageningen University, Wageningen, Gelderland, The Netherlands;
+$^{10}$Institute for Biology and Environmental Sciences, Carl von Ossietzky-University of Oldenburg, Carl von Ossietzky Str. 9-11, 26129 Oldenburg, Germany;
+$^{11}$Department of Biology, Ecology and Evolution, Vrije Universiteit Brussel, Pleinlaan 2, 1050 Brussels;
+$^{12}$Institute of Agricultural and Environmental Sciences, Estonian University of Life Sciences, Kreutzwaldi 1, 51014 Tartu, Estonia;
+$^{13}$Graduate School of Agriculture, Kyoto University, Kyoto, 606-8502, Japan;
+$^{14}$Embrapa Temperate Agriculture, Rodovia BR 392, Km 78, CEP 96010-971 Pelotas, RS, Brazil;
+$^{15}$Conservation Biology Department, Institute of Environmental Sciences, Leiden University, Leiden, The Netherlands
 

### 2018-01-03 manuscript-md/index.Rmd
Add Mike's suggestions to summary

@@
 
-* We investigated whether global leaf economic relationships across plant functional types (PFTs) area also present within-PFTs, and the extent to which this hierarchical structure can be used to constrain trait estimates.
+* We investigated whether global leaf economic relationships across plant functional types (PFTs) are also present within-PFTs, and the extent to which this hierarchical structure can be used to constrain trait estimates.
 
@@
 
+* Trait correlations were generally consistent in direction within and across PFTs, and consistent with predictions of the leaf economic spectrum.
+However, correlation strength varied substantially across PFTs indicating that leaf economic relationships within PFTs are often confounded by other biotic and abiotic factors.
+
 * Leveraging covariance in multivariate models reduced uncertainties in mean trait estimates, particularly for undersampled trait-PFT combinations.
 However, additional constraint from the across-PFT hierarchy was limited compared to the simple multivariate models.
-
-* Trait correlations were generally consistent in direction within and across PFTs, and consistent with predictions of the leaf economic spectrum.
-However, correlation strength varied substantially across PFTs indicating that leaf economic relationships within PFTs are often confounded by other biotic and abiotic factors.

### 2018-01-03 manuscript-md/index.Rmd
Address Niinemets comments on introduction

@@
 
-* We developed a hierarchical multivariate Bayesian model that assumes separate means and covariance structures within and across PFTs and fit this model to seven leaf traits from the TRY database: longevity, specific leaf area, nitrogen and phosphorus contents, and rates of dark respiration, maximum Rubisco carboxylation, and maximum electron transport.
+* We developed a hierarchical multivariate Bayesian model that assumes separate means and covariance structures within and across PFTs and fit this model to seven leaf traits from the TRY database: longevity, specific leaf area, nitrogen content, phosphorus content, dark respiration rate, maximum photosynthetic carboxylation rate ($V_{c,max}$), and maximum photosynthetic electron transport rate ($J_{max}$).
 We evaluated the ability of this model to estimate trait means and reproduce leaf economic trade-offs.
@@
 * Trait correlations were generally consistent in direction within and across PFTs, and consistent with predictions of the leaf economic spectrum.
-However, correlation strength varied substantially across PFTs indicating that leaf economic relationships within PFTs are often confounded by other biotic and abiotic factors.
+However, correlation strength varied substantially across PFTs indicating that leaf economic relationships within PFTs are often confounded by the unique physiology of certain plant types or environmental conditions in certain biomes.
 

### 2018-01-08 manuscript-md/02_methods.Rmd
Convert SLA units to kg m-2

@@
 Leaf longevity (months),
-specific leaf area (SLA, mm$^2$ mg$^{-1}$),
+specific leaf area (SLA, m$^2$ kg$^{-1}$),
 leaf nitrogen content ($N_{mass}$, mg N g$^{-1}$ or $N_{area}$, g m$^{-2}$),

### 2018-01-08 scripts/pft_abbreviations.R

@@ -57,7 +57,7 @@ param_markdown_nounit <- c(
 )
 param_markdown <- c(
   leaf_lifespan = "Leaf lifespan (months)",
-  SLA = "SLA (mm$^2$ mg$^{-1}$)",
+  SLA = "SLA (m$^2$ kg$^{-1}$)",
   Nmass = "$N_{mass}$ (mg g$^{-1}$)",
   Narea = "$N_{area}$ (g m$^{-2}$)",
   Pmass = "$P_{mass}$ (mg g$^{-1}$)",
@@ -74,7 +74,7 @@ param_fancy_chr <- c(
   leaf_lifespan = "Leaf~lifespan~(months)",
   Nmass = "N[mass]~(mg ~ g ^ -1)",
   Pmass = "P[mass]~(mg ~ g ^ -1)",
-  SLA = "SLA~(mm ^ 2 ~ mg ^ -1)",
+  SLA = "SLA~(m ^ 2 ~ kg ^ -1)",
   Rdmass = "R[list(d,mass)]~(mu * mol ~ g ^ -1 ~ s ^ -1)",
   Vcmax_mass = "V[list(c, max, mass)]~(mu * mol ~ g ^ -1 ~ s ^ -1)",
   Jmax_mass = "J[list(max, mass)]~(mu * mol ~ g ^ -1 ~ s ^ -1)",
@@ -92,7 +92,7 @@ param_fancy <- sapply(param_fancy_chr, formula)
   #leaf_lifespan = "Leaf lifespan (months)",
   #Nmass = expression(N[mass]~(
   #Pmass = expression((P[mass], " (%)")),
-  #SLA = expression(("SLA (", mm ^ 2 ~ mg ^ -1, ")")),
+  #SLA = expression(("SLA (", m ^ 2 ~ kg ^ -1, ")")),
   #Rdmass = expression((R[{list(d, mass)}], " (", mu * mol ~ g ^ -1 ~ s ^ -1, ")")),
   #Jmax_mass = expression((J[{list(max, mass)}], " (", mu * mol ~ g ^ -1 ~ s ^ -1, ")")),
   #Vcmax_mass = expression((V[{list(c, max, mass)}], " (", mu * mol ~ g ^ -1 ~ s ^ -1, ")")),

### 2018-01-08 manuscript-md/02_methods.Rmd
Fix `km^2` font

@@
 For biome specification, we collected all latitude and longitude data for each species,
-matched these data to 30 second ($\sim1 km^{2}$) mean annual temperature ($AMT$, averaged 1970-2000) data from WorldClim-2 [@worldclim],
+matched these data to 30 second ($\sim1$ km$^{2}$) mean annual temperature ($AMT$, averaged 1970-2000) data from WorldClim-2 [@worldclim],
 calculated the mean AMT for all sites at which that species was observed,

### 2018-01-08 manuscript-md/02_methods.Rmd
Clarify text on assigning PFTs (Dietze)

@@
 We obtained categorical data on growth form, leaf type, phenology, and photosynthetic pathway from the TRY database.
-Where the same species was reported to have different attributes in different datasets, we assigned the attribute that was observed most frequently, and, in cases of ties, selected the attribute that was more likely based on phylogenetically similar species or more likely to occur in nature.
+Where species attributes disagreed between datasets (e.g., categorized as a shrub in one dataset but a tree in another), we assigned the attribute that was observed most frequently between the datasets (e.g., if five datasets say "shrub" but only one says "tree", we would classify it as a shrub).
+Where species lacked data on certain attributes, we assigned those attributes based on higher order phylogeny where appropriate (e.g., \em{Poaceae} family are usually grasses, \em{Larix spp.} are deciduous needleleaved trees) or otherwise omitted the species from our analyses.
 For biome specification, we collected all latitude and longitude data for each species,

### 2018-01-08 manuscript-md/02_methods.Rmd
Fix equations (remove `...`) in methods

@@
 The simplest model was the ‘univariate’ model, in which each trait was modeled independently.
-For an observation $x_{i,t}$ of trait $t$ and sample $i$...
+For an observation $x_{i,t}$ of trait $t$ and sample $i$:
 
@@
 
-...where $N$ is the univariate normal (Gaussian) distribution with mean $\mu_t$ and standard deviation $\sigma_t$ for trait $t$.
+where $N$ is the univariate normal (Gaussian) distribution with mean $\mu_t$ and standard deviation $\sigma_t$ for trait $t$.
 
 The second-simplest model was the ‘multivariate’ model, in which traits were modeled as samples from a multivariate distribution with a single mean vector and covariance matrix.
-For the observed vector of traits ${\mathbf{x_i}}$ for sample $i$...
+For the observed vector of traits ${\mathbf{x_i}}$ for sample $i$:
 
@@
 
-...where $mvN$ is the multivariate normal (Gaussian) distribution with mean vector $\mathbf{\mu}$ and variance-covariance matrix $\mathbf{\Sigma}$.
+where $mvN$ is the multivariate normal (Gaussian) distribution with mean vector $\mathbf{\mu}$ and variance-covariance matrix $\mathbf{\Sigma}$.
 We ran both of these models independently for each PFT as well as for the entire dataset (as if every observation belonged to the same PFT).
@@
 The most complex model was the ‘hierarchical’ model, in which observed trait values were drawn from a PFT-specific multivariate normal distribution describing within-PFT variation and whose parameters were themselves sampled from a global multivariate distribution describing the variation among PFTs.
-For the observed vector of traits $\mathbf{x}_{i,p}$ for sample $i$ belonging to PFT $p$...
+For the observed vector of traits $\mathbf{x}_{i,p}$ for sample $i$ belonging to PFT $p$:
 
@@
 
-…where $\mathbf{\mu}_p$ and $\mathbf{\Sigma}_p$ are the mean vector and variance-covariance matrix describing variation within PFT $p$, and $\mathbf{\mu}_g$ and $\mathbf{\Sigma}_g$ are the mean vector and variance-covariance matrix describing among-PFT (global) variation.
+where $\mathbf{\mu}_p$ and $\mathbf{\Sigma}_p$ are the mean vector and variance-covariance matrix describing variation within PFT $p$, and $\mathbf{\mu}_g$ and $\mathbf{\Sigma}_g$ are the mean vector and variance-covariance matrix describing among-PFT (global) variation.
 
@@
 We fit the above models using a Gibbs sampling algorithm that leveraged known conjugate prior relationships for efficient exploration of the sampling space.
-For priors on all multivariate mean vectors ($\mathbf{\mu}$), we used normal distributions…
+For priors on all multivariate mean vectors ($\mathbf{\mu}$), we used normal distributions:
 
@@
 
-...which gives rise to the following expression for the posterior...
+This gives rise to the following expression for the posterior:
 
@@
 
-...where...
-
 $$\mathbf{\Sigma^*} = (\mathbf{\Sigma}_0^{-1} + n \mathbf{\Sigma}^{-1})^{-1}$$
@@
 
-...where ${\bar{{\mathbf{x}}}}$ are the sample means of the data and $n$ is the number of rows in the data.
+where ${\bar{{\mathbf{x}}}}$ are the sample means of the data and $n$ is the number of rows in the data.
 
-For priors on all multivariate variance-covariance matrices, we used the Wishart distribution ($W$)...
+For priors on all multivariate variance-covariance matrices, we used the Wishart distribution ($W$):
 
@@
 
-...where...
-
 $$\nu^* = 1 + \nu_0 + n + m$$
@@
 
-...where $n$ is the number of rows and $m$ is the number of columns in data matrix $x$. For further details on this derivation, see @gelman_bayesian.
+where $n$ is the number of rows and $m$ is the number of columns in data matrix $x$. For further details on this derivation, see @gelman_bayesian.
 
@@
 For a block of data $\mathbf{x'}$ containing missing observations in columns $\mathbf{m}$ and present observations in columns $\mathbf{p}$,
-the missing values $\mathbf{x'}[m]$ are drawn randomly from a marginal multivariate normal distribution at each iteration of the sampling algorithm...
+the missing values $\mathbf{x'}[m]$ are drawn randomly from a conditional multivariate normal distribution at each iteration of the sampling algorithm:
 
-$$\mathbf{x'}[m] \sim mvN(\mathbf{\mu}', \mathbf{\Sigma}')$$
-
-...where...
+$$\mathbf{x'}[m|p] \sim mvN(\mathbf{\mu}', \mathbf{\Sigma}')$$
 

### 2018-01-08 manuscript-md/02_methods.Rmd
Simplify mean and CI sentence in methods (Dietze)

@@
 To assess the impact of multivariate and hierarchical constraint on trait estimates,
-we compared the mean and size of the 95% confidence intervals of mean estimates of each trait for each PFT from each model (Fig.\ \@ref(fig:meancompare)).
-In addition, we report the mean and 95% confidence intervals of mean estimates for each trait and PFT from the hierarchical model in tabular form (Tab.\ 3 and 4).
+we compared the mean and 95% confidence intervals of trait estimates for each PFT from each model (Fig.\ \@ref(fig:meancompare)).
 

### 2018-01-08 manuscript-md/02_methods.Rmd
First pass at clarifying eigenvalue analysis

@@
 
-To assess the consistency of among- and within-PFT trait trade-offs, we calculated the eigenvalues from pairwise variance-covariance matrix estimates for each trait pair and, where the covariance was significantly different from zero ($p < 0.05$), plotted the dominant eigenvectors centered on the mean estimates (Fig.\ \@ref(fig:stickpairs)).
+To assess the consistency of among- and within-PFT trait trade-offs, we looked at covariance estimates for each trait pair and, where these values were significantly different from zero ($p < 0.05$),
+we calculated the eigenvalues from the pairwise variance-covariance matrix for that trait pair and plotted the corresponding dominant eigenvectors centered on the mean estimates (Fig.\ \@ref(fig:stickpairs)).
 This figure provides a visual representation of relative positions of PFTs in trait space and both the direction and extent of within-PFT trait covariance, and is directly analogous to conceptual figures describing hierarchical trait variability across environmental gradients as presented in, for instance, @cornwell_community_2009 and @albert_intraspecific_2010.

### 2018-01-08 manuscript-md/02_methods.Rmd
Add intercept analysis to methods

@@
 
-If all three models performed equally well at all sample sizes, their respective slope coefficients would be statistically indistinguishable.
-On the other hand, models that perform better should have lower slope ($b_1$) coefficients, indicating a slower increase in the uncertainty ($\alpha$) with decreasing sample size ($n$).
+If all three models performed equally well at all sample sizes, their respective slope and intercept coefficients would be statistically indistinguishable.
+On the other hand, models that perform better should have
+lower intercept ($b_0$) coefficients, indicating generally lower uncertainty,
+and
+lower slope ($b_1$) coefficients, indicating a reduced sensitivity of uncertainty ($\alpha$) to sample size ($n$).
 

### 2018-01-08 manuscript-md/03_results.Rmd
Add slope and intercept discussion to results

--- 2017-11-16 manuscript-md/03_results.Rmd
@@
 For many of these trait-PFT combinations, the additional constraint from trait covariance provided by the multivariate and hierarchical models substantially reduced error bars, making it possible to compare estimates against those of other PFTs.
-Our analysis of the relationship between model type, sample size, and estimate relative uncertainty found that this covariance-based constraint from the multivariate model becomes increasingly important at smaller sample sizes (\@ref(fig:civssamplesize)).
+Our analysis of the relationship between model type, sample size, and estimate relative uncertainty found that this covariance-based constraint from the multivariate model both reduced uncertainty overall (lower intercept) and reduced the sensitivity of estimate uncertainty to sample size (lower slope) compared to the univariate model (Fig.\ \@ref(fig:civssamplesize)).
 However, this analysis revealed no consistent significant benefit from the hierarchical model.

### 2018-01-08 manuscript-md/03_results.Rmd
Results: Fix contradictory correlation sentence

@@
 
-Where pairwise trait correlations were statistically significant, these correlations were generally consistent in direction both globally and within each PFT (Fig.\ \@ref(fig:stickpairs)).
+For all traits except leaf lifespan, pairwise trait correlations were generally consistent in direction both globally and within each PFT (Fig.\ \@ref(fig:stickpairs)).
 In particular, mass- and area-normalized traits were all positively correlated with each other and, respectively, positively and negatively correlated with SLA, both globally and within each PFT.

### 2018-01-08 R/pft_colors.R
First pass at OSF re-organization
- Moved global definitions (trait names, colors, etc.) into R package
(`R/` directory), and added documentation
- Extracted code from `Rmd` files into R scripts using `knitr::purl`.
These currently live in `purld`, but will be moved elsewhere at some
point.
- Changed package name to `shiklomanov2017np` to be less ambiguous.
- Added `here` package dependency, and `.here` file in package root.

@@ -1,9 +1,55 @@
+#' Paul Tol color schemes
+#'
+#' Effective color schemes for categorical data.
+#' Available for 14, 15, 18, and 21 levels.
+#'
+#' @export
 tol14rainbow <- c("#882E72", "#B178A6", "#D6C1DE", "#1965B0", "#5289C7", "#7BAFDE", "#4EB265", "#90C987", "#CAE0AB", "#F7EE55", "#F6C141", "#F1932D", "#E8601C", "#DC050C")
+
+#' @rdname tol14rainbow
+#' @export
 tol15rainbow <- c("#114477", "#4477AA", "#77AADD", "#117755", "#44AA88", "#99CCBB", "#777711", "#AAAA44", "#DDDD77", "#771111", "#AA4444", "#DD7777", "#771144", "#AA4477", "#DD77AA")
+
+#' @rdname tol14rainbow
+#' @export
 tol18rainbow <- c("#771155", "#AA4488", "#CC99BB", "#114477", "#4477AA", "#77AADD", "#117777", "#44AAAA", "#77CCCC", "#777711", "#AAAA44", "#DDDD77", "#774411", "#AA7744", "#DDAA77", "#771122", "#AA4455", "#DD7788")
-# ...and finally, the Paul Tol 21-color salute
+
+#' @rdname tol14rainbow
+#' @export
 tol21rainbow <- c("#771155", "#AA4488", "#CC99BB", "#114477", "#4477AA", "#77AADD", "#117777", "#44AAAA", "#77CCCC", "#117744", "#44AA77", "#88CCAA", "#777711", "#AAAA44", "#DDDD77", "#774411", "#AA7744", "#DDAA77", "#771122", "#AA4455", "#DD7788")
 
+#' Indices defining PFT colors
+#'
+#' Indices correspond to `tol21rainbow` color scheme.
+pft_inds <- c(
+  BlETr = 1,
+  BlETe = 3,
+  BlDTr = 4,
+  BlDTe = 5,
+  BlDBo = 6,
+  NlETe = 7,
+  NlEBo = 8,
+  NlD = 9,
+  ShE = 13,
+  ShDTe = 14,
+  ShDBo = 15,
+  C3GAr = 16,
+  C3GTe = 17,
+  C4G = 18
+)
+
+#' PFT color scheme
+#'
+#' @export
+pft_colors <- tol21rainbow[pft_inds]
+
+#' Quick function for examining color schemes
+#' 
[…]

### 2018-01-08 R/pft_names.R

@@ -1,0 +1,27 @@
+#' PFT abbreviations
+#'
+#' Dictionary from PFT abbreviations to PFTs and vice-versa.
+#'
+#' @export
+abbr2pft <- c(
+  GLOB = "global",
+  BlETr = "broadleaf_evergreen_tropical",
+  BlETe = "broadleaf_evergreen_temperate",
+  BlDTr = "broadleaf_deciduous_tropical",
+  BlDTe = "broadleaf_deciduous_temperate",
+  BlDBo = "broadleaf_deciduous_boreal",
+  NlETe = "needleleaf_evergreen_temperate",
+  NlEBo = "needleleaf_evergreen_boreal",
+  NlD = "needleleaf_deciduous",
+  ShE = "shrub_evergreen",
+  ShDTe = "shrub_deciduous_temperate",
+  ShDBo = "shrub_deciduous_boreal",
+  C3GAr = "c3_grass_arctic",
+  C3GTe = "c3_grass_temperate",
+  C4G = "c4_grass"
+)
+
+#' @rdname abbr2pft
+#' @export
+pft2abbr <- names(abbr2pft)
+names(pft2abbr) <- abbr2pft

### 2018-01-08 R/trait_names.R

@@ -1,0 +1,117 @@
+#' Trait names
+#'
+#' Names of traits used in analyses, useful for factor levels and for quick 
+#' selection from data frames.
+#'
+#' Contains the following:
+#' - `mass_params` -- Mass-based traits
+#' - `area_params` -- Area-based traits
+#' - `both_params` -- "Base" names
+#'
+#' @name trait_names
+NULL
+
+#' @rdname trait_names
+#' @export
+mass_params <- c("leaf_lifespan", "SLA", "Nmass", "Pmass", "Rdmass", "Vcmax_mass", "Jmax_mass")
+
+#' @rdname trait_names
+#' @export
+area_params <- gsub("mass$", "area", mass_params)
+
+#' @rdname trait_names
+#' @export
+both_params <- c("leaf_lifespan", "SLA", "Nmass", "Narea", "Pmass", "Parea",
+                 "Rdmass", "Rdarea", "Vcmax_mass", "Vcmax_area",
+                 "Jmax_mass", "Jmax_area")
+
+#' Trait labels
+#'
+#' Various vectors for providing formatted trait labels for figures and tables.
+#'
+#' Contains the following:
+#' - `param_labels` -- Base trait names, without mass/area suffix
+#' - `param_simple` -- R expression-formatted trait names with units, as character (used for `parse = TRUE`)
+#' - `param_fancy_chr` -- R expression-formatted trait names with units, as character (used for `parse = TRUE`)
+#' - `param_fancy` -- R expression-formatted trait names with units, as R `formula` vector
+#' - `param_markdown_nounit` -- LaTeX-formatted trait names without units, as character
+#' - `param_markdown` -- LaTeX-formatted trait names with units, as character
+#' @name trait_labels
+NULL
+
+#' @rdname trait_labels 
+#' @export
+param_labels <- c("Leaf lifespan", "SLA", "N", "P", "Rd", "Vcmax", "Jmax")
+
[…]

### 2018-01-08 purld/figure_correlation_boxplot.R

@@ -1,0 +1,103 @@
+library(shiklomanov2017np)
+
+## ----boxprep-------------------------------------------------------------
+outdir <- here::here("output")
+hier_mass_file <- tail(dir(outdir, "hier.mass.clm45", full.names = TRUE), 1)
+hier_area_file <- tail(dir(outdir, "hier.area.clm45", full.names = TRUE), 1)
+multi_mass_file <- tail(dir(outdir, "multi.mass.clm45", full.names = TRUE), 1)
+multi_area_file <- tail(dir(outdir, "multi.area.clm45", full.names = TRUE), 1)
+
+hier_summary_mass <- readRDS(hier_mass_file)[["summary_table"]] %>%
+  mutate(mass_area = "mass")
+hier_summary_area <- readRDS(hier_area_file)[["summary_table"]] %>%
+  mutate(mass_area = "area")
+multi_summary_mass <- readRDS(multi_mass_file)[["summary_table"]] %>%
+  mutate(mass_area = "mass")
+multi_summary_area <- readRDS(multi_area_file)[["summary_table"]] %>%
+  mutate(mass_area = "area")
+
+proc <- function(dat, hier = TRUE) {
+  if (hier) {
+    dat <- filter(dat, group != "global")
+  } else {
+    dat <- mutate(dat, group = "global")
+  }
+  dat <- dat %>%
+    filter(variable == "Corr") %>%
+    separate(index, c("xvar", "yvar"), sep = "\\.\\.") %>%
+    mutate(
+      pft = factor(group, abbr2pft) %>% forcats::lvls_revalue(pft2abbr)
+    ) %>%
+    select(mass_area, pft, xvar, yvar, Mean, `2.5%`, `97.5%`)
+}
+
+hsm <- proc(hier_summary_mass)
+hsa <- proc(hier_summary_area)
+msm <- proc(multi_summary_mass, FALSE)
+msa <- proc(multi_summary_area, FALSE)
+hs <- bind_rows(hsm, hsa, msm, msa)
+
+## ----corrbars, fig.width=7, fig.height=7, fig.cap='(ref:corrbarscap)'----
+.zzz <- close.screen(all.screens = TRUE)
+corr_bar <- function(dat, xvar, yvar, mass_area) {
+  dp <- dat %>%
+    filter(
+      mass_area == !!mass_area,
+      (xvar == !!xvar & yvar == !!yvar) | 
[…]

### 2018-01-08 purld/figure_samplesize.R

@@ -1,0 +1,39 @@
+# Chunk name: samplesize
+# Description: Figure of the sample sizes of each PFT
+
+## ----options, echo=FALSE, message=FALSE----------------------------------
+library(shiklomanov2017np)
+try_data <- readRDS("extdata/traits_analysis.rds")
+
+try_sub <- try_data %>%
+  select(pft = clm45, one_of(both_params),
+         -Latitude, -Longitude, -LMA)
+
+sample_size <- try_sub %>%
+  select(pft, one_of(both_params)) %>%
+  mutate(
+    pft = factor(pft, abbr2pft) %>% forcats::lvls_revalue(pft2abbr)
+  ) %>%
+  group_by(pft) %>%
+  summarize_all(~sum(!is.na(.))) %>%
+  gather("param", "sample_size", -pft) %>%
+  mutate(param = factor(param, both_params))
+
+## ----"samplesize", fig.cap='(ref:ssfigcap)', fig.width=7, fig.height=7----
+ss_plot <- sample_size %>%
+  mutate(param = forcats::lvls_revalue(param, param_simple[both_params])) %>%
+  ggplot() +
+  aes(x = pft, y = sample_size, fill = pft) +
+  geom_col() +
+  scale_y_log10(breaks = c(10, 100, 1000, 10000)) +
+  facet_wrap(~param, ncol = 2, labeller = label_parsed) +
+  scale_fill_manual(values = pft_colors) +
+  xlab("Plant functional type") +
+  ylab("Sample size") +
+  guides(fill = guide_legend(title = "PFT")) +
+  theme_bw() +
+  theme(
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank()
+  )
+ss_plot

### 2018-01-08 purld/many_results.R

@@ -1,0 +1,284 @@
+library(shiklomanov2017np)
+
+## ----prepresults---------------------------------------------------------
+cachefile <- here::here(".cache/mvtraits_results.rds")
+results_all <- readRDS(cachefile)
+
+max_ss <- max(sample_size$sample_size)
+
+multi_summary <- results_all %>%
+  filter(model_type %in% c("uni", "multi")) %>%
+  mutate(
+    mu_mean = map(data, c("mu", "Mean")) %>%
+      map(~as_tibble(as.list(.))),
+    mu_lo = map(data, c("mu", "2.5%")) %>%
+      map(~as_tibble(as.list(.))),
+    mu_hi = map(data, c("mu", "97.5%")) %>%
+      map(~as_tibble(as.list(.)))
+      ) %>%
+  select(model_type:pft, mu_mean:mu_hi) %>%
+  unnest(.sep = ".") %>%
+  gather("variable", "value", -(model_type:pft), na.rm = TRUE) %>%
+  separate(variable, c("stat", "param"), sep = "\\.") %>%
+  spread(stat, value) %>%
+  mutate(
+    pft = factor(pft, abbr2pft) %>% forcats::lvls_revalue(pft2abbr),
+    param = factor(param, both_params)
+  )
+
+hier_prep <- results_all %>%
+  filter(model_type == "hier") %>%
+  select(model_type:pft_type, data) %>%
+  mutate(
+    mu_vals = map(data, "mu_group"),
+    pft = map(mu_vals, names),
+    mu_df = map2(pft, mu_vals, ~tibble(pft = ..1, dat = ..2))
+  ) %>%
+  select(-data, -mu_vals, -pft) %>%
+  unnest()
+
+hier_summary <- hier_prep %>%
+  mutate(
+    mu_mean = map(dat, c("Mean")) %>%
+      map(~as_tibble(as.list(.))),
+    mu_lo = map(dat, c("2.5%")) %>%
+      map(~as_tibble(as.list(.))),
+    mu_hi = map(dat, c("97.5%")) %>%
+      map(~as_tibble(as.list(.)))
+      ) %>%
+  select(model_type:pft, mu_mean:mu_hi) %>%
+  unnest(.sep = ".") %>%
+  gather("variable", "value", -(model_type:pft), na.rm = TRUE) %>%
+  separate(variable, c("stat", "param"), sep = "\\.") %>%
+  spread(stat, value) %>%
+  mutate(
[…]

### 2018-01-08 purld/table_n_species.R

@@ -1,0 +1,21 @@
+# Chunk name: "pfts"
+# Description: Table of number of species in each PFT
+library(shiklomanov2017np)
+try_data <- readRDS("extdata/traits_analysis.rds")
+
+## ----"pfts"--------------------------------------------------------------
+n_species <- try_data %>%
+  distinct(clm45, AccSpeciesID) %>%
+  select(pft = clm45, species = AccSpeciesID) %>%
+  count(pft) %>%
+  mutate(
+    pft = as.character(pft),
+    label = pft2abbr[pft],
+    pft = tools::toTitleCase(gsub("_", " ", pft))
+  ) %>%
+  select(`Label` = label, `PFT` = pft, `Number of species` = n)
+species_caption <- paste(
+  "Names, labels, and species counts for plant functional types (PFTs)",
+  "used in this analysis."
+)
+kable(n_species, caption = species_caption, align = rep("c", ncol(n_species)))

### 2018-01-08 purld/try_ref_table.R

@@ -1,0 +1,26 @@
+# Chunk name: trydata
+# Description: Table of TRY data citations
+
+library(shiklomanov2017np)
+crtable <- readRDS("extdata/crossref_table.rds") %>%
+  mutate(
+    Reference = case_when(
+      Reference == "@try_data_03" ~ "@baraloto_decoupled_2010",
+      Reference == "@try_data_09" ~ "@wright_worldwide_2004",
+      Reference == "@try_data_14" ~ "@kattge_2009_vcmax",
+      Reference == "@try_data_15" ~ "@laughlin_multi-trait_2010",
+      Reference == "@try_data_26" ~ "@onoda_2011_global",
+      Reference == "@try_data_30" ~ "@messier_how_2010",
+      Reference == "@try_data_33" ~ "@pierce_allocating_2013",
+      Reference == "@try_data_41" ~ "@poorter_causes_2009",
+      Reference == "@try_data_44" ~ "@choat_global_2012",
+      Reference == "@try_data_59" ~ "@laughlin_climatic_2011",
+      Reference == "@try_data_67" ~ "@wright_does_2012",
+      Reference == "@try_data_80" ~ "@choat_global_2012",
+      Reference == "@try_data_114" ~ "@kichenin_2013_contrasting",
+      Reference == "@try_data_134" ~ "@diaz_plant_2004",
+      TRUE ~ Reference
+    )
+  )
+cap <- "Data sources and observation counts for foliar trait data"
+kable(crtable, caption = cap)

### 2018-01-08 purld/wordcount.R

@@ -1,0 +1,30 @@
+## ----wordcount, echo = FALSE---------------------------------------------
+fnames <- c(
+  introduction = "01_introduction.md",
+  methods = "02_methods.Rmd",
+  results = "03_results.Rmd",
+  discussion = "04_discussion.md",
+  conclusions = "05_conclusions.md"
+)
+files <- file.path(
+  "~/Projects/new-phytologist-traits/np-trait-analysis/manuscript-md",
+  fnames
+)
+names(files) <- names(fnames)
+
+getwc <- function(f) {
+  ct <- withr::with_options(
+    list(knitr.table.format = "pandoc"),
+    wordcountaddin::text_stats(f)
+  )
+  wline <- ct[3]
+  wlist <- strsplit(wline, split = "[[:space:]]+")
+  as.numeric(wlist[[1]][3])
+}
+
+wc_table <- data.frame(wordcount = vapply(files, getwc, numeric(1)))
+wc_total <- rbind(wc_table, sum(wc_table$wordcount))
+wc_total$percent <- 100 * wc_total$wordcount /
+  sum(wc_total$wordcount[-nrow(wc_total)])
+rownames(wc_total)[nrow(wc_total)] <- "total"
+knitr::kable(wc_total, digits = 1)

### 2018-01-08 scripts/pft_abbreviations.R

@@ -1,124 +1,0 @@
-abbr2pft <- c(
-  GLOB = "global",
-  BlETr = "broadleaf_evergreen_tropical",
-  BlETe = "broadleaf_evergreen_temperate",
-  BlDTr = "broadleaf_deciduous_tropical",
-  BlDTe = "broadleaf_deciduous_temperate",
-  BlDBo = "broadleaf_deciduous_boreal",
-  NlETe = "needleleaf_evergreen_temperate",
-  NlEBo = "needleleaf_evergreen_boreal",
-  NlD = "needleleaf_deciduous",
-  ShE = "shrub_evergreen",
-  ShDTe = "shrub_deciduous_temperate",
-  ShDBo = "shrub_deciduous_boreal",
-  C3GAr = "c3_grass_arctic",
-  C3GTe = "c3_grass_temperate",
-  C4G = "c4_grass"
-)
-
-pft2abbr <- names(abbr2pft)
-names(pft2abbr) <- abbr2pft
-
-mass_params <- c("leaf_lifespan", "SLA", "Nmass", "Pmass", "Rdmass", "Vcmax_mass", "Jmax_mass")
-area_params <- gsub("mass$", "area", mass_params)
-both_params <- c("leaf_lifespan", "SLA", "Nmass", "Narea", "Pmass", "Parea",
-                 "Rdmass", "Rdarea", "Vcmax_mass", "Vcmax_area",
-                 "Jmax_mass", "Jmax_area")
-param_labels <- c("Leaf lifespan", "SLA", "N", "P", "Rd", "Vcmax", "Jmax")
-
-param_simple <- c(
-  leaf_lifespan = "Leaf~lifespan",
-  SLA = "SLA",
-  Nmass = "N[mass]",
-  Narea = "N[area]",
-  Pmass = "P[mass]",
-  Parea = "P[area]",
-  Rdmass = "R[list(d, mass)]",
-  Rdarea = "R[list(d, area)]",
-  Vcmax_mass = "V[list(c, max, mass)]",
-  Vcmax_area = "V[list(c, max, area)]",
-  Jmax_mass = "J[list(max, mass)]",
-  Jmax_area = "J[list(max, area)]"
-)
-
[…]

### 2018-01-08 manuscript-md/06_figures_tables.Rmd
Isolate code from Rmd in separate file
This file will appear at the end of the manuscript, following the
formatting requirements of New Phytologist.

Ideally, I'd like to show only figure captions, but not the figures
themselves, so that the figure labels are correctly included in the
text.

--- 2017-11-17 manuscript-md/03_results.Rmd
@@
-# Results
+<!-- Methods -->
+
+```{r "pfts"}
+n_species <- try_data %>%
+  distinct(clm45, AccSpeciesID) %>%
+  select(pft = clm45, species = AccSpeciesID) %>%
+  count(pft) %>%
+  mutate(
+    pft = as.character(pft),
+    label = pft2abbr[pft],
+    pft = tools::toTitleCase(gsub("_", " ", pft))
+  ) %>%
+  select(`Label` = label, `PFT` = pft, `Number of species` = n)
+species_caption <- paste(
+  "Names, labels, and species counts for plant functional types (PFTs)",
+  "used in this analysis."
+)
+kable(n_species, caption = species_caption, align = rep("c", ncol(n_species)))
+```
+
+(ref:ssfigcap) Sample sizes for each trait-PFT pair. $y$ axis is scaled logarithmically.
+
+```{r "samplesize", fig.cap='(ref:ssfigcap)', fig.width=7, fig.height=7}
+ss_plot <- sample_size %>%
+  mutate(param = forcats::lvls_revalue(param, param_simple[both_params])) %>%
+  ggplot() + 
+  aes(x = pft, y = sample_size, fill = pft) + 
+  geom_col() +
+  scale_y_log10(breaks = c(10, 100, 1000, 10000)) +
+  facet_wrap(~param, ncol = 2, labeller = label_parsed) +
+  scale_fill_manual(values = pft_colors) +
+  xlab("Plant functional type") +
+  ylab("Sample size") +
+  guides(fill = guide_legend(title = "PFT")) +
+  theme_bw() +
+  theme(
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank()
+  )
+ss_plot
+```
+
+<!-- Results -->
 
@@
 ```
-
-## Estimates of PFT-level means
 
@@
 
-In general, leaf trait estimates from the univariate, multivariate, and hierarchical models were similar (Fig.\ \@ref(fig:meancompare)).
-Where we observed differences between models, the largest were between the univariate and multivariate models, while the additional constraint from the hierarchical model tended to have a minimal effect on trait estimates.
-Significant differences in trait estimates between univariate and multivariate models occurred even for traits with relatively large sample sizes, such as leaf nitrogen content.
-
-Evergreen PFTs had by far the largest leaf lifespan, with the longest lifespan observed for temperate and boreal needleleaf evergreen species.
-Meanwhile, all of deciduous species had lifespans shorter than 7 months.
-Among deciduous species, lifespan was generally longer in warmer biomes than colder ones.
-
-Across-PFT patterns in SLA and $N_{mass}$, $P_{mass}$, and $R_{d,mass}$ were similar.
-Temperate broadleaved deciduous trees and shrubs generally had among the highest values of these traits, while temperate evergreen trees and shrubs had generally among the lowest.
-However, none of these patterns were universal to all four traits.
-For example, tropical evergreen trees had relatively high $N_{mass}$ and average SLA and $R_{d,mass}$, but among the lowest $P_{mass}$.
-Similarly, temperate and boreal shrubs had higher $N_{mass}$ and $P_{mass}$ than any of the grasses, but comparable SLA.
-
-Across-PFT patterns in $N_{area}$, $P_{area}$, and $R_{d,area}$ were different from their mass-normalized counterparts.
-For example, tropical broadleaved evergreen and needleleaf evergreen trees had among the lowest $N_{mass}$ and $P_{mass}$ basis but among the highest $N_{area}$ and $P_{area}$, while the opposite was true of deciduous temperate trees and shrubs.
-Species with N contents near the middle of the observed range did not shift as dramatically depending on type of normalization.
-
-C3 grasses had both the highest $V_{c,max,mass}$ and $V_{c,max,area}$.
-Compared to broadleaved trees, temperate needleleaved evergreen trees had lower $V_{c,max,mass}$ but higher $V_{c,max,area}$.
-Among broadleaved trees, deciduous trees had higher $V_{c,max,mass}$ and slightly higher $V_{c,max,area}$ than evergreen trees.
-Between the deciduous and evergreen tree PFTs, we observed no significant trend by climate zone.
-
-C3 grasses and temperate needleleaved evergreen trees had the highest $J_{max,area}$, but temperate broadleaved deciduous trees had the highest $J_{max,mass}$.
-All of the shrub PFTs had the lowest $J_{max,mass}$ but average or above-average $J_{max,area}$, while the opposite was true of broadleaved tropical PFTs.
-Of the tree PFTs, needleleaved evergreen trees had the highest $J_{max,area}$ but the lowest $J_{max,mass}$.
-
-We observed clear differences in the relative uncertainties of mean estimates with respect to sample size.
-All of the high-latitude PFTs consistently had among the largest error bars around their mean estimates relative to other PFTs, while the traits with the largest uncertainties were dark respiration, $V_{c,max}$, and $J_{max}$.
-For many of these trait-PFT combinations, the additional constraint from trait covariance provided by the multivariate and hierarchical models substantially reduced error bars, making it possible to compare estimates against those of other PFTs.
-Our analysis of the relationship between model type, sample size, and estimate relative uncertainty found that this covariance-based constraint from the multivariate model becomes increasingly important at smaller sample sizes (Fig.\ \@ref(fig:civssamplesize)).
-However, this analysis revealed no consistent significant benefit from the hierarchical model.
-
-(ref:sscap) Relative uncertainty in PFT-level trait estimates as a function of sample size for each model type.
-Lines represent linear models ($\log(y) = b_0 + b_1 \log(x)$) fit independently for each model type.
-In general, differences in estimate uncertainty between the univariate and multivariate models were minimal at large sample sizes but increasingly important at low sample sizes.
-However, differences in estimate uncertainty between the multivariate and hierarchical models were consistently negligible.
-
 ```{r "civssamplesize", fig.width=4.4, fig.height=3, fig.cap='(ref:sscap)'}
@@
 
-## Trait correlation patterns among- and within-PFTs
-
 (ref:stickcap) Pairwise trait mean and covariance estimates for all data pooled globally (black) and for each PFT (colored).
@@
 ```
-
-Where pairwise trait correlations were statistically significant, these correlations were generally consistent in direction both globally and within each PFT (Fig.\ \@ref(fig:stickpairs)).
-In particular, mass- and area-normalized traits were all positively correlated with each other and, respectively, positively and negatively correlated with SLA, both globally and within each PFT.
-The same was generally true of correlations of mass-based traits with leaf lifespan, but correlations of leaf lifespan with area-normalized traits were more variable.
-The correlation between $N_{area}$ and leaf lifespan was positive globally and for evergreen shrubs but negative for broadleaved deciduous trees and not significant for any other PFTs.
-Similarly, the correlation between $P_{area}$ and leaf lifespan was positive for temperate broadleaved evergreen trees and needleleaved evergreen trees but negative for evergreen shrubs.
-The correlation between leaf lifespan and $R_{d,area}$ significant and positive globally, but was not significant within any PFTs.
-The correlations of leaf lifespan with $V_{c,max,area}$ and $J_{max,area}$ were negative for temperate broadleaved deciduous trees, but not significant for any other PFTs or globally.
-
-A large number of pairwise trait correlations were not significant.
-In some cases, this was driven by sample size (Fig.\ \@ref(fig:samplesize)).
-For instance, needleleaved deciduous trees were often the only PFT for which a correlation was not statistically significant.
-In other cases, though, PFTs with smaller sample sizes had significant pairwise correlations while PFTs with much larger sample sizes had none.
-For example, tropical broadleaved evergreen trees were relatively well-sampled for all traits, but none of their traits were well correlated with leaf lifespan.
-Meanwhile, temperate broadleaved evergreen trees had significant correlations between leaf lifespan and SLA, $P_{area}$, and $R_{d,area}$.
-In general, we observed fewer significant trait correlations among area-normalized traits than mass-normalized traits.
 
@@
 ```
-
-The strength of pairwise trait correlations varied substantially depending on scale, PFT, and trait (Fig.\ \@ref(fig:corrbars)).
-The two pairwise trait correlations that exhibited the most consistent strength globally and within each PFT were the correlation between SLA and $N_{area}$, and between $N_{mass}$ and $P_{mass}$.
-Correlation strength was often, but not always, related to sample size, with well-sampled PFTs exhibiting stronger correlations and undersampled PFTs exhibiting weaker correlations.

### 2018-01-12 scripts/01.prep_submit_df.R
Revise submit_df script. Promote cache_output.
`01.prep_submit_df.R`:
- Fix package reference
- Fix script directory

Output processing:
- Delete old `load_output` script. It's irrelevant with new results
format.
- Make `cache_output` new second script. Change directory from `.cache`
to `results`.

@@ -1,5 +1,5 @@
 #!/usr/bin/env Rscript
-library(nptraits)
+library(shiklomanov2017np)
 
 submit_tab_fname <- "submit_df.dat"
 array_submit_fname <- "array_submit.sh"
@@ -42,7 +42,7 @@ script <- c("#!/bin/bash -l",
             "#$ -l h_rt=11:59:00",
             paste("#$ -pe omp", nchains),
             paste0("#$ -t 1-", nrow(submit_df)),
-            paste("Rscript run_model.R",
+            paste("Rscript scripts/run_model.R",
                   "$(head -n ${SGE_TASK_ID} submit_df.dat ",
                   "| tail -n 1)")
             )

### 2018-01-12 scripts/02.load_output.R

@@ -7,13 +7,13 @@ if (is.na(outdir)) {
   outdir <- "output"
 }
 
-cachedir <- arg[2]
-if (is.na(cachedir)) {
-  cachedir <- ".cache"
+resultsdir <- arg[2]
+if (is.na(resultsdir)) {
+  resultsdir <- "results"
 }
 
-dir.create(cachedir, showWarnings = FALSE)
-cachefile <- file.path(cachedir, "mvtraits_results.rds")
+dir.create(resultsdir, showWarnings = FALSE)
+cachefile <- file.path(resultsdir, "mvtraits_results.rds")
 
 parse_filetag <- function(fname_list) {
   l <- lapply(fname_list, strsplit, split = "\\.")

### 2018-01-12 scripts/02.load_results.R

@@ -1,54 +1,0 @@
-library(mvtraits)
-library(magrittr)
-
-output_dir <- "output"
-summary_dir <- "summaries"
-dir.create(summary_dir, showWarnings = FALSE)
-outputs <- list.files(output_dir, recursive = TRUE)
-nfiles <- length(outputs)
-
-for (out_file in outputs[-1:-6]) {
-    message('Processing file: ', out_file)
-    nth_file <- grep(out_file, outputs)
-    message('File ', nth_file, ' of ', nfiles)
-
-    fname <- file.path(output_dir, out_file)
-
-    message('Reading file...')
-    raw_output <- readRDS(fname)
-    if (is.null(raw_output)) {
-        message('File ', fname, ' has NULL output. Moving on.')
-        next
-    }
-
-    model_type <- gsub('^(from_progress/)?(multi|hier).*', '\\2', out_file)
-    is_hier <- model_type == 'hier'
-    if (is_hier) {
-        nparam <- ncol(raw_output[[1]]$mu_global)
-        ngroups <- ncol(raw_output[[1]]$mu_group) / nparam
-    } else {
-        ngroups <- NULL
-    }
-
-    niter <- nrow(raw_output[[1]][[1]])
-    nburn <- pmax(niter - 5000, floor(0.5 * niter))
-
-    burn <- function(samps) lapply(samps, function(x) x[nburn:niter,])
-    raw_burned <- lapply(raw_output, burn)
-    rm(raw_output); gc()
-
-    # This step is slow!
-    message('Adding correlations...')
-    raw_withcorr <- add_correlations(raw_burned,
-                                     hier = is_hier,
-                                     ngroups = ngroups)
-
-    message('Converting to MCMC list...')
-    raw_mcmc <- results2mcmclist(raw_withcorr, model_type)
-    rm(raw_burned, raw_withcorr); gc()
-
-    message('Calculating summary...')
-    out_summary <- summary_df(raw_mcmc, is_hier)
-    saveRDS(out_summary, file.path(summary_dir, basename(out_file)))
-    rm(raw_mcmc); gc()
-}

### 2018-01-12 scripts/02.load_output.R
Add README. Tweak 02.load_results script

@@ -1,19 +1,20 @@
 library(tidyverse)
+library(here)
 
 arg <- commandArgs(trailingOnly = TRUE)
 
 outdir <- arg[1]
 if (is.na(outdir)) {
-  outdir <- "output"
+  outdir <- here("output")
 }
 
 resultsdir <- arg[2]
 if (is.na(resultsdir)) {
-  resultsdir <- "results"
+  resultsdir <- here("results")
 }
 
 dir.create(resultsdir, showWarnings = FALSE)
-cachefile <- file.path(resultsdir, "mvtraits_results.rds")
+resultsfile <- file.path(resultsdir, "mvtraits_results.rds")
 
 parse_filetag <- function(fname_list) {
   l <- lapply(fname_list, strsplit, split = "\\.")
@@ -42,8 +43,8 @@ read_output <- function(fname) {
   readRDS(fname)[["stats"]]
 }
 
-outfiles$data <- lapply(outfiles$fname, read_output)
+outfiles$data <- map(outfiles$fname, safely(read_output))
 
 outfiles %>%
   select(model_type, mass_area, pft_type, pft, date, fname, data) %>%
-  saveRDS(cachefile)
+  saveRDS(resultsfile)

### 2018-01-12 purld/figure_samplesize.R
First major scripts cleanup
- Remove a lot of deprecated scripts
- Add information about remaining scripts to README
- A few other minor updates and fixes

@@ -3,7 +3,7 @@
 
 ## ----options, echo=FALSE, message=FALSE----------------------------------
 library(shiklomanov2017np)
-try_data <- readRDS("extdata/traits_analysis.rds")
+try_data <- readRDS(here("extdata/traits_analysis.rds"))
 
 try_sub <- try_data %>%
   select(pft = clm45, one_of(both_params),

### 2018-01-12 scripts/03.load_summaries.R

@@ -1,25 +1,0 @@
-library(mvtraits)
-library(tidyverse)
-
-summaries_dir <- 'summaries'
-results_dir <- 'results'
-dir.create(results_dir, showWarnings = FALSE)
-
-summary_files <- list.files(summaries_dir)
-summary_file_pieces <- strsplit(summary_files, '\\.')
-summary_file_dat <- do.call(rbind, summary_file_pieces) %>% 
-    as_tibble() %>% 
-    select(model_type = V1, area_mass = V2, pft_scheme = V3, pft = V4) %>% 
-    mutate(fname = file.path(summaries_dir, summary_files))
-
-summary_file_full <- summary_file_dat %>% 
-    mutate(data = map(fname, readRDS))
-
-fix_groups <- function(dat) {
-    mutate(dat, group = as.character(group))
-}
-
-summary_file_fixed <- summary_file_full %>% 
-    mutate(data = map(data, fix_groups))
-
-saveRDS(summary_file_fixed, file = file.path(results_dir, 'summaries.rds'))

### 2018-01-12 scripts/clean_output.R

@@ -1,22 +1,0 @@
-library(mvtraits)
-
-try_dat <- readRDS("extdata/traits_analysis.rds")
-pfts <- levels(try_dat$clm45)
-group_names <- sprintf("group%02d", seq_along(pfts))
-
-fname <- "output/hier.area.clm45.NA.2017-10-25-1632.rds"
-
-out <- readRDS(fname)
-
-summary_table <- out$summary_table
-
-### Fix names
-group_vars <- c("mu_group", "Sigma_group", "Corr_group")
-clean_out$means[group_vars] <- lapply(
-  clean_out$means[group_vars],
-  "names<-",
-  pfts
-)
-names(out)[2] <- "stats"
-saveRDS(out, fname)
-

### 2018-01-12 scripts/compare_mean_variance.R

@@ -1,24 +1,0 @@
-library(nptraits)
-
-all_summaries <- readRDS('results/summaries_processed.rds')
-figures_dir <- 'figures'
-
-plot_data <- all_summaries %>%
-    distinct(area_mass, pft_scheme) %>%
-    mutate(plots = map2(area_mass, pft_scheme, compare_mv, summary_data = all_summaries)) %>%
-    mutate_if(is.factor, as.character) %>%
-    unnest()
-
-plot_data %$%
-    pwalk(list(area_mass, pft_scheme, mean_plot),
-          ~ ggsave(filename = file.path(
-              figures_dir, paste('compare_mean', ..1, ..2, 'png', sep = '.')
-          ),
-          plot = ..3))
-
-plot_data %$%
-    pwalk(list(area_mass, pft_scheme, var_plot),
-          ~ ggsave(filename = file.path(
-              figures_dir, paste('compare_variance', ..1, ..2, 'png', sep = '.')
-          ),
-          plot = ..3))

### 2018-01-12 scripts/correlation_barplots.R

@@ -1,49 +1,0 @@
-library(nptraits)
-
-all_summaries <- readRDS('results/summaries_processed.rds')
-figures_dir <- 'figures'
-
-prep_data <- function(dat) {
-    dat %>%
-        mutate(xparam = factor(xparam, tail(levels(param), -1)),
-               yparam = factor(yparam, head(levels(param), -1)))
-}
-
-custom_barplot <- function(..., pft_scheme) {
-    pft_cols <- pft_colors[[pft_scheme]]
-    summary_barplot(...) +
-        theme_bw() +
-        theme(axis.text.x = element_blank(),
-              axis.ticks.x = element_blank()) +
-        scale_fill_manual(values = pft_cols)
-}
-
-plots <- all_summaries %>%
-    filter(model_type == 'hier') %>%
-    mutate(
-        data = map(data, prep_data),
-        facet_list = ifelse(area_mass == 'mass', list(list()), list(list(facets = ~yparam + xparam))),
-        bar_plot = pmap(
-            list(
-                summary_data = data,
-                facet_list = facet_list,
-                pft_scheme = pft_scheme
-            ),
-            custom_barplot,
-            varname = 'Corr'
-        ),
-        filename = paste(
-            'correlation_barplot',
-            model_type,
-            area_mass,
-            pft_scheme,
-            'png',
-            sep = '.'
-        ),
-        filename = file.path(figures_dir, filename)
-    )
-
-# plots[[1, 'bar_plot']]
-
-plots %$%
-    pwalk(list(filename = filename, plot = bar_plot), ggsave, width = 9, height = 8)

### 2018-01-12 scripts/data_map.R

@@ -1,7 +1,6 @@
 library(tidyverse)
 
 try_dat <- readRDS("extdata/traits_analysis.rds")
-try_clim <- readRDS("extdata/traits_with_climate.rds")
 
 miss <- function(x) {
   as.integer(is.na(x))

### 2018-01-12 scripts/exploratory.R

@@ -1,13 +1,0 @@
-library(tidyverse)
-d <- readRDS("extdata/traits_analysis.rds")
-
-d %>%
-  filter(
-    #clm45 == "broadleaf_deciduous_temperate",
-    !is.na(Narea), !is.na(Vcmax_area)
-  ) %>%
-  ggplot() +
-  aes(x = Narea, y = Vcmax_area) +
-  geom_point() +
-  geom_smooth(method = "lm")
-

### 2018-01-12 scripts/fit_uni.R

@@ -1,8 +1,7 @@
 library(runjags)
-library(tidyverse)
+library(shiklomanov2017np)
 
 try_data <- readRDS("extdata/traits_analysis.rds")
-source("scripts/pft_abbreviations.R")
 source("scripts/informative_prior.R")
 
 jags_model <- "

### 2018-01-12 scripts/progress2output.R

@@ -1,43 +1,0 @@
-# Code for exploring status of runs in progress
-library(tidyverse)
-
-progress_dir <- "progress"
-output_dir <- "output"
-output_progress_dir <- file.path(output_dir, "from_progress")
-dir.create(output_progress_dir)
-nfiles <- 20
-
-submit_df <- read_delim('submit_df.dat', ' ', col_names = FALSE) %>%
-    setNames(c('model_type', 'area_mass', 'pft_scheme', 'pft'))
-
-get_fname <- function(fname) {
-    file_list <- list.files(output_dir, fname)
-    nfiles <- as.character(length(file_list))
-    switch(nfiles, `0` = NA_character_, `1` = file_list, "MANY")
-}
-
-find_files <- submit_df %>%
-    mutate(
-        file_string = paste(model_type, area_mass, pft_scheme, pft, sep = '.'),
-        file_name = map_chr(file_string, get_fname)
-    )
-
-missing_files <- find_files %>%
-    filter(is.na(file_name))
-print(missing_files)
-
-progress2output <- function(fname) {
-    print(fname)
-    all_files <- list.files(progress_dir, fname, full.names = TRUE)
-    sub_files <- tail(all_files, nfiles)
-    all_results <- Reduce(
-        mvtraits:::combine_results,
-        lapply(sub_files, readRDS)
-    )
-    out_fname <- file.path(output_progress_dir, paste0(fname, '.rds'))
-    saveRDS(all_results, out_fname)
-}
-
-missing_files %>%
-    pull(file_string) %>%
-    walk(progress2output)

### 2018-01-12 scripts/regression.R

@@ -1,20 +1,0 @@
-library(tidyverse)
-
-try_data <- readRDS('extdata/traits_analysis.rds')
-pft_data <- readRDS('extdata/all_pfts.rds')
-
-pft_sub <- pft_data %>%
-  select(AccSpeciesID:climate_zone)
-
-try_join <- try_data %>%
-  left_join(pft_sub)
-
-fit <- lm(log(SLA) ~ leaf_type * climate_zone * phenology * growth_form, data = try_join %>% mutate(species = factor(AccSpeciesID)))
-fit_anova <- anova(fit)
-fit_names <- rownames(fit_anova)
-fit_ss <- fit_anova[['Sum Sq']]
-fit_propvar <- fit_ss / sum(fit_ss)
-names(fit_propvar) <- fit_names
-sort(fit_propvar * 100, decreasing = TRUE)
-
-fit <- lm(Jmax_area ~ climate_zone + n_fixation + growth_form + ps_pathway + phenology + leaf_type + AccSpeciesID, data = try_join)

### 2018-01-12 scripts/results.2d.density.R

@@ -1,72 +1,0 @@
-library(mvtraits)
-
-# Plot settings
-cex <- 4
-cex.labels <- cex
-cex.axis <- cex * 0.3
-cex.main <- cex * 0.5
-pairs_density_settings <- function(uni, multi, hier, obs_means, ...) {
-    pairs_density(uni, multi, hier, obs_means,
-                  cex.axis = cex.axis,
-                  cex.labels = cex.labels,
-                  cex.main = cex.main, 
-                  ...)
-}
-
-# Draw global plot
-
-message("Loading univariate global...")
-uni.all.global <- readRDS("processed_output/sims.uni_global.rds")
-
-message("Loading multivariate global...")
-multi.all.global <- readRDS("processed_output/sims.multi_global.rds")
-
-message("Loading hierarchical output...")
-hier.all <- readRDS("processed_output/sims.hier.rds")
-
-# Get TRY data
-try_data <- loadTRYData("data/try.data.rds")
-obs.means.global <- try_data[, lapply(.SD, mean, na.rm=TRUE), 
-                           .SDcols = traits] %>% c() %>% unlist()
-
-pairs_path <- "figures/alexey_pairs"
-dir.create(pairs_path)
-message("Creating global figure...")
-graphic(file.path(pairs_path, "00.global"))
-pairs_density_settings(uni.all.global$mu, 
-              multi.all.global$mu, 
-              hier.all$mu_global, 
-              obs.means.global, 
-              main="Global")
-dev.off()
-
-# Draw plots by PFT
-uni.all.pft <- readRDS("processed_output/sims.uni_pft.rds")
-multi.all.pft <- readRDS("processed_output/sims.multi_pft.rds")
-for(i in 1:npft){
-    current_pft <- pft.names[i]
-    print(paste(i, current_pft))
-    uni.mus <- uni.all.pft[[current_pft]]$mu
-    multi.mus <- multi.all.pft[[current_pft]]$mu
-    hier.mus <- hier.all$mu_pft[,current_pft,]
-    obs.means <- try_data[pft == current_pft, 
-                          lapply(.SD, mean, na.rm=TRUE),
-                          .SDcols = traits] %>% c() %>% unlist()
-
-    graphic(file.path(pairs_path, sprintf("%02d.pft", i)))
-    pairs_density_settings(uni.mus, multi.mus, hier.mus, 
-                           obs.means, 
-                           main=paste(i, pft.names[i]))
-    dev.off()
-}
-
-# Create combined graphic
[…]

### 2018-01-12 scripts/run_model.R

@@ -2,6 +2,9 @@
 library(mvtraits)
 library(tidyverse)
 
+message("Current working directory: ", getwd())
+print(sessionInfo())
+
 try_data <- readRDS('extdata/traits_analysis.rds')
 
 # For testing
@@ -21,6 +24,8 @@ out_dir <- "output"
 dir.create(out_dir, showWarnings = FALSE)
 
 if (!exists("cmdargs")) cmdargs <- commandArgs(trailingOnly = TRUE)
+message("Running with the following arguments:")
+print(cmdargs)
 # Arguments:
 #   1. Model type. Must be either 'multi' or 'hier'
 #   2. Data type. Must be either 'area' or 'mass'

### 2018-01-12 scripts/summary_table.R

@@ -1,48 +1,0 @@
-library(mvtraits)
-library(tidyverse)
-
-source("scripts/pft_abbreviations.R")
-cachefile <- ".cache/mvtraits_results.rds"
-results_all <- readRDS(cachefile)
-
-results_hier <- results_all %>%
-  filter(model_type == "hier") %>%
-  select(mass_area, data) %>%
-  mutate(
-    mu_group = map(data, "mu_group"),
-    Sigma_group = map(data, "Sigma_group"),
-    pft_names = map(mu_group, names),
-    pft_dat = pmap(
-      list(mu_group, Sigma_group, pft_names),
-      ~tibble(pft = ..3, mu = ..1, Sigma = ..2)
-    )
-  ) %>%
-  unnest(pft_dat) %>%
-  mutate(
-    mu_mean = map(mu, "Mean"),
-    Sigma_mean = map(Sigma, "Mean") %>% map(diag),
-    param = map(mu_mean, names)
-  ) %>%
-  unnest(mu_mean, Sigma_mean, param)
-
-results_table <- results_hier %>%
-  mutate(
-    lo = 10 ^ (mu_mean - 1.96 * Sigma_mean),
-    mid = 10 ^ mu_mean,
-    hi = 10 ^ (mu_mean + 1.96 * Sigma_mean),
-    lo_f = formatC(lo, digits = 3),
-    mid_f = formatC(mid, digits = 3),
-    hi_f = formatC(hi, digits = 3),
-    value = sprintf("%s (%s - %s)", mid_f, lo_f, hi_f)
-  )
-results_table %>%
-  filter(!(param %in% c("SLA", "leaf_lifespan") & mass_area == "mass")) %>%
-  select(pft, param, value) %>%
-  mutate(
-    param = factor(param, both_params) %>% 
-      forcats::lvls_revalue(param_markdown[both_params]),
-    pft = factor(pft, abbr2pft) %>% forcats::lvls_revalue(pft2abbr)
-  ) %>%
-  spread(param, value) %>%
-  arrange(pft) %>%
-  pander::pandoc.table("")

### 2018-01-12 scripts/table_plot.R

@@ -1,48 +1,0 @@
-library(nptraits)
-library(magrittr)
-
-try_data <- readRDS("extdata/traits_analysis.rds")
-
-try_sub <- try_data %>%
-    select(-ObservationID, -AccSpeciesID, -LMA) %>%
-    select(jules1:custom, leaf_lifespan, SLA, Nmass, Narea, Pmass, Parea, Rdmass, Rdarea, Vcmax_mass, Vcmax_area, Jmax_mass, Jmax_area)
-
-count_nas <- function(dat, pft) {
-    dat %>%
-        mutate(pft_scheme = pft) %>%
-        rename_(.dots = list(pft = pft)) %>%
-        group_by(pft_scheme, pft) %>%
-        summarize_if(is.double, ~sum(!is.na(.))) %>%
-        mutate(pft = as.character(pft))
-}
-
-na_counts <- map_df(pft_schemes, count_nas, dat = try_sub)
-
-na_long <- na_counts %>%
-    gather(variable, value, -pft, -pft_scheme) %>%
-    mutate(variable = factor(variable) %>% forcats::fct_inorder())
-
-na_nested <- na_long %>%
-    group_by(pft_scheme) %>%
-    nest()
-
-missing_plot <- function(dat, pft_scheme, palette = "Paired") {
-    dat %>%
-        mutate(pft = factor(pft, pft_levels[[pft_scheme]]) %>% "levels<-"(names(pft_levels[[pft_scheme]]))) %>%
-        ggplot() +
-            aes(x = variable, y = value, color = pft) +
-            geom_point(position = position_dodge(width = 0.5), size = 2) +
-            scale_y_log10() +
-            labs(x = 'Trait', y = 'Number of non-missing observations', title = pft_scheme) +
-            theme(legend.position = 'bottom',
-                  axis.text.x = element_text(angle = 90, hjust = 1)) +
-            scale_color_brewer(palette = palette)
-}
-
-na_plots <- na_nested %>%
-    mutate(colors = c('Set1', 'Set1', 'Paired', 'Paired'),
-           plot = pmap(list(data, pft_scheme, colors), missing_plot),
-           filename = file.path('figures', paste('n_obs', pft_scheme, 'png', sep = '.')))
-
-na_plots %$%
-    pwalk(list(filename = filename, plot = plot), ggsave, width = 6, height = 4.5)

### 2018-01-12 manuscript/02_methods.Rmd
Move `manuscript-md` to `manuscript`

--- 2018-01-08 manuscript-md/02_methods.Rmd
@@
 # Materials and Methods
-
-```{r options, echo=FALSE, message=FALSE}
-library(knitr)
-opts_chunk$set(
-  echo = FALSE,
-  message = FALSE, 
-  warning = FALSE,
-  dpi = 300,
-  cache = TRUE
-)
-options(knitr.table.format = "pandoc")
-suppressPackageStartupMessages({
-  library(tidyverse)
-})
-source("scripts/pft_abbreviations.R")
-try_data <- readRDS("extdata/traits_analysis.rds")
-
-try_sub <- try_data %>%
-  select(pft = clm45, one_of(both_params),
-         -Latitude, -Longitude, -LMA)
-
-sample_size <- try_sub %>%
-  select(pft, one_of(both_params)) %>%
-  mutate(
-    pft = factor(pft, abbr2pft) %>% forcats::lvls_revalue(pft2abbr)
-  ) %>%
-  group_by(pft) %>%
-  summarize_all(~sum(!is.na(.))) %>%
-  gather("param", "sample_size", -pft) %>%
-  mutate(param = factor(param, both_params))
-```
 
@@
 
-```{r "trydata"}
-crtable <- readRDS("extdata/crossref_table.rds") %>%
-  mutate(
-    Reference = case_when(
-      Reference == "@try_data_03" ~ "@baraloto_decoupled_2010",
-      Reference == "@try_data_09" ~ "@wright_worldwide_2004",
-      Reference == "@try_data_14" ~ "@kattge_2009_vcmax",
-      Reference == "@try_data_15" ~ "@laughlin_multi-trait_2010",
-      Reference == "@try_data_26" ~ "@onoda_2011_global",
-      Reference == "@try_data_30" ~ "@messier_how_2010",
-      Reference == "@try_data_33" ~ "@pierce_allocating_2013",
-      Reference == "@try_data_41" ~ "@poorter_causes_2009",
-      Reference == "@try_data_44" ~ "@choat_global_2012",
-      Reference == "@try_data_59" ~ "@laughlin_climatic_2011",
-      Reference == "@try_data_67" ~ "@wright_does_2012",
-      Reference == "@try_data_80" ~ "@choat_global_2012",
-      Reference == "@try_data_114" ~ "@kichenin_2013_contrasting",
-      Reference == "@try_data_134" ~ "@diaz_plant_2004",
-      TRUE ~ Reference
-    )
-  )
-cap <- "Data sources and observation counts for foliar trait data"
-kable(crtable, caption = cap)
-```
-
 ## Plant functional types
-
-```{r "pfts"}
-n_species <- try_data %>%
-  distinct(clm45, AccSpeciesID) %>%
-  select(pft = clm45, species = AccSpeciesID) %>%
-  count(pft) %>%
-  mutate(
-    pft = as.character(pft),
-    label = pft2abbr[pft],
-    pft = tools::toTitleCase(gsub("_", " ", pft))
-  ) %>%
-  select(`Label` = label, `PFT` = pft, `Number of species` = n)
-species_caption <- paste(
-  "Names, labels, and species counts for plant functional types (PFTs)",
-  "used in this analysis."
-)
-kable(n_species, caption = species_caption, align = rep("c", ncol(n_species)))
-```
-
-(ref:ssfigcap) Sample sizes for each trait-PFT pair. $y$ axis is scaled logarithmically.
-
-```{r "samplesize", fig.cap='(ref:ssfigcap)', fig.width=7, fig.height=7}
-ss_plot <- sample_size %>%
-  mutate(param = forcats::lvls_revalue(param, param_simple[both_params])) %>%
-  ggplot() + 
-  aes(x = pft, y = sample_size, fill = pft) + 
-  geom_col() +
-  scale_y_log10(breaks = c(10, 100, 1000, 10000)) +
-  facet_wrap(~param, ncol = 2, labeller = label_parsed) +
-  scale_fill_manual(values = pft_colors) +
-  xlab("Plant functional type") +
-  ylab("Sample size") +
-  guides(fill = guide_legend(title = "PFT")) +
-  theme_bw() +
-  theme(
-    axis.text.x = element_blank(),
-    axis.ticks.x = element_blank()
-  )
-ss_plot
-```
 

### 2018-01-12 manuscript/03_results.Rmd
Move `manuscript-md` to `manuscript`

--- 2018-01-08 manuscript-md/03_results.Rmd
@@
 
-```{r prepresults}
-cachefile <- ".cache/mvtraits_results.rds"
-results_all <- readRDS(cachefile)
-
-max_ss <- max(sample_size$sample_size)
-
-multi_summary <- results_all %>%
-  filter(model_type %in% c("uni", "multi")) %>%
-  mutate(
-    mu_mean = map(data, c("mu", "Mean")) %>%
-      map(~as_tibble(as.list(.))),
-    mu_lo = map(data, c("mu", "2.5%")) %>%
-      map(~as_tibble(as.list(.))),
-    mu_hi = map(data, c("mu", "97.5%")) %>%
-      map(~as_tibble(as.list(.)))
-      ) %>%
-  select(model_type:pft, mu_mean:mu_hi) %>%
-  unnest(.sep = ".") %>%
-  gather("variable", "value", -(model_type:pft), na.rm = TRUE) %>%
-  separate(variable, c("stat", "param"), sep = "\\.") %>%
-  spread(stat, value) %>%
-  mutate(
-    pft = factor(pft, abbr2pft) %>% forcats::lvls_revalue(pft2abbr),
-    param = factor(param, both_params)
-  )
-
-hier_prep <- results_all %>%
-  filter(model_type == "hier") %>%
-  select(model_type:pft_type, data) %>%
-  mutate(
-    mu_vals = map(data, "mu_group"),
-    pft = map(mu_vals, names),
-    mu_df = map2(pft, mu_vals, ~tibble(pft = ..1, dat = ..2))
-  ) %>%
-  select(-data, -mu_vals, -pft) %>%
-  unnest()
-
-hier_summary <- hier_prep %>%
-  mutate(
-    mu_mean = map(dat, c("Mean")) %>%
-      map(~as_tibble(as.list(.))),
-    mu_lo = map(dat, c("2.5%")) %>%
-      map(~as_tibble(as.list(.))),
-    mu_hi = map(dat, c("97.5%")) %>%
-      map(~as_tibble(as.list(.)))
-      ) %>%
-  select(model_type:pft, mu_mean:mu_hi) %>%
-  unnest(.sep = ".") %>%
-  gather("variable", "value", -(model_type:pft), na.rm = TRUE) %>%
-  separate(variable, c("stat", "param"), sep = "\\.") %>%
-  spread(stat, value) %>%
-  mutate(
-    pft = if_else(is.na(pft), "global", pft),
-    pft = factor(pft, abbr2pft) %>% forcats::lvls_revalue(pft2abbr),
-    param = factor(param, both_params)
-  )
-
-# Absolute comparison of means
-means_dat <- multi_summary %>%
-  filter(pft != "GLOB", !is.na(pft)) %>%
-  bind_rows(hier_summary) %>%
-  left_join(sample_size) %>%
-  mutate(
-    model_type = factor(model_type, c("uni", "multi", "hier")) %>%
-      forcats::lvls_revalue(c("univariate", "multivariate", "hierarchical")),
-    mass_area = factor(mass_area, c("mass", "area")),
-  ) %>%
-  mutate_at(
-    c("mu_mean", "mu_lo", "mu_hi"),
-    ~10 ^ .
-  ) %>%
-  filter(
-    !(param %in% c("leaf_lifespan", "SLA") & mass_area == "area")
-  )
-```
-
 ## Estimates of PFT-level means
-
-```{r meanvaluetable, results = "asis"}
-## Write summary table
-table_cap_mass <- paste(
-  "Mean and 95\\% confidence interval of trait estimates",
-  "for mass-normalized traits from the hierarchical model."
-)
-table_cap_area <- gsub("mass", "area", table_cap_mass)
-
-censor <- function(rxp, pft, pfts, param, value) {
-  if_else(grepl(rxp, param) & pfts %in% pft, "---", value)
-}
-
-table_dat <- means_dat %>%
-  filter(
-    model_type == "hierarchical",
-    !(param %in% c("SLA", "leaf_lifespan") & mass_area == "area")
-  ) %>%
-  mutate(
-    mu_lo = if_else(param == "Rdmass", mu_lo * 1000, mu_lo),
-    mu_mean = if_else(param == "Rdmass", mu_mean * 1000, mu_mean),
-    mu_hi = if_else(param == "Rdmass", mu_hi * 1000, mu_hi),
-    mu_lo = if_else(param == "Rdarea", mu_lo * 1000, mu_lo),
-    mu_mean = if_else(param == "Rdarea", mu_mean * 1000, mu_mean),
-    mu_hi = if_else(param == "Rdarea", mu_hi * 1000, mu_hi),
-    lo_f = formatC(mu_lo, digits = 3),
-    mid_f = formatC(mu_mean, digits = 3),
-    hi_f = formatC(mu_hi, digits = 3),
-    value = sprintf("%s (%s,%s)", mid_f, lo_f, hi_f),
-    value = censor("Vcmax|Jmax", "C4G", pft, param, value),
-    value = censor("Rdmass", "ShDBo", pft, param, value),
-    value = censor("Jmax_mass", c("NlEBo", "NlD"), pft, param, value),
-    value = censor("Jmax_area", c("NlD"), pft, param, value)
-  ) %>%
-  select(pft, param, value) %>%
-  mutate(
-    param = factor(param, both_params) %>% 
-      forcats::lvls_revalue(param_markdown_nounit[both_params]),
-  )%>%
-  spread(param, value) %>%
-  arrange(pft)
-
-table_dat %>%
-  select(-matches("area")) %>%
-  kable(caption = table_cap_mass, format = "latex",
-        escape = FALSE, booktabs = TRUE) %>%
-  kableExtra::kable_styling(latex_options = c("scale_down"))
-table_dat %>%
-  select(-matches("mass")) %>%
-  kable(caption = table_cap_area, format = "latex",
-        escape = FALSE, booktabs = TRUE) %>%
-  kableExtra::kable_styling(latex_options = c("scale_down"))
-```
-
-(ref:mccap) Mean and 95% confidence interval on best estimates of traits for each plant functional type from the univariate, multivariate, and hierarchical models.
-For leaf lifespan and SLA, results were not significantly different between the mass- and area-based models, so only results from the mass-based model are shown.
-For some PFT-trait combinations, where large error bars resulting from the relatively uninformative priors are substantially larger than the variability among means, the $y$ axes are constrained to facilitate comparison.
-
-```{r meancompare, fig.width=7, fig.height=7, fig.cap='(ref:mccap)'}
-pclip <- function(p, pname, value, lim, hi = TRUE) {
-  f <- ifelse(hi, `>`, `<`)
-  if_else(p == pname & f(value, lim), lim, value)
-}
-
-plot_dat <- means_dat %>%
-  mutate(
-    #too_few = sample_size < 2,
-    #mu_hi = if_else(too_few, NA_real_, mu_hi),
-    #mu_lo = if_else(too_few, NA_real_, mu_lo),
-    #mu_mean = if_else(too_few, NA_real_, mu_mean),
-    irrelevant = grepl("Vcmax|Jmax", param) & pft == "C4G",
-    mu_hi = if_else(irrelevant, NA_real_, mu_hi),
-    mu_lo = if_else(irrelevant, NA_real_, mu_lo),
-    mu_mean = if_else(irrelevant, NA_real_, mu_mean),
-    ## TODO: Fix these in the prior and re-run
-    mu_hi = pclip(param, "Vcmax_mass", mu_hi, 1.5),
-    mu_hi = pclip(param, "Vcmax_area", mu_hi, 100),
-    mu_hi = pclip(param, "Jmax_mass", mu_hi, 1.8),
-    mu_hi = pclip(param, "Jmax_area", mu_hi, 150),
-    mu_mean = pclip(param, "Jmax_area", mu_mean, 150),
-    mu_hi = pclip(param, "Rdmass", mu_hi, 0.03),
-    mu_mean = pclip(param, "Rdmass", mu_mean, 0.03),
-    mu_lo = pclip(param, "Rdmass", mu_lo, 0.03),
-    mu_hi = pclip(param, "Rdarea", mu_hi, 0.0025),
-    mu_mean = pclip(param, "Rdarea", mu_mean, 0.0025),
-    mu_lo = pclip(param, "Rdarea", mu_lo, 0.0025),
-    mu_hi = pclip(param, "Parea", mu_hi, 0.285),
-    #mu_lo = pclip(param, "Jmax_area", mu_lo, 40, FALSE),
-    param = forcats::lvls_revalue(param, param_fancy_chr[both_params])
-  )
-
-p_means <- ggplot(plot_dat) +
-  aes(x = interaction(model_type, pft),
-      y = mu_mean, ymin = mu_lo, ymax = mu_hi,
-      color = pft, shape = model_type) +
-  geom_pointrange(size = 0.3) +
-  facet_wrap(~param, scales = "free", ncol = 2,
-              labeller = label_parsed) +
-  scale_y_continuous() +
-  scale_color_manual(values = pft_colors) +
-  guides(
-    shape = guide_legend(title = "Model type"),
-    color = guide_legend(title = "PFT")
-  ) +
-  ylab("Trait estimate mean and 95% CI") +
-  xlab("Model type and PFT") +
-  theme_bw() +
-  theme(
-    text = element_text(size = 12),
-    axis.text.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    panel.grid = element_blank()
-  )
-p_means
-```
 
@@
 
-```{r "civssamplesize", fig.width=4.4, fig.height=3, fig.cap='(ref:sscap)'}
-## Relative uncertainty
-cv_dat <- means_dat %>%
-  mutate(
-    cv = abs((mu_hi - mu_lo) / mu_mean)
-  ) %>%
-  filter(sample_size > 1)
-
-cv_mod <- cv_dat %>%
-  group_by(model_type) %>%
-  nest() %>%
-  mutate(
-    cvfit = map(data, ~lm(log10(cv) ~ log10(sample_size), data = .))
-  )
-
-cv_coef <- cv_mod %>%
-  mutate(
-    coef = map(cvfit, broom::tidy)
-  ) %>%
-  unnest(coef) %>%
-  select(model_type, term, estimate) %>%
-  spread(term, estimate) %>%
-  "names<-"(c("model_type", "intercept", "slope")) %>%
-  mutate(
-    eqn = sprintf("'%s: '~log[10](y) == %0.2f - %0.2f~log[10](x)", model_type, intercept, abs(slope))
-  )
-
-cv_pred <- cv_mod %>%
-  mutate(
-    grid = map(data, modelr::data_grid, sample_size = seq(5, max_ss)),
-    cvpred = map2(grid, cvfit, modelr::add_predictions)
-  ) %>%
-  unnest(cvpred) %>%
-  mutate(cv = 10 ^ pred)
-
-x_lims <- range(cv_pred$sample_size)
-y_lims <- range(cv_pred$cv)
-
-relative_ci <- ggplot(cv_dat) +
-  aes(x = sample_size, y = cv, color = model_type) +
-  geom_point(size = 0.5) +
-  geom_line(data = cv_pred) +
-  annotate("text", x = rep(2500, 3), y = c(1.4, 1.25, 1.1), 
-           label = cv_coef$eqn, parse = TRUE, 
-           size = 2, vjust = "right") +
-  scale_x_continuous(trans = "log10") +
-  coord_cartesian(xlim = x_lims, ylim = y_lims) +
-  xlab("Sample size") +
-  ylab("Relative width of 95% CI") +
-  scale_color_brewer(type = "qual", palette = 2) +
-  guides(color = guide_legend(title = "Model type")) +
-  theme_bw() +
-  theme(
-    legend.position = c(0.85, 0.75),
-    panel.grid = element_blank()
-  )
-relative_ci
-```
-
 ## Trait correlation patterns among- and within-PFTs
-
-(ref:stickcap) Pairwise trait mean and covariance estimates for all data pooled globally (black) and for each PFT (colored).
-Covariance estimates not significantly different from zero ($p < 0.05$) are indicated by x symbols at the mean estimate.
-$x$ and $y$ axes vary on a log scale, reflecting the fact that the model was fit using the base 10 log of all traits.
-With the exception of leaf lifespan, pairwise covariances are consistent in direction but vary somewhat in magnitude between PFTs, and when comparing PFT-level and global estimates.
-However, many pairwise covariances are not statistically significant, particularly (but not always) for undersampled traits and PFTs.
-
-```{r stickpairs, fig.width = 5, fig.height = 5, fig.cap = '(ref:stickcap)'}
-suppressPackageStartupMessages({
-  library(mvtraits)
-})
-results_sub <- results_all %>%
-  filter((model_type == "multi" & pft == "global") |
-         (model_type == "hier"))
-dat_list <- results_sub$data
-names(dat_list) <- with(results_sub, paste(model_type, mass_area, sep = "_"))
-
-stickplot_pairs(
-  #mu_global_lower = dat_list$hier_mass$mu_global,
-  #Sigma_global_lower = dat_list$hier_mass$Sigma_global,
-  mu_global_lower = dat_list$multi_mass$mu,
-  Sigma_global_lower = dat_list$multi_mass$Sigma,
-  mu_group_lower = dat_list$hier_mass$mu_group,
-  Sigma_group_lower = dat_list$hier_mass$Sigma_group,
-  vars_lower = mass_params,
-  group_names = pft2abbr[-1],
-  vars_label = param_labels,
-  unlog_axes = TRUE,
-  #mu_global_upper = dat_list$hier_area$mu_global,
-  #Sigma_global_upper = dat_list$hier_area$Sigma_global,
-  mu_global_upper = dat_list$multi_area$mu,
-  Sigma_global_upper = dat_list$multi_area$Sigma,
-  mu_group_upper = dat_list$hier_area$mu_group,
-  Sigma_group_upper = dat_list$hier_area$Sigma_group,
-  vars_upper = area_params,
-  col_group = pft_colors,
-  cex_g = 1.2,
-  lwd_g = 1,
-  lty_g = 1,
-  lwd_s = 1.3,
-  par_plot = list(oma = c(0, 5, 3, 5), cex = 0.4),
-  par_legplot = list(mar = c(0, 0, 0, 0)),
-  par_legend = list(ncol = 5, x = "center", cex = 0.6),
-  reorder_legend = TRUE
-)
-mtext("Mass-normalized", side = 2, cex = 0.7, line = 3)
-mtext("Area-normalized", side = 4, cex = 0.7, line = 1)
-```
 
@@
 
-```{r boxprep}
-outdir <- "../output"
-hier_mass_file <- tail(dir(outdir, "hier.mass.clm45", full.names = TRUE), 1)
-hier_area_file <- tail(dir(outdir, "hier.area.clm45", full.names = TRUE), 1)
-multi_mass_file <- tail(dir(outdir, "multi.mass.clm45", full.names = TRUE), 1)
-multi_area_file <- tail(dir(outdir, "multi.area.clm45", full.names = TRUE), 1)
-
-hier_summary_mass <- readRDS(hier_mass_file)[["summary_table"]] %>%
-  mutate(mass_area = "mass")
-hier_summary_area <- readRDS(hier_area_file)[["summary_table"]] %>%
-  mutate(mass_area = "area")
-multi_summary_mass <- readRDS(multi_mass_file)[["summary_table"]] %>%
-  mutate(mass_area = "mass")
-multi_summary_area <- readRDS(multi_area_file)[["summary_table"]] %>%
-  mutate(mass_area = "area")
-
-proc <- function(dat, hier = TRUE) {
-  if (hier) {
-    dat <- filter(dat, group != "global")
-  } else {
-    dat <- mutate(dat, group = "global")
-  }
-  dat <- dat %>%
-    filter(variable == "Corr") %>%
-    separate(index, c("xvar", "yvar"), sep = "\\.\\.") %>%
-    mutate(
-      pft = factor(group, abbr2pft) %>% forcats::lvls_revalue(pft2abbr)
-    ) %>%
-    select(mass_area, pft, xvar, yvar, Mean, `2.5%`, `97.5%`)
-}
-
-hsm <- proc(hier_summary_mass)
-hsa <- proc(hier_summary_area)
-msm <- proc(multi_summary_mass, FALSE)
-msa <- proc(multi_summary_area, FALSE)
-hs <- bind_rows(hsm, hsa, msm, msa)
-```
-
-(ref:corrbarscap) Mean and 95% CI on estimates of pairwise correlation coefficients
-for all data pooled globally (dark grey) and for each PFT (colored).
-For most PFT-trait pairs, correlations are mutually consistent in magnitude but vary in strength.
-
-```{r corrbars, fig.width=7, fig.height=7, fig.cap='(ref:corrbarscap)'}
-.zzz <- close.screen(all.screens = TRUE)
-corr_bar <- function(dat, xvar, yvar, mass_area) {
-  dp <- dat %>%
-    filter(
-      mass_area == !!mass_area,
-      (xvar == !!xvar & yvar == !!yvar) | 
-        (xvar == !!yvar & yvar == !!xvar),
-    ) %>%
-    arrange(pft) %>%
-    mutate(color = c("grey25", pft_colors))
-  x <- barplot(dp$Mean, col = dp$color, border = dp$color,
-               ylim = c(-1, 1), axes = FALSE)
-  arrows(x, dp[["2.5%"]], x, dp[["97.5%"]],
-         length = 0.01, angle = 90, code = 3)
-  box()
-}
-
-nv <- length(mass_params)
-main_screens <- split.screen(rbind(c(0, 1, 0.2, 1), c(0, 1, 0, 0.2)))
-par(mar = rep(0.1, 4), oma = c(0, 8, 1, 5), cex = 0.4)
-screens <- split.screen(c(nv, nv), screen = main_screens[1])
-for (i in seq_len(nv)) {
-  for (j in seq_len(nv)) {
-    k <- j + nv * (i - 1)
-    screen(screens[k])
-    if (i == j) {
-      # Label
-      plot(0:1, 0:1, type = "n", ann = FALSE, axes = FALSE)
-      box()
-      text(x = 0.5, y = 0.5, labels = param_labels[i], cex = 2)
-    } else {
-      if (j > i) {
-        # Upper triangle
-        mass_area <- "area"
-        xvar <- area_params[j]
-        yvar <- area_params[i]
-      } else {
-        # Lower triangle
-        mass_area <- "mass"
-        xvar <- mass_params[j]
-        yvar <- mass_params[i]
-      }
-      corr_bar(hs, xvar, yvar, mass_area)
-    }
-    if (i == 2 && j == 1) {
-      axis(side = 2)
-    }
-    if (i == 1 && j == nv) {
-      axis(side = 4)
-    }
-  }
-}
-## Draw legend
-screen(main_screens[2])
-par(mar = c(0, 0, 0, 0), oma = c(0, 0, 0, 0))
-plot(0, 0, type = "n", axes = FALSE, ann = FALSE)
-legend("center", legend = levels(hs$pft), fill = c("grey25", pft_colors), 
-       ncol = 5, bty = "n")
-close.screen(all.screens = TRUE)
-mtext("Mass-normalized", side = 2, cex = 0.7, line = 1)
-mtext("Mean and 95% CI of pairwise correlation", side = 2, cex = 1, line = 2)
-mtext("Area-normalized", side = 4, cex = 0.7, line = 0)
-```
-
 The strength of pairwise trait correlations varied substantially depending on scale, PFT, and trait (Fig.\ \@ref(fig:corrbars)).

### 2018-01-12 purld/wordcount.R

@@ -7,7 +7,7 @@ fnames <- c(
   conclusions = "05_conclusions.md"
 )
 files <- file.path(
-  "~/Projects/new-phytologist-traits/np-trait-analysis/manuscript-md",
+  "~/Projects/new-phytologist-traits/np-trait-analysis/manuscript",
   fnames
 )
 names(files) <- names(fnames)

### 2018-01-12 sh/delete.runs.sh
Remove deprecated `sh` directory

@@ -1,5 +1,0 @@
-#!/bin/bash
-
-while read id; do
-    qdel $id
-done < qstat.txt

### 2018-01-12 sh/purge_runs.sh

@@ -1,3 +1,0 @@
-#!/bin/bash
-rm -rf output
-rm -f logs/*

### 2018-01-12 sh/rsync.sh

@@ -1,7 +1,0 @@
-#!/bin/bash
-
-rsync -avz --progress geo:~/dietzelab/new_phytologist_traits/np-trait-analysis/results .
-#rsync -avz --progress geo:~/dietzelab/np_trait_analysis/processed_output/ processed_output/
-#rsync -avz geo:~/dietzelab/np_trait_analysis/output.n1/ output.n1/
-#rsync -avz geo:~/dl-full/ecowdery/NP_Trait_Analysis/output.n1/ output/
-#rsync -avz geo:~/dl-full/ecowdery/NP_Trait_Analysis/old.output/output/hier.trait.pft.na/ output/hier.trait.pft.na/

### 2018-01-12 R/clm45_parameters.R
Create template for CLM 4.5 data table

@@ -1,0 +1,7 @@
+#' CLM Table 8.1
+#'
+#' Community Land Model (CLM) version 4.5 parameters table (8.1), for 
+#' comparison with results.
+clm45_table81 <- tribble(
+  ~pft
+)

### 2018-02-13 scripts/manuscript-analyses/figure_correlation_boxplot.R
More script updates

@@ -55,6 +55,14 @@ corr_bar <- function(dat, xvar, yvar, mass_area) {
   box()
 }
 
+if (!interactive()) {
+  pdf(
+    here("figures", "manuscript", "correlation_boxplot.pdf"),
+    width = 7,
+    height = 7
+  )
+}
+
 nv <- length(mass_params)
 main_screens <- split.screen(rbind(c(0, 1, 0.2, 1), c(0, 1, 0, 0.2)))
 par(mar = rep(0.1, 4), oma = c(0, 8, 1, 5), cex = 0.4)
@@ -94,10 +102,12 @@ for (i in seq_len(nv)) {
 screen(main_screens[2])
 par(mar = c(0, 0, 0, 0), oma = c(0, 0, 0, 0))
 plot(0, 0, type = "n", axes = FALSE, ann = FALSE)
-legend("center", legend = levels(hs$pft), fill = c("grey25", pft_colors), 
+legend("center", legend = levels(hs$pft), fill = c("grey25", pft_colors),
        ncol = 5, bty = "n")
 close.screen(all.screens = TRUE)
 mtext("Mass-normalized", side = 2, cex = 0.7, line = 1)
 mtext("Mean and 95% CI of pairwise correlation", side = 2, cex = 1, line = 2)
 mtext("Area-normalized", side = 4, cex = 0.7, line = 0)
-
+if (!interactive()) {
+  dev.off()
+}

### 2018-02-13 scripts/manuscript-analyses/figure_samplesize.R

@@ -1,39 +1,0 @@
-# Chunk name: samplesize
-# Description: Figure of the sample sizes of each PFT
-
-## ----options, echo=FALSE, message=FALSE----------------------------------
-library(shiklomanov2017np)
-try_data <- readRDS(here("extdata/traits_analysis.rds"))
-
-try_sub <- try_data %>%
-  select(pft = clm45, one_of(both_params),
-         -Latitude, -Longitude, -LMA)
-
-sample_size <- try_sub %>%
-  select(pft, one_of(both_params)) %>%
-  mutate(
-    pft = factor(pft, abbr2pft) %>% forcats::lvls_revalue(pft2abbr)
-  ) %>%
-  group_by(pft) %>%
-  summarize_all(~sum(!is.na(.))) %>%
-  gather("param", "sample_size", -pft) %>%
-  mutate(param = factor(param, both_params))
-
-## ----"samplesize", fig.cap='(ref:ssfigcap)', fig.width=7, fig.height=7----
-ss_plot <- sample_size %>%
-  mutate(param = forcats::lvls_revalue(param, param_simple[both_params])) %>%
-  ggplot() +
-  aes(x = pft, y = sample_size, fill = pft) +
-  geom_col() +
-  scale_y_log10(breaks = c(10, 100, 1000, 10000)) +
-  facet_wrap(~param, ncol = 2, labeller = label_parsed) +
-  scale_fill_manual(values = pft_colors) +
-  xlab("Plant functional type") +
-  ylab("Sample size") +
-  guides(fill = guide_legend(title = "PFT")) +
-  theme_bw() +
-  theme(
-    axis.text.x = element_blank(),
-    axis.ticks.x = element_blank()
-  )
-ss_plot

### 2018-02-13 scripts/manuscript-analyses/many_results.R

@@ -1,19 +1,63 @@
+# Chunk name: samplesize
+# Description: Figure of the sample sizes of each PFT
+
+## ----options, echo=FALSE, message=FALSE----------------------------------
 library(shiklomanov2017np)
+manuscript_fig_dir <- here("figures", "manuscript")
+dir.create(manuscript_fig_dir, showWarnings = FALSE)
+
+try_data <- readRDS(here("extdata/traits_analysis.rds"))
+
+try_sub <- try_data %>%
+  select(pft = clm45, !!both_params)
+
+sample_size <- try_sub %>%
+  select(pft, one_of(both_params)) %>%
+  mutate(
+    pft = factor(pft, abbr2pft) %>% forcats::lvls_revalue(pft2abbr)
+  ) %>%
+  group_by(pft) %>%
+  summarize_all(~sum(!is.na(.))) %>%
+  gather("param", "sample_size", -pft) %>%
+  mutate(param = factor(param, both_params))
+
+## ----"samplesize", fig.cap='(ref:ssfigcap)', fig.width=7, fig.height=7----
+ss_plot <- sample_size %>%
+  mutate(param = forcats::lvls_revalue(param, param_simple[both_params])) %>%
+  ggplot() +
+  aes(x = pft, y = sample_size, fill = pft) +
+  geom_col() +
+  scale_y_log10(breaks = c(10, 100, 1000, 10000)) +
+  facet_wrap(~param, ncol = 2, labeller = label_parsed) +
+  scale_fill_manual(values = pft_colors) +
+  xlab("Plant functional type") +
+  ylab("Sample size") +
+  guides(fill = guide_legend(title = "PFT")) +
+  theme_bw() +
+  theme(
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank()
+  )
+if (interactive()) {
+  ss_plot
+}
+ggsave(file.path(manuscript_fig_dir, "sample_size.pdf"), ss_plot, 
+        width = 7, height = 7)
 
 ## ----prepresults---------------------------------------------------------
-cachefile <- here::here(".cache/mvtraits_results.rds")
-results_all <- readRDS(cachefile)
+resultsfile <- here("results/mvtraits_results.rds")
+results_all <- readRDS(resultsfile)
 
[…]

### 2018-02-13 scripts/manuscript-analyses/table_n_species.R

@@ -1,9 +1,9 @@
-# Chunk name: "pfts"
+# Chunk name: pfts
 # Description: Table of number of species in each PFT
+
+## ---- pfts --------------------------------------------------------------
 library(shiklomanov2017np)
 try_data <- readRDS("extdata/traits_analysis.rds")
-
-## ----"pfts"--------------------------------------------------------------
 n_species <- try_data %>%
   distinct(clm45, AccSpeciesID) %>%
   select(pft = clm45, species = AccSpeciesID) %>%

### 2018-02-13 scripts/manuscript-analyses/try_ref_table.R

@@ -1,8 +1,9 @@
 # Chunk name: trydata
 # Description: Table of TRY data citations
 
+## ---- trydata ----------
 library(shiklomanov2017np)
-crtable <- readRDS("extdata/crossref_table.rds") %>%
+crtable <- readRDS(here("extdata/crossref_table.rds")) %>%
   mutate(
     Reference = case_when(
       Reference == "@try_data_03" ~ "@baraloto_decoupled_2010",

### 2018-02-13 manuscript/02_methods.Rmd
Revise methods
- Add info on flux trait temperature corrections
- Add info on CLM values
- Replace repository link with Open Science Framework link
- Minor text revisions

@@
 leaf phosphorus content ($P_{mass}$, mg P g$^{-1}$ or $P_{area}$, g m$^{-2}$),
-leaf dark respiration at leaf temperature ($R_{d,mass}$, µmol g$^{-1}$ s$^{-1}$, or $R_{d,area}$, µmol m$^{-2}$ s$^{-1}$),
+leaf dark respiration at 25°C ($R_{d,mass}$, µmol g$^{-1}$ s$^{-1}$, or $R_{d,area}$, µmol m$^{-2}$ s$^{-1}$),
 maximum Rubisco carboxylation rate at 25°C ($V_{c,max,mass}$, µmol g$^{-1}$ s$^{-1}$, or $V_{c,max,area}$, µmol m$^{-2}$ s$^{-1}$),
-and maximum electron transport rate ($J_{max,mass}$, µmol g$^{-1}$ s$^{-1}$, or $J_{max,area}$, µmol m$^{-2}$ s$^{-1}$).
-We performed analyses separately for both mass- and area-normalized traits to avoid potential artifacts caused by different trait normalization [@osnas_global_2013; @lloyd_les].
+and maximum electron transport rate at 25°C ($J_{max,mass}$, µmol g$^{-1}$ s$^{-1}$, or $J_{max,area}$, µmol m$^{-2}$ s$^{-1}$).
+For $V_{c,max}$, we only used values already reported in TRY as being at 25°C.
+For $R_{d}$, we normalized the values to 25°C based on reported leaf temperature values following the same methods as @atkin_global_2015.
+For $J_{max}$, we normalized the values to 25°C based on reported leaf temperature values using the temperature response function described in @kattge_2007_temperature (Equation 1 therein).
+To avoid potential artifacts caused by different trait normalization, we performed analyses separately for both mass- and area-normalized traits [@osnas_global_2013; @lloyd_les].
 We restricted our analysis to TRY data that have been quality-controlled and for which adequate species information was provided for functional type classification [see @kattge_try_2011].
@@
 
-We assigned each species to a unique plant functional type (PFT) following the scheme in the Community Land Model [CLM4.5, @clm45_note; Tab.\ \@ref(tab:pfts), Fig.\ \@ref(fig:samplesize)].
+We assigned each species to a unique plant functional type (PFT) following the scheme in the Community Land Model [CLM4.5, @clm45_note; Tab.\ \@ref(tab:pfts), Fig.\ 1].
 We obtained categorical data on growth form, leaf type, phenology, and photosynthetic pathway from the TRY database.
@@
 
-For each model fit, we ran five parallel MCMC chains, automatically increasing the sample size until the final result achieved convergence as determined by a Gelman-Rubin potential scale reduction statistic less than 1.1 [@gelman_1992_inference].
-We implemented this sampling algorithm in an open source, publicly available R [version 3.4.2; @rstats] package (<http://github.com/ashiklom/mvtraits>).
+For each model fit, we ran five parallel MCMC chains, continuing the sampling until the final result achieved convergence as determined by a Gelman-Rubin potential scale reduction statistic less than 1.1 [@gelman_1992_inference].
+We implemented this sampling algorithm in an open source, publicly available R [version 3.4.3; @rstats] package (<http://github.com/ashiklom/mvtraits>).
 
@@
 To assess the impact of multivariate and hierarchical constraint on trait estimates,
-we compared the mean and 95% confidence intervals of trait estimates for each PFT from each model (Fig.\ \@ref(fig:meancompare)).
+we compared the mean and 95% confidence intervals of trait estimates for each PFT from each model (Fig.\ 2).
+For reference, we also added the default parameter values of CLM 4.5 [Table 8.1 in @clm45_note] for SLA, $N_{mass}$, $N_{area}$, $V_{c,max,mass}$, and $V_{c,max,area}$ to Fig.\ 2.
+To convert CLM's reported C:N ratio to $N_{mass}$, we assumed a uniform leaf C fraction of 0.46.
+We then divided this calculated $N_{mass}$ by the reported SLA to obtain $N_{area}$.
+We calculated $V_{c,max,mass}$ by multiplying the reported $V_{c,max,area}$ by the reported SLA.
 
@@
 
-We then fit a generalized linear model relating relative uncertainty to sample size ($n$) for each of the model types (univariate, multivariate, and hierarchical; Fig. \@ref(fig:civssamplesize)).
+We then fit a generalized linear model relating relative uncertainty to sample size ($n$) for each of the model types (univariate, multivariate, and hierarchical; Fig. 3).
 
@@
 To assess the consistency of among- and within-PFT trait trade-offs, we looked at covariance estimates for each trait pair and, where these values were significantly different from zero ($p < 0.05$),
-we calculated the eigenvalues from the pairwise variance-covariance matrix for that trait pair and plotted the corresponding dominant eigenvectors centered on the mean estimates (Fig.\ \@ref(fig:stickpairs)).
+we calculated the eigenvalues from the pairwise variance-covariance matrix for that trait pair and plotted the corresponding dominant eigenvectors centered on the mean estimates (Fig.\ 4).
 This figure provides a visual representation of relative positions of PFTs in trait space and both the direction and extent of within-PFT trait covariance, and is directly analogous to conceptual figures describing hierarchical trait variability across environmental gradients as presented in, for instance, @cornwell_community_2009 and @albert_intraspecific_2010.
@@
 the pairwise covariance normalized to the variances of the component variables).
-To do this, we plotted the mean and 95% confidence interval of the pairwise trait correlation coefficients for the global estimate from the pooled multivariate model and PFT-level estimates from the hierarchical model (Fig.\ \@ref(fig:corrbars)).
+To do this, we plotted the mean and 95% confidence interval of the pairwise trait correlation coefficients for the global estimate from the pooled multivariate model and PFT-level estimates from the hierarchical model (Fig.\ 5).
 
-The R code for running these analyses is publicly available online at <http://github.com/ashiklom/np_trait_analysis>.
+The R code and ancillary data for running these analyses is publicly available online via the Open Science Framework (OSF) at <https://osf.io/w8y73/>.
 The TRY data used for this analysis can be requested at <http://try-db.org>.

### 2018-02-13 scripts/manuscript-analyses/many_results.R
Minor revisions to `many_analyses.R`

@@ -41,7 +41,7 @@ ss_plot <- sample_size %>%
 if (interactive()) {
   ss_plot
 }
-ggsave(file.path(manuscript_fig_dir, "sample_size.pdf"), ss_plot, 
+ggsave(file.path(manuscript_fig_dir, "sample_size.pdf"), ss_plot,
         width = 7, height = 7)
 
 ## ----prepresults---------------------------------------------------------
@@ -183,11 +183,12 @@ clm_dat <- clm45_table81 %>%
   mutate(
     Nmass = 0.46 / CN_leaf,         # g N g-1 leaf
     Narea = Nmass / SLA,            # g N m-2 leaf
+    Vcmax_mass = Vcmax_25 * SLA,    # umol g-1 s-1
     SLA = SLA * 1000,               # m2 kg-1
     model_type = factor("CLM 4.5", model_type_levels),
     pft = factor(pft, abbr2pft) %>% lvls_revalue(pft2abbr)
   ) %>%
-  select(pft, model_type, SLA, Nmass, Narea, Vcmax_area = Vcmax_25) %>%
+  select(pft, model_type, SLA, Nmass, Narea, Vcmax_mass, Vcmax_area = Vcmax_25) %>%
   gather(key = "param", value = "mu_mean", -pft, -model_type) %>%
   mutate(
     param = factor(param, both_params),
@@ -307,7 +308,11 @@ relative_ci <- ggplot(cv_dat) +
     legend.position = c(0.85, 0.75),
     panel.grid = element_blank()
   )
-relative_ci
+if (interactive())
+  relative_ci
+
+ggsave(file.path(manuscript_fig_dir, "relative_ci_model.pdf"),
+       width = 4.4, height = 3)
 
 ## ----stickpairs, fig.width = 5, fig.height = 5, fig.cap = '(ref:stickcap)'----
 results_sub <- results_all %>%

### 2018-02-13 manuscript/03_results.Rmd
Revise results: PFT-level means -- Add CLM values

@@
 
-In general, leaf trait estimates from the univariate, multivariate, and hierarchical models were similar (Fig.\ \@ref(fig:meancompare)).
+In general, leaf trait estimates from the univariate, multivariate, and hierarchical models were similar (Fig.\ 2).
 Where we observed differences between models, the largest were between the univariate and multivariate models, while the additional constraint from the hierarchical model tended to have a minimal effect on trait estimates.
@@
 
+A key application of this study was to provide data-driven parameter estimates for Earth System Models.
+To this end, we compared our mean parameter estimates with corresponding default parameters in CLM 4.5 [@clm45_note] (Fig.\ 2).
+Our mean estimates of SLA agreed with CLM's defaults [@clm45_note, Table 8.1] only for tropical broadleaved evergreen trees, and for all other PFTs, our estimates are significantly lower.
+For $N_{mass}$, our estimates agreed reasonably well with CLM for evergreen temperate trees, needleleaved trees, and C3 arctic grasses, and were substantially different for all other PFTs.
+Our $N_{mass}$ estimates also varied much more across PFTs than CLM's parameters.
+For $N_{area}$, our estimates were significantly higher than CLM's for all PFTs, likely due to CLM's overestimates of SLA.
+Our mean estimates of $V_{c,max_area}$ (among those that were reasonably constrained by data) were lower than those of CLM for all PFTs except needleleaf evergreen trees and grasses, for which our estimates overlapped CLM's parameters.
+@kattge_2009_vcmax also found that $V_{c,max,area}$ was overestimated by Earth System models,
+but their estimates of $V_{c,max,area}$ and $N_{area}$ are generally slightly higher than ours.
+
 We observed clear differences in the relative uncertainties of mean estimates with respect to sample size.
@@
 For many of these trait-PFT combinations, the additional constraint from trait covariance provided by the multivariate and hierarchical models substantially reduced error bars, making it possible to compare estimates against those of other PFTs.
-Our analysis of the relationship between model type, sample size, and estimate relative uncertainty found that this covariance-based constraint from the multivariate model both reduced uncertainty overall (lower intercept) and reduced the sensitivity of estimate uncertainty to sample size (lower slope) compared to the univariate model (Fig.\ \@ref(fig:civssamplesize)).
+Our analysis of the relationship between model type, sample size, and estimate relative uncertainty found that this covariance-based constraint from the multivariate model both reduced uncertainty overall (lower intercept) and reduced the sensitivity of estimate uncertainty to sample size (lower slope) compared to the univariate model (Fig.\ 3).
 However, this analysis revealed no consistent significant benefit from the hierarchical model.
-
-(ref:sscap) Relative uncertainty in PFT-level trait estimates as a function of sample size for each model type.
-Lines represent linear models ($\log(y) = b_0 + b_1 \log(x)$) fit independently for each model type.
-In general, differences in estimate uncertainty between the univariate and multivariate models were minimal at large sample sizes but increasingly important at low sample sizes.
-However, differences in estimate uncertainty between the multivariate and hierarchical models were consistently negligible.
 
@@
 
-For all traits except leaf lifespan, pairwise trait correlations were generally consistent in direction both globally and within each PFT (Fig.\ \@ref(fig:stickpairs)).
+For all traits except leaf lifespan, pairwise trait correlations were generally consistent in direction both globally and within each PFT (Fig.\ 4).
 In particular, mass- and area-normalized traits were all positively correlated with each other and, respectively, positively and negatively correlated with SLA, both globally and within each PFT.
@@
 A large number of pairwise trait correlations were not significant.
-In some cases, this was driven by sample size (Fig.\ \@ref(fig:samplesize)).
+In some cases, this was driven by sample size (Fig.\ 1).
 For instance, needleleaved deciduous trees were often the only PFT for which a correlation was not statistically significant.
@@
 
-The strength of pairwise trait correlations varied substantially depending on scale, PFT, and trait (Fig.\ \@ref(fig:corrbars)).
+The strength of pairwise trait correlations varied substantially depending on scale, PFT, and trait (Fig.\ 5).
 The two pairwise trait correlations that exhibited the most consistent strength globally and within each PFT were the correlation between SLA and $N_{area}$, and between $N_{mass}$ and $P_{mass}$.

### 2018-02-13 manuscript/03_results.Rmd
Results: Revise CLM comparison
On closer examination...

@@
 For $N_{area}$, our estimates were significantly higher than CLM's for all PFTs, likely due to CLM's overestimates of SLA.
-Our mean estimates of $V_{c,max_area}$ (among those that were reasonably constrained by data) were lower than those of CLM for all PFTs except needleleaf evergreen trees and grasses, for which our estimates overlapped CLM's parameters.
-@kattge_2009_vcmax also found that $V_{c,max,area}$ was overestimated by Earth System models,
+Our estimates of $V_{c,max_mass}$ were lower across all PFTs, with particularly large differences for tropical and temeprate broadleaf deciduous trees and evergreen shrubs, and temperate C3 grasses.
+Our estimates of $V_{c,max,area}$ showed better agreement, though our values were still significantly lower for many PFTs.
+Like us, @kattge_2009_vcmax also found that $V_{c,max,area}$ was overestimated by Earth System models,
 but their estimates of $V_{c,max,area}$ and $N_{area}$ are generally slightly higher than ours.

### 2018-02-13 manuscript/03_results.Rmd
Results: Update remaining results with new runs

@@
 The same was generally true of correlations of mass-based traits with leaf lifespan, but correlations of leaf lifespan with area-normalized traits were more variable.
-The correlation between $N_{area}$ and leaf lifespan was positive globally and for evergreen shrubs but negative for broadleaved deciduous trees and not significant for any other PFTs.
-Similarly, the correlation between $P_{area}$ and leaf lifespan was positive for temperate broadleaved evergreen trees and needleleaved evergreen trees but negative for evergreen shrubs.
-The correlation between leaf lifespan and $R_{d,area}$ significant and positive globally, but was not significant within any PFTs.
-The correlations of leaf lifespan with $V_{c,max,area}$ and $J_{max,area}$ were negative for temperate broadleaved deciduous trees, but not significant for any other PFTs or globally.
+The correlation between $N_{area}$ and leaf lifespan was positive globally and for evergreen shrubs, tropical broadleaved deciduous trees, temperate needleleaved evergreen trees but negative for temperate and boreal broadleaved deciduous trees and not significant for any other PFTs.
+Similarly, the correlation between $P_{area}$ and leaf lifespan was positive globally but negative for evergreen shrubs and not significant for any other PFTs.
+The correlation between leaf lifespan and $R_{d,area}$ significant and negative globally, but was not significant within any PFTs.
+The only significant correlations of leaf lifespan with $V_{c,max,area}$ and $J_{max,area}$ were negative for temperate broadleaved deciduous trees.
 
@@
 In some cases, this was driven by sample size (Fig.\ 1).
-For instance, needleleaved deciduous trees were often the only PFT for which a correlation was not statistically significant.
+For instance, needleleaved deciduous trees, the most undersampled PFT in our analysis, were often the only PFT for which a correlation was not statistically significant.
 In other cases, though, PFTs with smaller sample sizes had significant pairwise correlations while PFTs with much larger sample sizes had none.
-For example, tropical broadleaved evergreen trees were relatively well-sampled for all traits, but none of their traits were well correlated with leaf lifespan.
-Meanwhile, temperate broadleaved evergreen trees had significant correlations between leaf lifespan and SLA, $P_{area}$, and $R_{d,area}$.
+For example, tropical broadleaved evergreen trees were relatively well-sampled for all traits, but none of their area-normalized traits were significantly correlated with leaf lifespan.
 In general, we observed fewer significant trait correlations among area-normalized traits than mass-normalized traits.

### 2018-02-13 scripts/manuscript-analyses/many_results.R
Stretch ylims on `many_results` means plot

@@ -209,9 +209,10 @@ plot_dat <- means_dat %>%
     mu_lo = if_else(irrelevant, NA_real_, mu_lo),
     mu_mean = if_else(irrelevant, NA_real_, mu_mean),
     ## TODO: Fix these in the prior and re-run
-    mu_hi = pclip(param, "Vcmax_mass", mu_hi, 1.5),
+    mu_hi = pclip(param, "Vcmax_mass", mu_hi, 2.5),
     mu_hi = pclip(param, "Vcmax_area", mu_hi, 100),
     mu_hi = pclip(param, "Jmax_mass", mu_hi, 1.8),
+    mu_mean = pclip(param, "Jmax_mass", mu_mean, 1.8),
     mu_hi = pclip(param, "Jmax_area", mu_hi, 150),
     mu_mean = pclip(param, "Jmax_area", mu_mean, 150),
     #mu_hi = pclip(param, "Rdmass", mu_hi, 0.03),

### 2018-02-13 manuscript/06_figures_tables.Rmd
Revise manuscript build process
- Use `doc` as default output
- Remove figures (draw them separately; only use figure captions in
manuscript)
- Render tables using `knitr`, but at end of manuscript

--- 2017-11-17 manuscript-md/03_results.Rmd
@@
-# Results
+# Tables {-}
 
-```{r prepresults}
-cachefile <- ".cache/mvtraits_results.rds"
-results_all <- readRDS(cachefile)
-
-max_ss <- max(sample_size$sample_size)
-
-multi_summary <- results_all %>%
-  filter(model_type %in% c("uni", "multi")) %>%
-  mutate(
-    mu_mean = map(data, c("mu", "Mean")) %>%
-      map(~as_tibble(as.list(.))),
-    mu_lo = map(data, c("mu", "2.5%")) %>%
-      map(~as_tibble(as.list(.))),
-    mu_hi = map(data, c("mu", "97.5%")) %>%
-      map(~as_tibble(as.list(.)))
-      ) %>%
-  select(model_type:pft, mu_mean:mu_hi) %>%
-  unnest(.sep = ".") %>%
-  gather("variable", "value", -(model_type:pft), na.rm = TRUE) %>%
-  separate(variable, c("stat", "param"), sep = "\\.") %>%
-  spread(stat, value) %>%
-  mutate(
-    pft = factor(pft, abbr2pft) %>% forcats::lvls_revalue(pft2abbr),
-    param = factor(param, both_params)
+```{r "knitr_options", echo = FALSE, message = FALSE, warning = FALSE, error = FALSE}
+library(knitr)
+library(here)
+opts_chunk$set(
+  echo = FALSE,
+  message = FALSE,
+  warning = FALSE,
+  error = FALSE
 )
-
-hier_prep <- results_all %>%
-  filter(model_type == "hier") %>%
-  select(model_type:pft_type, data) %>%
-  mutate(
-    mu_vals = map(data, "mu_group"),
-    pft = map(mu_vals, names),
-    mu_df = map2(pft, mu_vals, ~tibble(pft = ..1, dat = ..2))
-  ) %>%
-  select(-data, -mu_vals, -pft) %>%
-  unnest()
-
-hier_summary <- hier_prep %>%
-  mutate(
-    mu_mean = map(dat, c("Mean")) %>%
-      map(~as_tibble(as.list(.))),
-    mu_lo = map(dat, c("2.5%")) %>%
-      map(~as_tibble(as.list(.))),
-    mu_hi = map(dat, c("97.5%")) %>%
-      map(~as_tibble(as.list(.)))
-      ) %>%
-  select(model_type:pft, mu_mean:mu_hi) %>%
-  unnest(.sep = ".") %>%
-  gather("variable", "value", -(model_type:pft), na.rm = TRUE) %>%
-  separate(variable, c("stat", "param"), sep = "\\.") %>%
-  spread(stat, value) %>%
-  mutate(
-    pft = if_else(is.na(pft), "global", pft),
-    pft = factor(pft, abbr2pft) %>% forcats::lvls_revalue(pft2abbr),
-    param = factor(param, both_params)
-  )
-
-# Absolute comparison of means
-means_dat <- multi_summary %>%
-  filter(pft != "GLOB", !is.na(pft)) %>%
-  bind_rows(hier_summary) %>%
-  left_join(sample_size) %>%
-  mutate(
-    model_type = factor(model_type, c("uni", "multi", "hier")) %>%
-      forcats::lvls_revalue(c("univariate", "multivariate", "hierarchical")),
-    mass_area = factor(mass_area, c("mass", "area")),
-  ) %>%
-  mutate_at(
-    c("mu_mean", "mu_lo", "mu_hi"),
-    ~10 ^ .
-  ) %>%
-  filter(
-    !(param %in% c("leaf_lifespan", "SLA") & mass_area == "area")
-  )
+options(knitr.table.format = "pandoc")
 ```
 
-## Estimates of PFT-level means
-
-```{r meanvaluetable, results = "asis"}
-## Write summary table
-table_cap_mass <- paste(
-  "Mean and 95\\% confidence interval of trait estimates",
-  "for mass-normalized traits from the hierarchical model."
-)
-table_cap_area <- gsub("mass", "area", table_cap_mass)
-
-censor <- function(rxp, pft, pfts, param, value) {
-  if_else(grepl(rxp, param) & pfts %in% pft, "---", value)
-}
-
-table_dat <- means_dat %>%
-  filter(
-    model_type == "hierarchical",
-    !(param %in% c("SLA", "leaf_lifespan") & mass_area == "area")
-  ) %>%
-  mutate(
-    mu_lo = if_else(param == "Rdmass", mu_lo * 1000, mu_lo),
-    mu_mean = if_else(param == "Rdmass", mu_mean * 1000, mu_mean),
-    mu_hi = if_else(param == "Rdmass", mu_hi * 1000, mu_hi),
-    mu_lo = if_else(param == "Rdarea", mu_lo * 1000, mu_lo),
-    mu_mean = if_else(param == "Rdarea", mu_mean * 1000, mu_mean),
-    mu_hi = if_else(param == "Rdarea", mu_hi * 1000, mu_hi),
-    lo_f = formatC(mu_lo, digits = 3),
-    mid_f = formatC(mu_mean, digits = 3),
-    hi_f = formatC(mu_hi, digits = 3),
-    value = sprintf("%s (%s,%s)", mid_f, lo_f, hi_f),
-    value = censor("Vcmax|Jmax", "C4G", pft, param, value),
-    value = censor("Rdmass", "ShDBo", pft, param, value),
-    value = censor("Jmax_mass", c("NlEBo", "NlD"), pft, param, value),
-    value = censor("Jmax_area", c("NlD"), pft, param, value)
-  ) %>%
-  select(pft, param, value) %>%
-  mutate(
-    param = factor(param, both_params) %>% 
-      forcats::lvls_revalue(param_markdown_nounit[both_params]),
-  )%>%
-  spread(param, value) %>%
-  arrange(pft)
-
-table_dat %>%
-  select(-matches("area")) %>%
-  kable(caption = table_cap_mass, format = "latex",
-        escape = FALSE, booktabs = TRUE) %>%
-  kableExtra::kable_styling(latex_options = c("scale_down"))
-table_dat %>%
-  select(-matches("mass")) %>%
-  kable(caption = table_cap_area, format = "latex",
-        escape = FALSE, booktabs = TRUE) %>%
-  kableExtra::kable_styling(latex_options = c("scale_down"))
+<!-- Tables ---------------------------->
+```{r "read_chunks"}
+knitr::read_chunk(here::here("scripts/manuscript-analyses/try_ref_table.R"))
+knitr::read_chunk(here::here("scripts/manuscript-analyses/table_n_species.R"))
 ```
 
-(ref:mccap) Mean and 95% confidence interval on best estimates of traits for each plant functional type from the univariate, multivariate, and hierarchical models.
+<!-- TRY data citations -->
+```{r "trydata"}
+```
+
+<!-- Number of species in each PFT -->
+```{r "pfts"}
+```
+
+# Figure captions {-}
+
+**Figure 1**: Sample sizes for each trait-PFT pair. $y$ axis is scaled logarithmically.
+
+**Figure 2**: Mean and 95% confidence interval on best estimates of traits for each plant functional type from the univariate, multivariate, and hierarchical models.
 For leaf lifespan and SLA, results were not significantly different between the mass- and area-based models, so only results from the mass-based model are shown.
@@
 
-```{r meancompare, fig.width=7, fig.height=7, fig.cap='(ref:mccap)'}
-pclip <- function(p, pname, value, lim, hi = TRUE) {
-  f <- ifelse(hi, `>`, `<`)
-  if_else(p == pname & f(value, lim), lim, value)
-}
-
-plot_dat <- means_dat %>%
-  mutate(
-    #too_few = sample_size < 2,
-    #mu_hi = if_else(too_few, NA_real_, mu_hi),
-    #mu_lo = if_else(too_few, NA_real_, mu_lo),
-    #mu_mean = if_else(too_few, NA_real_, mu_mean),
-    irrelevant = grepl("Vcmax|Jmax", param) & pft == "C4G",
-    mu_hi = if_else(irrelevant, NA_real_, mu_hi),
-    mu_lo = if_else(irrelevant, NA_real_, mu_lo),
-    mu_mean = if_else(irrelevant, NA_real_, mu_mean),
-    ## TODO: Fix these in the prior and re-run
-    mu_hi = pclip(param, "Vcmax_mass", mu_hi, 1.5),
-    mu_hi = pclip(param, "Vcmax_area", mu_hi, 100),
-    mu_hi = pclip(param, "Jmax_mass", mu_hi, 1.8),
-    mu_hi = pclip(param, "Jmax_area", mu_hi, 150),
-    mu_mean = pclip(param, "Jmax_area", mu_mean, 150),
-    mu_hi = pclip(param, "Rdmass", mu_hi, 0.03),
-    mu_mean = pclip(param, "Rdmass", mu_mean, 0.03),
-    mu_lo = pclip(param, "Rdmass", mu_lo, 0.03),
-    mu_hi = pclip(param, "Rdarea", mu_hi, 0.0025),
-    mu_mean = pclip(param, "Rdarea", mu_mean, 0.0025),
-    mu_lo = pclip(param, "Rdarea", mu_lo, 0.0025),
-    mu_hi = pclip(param, "Parea", mu_hi, 0.285),
-    #mu_lo = pclip(param, "Jmax_area", mu_lo, 40, FALSE),
-    param = forcats::lvls_revalue(param, param_fancy_chr[both_params])
-  )
-
-p_means <- ggplot(plot_dat) +
-  aes(x = interaction(model_type, pft),
-      y = mu_mean, ymin = mu_lo, ymax = mu_hi,
-      color = pft, shape = model_type) +
-  geom_pointrange(size = 0.3) +
-  facet_wrap(~param, scales = "free", ncol = 2,
-              labeller = label_parsed) +
-  scale_y_continuous() +
-  scale_color_manual(values = pft_colors) +
-  guides(
-    shape = guide_legend(title = "Model type"),
-    color = guide_legend(title = "PFT")
-  ) +
-  ylab("Trait estimate mean and 95% CI") +
-  xlab("Model type and PFT") +
-  theme_bw() +
-  theme(
-    text = element_text(size = 12),
-    axis.text.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    panel.grid = element_blank()
-  )
-p_means
-```
-
-In general, leaf trait estimates from the univariate, multivariate, and hierarchical models were similar (Fig.\ \@ref(fig:meancompare)).
-Where we observed differences between models, the largest were between the univariate and multivariate models, while the additional constraint from the hierarchical model tended to have a minimal effect on trait estimates.
-Significant differences in trait estimates between univariate and multivariate models occurred even for traits with relatively large sample sizes, such as leaf nitrogen content.
-
-Evergreen PFTs had by far the largest leaf lifespan, with the longest lifespan observed for temperate and boreal needleleaf evergreen species.
-Meanwhile, all of deciduous species had lifespans shorter than 7 months.
-Among deciduous species, lifespan was generally longer in warmer biomes than colder ones.
-
-Across-PFT patterns in SLA and $N_{mass}$, $P_{mass}$, and $R_{d,mass}$ were similar.
-Temperate broadleaved deciduous trees and shrubs generally had among the highest values of these traits, while temperate evergreen trees and shrubs had generally among the lowest.
-However, none of these patterns were universal to all four traits.
-For example, tropical evergreen trees had relatively high $N_{mass}$ and average SLA and $R_{d,mass}$, but among the lowest $P_{mass}$.
-Similarly, temperate and boreal shrubs had higher $N_{mass}$ and $P_{mass}$ than any of the grasses, but comparable SLA.
-
-Across-PFT patterns in $N_{area}$, $P_{area}$, and $R_{d,area}$ were different from their mass-normalized counterparts.
-For example, tropical broadleaved evergreen and needleleaf evergreen trees had among the lowest $N_{mass}$ and $P_{mass}$ basis but among the highest $N_{area}$ and $P_{area}$, while the opposite was true of deciduous temperate trees and shrubs.
-Species with N contents near the middle of the observed range did not shift as dramatically depending on type of normalization.
-
-C3 grasses had both the highest $V_{c,max,mass}$ and $V_{c,max,area}$.
-Compared to broadleaved trees, temperate needleleaved evergreen trees had lower $V_{c,max,mass}$ but higher $V_{c,max,area}$.
-Among broadleaved trees, deciduous trees had higher $V_{c,max,mass}$ and slightly higher $V_{c,max,area}$ than evergreen trees.
-Between the deciduous and evergreen tree PFTs, we observed no significant trend by climate zone.
-
-C3 grasses and temperate needleleaved evergreen trees had the highest $J_{max,area}$, but temperate broadleaved deciduous trees had the highest $J_{max,mass}$.
-All of the shrub PFTs had the lowest $J_{max,mass}$ but average or above-average $J_{max,area}$, while the opposite was true of broadleaved tropical PFTs.
-Of the tree PFTs, needleleaved evergreen trees had the highest $J_{max,area}$ but the lowest $J_{max,mass}$.
-
-We observed clear differences in the relative uncertainties of mean estimates with respect to sample size.
-All of the high-latitude PFTs consistently had among the largest error bars around their mean estimates relative to other PFTs, while the traits with the largest uncertainties were dark respiration, $V_{c,max}$, and $J_{max}$.
-For many of these trait-PFT combinations, the additional constraint from trait covariance provided by the multivariate and hierarchical models substantially reduced error bars, making it possible to compare estimates against those of other PFTs.
-Our analysis of the relationship between model type, sample size, and estimate relative uncertainty found that this covariance-based constraint from the multivariate model becomes increasingly important at smaller sample sizes (Fig.\ \@ref(fig:civssamplesize)).
-However, this analysis revealed no consistent significant benefit from the hierarchical model.
-
-(ref:sscap) Relative uncertainty in PFT-level trait estimates as a function of sample size for each model type.
+**Figure 3**: Relative uncertainty in PFT-level trait estimates as a function of sample size for each model type.
 Lines represent linear models ($\log(y) = b_0 + b_1 \log(x)$) fit independently for each model type.
@@
 
-```{r "civssamplesize", fig.width=4.4, fig.height=3, fig.cap='(ref:sscap)'}
-## Relative uncertainty
-cv_dat <- means_dat %>%
-  mutate(
-    cv = abs((mu_hi - mu_lo) / mu_mean)
-  ) %>%
-  filter(sample_size > 1)
-
-cv_mod <- cv_dat %>%
-  group_by(model_type) %>%
-  nest() %>%
-  mutate(
-    cvfit = map(data, ~lm(log10(cv) ~ log10(sample_size), data = .))
-  )
-
-cv_coef <- cv_mod %>%
-  mutate(
-    coef = map(cvfit, broom::tidy)
-  ) %>%
-  unnest(coef) %>%
-  select(model_type, term, estimate) %>%
-  spread(term, estimate) %>%
-  "names<-"(c("model_type", "intercept", "slope")) %>%
-  mutate(
-    eqn = sprintf("'%s: '~log[10](y) == %0.2f - %0.2f~log[10](x)", model_type, intercept, abs(slope))
-  )
-
-cv_pred <- cv_mod %>%
-  mutate(
-    grid = map(data, modelr::data_grid, sample_size = seq(5, max_ss)),
-    cvpred = map2(grid, cvfit, modelr::add_predictions)
-  ) %>%
-  unnest(cvpred) %>%
-  mutate(cv = 10 ^ pred)
-
-x_lims <- range(cv_pred$sample_size)
-y_lims <- range(cv_pred$cv)
-
-relative_ci <- ggplot(cv_dat) +
-  aes(x = sample_size, y = cv, color = model_type) +
-  geom_point(size = 0.5) +
-  geom_line(data = cv_pred) +
-  annotate("text", x = rep(2500, 3), y = c(1.4, 1.25, 1.1), 
-           label = cv_coef$eqn, parse = TRUE, 
-           size = 2, vjust = "right") +
-  scale_x_continuous(trans = "log10") +
-  coord_cartesian(xlim = x_lims, ylim = y_lims) +
-  xlab("Sample size") +
-  ylab("Relative width of 95% CI") +
-  scale_color_brewer(type = "qual", palette = 2) +
-  guides(color = guide_legend(title = "Model type")) +
-  theme_bw() +
-  theme(
-    legend.position = c(0.85, 0.75),
-    panel.grid = element_blank()
-  )
-relative_ci
-```
-
-## Trait correlation patterns among- and within-PFTs
-
-(ref:stickcap) Pairwise trait mean and covariance estimates for all data pooled globally (black) and for each PFT (colored).
+**Figure 4**: Pairwise trait mean and covariance estimates for all data pooled globally (black) and for each PFT (colored).
 Covariance estimates not significantly different from zero ($p < 0.05$) are indicated by x symbols at the mean estimate.
@@
 
-```{r stickpairs, fig.width = 5, fig.height = 5, fig.cap = '(ref:stickcap)'}
-suppressPackageStartupMessages({
-  library(mvtraits)
-})
-results_sub <- results_all %>%
-  filter((model_type == "multi" & pft == "global") |
-         (model_type == "hier"))
-dat_list <- results_sub$data
-names(dat_list) <- with(results_sub, paste(model_type, mass_area, sep = "_"))
-
-stickplot_pairs(
-  #mu_global_lower = dat_list$hier_mass$mu_global,
-  #Sigma_global_lower = dat_list$hier_mass$Sigma_global,
-  mu_global_lower = dat_list$multi_mass$mu,
-  Sigma_global_lower = dat_list$multi_mass$Sigma,
-  mu_group_lower = dat_list$hier_mass$mu_group,
-  Sigma_group_lower = dat_list$hier_mass$Sigma_group,
-  vars_lower = mass_params,
-  group_names = pft2abbr[-1],
-  vars_label = param_labels,
-  unlog_axes = TRUE,
-  #mu_global_upper = dat_list$hier_area$mu_global,
-  #Sigma_global_upper = dat_list$hier_area$Sigma_global,
-  mu_global_upper = dat_list$multi_area$mu,
-  Sigma_global_upper = dat_list$multi_area$Sigma,
-  mu_group_upper = dat_list$hier_area$mu_group,
-  Sigma_group_upper = dat_list$hier_area$Sigma_group,
-  vars_upper = area_params,
-  col_group = pft_colors,
-  cex_g = 1.2,
-  lwd_g = 1,
-  lty_g = 1,
-  lwd_s = 1.3,
-  par_plot = list(oma = c(0, 5, 3, 5), cex = 0.4),
-  par_legplot = list(mar = c(0, 0, 0, 0)),
-  par_legend = list(ncol = 5, x = "center", cex = 0.6),
-  reorder_legend = TRUE
-)
-mtext("Mass-normalized", side = 2, cex = 0.7, line = 3)
-mtext("Area-normalized", side = 4, cex = 0.7, line = 1)
-```
-
-Where pairwise trait correlations were statistically significant, these correlations were generally consistent in direction both globally and within each PFT (Fig.\ \@ref(fig:stickpairs)).
-In particular, mass- and area-normalized traits were all positively correlated with each other and, respectively, positively and negatively correlated with SLA, both globally and within each PFT.
-The same was generally true of correlations of mass-based traits with leaf lifespan, but correlations of leaf lifespan with area-normalized traits were more variable.
-The correlation between $N_{area}$ and leaf lifespan was positive globally and for evergreen shrubs but negative for broadleaved deciduous trees and not significant for any other PFTs.
-Similarly, the correlation between $P_{area}$ and leaf lifespan was positive for temperate broadleaved evergreen trees and needleleaved evergreen trees but negative for evergreen shrubs.
-The correlation between leaf lifespan and $R_{d,area}$ significant and positive globally, but was not significant within any PFTs.
-The correlations of leaf lifespan with $V_{c,max,area}$ and $J_{max,area}$ were negative for temperate broadleaved deciduous trees, but not significant for any other PFTs or globally.
-
-A large number of pairwise trait correlations were not significant.
-In some cases, this was driven by sample size (Fig.\ \@ref(fig:samplesize)).
-For instance, needleleaved deciduous trees were often the only PFT for which a correlation was not statistically significant.
-In other cases, though, PFTs with smaller sample sizes had significant pairwise correlations while PFTs with much larger sample sizes had none.
-For example, tropical broadleaved evergreen trees were relatively well-sampled for all traits, but none of their traits were well correlated with leaf lifespan.
-Meanwhile, temperate broadleaved evergreen trees had significant correlations between leaf lifespan and SLA, $P_{area}$, and $R_{d,area}$.
-In general, we observed fewer significant trait correlations among area-normalized traits than mass-normalized traits.
-
-```{r boxprep}
-outdir <- "../output"
-hier_mass_file <- tail(dir(outdir, "hier.mass.clm45", full.names = TRUE), 1)
-hier_area_file <- tail(dir(outdir, "hier.area.clm45", full.names = TRUE), 1)
-multi_mass_file <- tail(dir(outdir, "multi.mass.clm45", full.names = TRUE), 1)
-multi_area_file <- tail(dir(outdir, "multi.area.clm45", full.names = TRUE), 1)
-
-hier_summary_mass <- readRDS(hier_mass_file)[["summary_table"]] %>%
-  mutate(mass_area = "mass")
-hier_summary_area <- readRDS(hier_area_file)[["summary_table"]] %>%
-  mutate(mass_area = "area")
-multi_summary_mass <- readRDS(multi_mass_file)[["summary_table"]] %>%
-  mutate(mass_area = "mass")
-multi_summary_area <- readRDS(multi_area_file)[["summary_table"]] %>%
-  mutate(mass_area = "area")
-
-proc <- function(dat, hier = TRUE) {
-  if (hier) {
-    dat <- filter(dat, group != "global")
-  } else {
-    dat <- mutate(dat, group = "global")
-  }
-  dat <- dat %>%
-    filter(variable == "Corr") %>%
-    separate(index, c("xvar", "yvar"), sep = "\\.\\.") %>%
-    mutate(
-      pft = factor(group, abbr2pft) %>% forcats::lvls_revalue(pft2abbr)
-    ) %>%
-    select(mass_area, pft, xvar, yvar, Mean, `2.5%`, `97.5%`)
-}
-
-hsm <- proc(hier_summary_mass)
-hsa <- proc(hier_summary_area)
-msm <- proc(multi_summary_mass, FALSE)
-msa <- proc(multi_summary_area, FALSE)
-hs <- bind_rows(hsm, hsa, msm, msa)
-```
-
-(ref:corrbarscap) Mean and 95% CI on estimates of pairwise correlation coefficients
+**Figure 5**: Mean and 95% CI on estimates of pairwise correlation coefficients
 for all data pooled globally (dark grey) and for each PFT (colored).
@@
 
-```{r corrbars, fig.width=7, fig.height=7, fig.cap='(ref:corrbarscap)'}
-.zzz <- close.screen(all.screens = TRUE)
-corr_bar <- function(dat, xvar, yvar, mass_area) {
-  dp <- dat %>%
-    filter(
-      mass_area == !!mass_area,
-      (xvar == !!xvar & yvar == !!yvar) | 
-        (xvar == !!yvar & yvar == !!xvar),
-    ) %>%
-    arrange(pft) %>%
-    mutate(color = c("grey25", pft_colors))
-  x <- barplot(dp$Mean, col = dp$color, border = dp$color,
-               ylim = c(-1, 1), axes = FALSE)
-  arrows(x, dp[["2.5%"]], x, dp[["97.5%"]],
-         length = 0.01, angle = 90, code = 3)
-  box()
-}
-
-nv <- length(mass_params)
-main_screens <- split.screen(rbind(c(0, 1, 0.2, 1), c(0, 1, 0, 0.2)))
-par(mar = rep(0.1, 4), oma = c(0, 8, 1, 5), cex = 0.4)
-screens <- split.screen(c(nv, nv), screen = main_screens[1])
-for (i in seq_len(nv)) {
-  for (j in seq_len(nv)) {
-    k <- j + nv * (i - 1)
-    screen(screens[k])
-    if (i == j) {
-      # Label
-      plot(0:1, 0:1, type = "n", ann = FALSE, axes = FALSE)
-      box()
-      text(x = 0.5, y = 0.5, labels = param_labels[i], cex = 2)
-    } else {
-      if (j > i) {
-        # Upper triangle
-        mass_area <- "area"
-        xvar <- area_params[j]
-        yvar <- area_params[i]
-      } else {
-        # Lower triangle
-        mass_area <- "mass"
-        xvar <- mass_params[j]
-        yvar <- mass_params[i]
-      }
-      corr_bar(hs, xvar, yvar, mass_area)
-    }
-    if (i == 2 && j == 1) {
-      axis(side = 2)
-    }
-    if (i == 1 && j == nv) {
-      axis(side = 4)
-    }
-  }
-}
-## Draw legend
-screen(main_screens[2])
-par(mar = c(0, 0, 0, 0), oma = c(0, 0, 0, 0))
-plot(0, 0, type = "n", axes = FALSE, ann = FALSE)
-legend("center", legend = levels(hs$pft), fill = c("grey25", pft_colors), 
-       ncol = 5, bty = "n")
-close.screen(all.screens = TRUE)
-mtext("Mass-normalized", side = 2, cex = 0.7, line = 1)
-mtext("Mean and 95% CI of pairwise correlation", side = 2, cex = 1, line = 2)
-mtext("Area-normalized", side = 4, cex = 0.7, line = 0)
-```
-
-The strength of pairwise trait correlations varied substantially depending on scale, PFT, and trait (Fig.\ \@ref(fig:corrbars)).
-The two pairwise trait correlations that exhibited the most consistent strength globally and within each PFT were the correlation between SLA and $N_{area}$, and between $N_{mass}$ and $P_{mass}$.
-Correlation strength was often, but not always, related to sample size, with well-sampled PFTs exhibiting stronger correlations and undersampled PFTs exhibiting weaker correlations.
+# References {-}

### 2018-02-13 manuscript/index.Rmd
Revise manuscript build process
- Use `doc` as default output
- Remove figures (draw them separately; only use figure captions in
manuscript)
- Render tables using `knitr`, but at end of manuscript

--- 2018-01-03 manuscript-md/index.Rmd
@@
 ---
-title: "Covariance as constraint: Leveraging leaf economic relationships to constrain plant functional type trait estimates"
-author: >
-  Alexey N. Shiklomanov$^1$,
+title: "Does the leaf economic spectrum hold within plant functional types? A Bayesian hierarchical meta-analysis"
+author:
+  - >
+    Alexey N. Shiklomanov$^1$
     Elizabeth M. Cowdery$^1$,
@@
     Koen Kramer$^{9}$,
-  Vanessa Minden$^{10,11}$
+    Vanessa Minden$^{10,11}$,
     Ülo Niinemets$^{12}$,
@@
 output:
+  bookdown::word_document2:
+    toc: no
+    reference_docx: ~/Projects/misc/pandoc-docx-template/academic-numbered.docx
   bookdown::pdf_document2:
@@
 ---
+<div custom-style="Affiliation">
 $^1$Department of Earth & Environment, Boston University, Boston, MA 02215, USA;
@@
 $^{15}$Conservation Biology Department, Institute of Environmental Sciences, Leiden University, Leiden, The Netherlands
+</div>
 
+<div custom-style="Single space">
 Corresponding author:
+
 Alexey Shiklomanov,
+
 telephone: +1 (302) 897-3348,
+
 email: ashiklom\@bu.edu
@@
 Key words: Functional trade-off; hierarchical modeling; trait variation; ecological modeling; leaf morphology; leaf biochemistry
+</div>
 
@@
 )
-files <- file.path(
-  "~/Projects/new-phytologist-traits/np-trait-analysis/manuscript-md", 
+files <- here::here(
+  "manuscript", 
   fnames

### 2018-02-16 manuscript/02_methods.Rmd
Consistent use of across-PFT (not among-PFT)
Per recommendation of Kramer.

@@
 
-The most complex model was the ‘hierarchical’ model, in which observed trait values were drawn from a PFT-specific multivariate normal distribution describing within-PFT variation and whose parameters were themselves sampled from a global multivariate distribution describing the variation among PFTs.
+The most complex model was the ‘hierarchical’ model, in which observed trait values were drawn from a PFT-specific multivariate normal distribution describing within-PFT variation and whose parameters were themselves sampled from a global multivariate distribution describing the variation across PFTs.
 For the observed vector of traits $\mathbf{x}_{i,p}$ for sample $i$ belonging to PFT $p$:
@@
 
-where $\mathbf{\mu}_p$ and $\mathbf{\Sigma}_p$ are the mean vector and variance-covariance matrix describing variation within PFT $p$, and $\mathbf{\mu}_g$ and $\mathbf{\Sigma}_g$ are the mean vector and variance-covariance matrix describing among-PFT (global) variation.
+where $\mathbf{\mu}_p$ and $\mathbf{\Sigma}_p$ are the mean vector and variance-covariance matrix describing variation within PFT $p$, and $\mathbf{\mu}_g$ and $\mathbf{\Sigma}_g$ are the mean vector and variance-covariance matrix describing across-PFT (global) variation.
 
@@
 
-To assess the consistency of among- and within-PFT trait trade-offs, we looked at covariance estimates for each trait pair and, where these values were significantly different from zero ($p < 0.05$),
+To assess the consistency of within- and across-PFT trait trade-offs, we looked at covariance estimates for each trait pair and, where these values were significantly different from zero ($p < 0.05$),
 we calculated the eigenvalues from the pairwise variance-covariance matrix for that trait pair and plotted the corresponding dominant eigenvectors centered on the mean estimates (Fig.\ 4).

### 2018-02-16 scripts/manuscript-analyses/figure_correlation_boxplot.R
Bugfix in correlation boxplot figure

@@ -4,8 +4,8 @@ library(shiklomanov2017np)
 outdir <- here::here("output")
 hier_mass_file <- tail(dir(outdir, "hier.mass.clm45", full.names = TRUE), 1)
 hier_area_file <- tail(dir(outdir, "hier.area.clm45", full.names = TRUE), 1)
-multi_mass_file <- tail(dir(outdir, "multi.mass.clm45", full.names = TRUE), 1)
-multi_area_file <- tail(dir(outdir, "multi.area.clm45", full.names = TRUE), 1)
+multi_mass_file <- tail(dir(outdir, "multi.mass.clm45.NA", full.names = TRUE), 1)
+multi_area_file <- tail(dir(outdir, "multi.area.clm45.NA", full.names = TRUE), 1)
 
 hier_summary_mass <- readRDS(hier_mass_file)[["summary_table"]] %>%
   mutate(mass_area = "mass")
@@ -55,13 +55,11 @@ corr_bar <- function(dat, xvar, yvar, mass_area) {
   box()
 }
 
-if (!interactive()) {
-  pdf(
-    here("figures", "manuscript", "correlation_boxplot.pdf"),
-    width = 7,
-    height = 7
-  )
-}
+pdf(
+  here("figures", "manuscript", "correlation_boxplot.pdf"),
+  width = 7,
+  height = 7
+)
 
 nv <- length(mass_params)
 main_screens <- split.screen(rbind(c(0, 1, 0.2, 1), c(0, 1, 0, 0.2)))
@@ -108,6 +106,5 @@ close.screen(all.screens = TRUE)
 mtext("Mass-normalized", side = 2, cex = 0.7, line = 1)
 mtext("Mean and 95% CI of pairwise correlation", side = 2, cex = 1, line = 2)
 mtext("Area-normalized", side = 4, cex = 0.7, line = 0)
-if (!interactive()) {
-  dev.off()
-}
+
+dev.off()

### 2018-02-16 scripts/manuscript-analyses/many_results.R
Revise the discussion to reflect new results
Also, clean up some of the language a little bit.

@@ -325,8 +325,6 @@ names(dat_list) <- with(results_sub, paste(model_type, mass_area, sep = "_"))
 if (!interactive()) pdf(file.path(manuscript_fig_dir, "stick_pairs.pdf"),
                         width = 5, height = 5)
 stickplot_pairs(
-  #mu_global_lower = dat_list$hier_mass$mu_global,
-  #Sigma_global_lower = dat_list$hier_mass$Sigma_global,
   mu_global_lower = dat_list$multi_mass$mu,
   Sigma_global_lower = dat_list$multi_mass$Sigma,
   mu_group_lower = dat_list$hier_mass$mu_group,
@@ -335,8 +333,6 @@ stickplot_pairs(
   group_names = pft2abbr[-1],
   vars_label = param_labels,
   unlog_axes = TRUE,
-  #mu_global_upper = dat_list$hier_area$mu_global,
-  #Sigma_global_upper = dat_list$hier_area$Sigma_global,
   mu_global_upper = dat_list$multi_area$mu,
   Sigma_global_upper = dat_list$multi_area$Sigma,
   mu_group_upper = dat_list$hier_area$mu_group,

### 2018-02-16 manuscript/03_results.Rmd
Consistent use of across-PFT (not among-PFT)
Per recommendation of Kramer.
Address most remaining comments

@@
 However, none of these patterns were universal to all four traits.
-For example, tropical evergreen trees had relatively high $N_{mass}$ and average SLA and $R_{d,mass}$, but among the lowest $P_{mass}$.
+For example, tropical evergreen trees had relatively high $N_{mass}$ and mean SLA and $R_{d,mass}$, but among the lowest $P_{mass}$.
 Similarly, temperate and boreal shrubs had higher $N_{mass}$ and $P_{mass}$ than any of the grasses, but comparable SLA.
@@
 
-## Trait correlation patterns among- and within-PFTs
+## Trait correlation patterns across- and within-PFTs
 

### 2018-02-16 manuscript/02_methods.Rmd
Revise formatting for submission

@@
 Where species attributes disagreed between datasets (e.g., categorized as a shrub in one dataset but a tree in another), we assigned the attribute that was observed most frequently between the datasets (e.g., if five datasets say "shrub" but only one says "tree", we would classify it as a shrub).
-Where species lacked data on certain attributes, we assigned those attributes based on higher order phylogeny where appropriate (e.g., \em{Poaceae} family are usually grasses, \em{Larix spp.} are deciduous needleleaved trees) or otherwise omitted the species from our analyses.
+Where species lacked data on certain attributes, we assigned those attributes based on higher order phylogeny where appropriate (e.g., *Poaceae* family are usually grasses, *Larix spp.* are deciduous needleleaved trees) or otherwise omitted the species from our analyses.
 For biome specification, we collected all latitude and longitude data for each species,

### 2018-02-16 manuscript/index.Rmd
Revise formatting for submission

@@
     toc: no
-    reference_docx: ~/Projects/misc/pandoc-docx-template/academic-numbered.docx
+    reference_docx: ~/Projects/misc/pandoc-docx-template/custom.docx
   bookdown::pdf_document2:
@@
   results = "03_results.Rmd",
-  discussion = "04_discussion.md",
-  conclusions = "05_conclusions.md"
+  discussion = "04_discussion.md"
 )
@@
 
-# Summary {-}
+<div custom-style="Heading, no number">Summary</div>
 

### 2018-02-16 manuscript/index.Rmd
Revise author list

@@
     Andrés Gonzalez-Melo$^5$,
-    Alvaro G. Gutiérrez$^6$,
-    Steven Jansen$^7$,
-    Nathan Kraft$^8$,
-    Koen Kramer$^{9}$,
-    Vanessa Minden$^{10,11}$,
-    Ülo Niinemets$^{12}$,
-    Yusuke Onoda$^{13}$,
-    Enio Egon Sosinski$^{14}$,
-    Nadejda A. Soudzilovskaia$^{15}$,
+    Steven Jansen$^6$,
+    Nathan Kraft$^7$,
+    Koen Kramer$^{8}$,
+    Vanessa Minden$^{9,10}$,
+    Ülo Niinemets$^{11}$,
+    Yusuke Onoda$^{12}$,
+    Enio Egon Sosinski$^{13}$,
+    Nadejda A. Soudzilovskaia$^{14}$,
     Michael C. Dietze$^1$
@@
 $^5$Universidad del Rosario. Facultad de Ciencias Naturales y Matemáticas. Carrera 26 #63B-48. Bogotá, Colombia;
-$^6$Departamento de Ciencias Ambientales y Recursos Naturales, Facultad de Ciencias Agronomicas, Universidad de Chile, Av. Santa Rosa 11315, La Pintana 8820808, Santiago, Chile;
-$^7$Institute of Systematic Botany and Ecology, Ulm University, Ulm, Germany;
-$^8$Department of Ecology and Evolutionary Biology, University of California Los Angeles, Los Angeles, CA 90095, USA;
-$^{9}$Vegetation, Forest, and Landscape Ecology, Wageningen Environmental Research and Wageningen University, Wageningen, Gelderland, The Netherlands;
-$^{10}$Institute for Biology and Environmental Sciences, Carl von Ossietzky-University of Oldenburg, Carl von Ossietzky Str. 9-11, 26129 Oldenburg, Germany;
-$^{11}$Department of Biology, Ecology and Evolution, Vrije Universiteit Brussel, Pleinlaan 2, 1050 Brussels;
-$^{12}$Institute of Agricultural and Environmental Sciences, Estonian University of Life Sciences, Kreutzwaldi 1, 51014 Tartu, Estonia;
-$^{13}$Graduate School of Agriculture, Kyoto University, Kyoto, 606-8502, Japan;
-$^{14}$Embrapa Temperate Agriculture, Rodovia BR 392, Km 78, CEP 96010-971 Pelotas, RS, Brazil;
-$^{15}$Conservation Biology Department, Institute of Environmental Sciences, Leiden University, Leiden, The Netherlands
+$^6$Institute of Systematic Botany and Ecology, Ulm University, Ulm, Germany;
+$^7$Department of Ecology and Evolutionary Biology, University of California Los Angeles, Los Angeles, CA 90095, USA;
+$^{8}$Vegetation, Forest, and Landscape Ecology, Wageningen Environmental Research and Wageningen University, Wageningen, Gelderland, The Netherlands;
+$^{9}$Institute for Biology and Environmental Sciences, Carl von Ossietzky-University of Oldenburg, Carl von Ossietzky Str. 9-11, 26129 Oldenburg, Germany;
+$^{10}$Department of Biology, Ecology and Evolution, Vrije Universiteit Brussel, Pleinlaan 2, 1050 Brussels;
+$^{11}$Institute of Agricultural and Environmental Sciences, Estonian University of Life Sciences, Kreutzwaldi 1, 51014 Tartu, Estonia;
+$^{12}$Graduate School of Agriculture, Kyoto University, Kyoto, 606-8502, Japan;
+$^{13}$Embrapa Temperate Agriculture, Rodovia BR 392, Km 78, CEP 96010-971 Pelotas, RS, Brazil;
+$^{14}$Conservation Biology Department, Institute of Environmental Sciences, Leiden University, Leiden, The Netherlands
 </div>

### 2018-02-16 manuscript/index.Rmd
Add acknowledgements and author contributions
Also, count words manually. R package hugely overestimates words for
some reason.

@@
 
-```{r wordcount, echo = FALSE}
-fnames <- c(
-  introduction = "01_introduction.md",
-  methods = "02_methods.Rmd",
-  results = "03_results.Rmd",
-  discussion = "04_discussion.md"
-)
-files <- here::here(
-  "manuscript", 
-  fnames
-)
-names(files) <- names(fnames)
-
-getwc <- function(f) {
-  ct <- withr::with_options(
-    list(knitr.table.format = "pandoc"),
-    wordcountaddin::text_stats(f)
-  )
-  wline <- ct[3]
-  wlist <- strsplit(wline, split = "[[:space:]]+")
-  as.numeric(wlist[[1]][3])
-}
-
-wc_table <- data.frame(wordcount = vapply(files, getwc, numeric(1)))
-wc_total <- rbind(wc_table, sum(wc_table$wordcount))
-wc_total$percent <- 100 * wc_total$wordcount / 
-  sum(wc_total$wordcount[-nrow(wc_total)])
-rownames(wc_total)[nrow(wc_total)] <- "total"
-knitr::kable(wc_total, digits = 1)
-```
+Section                   Word count          
+--------                  --------------      ------------
+Introduction              1060                Figures:
+Materials and Methods     1603                6 (all colored)
+Results                   1073
+Discussion                1789                Tables:
+Acknowledgements          88                  2
+Total                     5613
 

### 2018-02-16 manuscript/index.Rmd
Fix manuscript title and figure count

@@
 ---
-title: "Does the leaf economic spectrum hold within plant functional types? A Bayesian hierarchical meta-analysis"
+title: "Does the leaf economic spectrum hold within plant functional types? A Bayesian multivariate trait meta-analysis."
 author:
@@
 Introduction              1060                Figures:
-Materials and Methods     1603                6 (all colored)
+Materials and Methods     1603                5 (all colored)
 Results                   1073

### 2018-06-11 supplementary_information.Rmd
First pass at multiple imputation vignette

---
title: Estimating trait means and covariances through multiple imputation
author:
  - >
    Alexey N. Shiklomanov,
    Elizabeth M. Cowdery,
    Michael Bahn,
    Chaeho Byun,
    Joseph Craine,
    Andrés Gonzalez-Melo,
    Steven Jansen,
    Nathan Kraft,
    Koen Kramer,
    Vanessa Minden,
    Ülo Niinemets,
    Yusuke Onoda,
    Enio Egon Sosinski,
    Nadejda A. Soudzilovskaia,
    Michael C. Dietze
output: pdf_document
---

# Introduction

Under _single_ imputation, the imputation step occurs once outside of the MCMC loop.
In _multiple_ imputation, as implemented in this paper, the imputation step is part of the MCMC loop,
such that imputed data values are conditioned on the current state of the parameters and vice-versa.
In the figure below, $Y$ is the original data (with missing values), $Y^*$ is the original data with missing values imputed, $\mu$ and $\Sigma$ are the estimates of the

```{tikz diagram, echo = FALSE}
\usetikzlibrary{positioning, calc, fit}
\begin{tikzpicture}[
	every node/.style = {shape = rectangle, draw = black},
	scale = 1.1
]
\node (Y) at (-1, 3) {$Y$};
\node (data params) at (-1, 2) {$\mu_p$, $\Sigma_p$};
\node (Ystar) at (-1, 1) {$Y^*$};
\node (guess params) [align = center] at (1, 1) {Initial guess\\$\mu^*_0$ $\Sigma^*_0$};
\node (mu star) at (0, 0) {$\mu^*_t \mid Y^*, \Sigma^*_{t-1}$};
\node (sigma star) at (0, -1) {$\Sigma^*_t \mid Y^*, \mu^*_t$};
\path [->] (Y) edge (data params);
\path [->] (data params) edge (Ystar);
\path [->] (Ystar) edge (mu star);
\path [->] (guess params) edge (mu star);
\path [->] (mu star) edge (sigma star);
\path [->, out = 0, in = 0] (sigma star.east) edge
	node (t plus one) [right, draw = none, font = {\footnotesize}] {$t = t + 1$}
	(mu star.east);
\node (mcmc loop) [fit = (mu star)(sigma star)(t plus one), shape = rectangle, draw = black!50] {};
\node (mcmc loop text) [align = center, anchor = west, draw = none] at (mcmc loop.east) {MCMC\\loop};
\node (title) [anchor = north, draw = none] at (mcmc loop.south) {Single imputation};
\end{tikzpicture}
\hspace{3cm}
\begin{tikzpicture}[
	every node/.style = {shape = rectangle, draw = black},
	scale = 1.1
]
\node (Y) at (-1, 1) {$Y$};
\node (guess params) [align = center] at (1, 1) {Initial guess\\$\mu^*_0$ $\Sigma^*_0$};
\node (Ystar) at (0, 0) {$Y^*_t \mid Y, \mu^*_{t-1}, \Sigma^*_{t-1}$};
\node (mu star) at (0, -1) {$\mu^*_t \mid Y^*_t, \Sigma^*_{t-1}$};
\node (sigma star) at (0, -2) {$\Sigma^*_t \mid Y^*_t, \mu^*_t$};
\path [->] (Y) edge (Ystar);
\path [->] (guess params) edge (Ystar);
\path [->] (Ystar) edge (mu star);
\path [->] (mu star) edge (sigma star);
\path [->, out = 0, in = 0] (sigma star.east) edge
	node (t plus one) [right, draw = none, font = {\footnotesize}] {$t = t + 1$}
	(Ystar.east);
\node (mcmc loop) [fit = (Ystar)(mu star)(sigma star)(t plus one), shape = rectangle, draw = black!50] {};
\node (mcmc loop text) [align = center, anchor = west, draw = none] at (mcmc loop.east) {MCMC\\loop};
\node (title) [anchor = north, draw = none] at (mcmc loop.south) {Multiple imputation};
\end{tikzpicture}
```

# Demonstration

Consider two positively correlated traits, A and B.

```{r create_traits}
library(mvtraits)

true_mu <- c(0, 0)
true_cov <- matrix(c(1, 0.8, 0.8, 1), nrow = 2)
Y_all <- random_mvnorm(1000, true_mu, true_cov)
colnames(Y_all) <- c("A", "B")
plot(Y_all, pch = 19)
```

Randomly remove some values from the data.

```{r remove_na}
Y <- Y_all
Y[sample.int(700)] <- NA
plot(Y_all, pch = 19, col = "grey")
points(Y, pch = 19, col = "black")
```

Set an uninformative multivariate normal prior on $\mu$...
```{r prior_mu}
mu_0 <- c(0, 0)
Sigma_0 <- diag(10, 2)
```

...and an uninformative Wishart prior on $\Sigma$.
```{r prior_sigma}
v0 <- 2
S0 <- diag(10, 2)
```

Take an initial guess at the mean and covariance by drawing from their priors.

```{r initial}
mu_star <- random_mvnorm(1, mu_0, Sigma_0)[1,]
Sigma_star <- rWishart(1, v0, S0)[,,1]
```

Impute values based on this initial guess.

```{r impute}
Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)
plot(Y_all, pch = 19, col = "grey")
points(Ystar, pch = 19, col = "red")
points(Y, pch = 19, col = "black")
```

Clearly, these initial values are a bad fit to the data.
Nevertheless, draw $\mu$ and $\Sigma$ conditioned on this set of imputed values.

```{r draw}
mu_star <- draw_mu(Ystar, Sigma_star, mu_0, Sigma_0)
Sigma_star <- draw_Sigma(Ystar, mu_star, v0, S0)
```

Draw a new set of imputed values, conditioned on the current $\mu^*$ and $\Sigma^*$.

```{r impute_2}
Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)
plot(Y_all, pch = 19, col = "grey")
points(Ystar, pch = 19, col = "red")
points(Y, pch = 19, col = "black")
```

Doing this repeatedly (in a loop) gradually improves the estimates of the covariance.

```{r loop}
par(mfrow = c(5, 4), mar = c(3, 3, 2, 0.1), cex = 0.4)
for (i in 1:20) {
  mu_star <- draw_mu(Ystar, Sigma_star, mu_0, Sigma_0)[1,]
  Sigma_star <- draw_Sigma(Ystar, mu_star, v0, S0)
  Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)
  title <- sprintf("iteration: %d, covariance: %.2f", i, Sigma_star[2, 1])
  plot(Y_all, pch = 19, col = "grey", xlab = "", ylab = "", main = title)
  points(Ystar, pch = 19, col = "red")
  points(Y, pch = 19, col = "black")
}
```

After a lot of MCMC samples, we can generate a distribution of covariance values, which provides an uncertainty estimate on the covariance.

```{r cov_estimate}
cov_ab <- numeric(1000)
for (i in seq_along(cov_ab)) {
  mu_star <- draw_mu(Ystar, Sigma_star, mu_0, Sigma_0)[1,]
  Sigma_star <- draw_Sigma(Ystar, mu_star, v0, S0)
  Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)
  cov_ab[i] <- Sigma_star[2, 1]
}
hist(cov_ab, xlab = "Covariance estimate", ylab = "Count", main = "Covariance estimate")
abline(v = true_cov[2, 1], col = "red", lwd = 2)
```

### 2018-06-12 supplementary_information.Rmd
Setup drake. Work on supplementary materials.

@@
 ---
-title: Estimating trait means and covariances through multiple imputation
+title: Supplementary information
 author:
@@
 output: pdf_document
+header-includes:
+  - \usepackage{booktabs}
+  - \usepackage{longtable}
+  - \usepackage{array}
+  - \usepackage{multirow}
+  - \usepackage[table]{xcolor}
+  - \usepackage{wrapfig}
+  - \usepackage{float}
+  - \usepackage{colortbl}
+  - \usepackage{pdflscape}
+  - \usepackage{tabu}
+  - \usepackage{threeparttable}
+  - \usepackage{threeparttablex}
+  - \usepackage[normalem]{ulem}
+  - \usepackage{makecell}
 ---
 
-# Introduction
-
-Under _single_ imputation, the imputation step occurs once outside of the MCMC loop.
-In _multiple_ imputation, as implemented in this paper, the imputation step is part of the MCMC loop,
-such that imputed data values are conditioned on the current state of the parameters and vice-versa.
-In the figure below, $Y$ is the original data (with missing values), $Y^*$ is the original data with missing values imputed, $\mu$ and $\Sigma$ are the estimates of the
-
-```{tikz diagram, echo = FALSE}
-\usetikzlibrary{positioning, calc, fit}
-\begin{tikzpicture}[
-	every node/.style = {shape = rectangle, draw = black},
-	scale = 1.1
-]
-\node (Y) at (-1, 3) {$Y$};
-\node (data params) at (-1, 2) {$\mu_p$, $\Sigma_p$};
-\node (Ystar) at (-1, 1) {$Y^*$};
-\node (guess params) [align = center] at (1, 1) {Initial guess\\$\mu^*_0$ $\Sigma^*_0$};
-\node (mu star) at (0, 0) {$\mu^*_t \mid Y^*, \Sigma^*_{t-1}$};
-\node (sigma star) at (0, -1) {$\Sigma^*_t \mid Y^*, \mu^*_t$};
-\path [->] (Y) edge (data params);
-\path [->] (data params) edge (Ystar);
-\path [->] (Ystar) edge (mu star);
-\path [->] (guess params) edge (mu star);
-\path [->] (mu star) edge (sigma star);
-\path [->, out = 0, in = 0] (sigma star.east) edge
-	node (t plus one) [right, draw = none, font = {\footnotesize}] {$t = t + 1$}
-	(mu star.east);
-\node (mcmc loop) [fit = (mu star)(sigma star)(t plus one), shape = rectangle, draw = black!50] {};
-\node (mcmc loop text) [align = center, anchor = west, draw = none] at (mcmc loop.east) {MCMC\\loop};
-\node (title) [anchor = north, draw = none] at (mcmc loop.south) {Single imputation};
-\end{tikzpicture}
-\hspace{3cm}
-\begin{tikzpicture}[
-	every node/.style = {shape = rectangle, draw = black},
-	scale = 1.1
-]
-\node (Y) at (-1, 1) {$Y$};
-\node (guess params) [align = center] at (1, 1) {Initial guess\\$\mu^*_0$ $\Sigma^*_0$};
-\node (Ystar) at (0, 0) {$Y^*_t \mid Y, \mu^*_{t-1}, \Sigma^*_{t-1}$};
-\node (mu star) at (0, -1) {$\mu^*_t \mid Y^*_t, \Sigma^*_{t-1}$};
-\node (sigma star) at (0, -2) {$\Sigma^*_t \mid Y^*_t, \mu^*_t$};
-\path [->] (Y) edge (Ystar);
-\path [->] (guess params) edge (Ystar);
-\path [->] (Ystar) edge (mu star);
-\path [->] (mu star) edge (sigma star);
-\path [->, out = 0, in = 0] (sigma star.east) edge
-	node (t plus one) [right, draw = none, font = {\footnotesize}] {$t = t + 1$}
-	(Ystar.east);
-\node (mcmc loop) [fit = (Ystar)(mu star)(sigma star)(t plus one), shape = rectangle, draw = black!50] {};
-\node (mcmc loop text) [align = center, anchor = west, draw = none] at (mcmc loop.east) {MCMC\\loop};
-\node (title) [anchor = north, draw = none] at (mcmc loop.south) {Multiple imputation};
-\end{tikzpicture}
+```{r libraries, include = FALSE}
+library(knitr)
+library(kableExtra)
+library(drake)
 ```
 
-# Demonstration
-
-Consider two positively correlated traits, A and B.
-
-```{r create_traits}
-library(mvtraits)
-
-true_mu <- c(0, 0)
-true_cov <- matrix(c(1, 0.8, 0.8, 1), nrow = 2)
-Y_all <- random_mvnorm(1000, true_mu, true_cov)
-colnames(Y_all) <- c("A", "B")
-plot(Y_all, pch = 19)
+```{r sample_size_table, echo = FALSE, results = "asis"}
+kable(
+  readd(sample_size_wide),
+  col.names = c(
+    "PFT", "Leaf lifespan", "SLA",
+    "$N_\\textrm{mass}$", "$N_\\textrm{area}$",
+    "$P_\\textrm{mass}$", "$P_\\textrm{area}$",
+    "$R_\\textrm{d, mass}$", "$R_\\textrm{d, area}$",
+    "$V_\\textrm{c, max, mass}$", "$V_\\textrm{c, max, area}$",
+    "$J_\\textrm{max, mass}$", "$J_\\textrm{max, area}$"
+  ),
+  format = "latex",
+  booktabs = TRUE,
+  escape = FALSE,
+  caption = "The number of non-missing (outside parentheses) and imputed (inside parentheses) observations of each trait for each plant functional type used in this analysis."
+) %>%
+  kable_styling(latex_options = c("scale_down", "hold_position")) %>%
+  group_rows("Broadleaf", 1, 5) %>%
+  group_rows("Needleleaf", 6, 8) %>%
+  group_rows("Shrub", 9, 11) %>%
+  group_rows("Grass", 12, 14)
 ```
 
-Randomly remove some values from the data.
-
-```{r remove_na}
-Y <- Y_all
-Y[sample.int(700)] <- NA
-plot(Y_all, pch = 19, col = "grey")
-points(Y, pch = 19, col = "black")
+```{r pairwise_correlation_table, echo = FALSE, results = "asis"}
+kable(
+  readd(corr_processed), "latex",
+  booktabs = TRUE, longtable = TRUE, escape = FALSE,
+  col.names = c("PFT", "Trait 1", "Trait 2", "Correlation (95\\% CI)"),
+  caption = "Pairwise trait correlation values for each plant functional type. Values with asterisks indicate correlations significantly different from zero ($p < 0.05$)."
+)
 ```
-
-Set an uninformative multivariate normal prior on $\mu$...
-```{r prior_mu}
-mu_0 <- c(0, 0)
-Sigma_0 <- diag(10, 2)
-```
-
-...and an uninformative Wishart prior on $\Sigma$.
-```{r prior_sigma}
-v0 <- 2
-S0 <- diag(10, 2)
-```
-
-Take an initial guess at the mean and covariance by drawing from their priors.
-
-```{r initial}
-mu_star <- random_mvnorm(1, mu_0, Sigma_0)[1,]
-Sigma_star <- rWishart(1, v0, S0)[,,1]
-```
-
-Impute values based on this initial guess.
-
-```{r impute}
-Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)
-plot(Y_all, pch = 19, col = "grey")
-points(Ystar, pch = 19, col = "red")
-points(Y, pch = 19, col = "black")
-```
-
-Clearly, these initial values are a bad fit to the data.
-Nevertheless, draw $\mu$ and $\Sigma$ conditioned on this set of imputed values.
-
-```{r draw}
-mu_star <- draw_mu(Ystar, Sigma_star, mu_0, Sigma_0)
-Sigma_star <- draw_Sigma(Ystar, mu_star, v0, S0)
-```
-
-Draw a new set of imputed values, conditioned on the current $\mu^*$ and $\Sigma^*$.
-
-```{r impute_2}
-Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)
-plot(Y_all, pch = 19, col = "grey")
-points(Ystar, pch = 19, col = "red")
-points(Y, pch = 19, col = "black")
-```
-
-Doing this repeatedly (in a loop) gradually improves the estimates of the covariance.
-
-```{r loop}
-par(mfrow = c(5, 4), mar = c(3, 3, 2, 0.1), cex = 0.4)
-for (i in 1:20) {
-  mu_star <- draw_mu(Ystar, Sigma_star, mu_0, Sigma_0)[1,]
-  Sigma_star <- draw_Sigma(Ystar, mu_star, v0, S0)
-  Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)
-  title <- sprintf("iteration: %d, covariance: %.2f", i, Sigma_star[2, 1])
-  plot(Y_all, pch = 19, col = "grey", xlab = "", ylab = "", main = title)
-  points(Ystar, pch = 19, col = "red")
-  points(Y, pch = 19, col = "black")
-}
-```
-
-After a lot of MCMC samples, we can generate a distribution of covariance values, which provides an uncertainty estimate on the covariance.
-
-```{r cov_estimate}
-cov_ab <- numeric(1000)
-for (i in seq_along(cov_ab)) {
-  mu_star <- draw_mu(Ystar, Sigma_star, mu_0, Sigma_0)[1,]
-  Sigma_star <- draw_Sigma(Ystar, mu_star, v0, S0)
-  Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)
-  cov_ab[i] <- Sigma_star[2, 1]
-}
-hist(cov_ab, xlab = "Covariance estimate", ylab = "Count", main = "Covariance estimate")
-abline(v = true_cov[2, 1], col = "red", lwd = 2)
-```

### 2018-06-12 scripts/drake.R

@@ -1,0 +1,147 @@
+library(shiklomanov2017np)
+library(drake)
+library(knitr)
+library(kableExtra)
+
+import::from(here, inhere = here)
+
+get_sample_size_long <- function(try_dat) {
+  try_dat %>%
+    dplyr::select(pft, dplyr::one_of(both_params)) %>%
+    tidyr::gather(trait, value, -pft, na.rm = FALSE) %>%
+    dplyr::group_by(pft, trait) %>%
+    dplyr::summarize(
+      present = sum(!is.na(value)),
+      missing = sum(is.na(value))
+    ) %>%
+    dplyr::ungroup()
+}
+
+base_plan <- drake_plan(
+  try_data_all = readRDS(inhere("extdata/traits_analysis.rds")),
+  try_dat = try_data_all %>%
+    dplyr::select(pft = clm45, !!both_params) %>%
+    dplyr::mutate(
+      pft = factor(pft, abbr2pft) %>% forcats::lvls_revalue(pft2abbr)
+    ),
+  sample_size_long = get_sample_size_long(try_dat),
+  sample_size_wide = sample_size_long %>%
+    dplyr::mutate(display = sprintf("%d (%d)", present, missing)) %>%
+    dplyr::select(pft, trait, display) %>%
+    tidyr::spread(trait, display),
+  rmarkdown::render(
+    knitr_in("supplementary_information.Rmd"),
+    output_file = file_out("supplementary_information.pdf")
+  ),
+  strings_in_dots = "literals"
+)
+
+corr_data <- drake_plan(
+  hier_summary_mass = readRDS(file_in("output/hier.mass.clm45.NA.2018-02-12-1154.rds"))[["summary_table"]] %>%
+    mutate(mass_area = "mass", modeltype = "hier"),
+  hier_summary_area = readRDS(file_in("output/hier.area.clm45.NA.2018-02-12-1227.rds"))[["summary_table"]] %>%
+    mutate(mass_area = "area", modeltype = "hier"),
+  multi_summary_mass = readRDS(file_in("output/multi.mass.clm45.NA.2018-02-12-1127.rds"))[["summary_table"]] %>%
+    mutate(mass_area = "mass", modeltype = "multi"),
+  multi_summary_area = readRDS(file_in("output/multi.area.clm45.NA.2018-02-12-1146.rds"))[["summary_table"]] %>%
[…]

### 2018-06-12 manuscript/index.Rmd
Fix correlation-sample size table in SI

@@
     latex_engine: xelatex
+    keep_tex: yes
     toc: no

### 2018-06-12 supplementary_information.Rmd
Fix correlation-sample size table in SI

@@
 kable(
-  readd(corr_processed), "latex",
+  readd(corr_processed) %>%
+    dplyr::select(pft, yvar, xvar, corr_value, present, missing),
+  "latex",
   booktabs = TRUE, longtable = TRUE, escape = FALSE,
-  col.names = c("PFT", "Trait 1", "Trait 2", "Correlation (95\\% CI)"),
+  col.names = c("PFT", "Trait 1", "Trait 2", "Correlation (95\\% CI)", "present", "missing"),
   caption = "Pairwise trait correlation values for each plant functional type. Values with asterisks indicate correlations significantly different from zero ($p < 0.05$)."

### 2018-06-12 scripts/drake.R

@@ -144,4 +144,9 @@ if (FALSE) {
   readd(sample_size_wide)
   readd(pairwise_missing) %>% count(pft)
   readd(corr_processed)
+
+  ggplot(readd(corr_processed)) +
+    aes(x = present, y = abs(Mean), color = pft) +
+    geom_point() +
+    scale_x_log10(breaks = c(10, 100, 1000, 10000))
 }

### 2018-06-13 manuscript/02_methods.Rmd
Update reviewer response
Include sample size analysis.

Fix other tables.

Do some text revisions.

@@
 To avoid potential artifacts caused by different trait normalization, we performed analyses separately for both mass- and area-normalized traits [@osnas_global_2013; @lloyd_les].
-We restricted our analysis to TRY data that have been quality-controlled and for which adequate species information was provided for functional type classification [see @kattge_try_2011].
+We restricted our analysis to TRY data that have been quality-controlled, where at least one of the above traits was observed, and for which adequate species information was provided for functional type classification [see @kattge_try_2011].
 
@@
 The fundamentally multivariate nature of the sampling procedure described above makes it incapable of accommodating partially missing observations.
-Therefore, our algorithm also included imputation of partially missing data, which proceeded as follows:
+Therefore, our algorithm also included multiple imputation of partially missing data, which proceeded as follows:
 For a block of data $\mathbf{x'}$ containing missing observations in columns $\mathbf{m}$ and present observations in columns $\mathbf{p}$,
@@
   (\mathbf{\Sigma}[p,p]^{-1} \mathbf{\Sigma}[p,m])$$
+
+We emphasize that the imputation of missing data is performed iteratively as parameters are being estimated, such that each set of imputed values is conditioned on the current sampled state of the parameters.
+This approach is therefore distinct from single imputation, where data are imputed first in a separate step prior to parameter estimation.
+A demonstration of this multiple imputation approach and how it is used to estimate trait covariance is provided in a literate programming vignette in the supplementary information to this manuscript.
 

### 2018-06-13 supplementary_information.Rmd
Update reviewer response
Include sample size analysis.

Fix other tables.

Do some text revisions.

--- 2018-06-12 supplementary_information.Rmd
@@
 library(drake)
+library(ggplot2)
 ```
@@
 ```{r sample_size_table, echo = FALSE, results = "asis"}
+ssw <- readd(sample_size_wide)
 kable(
-  readd(sample_size_wide),
+  ssw,
   col.names = c(
@@
   kable_styling(latex_options = c("scale_down", "hold_position")) %>%
-  group_rows("Broadleaf", 1, 5) %>%
-  group_rows("Needleleaf", 6, 8) %>%
-  group_rows("Shrub", 9, 11) %>%
-  group_rows("Grass", 12, 14)
+  group_rows("Broadleaf", 2, 6) %>%
+  group_rows("Needleleaf", 7, 9) %>%
+  group_rows("Shrub", 10, 12) %>%
+  group_rows("Grass", 13, 15) %>%
+  landscape()
 ```
@@
 ```{r pairwise_correlation_table, echo = FALSE, results = "asis"}
+corr_results <- readd(corr_processed)
 kable(
-  readd(corr_processed), "latex",
+  corr_results %>%
+    dplyr::select(pft, yvar, xvar, corr_value, present, missing),
+  "latex",
   booktabs = TRUE, longtable = TRUE, escape = FALSE,
-  col.names = c("PFT", "Trait 1", "Trait 2", "Correlation (95\\% CI)"),
+  col.names = c("PFT", "Trait 1", "Trait 2", "Correlation (95\\% CI)", "present", "missing"),
   caption = "Pairwise trait correlation values for each plant functional type. Values with asterisks indicate correlations significantly different from zero ($p < 0.05$)."
 )
+
+correlation_scatter_cap <- paste(
+  "Pairwise mean squared trait correlation values as a function of sample size.",
+  "Note that the x-axis uses a $\\log_10$ scale."
+)
+correlation_fig_cap <- paste(
+  "Fraction of statistically significant ($p < 0.05$) pairwise trait correlation estimates",
+  "as a function of sample size for each plant functional type."
+)
 ```
+
+```{r correlation_sample_size_pairs, echo = FALSE, fig.cap = correlation_scatter_cap}
+corr_results %>%
+  mutate(
+    significant = factor(significant, c("TRUE", "FALSE")) %>%
+      forcats::lvls_revalue(c("yes", "no"))
+  ) %>%
+  ggplot() +
+  aes(x = present + 0.1, y = Mean, color = significant) +
+  geom_point() +
+  geom_hline(yintercept = 0, linetype = "dashed") +
+  facet_wrap(~ pft) +
+  scale_x_log10() +
+  labs(x = "Sample size", y = "Correlation") +
+  theme_bw() +
+  theme(
+	legend.position = "bottom"
+  )
+```
+
+```{r correlation_sample_size_analysis, echo = FALSE, fig.cap = correlation_fig_cap}
+corr_binned <- corr_results %>%
+  dplyr::mutate(
+    ss_bin = cut(
+      present,
+      c(0, 1, 10, 25, 50, 100, 500, 1000, Inf),
+      c("0", "< 10", "< 25", "< 50", "< 100", "< 500", "< 1000", "> 1000"),
+      include.lowest = TRUE,
+      right = FALSE
+    )
+  ) %>%
+  group_by(pft, ss_bin) %>%
+  summarize(frac_significant = mean(significant))
+
+ggplot(corr_binned) +
+  aes(x = ss_bin, y = frac_significant) +
+  geom_segment(aes(xend = ss_bin, yend = 0)) +
+  geom_point() +
+  facet_wrap(~ pft) +
+  labs(x = "Sample size", y = "Fraction of significant correlations") +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90, hjust = 0.9))
+```

### 2018-06-13 scripts/drake.R

@@ -6,15 +6,26 @@ library(kableExtra)
 import::from(here, inhere = here)
 
 get_sample_size_long <- function(try_dat) {
-  try_dat %>%
-    dplyr::select(pft, dplyr::one_of(both_params)) %>%
-    tidyr::gather(trait, value, -pft, na.rm = FALSE) %>%
-    dplyr::group_by(pft, trait) %>%
+  count_missing <- . %>%
     dplyr::summarize(
       present = sum(!is.na(value)),
       missing = sum(is.na(value))
-    ) %>%
+    )
+  try_long <- try_dat %>%
+    dplyr::select(pft, dplyr::one_of(both_params)) %>%
+    tidyr::gather(trait, value, -pft, na.rm = FALSE)
+  by_pft <- try_long %>%
+    dplyr::group_by(pft, trait) %>%
+    count_missing %>%
     dplyr::ungroup()
+  global <- try_long %>%
+    dplyr::group_by(trait) %>%
+    count_missing %>%
+    dplyr::mutate(
+      pft = factor("GLOB", pft2abbr) %>%
+        forcats::fct_recode("Across-PFT" = "GLOB")
+    )
+  dplyr::bind_rows(global, by_pft)
 }
 
 base_plan <- drake_plan(
@@ -140,13 +151,5 @@ dconfig <- drake_config(plan)
 make(plan)
 
 if (FALSE) {
-  loadd(plan)
-  readd(sample_size_wide)
-  readd(pairwise_missing) %>% count(pft)
-  readd(corr_processed)
-
-  ggplot(readd(corr_processed)) +
-    aes(x = present, y = abs(Mean), color = pft) +
-    geom_point() +
-    scale_x_log10(breaks = c(10, 100, 1000, 10000))
+  readd(sample_size_long)
 }

### 2018-06-13 supplementary_information.Rmd
Minor fixes to comments and supplementary info

@@
   "Pairwise mean squared trait correlation values as a function of sample size.",
-  "Note that the x-axis uses a $\\log_10$ scale."
+  "Note that the x-axis uses a $\\log_{10}$ scale."
 )

### 2018-06-18 supplementary_information.Rmd
Add data range analysis and some more review responses

--- 2018-06-13 supplementary_information.Rmd
@@
 library(ggplot2)
+
+correlation_scatter_cap <- paste(
+  "Pairwise mean squared trait correlation values as a function of sample size.",
+  "Note that the x-axis uses a $\\log_{10}$ scale."
+)
+
+correlation_fig_cap <- paste(
+  "Fraction of statistically significant ($p < 0.05$) pairwise trait correlation estimates",
+  "as a function of sample size for each plant functional type."
+)
+
+correlation_range_cap <- paste(
+  "Relationship between pairwise correlation strength and data range for each PFT.",
+  "In each pair of sub-panels, the left panel is relationship between pairwise correlation",
+  "and the range of the variable on the x axis",
+  "while the right panel is the relationship between pairwise correlation",
+  "and the y variable",
+  "(For instance, in the two-panel figure in row 1 column 2,",
+  "the left panel is the SLA-leaf lifespan correlation as a function of the range of SLA",
+  "and the right panel is the same correlation as a function of the range of leaf lifespan).",
+  "Data range is expressed as the difference between the 2.5% and 97.5% quantiles of each normalized trait (i.e. $\\frac{X}{E[X]}$).",
+  "Grey shading is the 95% confidence interval around a local regression (LOESS).",
+  "Upper triangle is area-normalized values, and lower triangle is mass-normalized values."
+)
+```
+
+```{r mean_value_table, echo = FALSE, results = "asis"}
+vals <- readd()
 ```
@@
 )
-
-correlation_scatter_cap <- paste(
-  "Pairwise mean squared trait correlation values as a function of sample size.",
-  "Note that the x-axis uses a $\\log_10$ scale."
-)
-correlation_fig_cap <- paste(
-  "Fraction of statistically significant ($p < 0.05$) pairwise trait correlation estimates",
-  "as a function of sample size for each plant functional type."
-)
 ```
@@
 
-```{r correlation_sample_size_analysis, echo = FALSE, fig.cap = correlation_fig_cap}
+```{r correlation_sample_size_analysis, echo = FALSE, include = FALSE, eval = FALSE, fig.cap = correlation_fig_cap}
 corr_binned <- corr_results %>%
@@
 ```
+
+```{r correlation_range_analysis, echo = FALSE, results = "hide", warning = FALSE, fig.cap = correlation_range_cap, fig.dim = c(12, 12)}
+readd(corr_range_plot_gg)
+```

### 2018-06-18 scripts/drake.R

@@ -5,6 +5,8 @@ library(kableExtra)
 
 import::from(here, inhere = here)
 
+pkgconfig::set_config("drake::strings_in_dots" = "literals")
+
 get_sample_size_long <- function(try_dat) {
   count_missing <- . %>%
     dplyr::summarize(
@@ -43,8 +45,7 @@ base_plan <- drake_plan(
   rmarkdown::render(
     knitr_in("supplementary_information.Rmd"),
     output_file = file_out("supplementary_information.pdf")
-  ),
-  strings_in_dots = "literals"
+  )
 )
 
 corr_data <- drake_plan(
@@ -55,8 +56,7 @@ corr_data <- drake_plan(
   multi_summary_mass = readRDS(file_in("output/multi.mass.clm45.NA.2018-02-12-1127.rds"))[["summary_table"]] %>%
     mutate(mass_area = "mass", modeltype = "multi"),
   multi_summary_area = readRDS(file_in("output/multi.area.clm45.NA.2018-02-12-1146.rds"))[["summary_table"]] %>%
-    mutate(mass_area = "area", modeltype = "multi"),
-  strings_in_dots = "literals"
+    mutate(mass_area = "area", modeltype = "multi")
 )
 
 corr_proc <- function(dat) {
@@ -135,21 +135,119 @@ corr_processed_plan <- drake_plan(
       pft = forcats::fct_recode(pft, "Across-PFT" = "GLOB")
     ) %>%
     ## dplyr::select(pft, yvar, xvar, corr_value) %>%
-    dplyr::arrange(pft),
-  strings_in_dots = "literals"
+    dplyr::arrange(pft)
 )
+
+corr_plot <- function(data, ma) {
+  data %>%
+    tidyr::gather(variable, z_range_95, x_zrange, y_zrange) %>%
+    dplyr::mutate(
+      variable = stringr::str_remove(variable, "_zrange"),
+      variable = dplyr::case_when(
+        ma == "mass" & variable == "x" ~ "y",
+        ma == "mass" & variable == "y" ~ "x",
+        TRUE ~ variable
+      )
+    ) %>%
+    ggplot() +
+    aes(x = z_range_95, y = Mean2, color = pft) +
+    geom_smooth(color = NA, method = "loess") +
+    geom_point() +
+    facet_grid(variable ~ .) +
+    scale_color_manual(values = c("black", pft_colors)) +
[…]

### 2018-06-20 manuscript/02_methods.Rmd
Minor text revisions

@@
 The R code and ancillary data for running these analyses is publicly available online via the Open Science Framework (OSF) at <https://osf.io/w8y73/>.
-The TRY data used for this analysis can be requested at <http://try-db.org>.
+The TRY data request used for this analysis has been archived at <http://try-db.org>,
+and can be retrieved by providing the request ID (#1584).
+Alternatively, the exact pre-formatted data used in this analysis are available on request to Alexey Shiklomanov (ashiklom@bu.edu).

### 2018-06-20 supplementary_information.Rmd
Fix correlation tables

@@
   - \usepackage{makecell}
+  - \renewcommand{\thetable}{S\arabic{table}}
+  - \renewcommand{\thefigure}{S\arabic{figure}}
 ---
@@
 ```{r mean_value_table, echo = FALSE, results = "asis"}
-vals <- readd()
+table_cap_mass <- paste(
+  "Mean and 95\\% confidence interval of trait estimates",
+  "for mass-normalized traits from the hierarchical model."
+)
+table_cap_area <- gsub("mass", "area", table_cap_mass)
+
+readd(table_dat) %>%
+  select(-matches("area")) %>%
+  kable(
+    caption = table_cap_mass,
+    format = "latex",
+    col.names = linebreak(c(
+      "PFT",
+      "Leaf lifespan\n(months)",
+      "SLA\n(kg m$^{-2}$)",
+      "$N_\\textrm{mass}$\n(mg g$^{-1}$)",
+      "$P_\\textrm{mass}$\n(mg g$^{-1}$)",
+      "$R_\\textrm{d, mass}$\n($\\mu$mol g$^{-1}$ s$^{-1}$)",
+      "$V_\\textrm{c, max, mass}$\n($\\mu$mol g$^{-1}$ s$^{-1}$)",
+      "$J_\\textrm{max, mass}$\n($\\mu$mol g$^{-1}$ s$^{-1}$)"
+    )),
+    escape = FALSE,
+    booktabs = TRUE
+  ) %>%
+  kableExtra::kable_styling(latex_options = c("scale_down", "hold_position"))
+
+readd(table_dat) %>%
+  select(-matches("mass")) %>%
+  kable(
+	caption = table_cap_area,
+	format = "latex",
+    col.names = linebreak(c(
+      "PFT",
+      "Leaf lifespan\n(months)",
+      "SLA\n(kg m$^{-2}$)",
+      "$N_\\textrm{area}$\n(g m$^{-2}$)",
+      "$P_\\textrm{area}$\n(g m$^{-2}$)",
+      "$R_\\textrm{d, area}$\n($\\mu$mol m$^{-2}$ s$^{-1}$)",
+      "$V_\\textrm{c, max, area}$\n($\\mu$mol m$^{-2}$ s$^{-1}$)",
+      "$J_\\textrm{max, area}$\n($\\mu$mol m$^{-2}$ s$^{-1}$)"
+    )),
+	escape = FALSE,
+	booktabs = TRUE
+) %>%
+  kableExtra::kable_styling(latex_options = c("scale_down", "hold_position"))
 ```
@@
   "latex",
-  booktabs = TRUE, longtable = TRUE, escape = FALSE,
-  col.names = c("PFT", "Trait 1", "Trait 2", "Correlation (95\\% CI)", "present", "missing"),
+  booktabs = TRUE, longtable = TRUE, escape = TRUE,
+  col.names = c("PFT", "Trait 1", "Trait 2", "Correlation (95% CI)", "present", "missing"),
   caption = "Pairwise trait correlation values for each plant functional type. Values with asterisks indicate correlations significantly different from zero ($p < 0.05$)."

### 2018-06-20 scripts/drake.R

@@ -130,9 +130,9 @@ corr_processed_plan <- drake_plan(
       significant = sign(`2.5%`) == sign(`97.5%`),
       sig_star = dplyr::if_else(significant, "*", ""),
       corr_value = sprintf("%.3f (%.3f, %.3f)%s", Mean, `2.5%`, `97.5%`, sig_star),
-      xvar = factor(xvar, names(traits_latex)) %>% forcats::lvls_revalue(unname(traits_latex)),
-      yvar = factor(yvar, names(traits_latex)) %>% forcats::lvls_revalue(unname(traits_latex)),
-      pft = forcats::fct_recode(pft, "Across-PFT" = "GLOB")
+      xvar = factor(xvar, names(traits_latex)),# %>% forcats::lvls_revalue(unname(traits_latex)),
+      yvar = factor(yvar, names(traits_latex))# %>% forcats::lvls_revalue(unname(traits_latex)),
+      ## pft = forcats::fct_recode(pft, "Across-PFT" = "GLOB")
     ) %>%
     ## dplyr::select(pft, yvar, xvar, corr_value) %>%
     dplyr::arrange(pft)
@@ -160,6 +160,28 @@ corr_plot <- function(data, ma) {
     labs(color = "PFT")
 }
 
+corr_grid <- function(data) {
+  data %>%
+    mutate(
+      significant = factor(significant, c("TRUE", "FALSE")) %>%
+        forcats::lvls_revalue(c("yes", "no"))
+    ) %>%
+  ggplot() +
+    aes(x = present + 0.1, y = Mean ^ 2, color = pft) + # size = significant
+    geom_point() +
+    geom_smooth(aes(group = 1), method = "loess", color = NA, span = 2) +
+    scale_x_log10(breaks = c(1, 100, 10000), minor_breaks = c(10, 1000), limits = c(0.1, 30000)) +
+    coord_cartesian(ylim = c(0, 1)) +
+    ## scale_size_manual(values = c("yes" = 3, "no" = 1)) +
+    scale_color_manual(name = "PFT", values = c("black", pft_colors)) +
+    theme_bw()
+}
+
+if (FALSE) {
+
+
+}
+
 text_plot <- function(text) {
   ggplot() +
     geom_text(aes(label = text, x = 0, y = 0)) +
[…]

### 2018-06-20 supplementary_information.Rmd
Remove deprecated plots. Add new correlation plot.

@@
 
-```{r correlation_sample_size_pairs, echo = FALSE, fig.cap = correlation_scatter_cap}
-corr_results %>%
-  mutate(
-    significant = factor(significant, c("TRUE", "FALSE")) %>%
-      forcats::lvls_revalue(c("yes", "no"))
-  ) %>%
-  ggplot() +
-  aes(x = present + 0.1, y = Mean, color = significant) +
-  geom_point() +
-  geom_hline(yintercept = 0, linetype = "dashed") +
-  facet_wrap(~ pft) +
-  scale_x_log10() +
-  labs(x = "Sample size", y = "Correlation") +
-  theme_bw() +
-  theme(
-	legend.position = "bottom"
-  )
-```
-
-```{r correlation_sample_size_analysis, echo = FALSE, include = FALSE, eval = FALSE, fig.cap = correlation_fig_cap}
-corr_binned <- corr_results %>%
-  dplyr::mutate(
-    ss_bin = cut(
-      present,
-      c(0, 1, 10, 25, 50, 100, 500, 1000, Inf),
-      c("0", "< 10", "< 25", "< 50", "< 100", "< 500", "< 1000", "> 1000"),
-      include.lowest = TRUE,
-      right = FALSE
-    )
-  ) %>%
-  group_by(pft, ss_bin) %>%
-  summarize(frac_significant = mean(significant))
-
-ggplot(corr_binned) +
-  aes(x = ss_bin, y = frac_significant) +
-  geom_segment(aes(xend = ss_bin, yend = 0)) +
-  geom_point() +
-  facet_wrap(~ pft) +
-  labs(x = "Sample size", y = "Fraction of significant correlations") +
-  theme_bw() +
-  theme(axis.text.x = element_text(angle = 90, hjust = 0.9))
+```{r correlation_sample_size_pairs, echo = FALSE, results = "hide", warning = FALSE, fig.cap = correlation_scatter_cap, fig.dim = c(12, 12)}
+readd(corr_ss_plot_gg)
 ```

### 2018-06-21 manuscript/02_methods.Rmd
Add some correlation text to manuscript

@@
 
-The R code and ancillary data for running these analyses is publicly available online via the Open Science Framework (OSF) at <https://osf.io/w8y73/>.
+The magnitude of correlation coefficients can be strongly influenced by data sampling, particularly sample size and range (correlations are likely to be stronger when data have more samples and a larger range).
+To evaluate the contribution of uneven data sampling to our correlation predictions, we plotted the square of each pairwise correlation coefficient against pairwise sample size and normalized data range (Fig.\ S1 and S2).
+
+The R code and ancillary data for running these analyses is publicly available online via the Open Science Framework at <https://osf.io/w8y73/>.
 The TRY data request used for this analysis has been archived at <http://try-db.org>,
-and can be retrieved by providing the request ID (#1584).
+and can be retrieved by providing the TRY data request ID (#1584).
 Alternatively, the exact pre-formatted data used in this analysis are available on request to Alexey Shiklomanov (ashiklom@bu.edu).

### 2018-06-21 manuscript/03_results.Rmd
Fix manuscript typos. Add cover letter.
Add some correlation text to manuscript

@@
 Similarly, the correlation between $P_{area}$ and leaf lifespan was positive globally but negative for evergreen shrubs and not significant for any other PFTs.
-The correlation between leaf lifespan and $R_{d,area}$ significant and negative globally, but was not significant within any PFTs.
-The only significant correlations of leaf lifespan with $V_{c,max,area}$ and $J_{max,area}$ were negative for temperate broadleaved deciduous trees.
-
-A large number of pairwise trait correlations were not significant.
-In some cases, this was driven by sample size (Fig.\ 1).
-For instance, needleleaved deciduous trees, the most undersampled PFT in our analysis, were often the only PFT for which a correlation was not statistically significant.
-In other cases, though, PFTs with smaller sample sizes had significant pairwise correlations while PFTs with much larger sample sizes had none.
-For example, tropical broadleaved evergreen trees were relatively well-sampled for all traits, but none of their area-normalized traits were significantly correlated with leaf lifespan.
-In general, we observed fewer significant trait correlations among area-normalized traits than mass-normalized traits.
+The correlation between leaf lifespan and $R_{d,area}$ was significant and negative globally, but was not significant within any PFTs.
+Finally, although we did not include maximum photosynthesis rate ($A_{max}$)---a key trait in the canonical leaf economic spectrum (Wright et al.\ 2004)---$V_{c,max,mass}$ and $J_{max,mass}$ generally exhibited the expected positive correlations with SLA and negative correlations with leaf lifespan, as did $R_{d,mass}$, though many correlations were not significant, primarily due to sample size.
 
 The strength of pairwise trait correlations varied substantially depending on scale, PFT, and trait (Fig.\ 5).
+For example, the correlations of SLA with $N_{mass}$ and $P_{mass}$ were generally weaker in needleleaved PFTs than in broadleaved PFTs.
+On the other hand, relationships of SLA and $N_{area}$ were strongly negative for all PFTs (except the data-limited needleleaved deciduous trees), and temperate needleleaved species in particular had among the strongest SLA-$N_{area}$ and SLA-$P_{area}$ correlations of any PFT.
+
+The strength of the correlations of $N$ with photosynthetic parameters $V_{c,max}$ and $J_{max}$ varied by plant functional type, regardless of whether the traits were normalized by mass or area.
+
+(TODO Flesh out this text).
+
 The two pairwise trait correlations that exhibited the most consistent strength globally and within each PFT were the correlation between SLA and $N_{area}$, and between $N_{mass}$ and $P_{mass}$.
-Correlation strength was often, but not always, related to sample size, with well-sampled PFTs exhibiting stronger correlations and undersampled PFTs exhibiting weaker correlations.
+
+In some cases, this variability was driven by low sample sizes (Fig.\ 1, S1; Tab.\ S3, S4).
+For instance, needleleaved deciduous trees, the most undersampled PFT in our analysis, were often the only PFT for which a correlation was not statistically significant.
+Similarly, we had no observations of dark respirations for deciduous boreal shrubs, which explains why we found no significant correlations of dark respiration with any other trait for that PFT.
+That being said, the relationship between correlations strength and sample size was very inconsistent (Fig.\ S1; Tab.\ S4).
+For example, temperate needleleaf evergreen trees and C3 arctic grasses showed the strongest $N_{area}$-$P_{area}$ correlations despite having relatively average sample sizes for both traits.
+In fact, every trait pair we investigated had at least one (and often several) cases where a PFT with a larger sample size showed weaker correlations than PFTs with lower sample sizes, or where correlation strength varied significantly among PFTs with similar sample sizes (Fig.\ S1).
+Relationships between correlation strength and data range were even less consistent (Fig.\ S2).
+Based on these results, we conclude that the variation we observed in our correlation results is not driven primarily by sampling artifacts and at least partially captures some underlying ecophysiological differences between PFTs.

### 2018-06-21 supplementary_information.Rmd
Add some correlation text to manuscript

@@
 correlation_scatter_cap <- paste(
-  "Pairwise mean squared trait correlation values as a function of sample size.",
+  "Relationship between squared pairwise correlation and sample size of pairwise observations for each PFT.",
+  "Grey shading is the 95% confidence interval around a local regression (LOESS).",
+  "Upper triangle is area-normalized values, and lower triangle is mass-normalized values.",
   "Note that the x-axis uses a $\\log_{10}$ scale."
@@
   "Mean and 95\\% confidence interval of trait estimates",
-  "for mass-normalized traits from the hierarchical model."
+  "for mass-normalized traits from the hierarchical model.",
+  "Note that some traits for some PFTs have effectively no constraint relative to the mostly uninformative prior (as evidenced by their very large confidence intervals),",
+  "so their values should be used with caution."
 )
@@
 kable(
-  ssw,
+  ssw %>%
+    dplyr::select(pft, leaf_lifespan, SLA,
+                  Nmass, Narea, Pmass, Parea, Rdmass, Rdarea,
+                  Vcmax_mass, Vcmax_area, Jmax_mass, Jmax_area),
   col.names = c(

### 2018-06-21 scripts/drake.R

@@ -228,7 +228,7 @@ range_plan <- drake_plan(
     yvar = yvar_grid
   ) %>%
     dplyr::mutate(
-      mass_area = factor(mass_area) %>% forcats::lvls_revalue(c("mass", "area")),
+      mass_area = factor(mass_area) %>% forcats::lvls_revalue(c("area", "mass")),
       xvar = factor(xvar, 1:7) %>% forcats::lvls_revalue(trait_names),
       yvar = factor(yvar, 1:7) %>% forcats::lvls_revalue(trait_names),
     ),
@@ -281,7 +281,7 @@ range_plan <- drake_plan(
       xvar = factor(xvar, names(traits_latex)),
       yvar = factor(yvar, names(traits_latex))
     ),
-  corr_plots = corr_range_data %>%
+  corr_plots_raw = corr_range_data %>%
     dplyr::group_by(mass_area, xvar, yvar) %>%
     tidyr::nest() %>%
     dplyr::mutate(
@@ -289,8 +289,8 @@ range_plan <- drake_plan(
       plt = purrr::map2(data, ma, corr_plot),
       xvar = forcats::fct_relabel(xvar, ~stringr::str_remove(., "_?(mass|area)")),
       yvar = forcats::fct_relabel(yvar, ~stringr::str_remove(., "_?(mass|area)"))
-    ) %>%
-    dplyr::arrange(yvar, mass_area, xvar),
+    ),
+  corr_plots = dplyr::left_join(trait_grid_table, corr_plots_raw),
   corr_plot_list = corr_plots$plt %>%
     append(list(text_plot("Leaf lifespan")), 0) %>%
     append(list(text_plot("SLA")), 8) %>%
@@ -313,10 +313,6 @@ range_plan <- drake_plan(
 censor <- function(rxp, pft, pfts, param, value) {
   if_else(grepl(rxp, param) & pfts %in% pft, "---", value)
 }
-
-## param_tex <- param_markdown
-## ## stringr::str_replace(param_markdown, " \\(", "\\\\(")
-## names(param_tex) <- names(param_markdown)
 
 results_plan <- drake_plan(
   results_all = readRDS(file_in("results/mvtraits_results.rds")),
@@ -396,11 +392,11 @@ results_plan <- drake_plan(
       lo_f = formatC(mu_lo, digits = 3),
       mid_f = formatC(mu_mean, digits = 3),
       hi_f = formatC(mu_hi, digits = 3),
[…]

### 2018-06-22 manuscript/03_results.Rmd
Minor tweaks to results and discussion text on Vcmax/Jmax

@@
 
-The strength of the correlations of $N$ with photosynthetic parameters $V_{c,max}$ and $J_{max}$ varied by plant functional type, regardless of whether the traits were normalized by mass or area.
+The strength of the correlations of $N$ with photosynthetic parameters $V_{c,max}$ and $J_{max}$ varied by PFT, regardless of whether the traits were normalized by mass or area.
 

### 2018-06-22 manuscript/03_results.Rmd
Partial revert of "Add some correlation text to manuscript"
This partially reverts commit
22382ee156d621251337e5d9b9921b31d5a1b6b9.

Decided to stick mostly to old correlation text.

@@
 The correlation between leaf lifespan and $R_{d,area}$ was significant and negative globally, but was not significant within any PFTs.
-Finally, although we did not include maximum photosynthesis rate ($A_{max}$)---a key trait in the canonical leaf economic spectrum (Wright et al.\ 2004)---$V_{c,max,mass}$ and $J_{max,mass}$ generally exhibited the expected positive correlations with SLA and negative correlations with leaf lifespan, as did $R_{d,mass}$, though many correlations were not significant, primarily due to sample size.
+The only significant correlations of leaf lifespan with $V_{c,max,area}$ and $J_{max,area}$ were negative for temperate broadleaved deciduous trees.
 
 The strength of pairwise trait correlations varied substantially depending on scale, PFT, and trait (Fig.\ 5).
-For example, the correlations of SLA with $N_{mass}$ and $P_{mass}$ were generally weaker in needleleaved PFTs than in broadleaved PFTs.
-On the other hand, relationships of SLA and $N_{area}$ were strongly negative for all PFTs (except the data-limited needleleaved deciduous trees), and temperate needleleaved species in particular had among the strongest SLA-$N_{area}$ and SLA-$P_{area}$ correlations of any PFT.
-
-The strength of the correlations of $N$ with photosynthetic parameters $V_{c,max}$ and $J_{max}$ varied by PFT, regardless of whether the traits were normalized by mass or area.
-
-(TODO Flesh out this text).
-
-The two pairwise trait correlations that exhibited the most consistent strength globally and within each PFT were the correlation between SLA and $N_{area}$, and between $N_{mass}$ and $P_{mass}$.
-
 In some cases, this variability was driven by low sample sizes (Fig.\ 1, S1; Tab.\ S3, S4).

### 2018-06-22 manuscript/03_results.Rmd
Minor tweaks to results and discussion text

@@
 For example, temperate needleleaf evergreen trees and C3 arctic grasses showed the strongest $N_{area}$-$P_{area}$ correlations despite having relatively average sample sizes for both traits.
-In fact, every trait pair we investigated had at least one (and often several) cases where a PFT with a larger sample size showed weaker correlations than PFTs with lower sample sizes, or where correlation strength varied significantly among PFTs with similar sample sizes (Fig.\ S1).
+In fact, every trait pair we investigated had at least one case (and often several cases) where a PFT with a larger sample size showed weaker correlations than PFTs with lower sample sizes, or where correlation strength varied significantly among PFTs with similar sample sizes (Fig.\ S1).
 Relationships between correlation strength and data range were even less consistent (Fig.\ S2).

### 2018-06-22 manuscript/index.Rmd
Update word count and imputation vignette

@@
 Introduction              1060                Figures:
-Materials and Methods     1603                5 (all colored)
-Results                   1073
-Discussion                1789                Tables:
+Materials and Methods     1695                5 (all colored)
+Results                   1108
+Discussion                1907                Tables:
 Acknowledgements          88                  2
-Total                     5613
+Total                     5858
 

### 2018-06-22 second-reviewer-response/multiple_imputation.Rmd
Update word count and imputation vignette

---
title: Estimating trait means and covariances through multiple imputation
author:
  - >
    Alexey N. Shiklomanov,
    Elizabeth M. Cowdery,
    Michael Bahn,
    Chaeho Byun,
    Joseph Craine,
    Andrés Gonzalez-Melo,
    Steven Jansen,
    Nathan Kraft,
    Koen Kramer,
    Vanessa Minden,
    Ülo Niinemets,
    Yusuke Onoda,
    Enio Egon Sosinski,
    Nadejda A. Soudzilovskaia,
    Michael C. Dietze
output: pdf_document
---

# Introduction

Under _single_ imputation, the imputation step occurs once outside of the MCMC loop.
In _multiple_ imputation, as implemented in this paper, the imputation step is part of the MCMC loop,
such that imputed data values are conditioned on the current state of the parameters and vice-versa.
In the figure below,
$Y$ is the original data (with missing values),
$Y^*$ is the original data with missing values imputed,
$\mu_p$ and $\Sigma_p$ are values of the mean vector and covariance matrix (respectively) calculated only from the original data,
and $\mu^*_t$ and $\Sigma^*_t$ are the draws of the mean vector and covariance matrix (respectively) at MCMC iteration $t$.

```{tikz diagram, echo = FALSE}
\usetikzlibrary{positioning, calc, fit}
\begin{tikzpicture}[
	every node/.style = {shape = rectangle, draw = black},
	scale = 1.1
]
\node (Y) at (-1, 3) {Data $Y$};
\node (data params) at (-1, 2) {Calculate $\mu_p, \Sigma_p$};
\node (Ystar) at (-1, 1) {Impute $Y^* \mid \mu_p, \Sigma_p$};
\node (guess params) [align = center, right = of Ystar] {Initial guess\\$\mu^*_0$ $\Sigma^*_0$};
\node (mu star) at (0, 0) {Draw $\mu^*_t \mid Y^*, \Sigma^*_{t-1}$};
\node (sigma star) at (0, -1) {Draw $\Sigma^*_t \mid Y^*, \mu^*_t$};
\path [->] (Y) edge (data params);
\path [->] (data params) edge (Ystar);
\path [->] (Ystar) edge (mu star);
\path [->] (guess params) edge (mu star);
\path [->] (mu star) edge (sigma star);
\path [->, out = 0, in = 0] (sigma star.east) edge
	node (t plus one) [right, draw = none, font = {\footnotesize}] {$t = t + 1$}
	(mu star.east);
\node (mcmc loop) [fit = (mu star)(sigma star)(t plus one), shape = rectangle, draw = black!50] {};
\node (mcmc loop text) [align = center, anchor = west, draw = none] at (mcmc loop.east) {MCMC\\loop};
\node (title) [anchor = north, draw = none] at (mcmc loop.south) {Single imputation};
\end{tikzpicture}
\hspace{1cm}
\begin{tikzpicture}[
	every node/.style = {shape = rectangle, draw = black},
	scale = 1.1
]
\node (Y) at (-1, 1) {Data $Y$};
\node (guess params) [align = center] at (1, 1) {Initial guess\\$\mu^*_0$ $\Sigma^*_0$};
\node (Ystar) at (0, 0) {Impute $Y^*_t \mid Y, \mu^*_{t-1}, \Sigma^*_{t-1}$};
\node (mu star) at (0, -1) {Draw $\mu^*_t \mid Y^*_t, \Sigma^*_{t-1}$};
\node (sigma star) at (0, -2) {Draw $\Sigma^*_t \mid Y^*_t, \mu^*_t$};
\path [->] (Y) edge (Ystar);
\path [->] (guess params) edge (Ystar);
\path [->] (Ystar) edge (mu star);
\path [->] (mu star) edge (sigma star);
\path [->, out = 0, in = 0] (sigma star.east) edge
	node (t plus one) [right, draw = none, font = {\footnotesize}] {$t = t + 1$}
	(Ystar.east);
\node (mcmc loop) [fit = (Ystar)(mu star)(sigma star)(t plus one), shape = rectangle, draw = black!50] {};
\node (mcmc loop text) [align = center, anchor = west, draw = none] at (mcmc loop.east) {MCMC\\loop};
\node (title) [anchor = north, draw = none] at (mcmc loop.south) {Multiple imputation};
\end{tikzpicture}
```

# Demonstration

Consider two positively correlated traits, A and B.

```{r create_traits}
set.seed(12345) # For reproducibility
library(mvtraits)

true_mu <- c(0, 0)
true_cov <- matrix(c(1, 0.8, 0.8, 1), nrow = 2)
Y_all <- random_mvnorm(1000, true_mu, true_cov)
colnames(Y_all) <- c("A", "B")
plot(Y_all, pch = 19)
```

Simulate missingness by randomly removing half of the data.

```{r remove_na}
Y <- Y_all
Y[sample.int(2000, 1000)] <- NA
plot(Y_all, pch = 19, col = "grey")
points(Y, pch = 19, col = "black")
legend(
  "topleft",
  legend = c("True missing", "Present"),
  col = c("grey", "black"),
  pch = 19,
  bg = "white"
)
```

Set an uninformative multivariate normal prior on $\mu$...
```{r prior_mu}
mu_0 <- c(0, 0)
Sigma_0 <- diag(10, 2)
```

...and an uninformative Wishart prior on $\Sigma$.
```{r prior_sigma}
v0 <- 2
S0 <- diag(10, 2)
```

Take an initial guess at the mean and covariance by drawing from their priors.

```{r initial}
mu_star <- random_mvnorm(1, mu_0, Sigma_0)[1,]
mu_star
Sigma_star <- rWishart(1, v0, S0)[,,1]
Sigma_star
```

Impute values based on this initial guess.

```{r impute}
Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)

mylegend <- function() {
  legend(
    "topleft",
    legend = c("True missing", "Present", "Imputed"),
    col = c("grey", "black", "red"),
    pch = 19,
	bg = "white"
  )
}
plot(Y_all, pch = 19, col = "grey")
points(Ystar, pch = 19, col = "red")
points(Y, pch = 19, col = "black")
mylegend()
```

Clearly, these initial values are a bad fit to the data.
Nevertheless, draw $\mu$ and $\Sigma$ conditioned on this set of imputed values.

```{r draw}
mu_star <- draw_mu(Ystar, Sigma_star, mu_0, Sigma_0)
mu_star
Sigma_star <- draw_Sigma(Ystar, mu_star, v0, S0)
Sigma_star
```

Draw a new set of imputed values, conditioned on the current $\mu^*$ and $\Sigma^*$.

```{r impute_2}
Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)
plot(Y_all, pch = 19, col = "grey")
points(Ystar, pch = 19, col = "red")
points(Y, pch = 19, col = "black")
mylegend()
```

This is still a bad fit to the data, but, though it is hard to tell, it has improved slightly.
Doing this repeatedly (in a loop) gradually improves the estimates of the covariance.

```{r loop}
par(mfrow = c(5, 4), mar = c(3, 3, 2, 0.1), cex = 0.4)
for (i in 1:20) {
  mu_star <- draw_mu(Ystar, Sigma_star, mu_0, Sigma_0)[1,]
  Sigma_star <- draw_Sigma(Ystar, mu_star, v0, S0)
  Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)
  title <- sprintf("iteration: %d, covariance: %.2f", i, Sigma_star[2, 1])
  plot(Y_all, pch = 19, col = "grey", xlab = "", ylab = "", main = title)
  points(Ystar, pch = 19, col = "red")
  points(Y, pch = 19, col = "black")
}
```

After a lot of MCMC samples, we can generate a distribution of covariance values, which provides an uncertainty estimate on the covariance.

```{r cov_estimate}
cov_ab <- numeric(1000)
for (i in seq_along(cov_ab)) {
  mu_star <- draw_mu(Ystar, Sigma_star, mu_0, Sigma_0)[1,]
  Sigma_star <- draw_Sigma(Ystar, mu_star, v0, S0)
  Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)
  cov_ab[i] <- Sigma_star[2, 1]
}
hist(cov_ab, xlab = "Covariance estimate", ylab = "Count", main = "Covariance estimate")
abline(v = true_cov[2, 1], col = "red", lwd = 2)
legend("topright", legend = "True cov.", lty = "solid", lwd = 2, col = "red")
```

### 2018-09-07 second-reviewer-response/response_summary.txt
Previously unstaged changes from second revision

Dear Dr. David Ackerly,

Thank you for your detailed and thoughtful review of our manuscript. We have considered all of the comments provided by you and the two reviewers, and are pleased to present a revised version of our manuscript in response. I have included a point-by-point response to all comments in a separate document, but I will summarize the two largest changes here.

The primary concern expressed by you and reviewer 1 was the imputation of trait data in our analysis. We believe that these concerns stem primarily from a misunderstanding of our methods. Namely, unlike single imputation, where the imputation is a separate step in the analysis performed before the main parameter estimation, our manuscript uses multiple imputation, where imputation and parameter estimation happen simultaneously throughout the sampling procedure. We have clarified this distinction in the Methods, and included an extended demonstration of this technique in the Supporting Information (Method S1). In addition, we have provided data tables indicating the sample sizes and extent of imputation throughout our analysis as requested, but we did not perform the jackknifing analysis you requested because we did not feel it made sense in the context of multiple imputation as implemented in our manuscript.

An additional concern you expressed was the confounding effect of sampling---particularly, sample size and data range---on our correlation estimates. We agree that we did not give this sufficient attention in our initial draft, and we performed all of the analyses you suggested in this revision. These analyses show that although our results are not entirely independent of sampling artifacts, there is substantial variability in trait correlations that cannot be explained by sample size and/or data range, indicating that our results are in fact biologically relevant. We present these analyses in full in the supplementary information (Tables S3 and S4, Figures S1 and S2) and discuss them in the revised "Methods" and "Results" sections.

We hope that this revision is a step in the right direction in moving this manuscript towards publication at New Phytologist.

Best,
Alexey Shiklomanov

### 2018-09-07 manuscript/02_methods.Rmd
Previously unstaged changes from second revision

@@
 This approach is therefore distinct from single imputation, where data are imputed first in a separate step prior to parameter estimation.
-A demonstration of this multiple imputation approach and how it is used to estimate trait covariance is provided in a literate programming vignette in the supplementary information to this manuscript.
+A demonstration of this multiple imputation approach and how it is used to estimate trait covariance is provided Supporting Information Method S1.
 
@@
 To assess the impact of multivariate and hierarchical constraint on trait estimates,
-we compared the mean and 95% confidence intervals of trait estimates for each PFT from each model (Fig.\ 2).
+we compared the mean and 95% confidence intervals of trait estimates for each PFT from each model (Fig.\ 2, Tab.\ S1 and S2).
 For reference, we also added the default parameter values of CLM 4.5 [Table 8.1 in @clm45_note] for SLA, $N_{mass}$, $N_{area}$, $V_{c,max,mass}$, and $V_{c,max,area}$ to Fig.\ 2.

### 2018-09-07 manuscript/05_figures_tables.Rmd
Revise formatting for submission
Previously unstaged changes from second revision

--- 2018-02-13 manuscript/06_figures_tables.Rmd
@@
-# Tables {-}
+<div custom-style="Heading, no number">Tables</div>
 
@@
   message = FALSE,
-  warning = FALSE,
-  error = FALSE
+  warning = FALSE, error = FALSE
 )
@@
 ```
-
-# Figure captions {-}
-
-**Figure 1**: Sample sizes for each trait-PFT pair. $y$ axis is scaled logarithmically.
-
-**Figure 2**: Mean and 95% confidence interval on best estimates of traits for each plant functional type from the univariate, multivariate, and hierarchical models.
-For leaf lifespan and SLA, results were not significantly different between the mass- and area-based models, so only results from the mass-based model are shown.
-For some PFT-trait combinations, where large error bars resulting from the relatively uninformative priors are substantially larger than the variability among means, the $y$ axes are constrained to facilitate comparison.
-
-**Figure 3**: Relative uncertainty in PFT-level trait estimates as a function of sample size for each model type.
-Lines represent linear models ($\log(y) = b_0 + b_1 \log(x)$) fit independently for each model type.
-In general, differences in estimate uncertainty between the univariate and multivariate models were minimal at large sample sizes but increasingly important at low sample sizes.
-However, differences in estimate uncertainty between the multivariate and hierarchical models were consistently negligible.
-
-**Figure 4**: Pairwise trait mean and covariance estimates for all data pooled globally (black) and for each PFT (colored).
-Covariance estimates not significantly different from zero ($p < 0.05$) are indicated by x symbols at the mean estimate.
-$x$ and $y$ axes vary on a log scale, reflecting the fact that the model was fit using the base 10 log of all traits.
-With the exception of leaf lifespan, pairwise covariances are consistent in direction but vary somewhat in magnitude between PFTs, and when comparing PFT-level and global estimates.
-However, many pairwise covariances are not statistically significant, particularly (but not always) for undersampled traits and PFTs.
-
-**Figure 5**: Mean and 95% CI on estimates of pairwise correlation coefficients
-for all data pooled globally (dark grey) and for each PFT (colored).
-For most PFT-trait pairs, correlations are mutually consistent in magnitude but vary in strength.
-
-# References {-}

### 2018-09-07 manuscript/index.Rmd
Previously unstaged changes from second revision

@@
 --------                  --------------      ------------
-Introduction              1060                Figures:
-Materials and Methods     1695                5 (all colored)
-Results                   1108
-Discussion                1907                Tables:
-Acknowledgements          88                  2
+Introduction              1060                Figures: 5 (all colored)
+Materials and Methods     1695                Tables: 2
+Results                   1108                Supplementary figures: 2
+Discussion                1907                Supplementary tables: 4
+Acknowledgements          88                  Supplementary methods: 1
 Total                     5858

### 2018-09-07 second-reviewer-response/multiple_imputation.Rmd
Previously unstaged changes from second revision

@@
 ---
-
-# Introduction
-
-Under _single_ imputation, the imputation step occurs once outside of the MCMC loop.
-In _multiple_ imputation, as implemented in this paper, the imputation step is part of the MCMC loop,
-such that imputed data values are conditioned on the current state of the parameters and vice-versa.
-In the figure below,
-$Y$ is the original data (with missing values),
-$Y^*$ is the original data with missing values imputed,
-$\mu_p$ and $\Sigma_p$ are values of the mean vector and covariance matrix (respectively) calculated only from the original data,
-and $\mu^*_t$ and $\Sigma^*_t$ are the draws of the mean vector and covariance matrix (respectively) at MCMC iteration $t$.
-
-```{tikz diagram, echo = FALSE}
-\usetikzlibrary{positioning, calc, fit}
-\begin{tikzpicture}[
-	every node/.style = {shape = rectangle, draw = black},
-	scale = 1.1
-]
-\node (Y) at (-1, 3) {Data $Y$};
-\node (data params) at (-1, 2) {Calculate $\mu_p, \Sigma_p$};
-\node (Ystar) at (-1, 1) {Impute $Y^* \mid \mu_p, \Sigma_p$};
-\node (guess params) [align = center, right = of Ystar] {Initial guess\\$\mu^*_0$ $\Sigma^*_0$};
-\node (mu star) at (0, 0) {Draw $\mu^*_t \mid Y^*, \Sigma^*_{t-1}$};
-\node (sigma star) at (0, -1) {Draw $\Sigma^*_t \mid Y^*, \mu^*_t$};
-\path [->] (Y) edge (data params);
-\path [->] (data params) edge (Ystar);
-\path [->] (Ystar) edge (mu star);
-\path [->] (guess params) edge (mu star);
-\path [->] (mu star) edge (sigma star);
-\path [->, out = 0, in = 0] (sigma star.east) edge
-	node (t plus one) [right, draw = none, font = {\footnotesize}] {$t = t + 1$}
-	(mu star.east);
-\node (mcmc loop) [fit = (mu star)(sigma star)(t plus one), shape = rectangle, draw = black!50] {};
-\node (mcmc loop text) [align = center, anchor = west, draw = none] at (mcmc loop.east) {MCMC\\loop};
-\node (title) [anchor = north, draw = none] at (mcmc loop.south) {Single imputation};
-\end{tikzpicture}
-\hspace{1cm}
-\begin{tikzpicture}[
-	every node/.style = {shape = rectangle, draw = black},
-	scale = 1.1
-]
-\node (Y) at (-1, 1) {Data $Y$};
-\node (guess params) [align = center] at (1, 1) {Initial guess\\$\mu^*_0$ $\Sigma^*_0$};
-\node (Ystar) at (0, 0) {Impute $Y^*_t \mid Y, \mu^*_{t-1}, \Sigma^*_{t-1}$};
-\node (mu star) at (0, -1) {Draw $\mu^*_t \mid Y^*_t, \Sigma^*_{t-1}$};
-\node (sigma star) at (0, -2) {Draw $\Sigma^*_t \mid Y^*_t, \mu^*_t$};
-\path [->] (Y) edge (Ystar);
-\path [->] (guess params) edge (Ystar);
-\path [->] (Ystar) edge (mu star);
-\path [->] (mu star) edge (sigma star);
-\path [->, out = 0, in = 0] (sigma star.east) edge
-	node (t plus one) [right, draw = none, font = {\footnotesize}] {$t = t + 1$}
-	(Ystar.east);
-\node (mcmc loop) [fit = (Ystar)(mu star)(sigma star)(t plus one), shape = rectangle, draw = black!50] {};
-\node (mcmc loop text) [align = center, anchor = west, draw = none] at (mcmc loop.east) {MCMC\\loop};
-\node (title) [anchor = north, draw = none] at (mcmc loop.south) {Multiple imputation};
-\end{tikzpicture}
-```
-
-# Demonstration
-
-Consider two positively correlated traits, A and B.
-
-```{r create_traits}
-set.seed(12345) # For reproducibility
-library(mvtraits)
-
-true_mu <- c(0, 0)
-true_cov <- matrix(c(1, 0.8, 0.8, 1), nrow = 2)
-Y_all <- random_mvnorm(1000, true_mu, true_cov)
-colnames(Y_all) <- c("A", "B")
-plot(Y_all, pch = 19)
-```
-
-Simulate missingness by randomly removing half of the data.
-
-```{r remove_na}
-Y <- Y_all
-Y[sample.int(2000, 1000)] <- NA
-plot(Y_all, pch = 19, col = "grey")
-points(Y, pch = 19, col = "black")
-legend(
-  "topleft",
-  legend = c("True missing", "Present"),
-  col = c("grey", "black"),
-  pch = 19,
-  bg = "white"
-)
-```
-
-Set an uninformative multivariate normal prior on $\mu$...
-```{r prior_mu}
-mu_0 <- c(0, 0)
-Sigma_0 <- diag(10, 2)
-```
-
-...and an uninformative Wishart prior on $\Sigma$.
-```{r prior_sigma}
-v0 <- 2
-S0 <- diag(10, 2)
-```
-
-Take an initial guess at the mean and covariance by drawing from their priors.
-
-```{r initial}
-mu_star <- random_mvnorm(1, mu_0, Sigma_0)[1,]
-mu_star
-Sigma_star <- rWishart(1, v0, S0)[,,1]
-Sigma_star
-```
-
-Impute values based on this initial guess.
-
-```{r impute}
-Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)
-
-mylegend <- function() {
-  legend(
-    "topleft",
-    legend = c("True missing", "Present", "Imputed"),
-    col = c("grey", "black", "red"),
-    pch = 19,
-	bg = "white"
-  )
-}
-plot(Y_all, pch = 19, col = "grey")
-points(Ystar, pch = 19, col = "red")
-points(Y, pch = 19, col = "black")
-mylegend()
-```
-
-Clearly, these initial values are a bad fit to the data.
-Nevertheless, draw $\mu$ and $\Sigma$ conditioned on this set of imputed values.
-
-```{r draw}
-mu_star <- draw_mu(Ystar, Sigma_star, mu_0, Sigma_0)
-mu_star
-Sigma_star <- draw_Sigma(Ystar, mu_star, v0, S0)
-Sigma_star
-```
-
-Draw a new set of imputed values, conditioned on the current $\mu^*$ and $\Sigma^*$.
-
-```{r impute_2}
-Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)
-plot(Y_all, pch = 19, col = "grey")
-points(Ystar, pch = 19, col = "red")
-points(Y, pch = 19, col = "black")
-mylegend()
-```
-
-This is still a bad fit to the data, but, though it is hard to tell, it has improved slightly.
-Doing this repeatedly (in a loop) gradually improves the estimates of the covariance.
-
-```{r loop}
-par(mfrow = c(5, 4), mar = c(3, 3, 2, 0.1), cex = 0.4)
-for (i in 1:20) {
-  mu_star <- draw_mu(Ystar, Sigma_star, mu_0, Sigma_0)[1,]
-  Sigma_star <- draw_Sigma(Ystar, mu_star, v0, S0)
-  Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)
-  title <- sprintf("iteration: %d, covariance: %.2f", i, Sigma_star[2, 1])
-  plot(Y_all, pch = 19, col = "grey", xlab = "", ylab = "", main = title)
-  points(Ystar, pch = 19, col = "red")
-  points(Y, pch = 19, col = "black")
-}
-```
-
-After a lot of MCMC samples, we can generate a distribution of covariance values, which provides an uncertainty estimate on the covariance.
-
-```{r cov_estimate}
-cov_ab <- numeric(1000)
-for (i in seq_along(cov_ab)) {
-  mu_star <- draw_mu(Ystar, Sigma_star, mu_0, Sigma_0)[1,]
-  Sigma_star <- draw_Sigma(Ystar, mu_star, v0, S0)
-  Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)
-  cov_ab[i] <- Sigma_star[2, 1]
-}
-hist(cov_ab, xlab = "Covariance estimate", ylab = "Count", main = "Covariance estimate")
-abline(v = true_cov[2, 1], col = "red", lwd = 2)
-legend("topright", legend = "True cov.", lty = "solid", lwd = 2, col = "red")
-```

### 2018-09-07 second-reviewer-response/multiple_imputation.tex
Previously unstaged changes from second revision

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Estimating trait means and covariances through multiple imputation},
            pdfauthor={Alexey N. Shiklomanov, Elizabeth M. Cowdery, Michael Bahn, Chaeho Byun, Joseph Craine, Andrés Gonzalez-Melo, Steven Jansen, Nathan Kraft, Koen Kramer, Vanessa Minden, Ülo Niinemets, Yusuke Onoda, Enio Egon Sosinski, Nadejda A. Soudzilovskaia, Michael C. Dietze},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Estimating trait means and covariances through multiple imputation}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Alexey N. Shiklomanov, Elizabeth M. Cowdery, Michael Bahn, Chaeho Byun,
Joseph Craine, Andrés Gonzalez-Melo, Steven Jansen, Nathan Kraft, Koen
Kramer, Vanessa Minden, Ülo Niinemets, Yusuke Onoda, Enio Egon Sosinski,
Nadejda A. Soudzilovskaia, Michael C. Dietze}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
    \date{}
    \predate{}\postdate{}

\begin{document}
\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Under \emph{single} imputation, the imputation step occurs once outside
of the MCMC loop. In \emph{multiple} imputation, as implemented in this
paper, the imputation step is part of the MCMC loop, such that imputed
data values are conditioned on the current state of the parameters and
vice-versa. In the figure below, \(Y\) is the original data (with
missing values), \(Y^*\) is the original data with missing values
imputed, \(\mu_p\) and \(\Sigma_p\) are values of the mean vector and
covariance matrix (respectively) calculated only from the original data,
and \(\mu^*_t\) and \(\Sigma^*_t\) are the draws of the mean vector and
covariance matrix (respectively) at MCMC iteration \(t\).

\includegraphics{multiple_imputation_files/figure-latex/diagram-1.pdf}

\hypertarget{demonstration}{%
\section{Demonstration}\label{demonstration}}

Consider two positively correlated traits, A and B.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{12345}\NormalTok{) }\CommentTok{# For reproducibility}
\KeywordTok{library}\NormalTok{(mvtraits)}

\NormalTok{true_mu <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{)}
\NormalTok{true_cov <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\FloatTok{0.8}\NormalTok{, }\FloatTok{0.8}\NormalTok{, }\DecValTok{1}\NormalTok{), }\DataTypeTok{nrow =} \DecValTok{2}\NormalTok{)}
\NormalTok{Y_all <-}\StringTok{ }\KeywordTok{random_mvnorm}\NormalTok{(}\DecValTok{1000}\NormalTok{, true_mu, true_cov)}
\KeywordTok{colnames}\NormalTok{(Y_all) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"A"}\NormalTok{, }\StringTok{"B"}\NormalTok{)}
\KeywordTok{plot}\NormalTok{(Y_all, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{multiple_imputation_files/figure-latex/create_traits-1.pdf}

Simulate missingness by randomly removing half of the data.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Y <-}\StringTok{ }\NormalTok{Y_all}
\NormalTok{Y[}\KeywordTok{sample.int}\NormalTok{(}\DecValTok{2000}\NormalTok{, }\DecValTok{1000}\NormalTok{)] <-}\StringTok{ }\OtherTok{NA}
\KeywordTok{plot}\NormalTok{(Y_all, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \StringTok{"grey"}\NormalTok{)}
\KeywordTok{points}\NormalTok{(Y, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \StringTok{"black"}\NormalTok{)}
\KeywordTok{legend}\NormalTok{(}
  \StringTok{"topleft"}\NormalTok{,}
  \DataTypeTok{legend =} \KeywordTok{c}\NormalTok{(}\StringTok{"True missing"}\NormalTok{, }\StringTok{"Present"}\NormalTok{),}
  \DataTypeTok{col =} \KeywordTok{c}\NormalTok{(}\StringTok{"grey"}\NormalTok{, }\StringTok{"black"}\NormalTok{),}
  \DataTypeTok{pch =} \DecValTok{19}\NormalTok{,}
  \DataTypeTok{bg =} \StringTok{"white"}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{multiple_imputation_files/figure-latex/remove_na-1.pdf}

Set an uninformative multivariate normal prior on \(\mu\)\ldots{}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{mu_}\DecValTok{0}\NormalTok{ <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{)}
\NormalTok{Sigma_}\DecValTok{0}\NormalTok{ <-}\StringTok{ }\KeywordTok{diag}\NormalTok{(}\DecValTok{10}\NormalTok{, }\DecValTok{2}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\ldots{}and an uninformative Wishart prior on \(\Sigma\).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{v0 <-}\StringTok{ }\DecValTok{2}
\NormalTok{S0 <-}\StringTok{ }\KeywordTok{diag}\NormalTok{(}\DecValTok{10}\NormalTok{, }\DecValTok{2}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Take an initial guess at the mean and covariance by drawing from their
priors.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{mu_star <-}\StringTok{ }\KeywordTok{random_mvnorm}\NormalTok{(}\DecValTok{1}\NormalTok{, mu_}\DecValTok{0}\NormalTok{, Sigma_}\DecValTok{0}\NormalTok{)[}\DecValTok{1}\NormalTok{,]}
\NormalTok{mu_star}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] -7.3409439  0.7029685
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Sigma_star <-}\StringTok{ }\KeywordTok{rWishart}\NormalTok{(}\DecValTok{1}\NormalTok{, v0, S0)[,,}\DecValTok{1}\NormalTok{]}
\NormalTok{Sigma_star}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##          [,1]      [,2]
## [1,] 81.17255 15.526650
## [2,] 15.52665  4.779821
\end{verbatim}

Impute values based on this initial guess.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Ystar <-}\StringTok{ }\KeywordTok{mvnorm_fill_missing}\NormalTok{(Y, mu_star, Sigma_star)}

\NormalTok{mylegend <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{() \{}
  \KeywordTok{legend}\NormalTok{(}
    \StringTok{"topleft"}\NormalTok{,}
    \DataTypeTok{legend =} \KeywordTok{c}\NormalTok{(}\StringTok{"True missing"}\NormalTok{, }\StringTok{"Present"}\NormalTok{, }\StringTok{"Imputed"}\NormalTok{),}
    \DataTypeTok{col =} \KeywordTok{c}\NormalTok{(}\StringTok{"grey"}\NormalTok{, }\StringTok{"black"}\NormalTok{, }\StringTok{"red"}\NormalTok{),}
    \DataTypeTok{pch =} \DecValTok{19}\NormalTok{,}
    \DataTypeTok{bg =} \StringTok{"white"}
\NormalTok{  )}
\NormalTok{\}}
\KeywordTok{plot}\NormalTok{(Y_all, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \StringTok{"grey"}\NormalTok{)}
\KeywordTok{points}\NormalTok{(Ystar, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \StringTok{"red"}\NormalTok{)}
\KeywordTok{points}\NormalTok{(Y, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \StringTok{"black"}\NormalTok{)}
\KeywordTok{mylegend}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{multiple_imputation_files/figure-latex/impute-1.pdf}

Clearly, these initial values are a bad fit to the data. Nevertheless,
draw \(\mu\) and \(\Sigma\) conditioned on this set of imputed values.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{mu_star <-}\StringTok{ }\KeywordTok{draw_mu}\NormalTok{(Ystar, Sigma_star, mu_}\DecValTok{0}\NormalTok{, Sigma_}\DecValTok{0}\NormalTok{)}
\NormalTok{mu_star}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##           [,1]    [,2]
## [1,] -2.319796 1.05317
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Sigma_star <-}\StringTok{ }\KeywordTok{draw_Sigma}\NormalTok{(Ystar, mu_star, v0, S0)}
\NormalTok{Sigma_star}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##           [,1]     [,2]
## [1,] 48.234522 6.449814
## [2,]  6.449814 2.684697
\end{verbatim}

Draw a new set of imputed values, conditioned on the current \(\mu^*\)
and \(\Sigma^*\).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Ystar <-}\StringTok{ }\KeywordTok{mvnorm_fill_missing}\NormalTok{(Y, mu_star, Sigma_star)}
\KeywordTok{plot}\NormalTok{(Y_all, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \StringTok{"grey"}\NormalTok{)}
\KeywordTok{points}\NormalTok{(Ystar, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \StringTok{"red"}\NormalTok{)}
\KeywordTok{points}\NormalTok{(Y, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \StringTok{"black"}\NormalTok{)}
\KeywordTok{mylegend}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{multiple_imputation_files/figure-latex/impute_2-1.pdf}

This is still a bad fit to the data, but, though it is hard to tell, it
has improved slightly. Doing this repeatedly (in a loop) gradually
improves the estimates of the covariance.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =} \KeywordTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{4}\NormalTok{), }\DataTypeTok{mar =} \KeywordTok{c}\NormalTok{(}\DecValTok{3}\NormalTok{, }\DecValTok{3}\NormalTok{, }\DecValTok{2}\NormalTok{, }\FloatTok{0.1}\NormalTok{), }\DataTypeTok{cex =} \FloatTok{0.4}\NormalTok{)}
\ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\DecValTok{20}\NormalTok{) \{}
\NormalTok{  mu_star <-}\StringTok{ }\KeywordTok{draw_mu}\NormalTok{(Ystar, Sigma_star, mu_}\DecValTok{0}\NormalTok{, Sigma_}\DecValTok{0}\NormalTok{)[}\DecValTok{1}\NormalTok{,]}
\NormalTok{  Sigma_star <-}\StringTok{ }\KeywordTok{draw_Sigma}\NormalTok{(Ystar, mu_star, v0, S0)}
\NormalTok{  Ystar <-}\StringTok{ }\KeywordTok{mvnorm_fill_missing}\NormalTok{(Y, mu_star, Sigma_star)}
\NormalTok{  title <-}\StringTok{ }\KeywordTok{sprintf}\NormalTok{(}\StringTok{"iteration: %d, covariance: %.2f"}\NormalTok{, i, Sigma_star[}\DecValTok{2}\NormalTok{, }\DecValTok{1}\NormalTok{])}
  \KeywordTok{plot}\NormalTok{(Y_all, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \StringTok{"grey"}\NormalTok{, }\DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{main =}\NormalTok{ title)}
  \KeywordTok{points}\NormalTok{(Ystar, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \StringTok{"red"}\NormalTok{)}
  \KeywordTok{points}\NormalTok{(Y, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \StringTok{"black"}\NormalTok{)}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\includegraphics{multiple_imputation_files/figure-latex/loop-1.pdf}

After a lot of MCMC samples, we can generate a distribution of
covariance values, which provides an uncertainty estimate on the
covariance.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{cov_ab <-}\StringTok{ }\KeywordTok{numeric}\NormalTok{(}\DecValTok{1000}\NormalTok{)}
\ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \KeywordTok{seq_along}\NormalTok{(cov_ab)) \{}
\NormalTok{  mu_star <-}\StringTok{ }\KeywordTok{draw_mu}\NormalTok{(Ystar, Sigma_star, mu_}\DecValTok{0}\NormalTok{, Sigma_}\DecValTok{0}\NormalTok{)[}\DecValTok{1}\NormalTok{,]}
\NormalTok{  Sigma_star <-}\StringTok{ }\KeywordTok{draw_Sigma}\NormalTok{(Ystar, mu_star, v0, S0)}
\NormalTok{  Ystar <-}\StringTok{ }\KeywordTok{mvnorm_fill_missing}\NormalTok{(Y, mu_star, Sigma_star)}
\NormalTok{  cov_ab[i] <-}\StringTok{ }\NormalTok{Sigma_star[}\DecValTok{2}\NormalTok{, }\DecValTok{1}\NormalTok{]}
\NormalTok{\}}
\KeywordTok{hist}\NormalTok{(cov_ab, }\DataTypeTok{xlab =} \StringTok{"Covariance estimate"}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{"Count"}\NormalTok{, }\DataTypeTok{main =} \StringTok{"Covariance estimate"}\NormalTok{)}
\KeywordTok{abline}\NormalTok{(}\DataTypeTok{v =}\NormalTok{ true_cov[}\DecValTok{2}\NormalTok{, }\DecValTok{1}\NormalTok{], }\DataTypeTok{col =} \StringTok{"red"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{2}\NormalTok{)}
\KeywordTok{legend}\NormalTok{(}\StringTok{"topright"}\NormalTok{, }\DataTypeTok{legend =} \StringTok{"True cov."}\NormalTok{, }\DataTypeTok{lty =} \StringTok{"solid"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{2}\NormalTok{, }\DataTypeTok{col =} \StringTok{"red"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{multiple_imputation_files/figure-latex/cov_estimate-1.pdf}

\end{document}

### 2018-09-07 supplementary_information.Rmd
Previously unstaged changes from second revision

@@
 ---
-title: Supplementary information
+title: Supporting information
 author:
@@
 ```
+
+\clearpage
+
+# Method S1: Estimating trait means and covariances through multiple imputation {-}
+
+## Introduction {-}
+
+Under _single_ imputation, the imputation step occurs once outside of the MCMC loop.
+In _multiple_ imputation, as implemented in this paper, the imputation step is part of the MCMC loop,
+such that imputed data values are conditioned on the current state of the parameters and vice-versa.
+In the figure below,
+$Y$ is the original data (with missing values),
+$Y^*$ is the original data with missing values imputed,
+$\mu_p$ and $\Sigma_p$ are values of the mean vector and covariance matrix (respectively) calculated only from the original data,
+and $\mu^*_t$ and $\Sigma^*_t$ are the draws of the mean vector and covariance matrix (respectively) at MCMC iteration $t$.
+
+```{tikz diagram, echo = FALSE}
+\usetikzlibrary{positioning, calc, fit}
+\begin{tikzpicture}[
+	every node/.style = {shape = rectangle, draw = black},
+	scale = 1.1
+]
+\node (Y) at (-1, 3) {Data $Y$};
+\node (data params) at (-1, 2) {Calculate $\mu_p, \Sigma_p$};
+\node (Ystar) at (-1, 1) {Impute $Y^* \mid \mu_p, \Sigma_p$};
+\node (guess params) [align = center, right = of Ystar] {Initial guess\\$\mu^*_0$ $\Sigma^*_0$};
+\node (mu star) at (0, 0) {Draw $\mu^*_t \mid Y^*, \Sigma^*_{t-1}$};
+\node (sigma star) at (0, -1) {Draw $\Sigma^*_t \mid Y^*, \mu^*_t$};
+\path [->] (Y) edge (data params);
+\path [->] (data params) edge (Ystar);
+\path [->] (Ystar) edge (mu star);
+\path [->] (guess params) edge (mu star);
+\path [->] (mu star) edge (sigma star);
+\path [->, out = 0, in = 0] (sigma star.east) edge
+	node (t plus one) [right, draw = none, font = {\footnotesize}] {$t = t + 1$}
+	(mu star.east);
+\node (mcmc loop) [fit = (mu star)(sigma star)(t plus one), shape = rectangle, draw = black!50] {};
+\node (mcmc loop text) [align = center, anchor = west, draw = none] at (mcmc loop.east) {MCMC\\loop};
+\node (title) [anchor = north, draw = none] at (mcmc loop.south) {Single imputation};
+\end{tikzpicture}
+\hspace{1cm}
+\begin{tikzpicture}[
+	every node/.style = {shape = rectangle, draw = black},
+	scale = 1.1
+]
+\node (Y) at (-1, 1) {Data $Y$};
+\node (guess params) [align = center] at (1, 1) {Initial guess\\$\mu^*_0$ $\Sigma^*_0$};
+\node (Ystar) at (0, 0) {Impute $Y^*_t \mid Y, \mu^*_{t-1}, \Sigma^*_{t-1}$};
+\node (mu star) at (0, -1) {Draw $\mu^*_t \mid Y^*_t, \Sigma^*_{t-1}$};
+\node (sigma star) at (0, -2) {Draw $\Sigma^*_t \mid Y^*_t, \mu^*_t$};
+\path [->] (Y) edge (Ystar);
+\path [->] (guess params) edge (Ystar);
+\path [->] (Ystar) edge (mu star);
+\path [->] (mu star) edge (sigma star);
+\path [->, out = 0, in = 0] (sigma star.east) edge
+	node (t plus one) [right, draw = none, font = {\footnotesize}] {$t = t + 1$}
+	(Ystar.east);
+\node (mcmc loop) [fit = (Ystar)(mu star)(sigma star)(t plus one), shape = rectangle, draw = black!50] {};
+\node (mcmc loop text) [align = center, anchor = west, draw = none] at (mcmc loop.east) {MCMC\\loop};
+\node (title) [anchor = north, draw = none] at (mcmc loop.south) {Multiple imputation};
+\end{tikzpicture}
+```
+
+## Demonstration {-}
+
+Consider two positively correlated traits, A and B.
+
+```{r create_traits}
+set.seed(12345) # For reproducibility
+library(mvtraits)
+
+true_mu <- c(0, 0)
+true_cov <- matrix(c(1, 0.8, 0.8, 1), nrow = 2)
+Y_all <- random_mvnorm(1000, true_mu, true_cov)
+colnames(Y_all) <- c("A", "B")
+plot(Y_all, pch = 19)
+```
+
+Simulate missingness by randomly removing half of the data.
+
+```{r remove_na}
+Y <- Y_all
+Y[sample.int(2000, 1000)] <- NA
+plot(Y_all, pch = 19, col = "grey")
+points(Y, pch = 19, col = "black")
+legend(
+  "topleft",
+  legend = c("True missing", "Present"),
+  col = c("grey", "black"),
+  pch = 19,
+  bg = "white"
+)
+```
+
+Set an uninformative multivariate normal prior on $\mu$...
+```{r prior_mu}
+mu_0 <- c(0, 0)
+Sigma_0 <- diag(10, 2)
+```
+
+...and an uninformative Wishart prior on $\Sigma$.
+```{r prior_sigma}
+v0 <- 2
+S0 <- diag(10, 2)
+```
+
+Take an initial guess at the mean and covariance by drawing from their priors.
+
+```{r initial}
+mu_star <- random_mvnorm(1, mu_0, Sigma_0)[1,]
+mu_star
+Sigma_star <- rWishart(1, v0, S0)[,,1]
+Sigma_star
+```
+
+Impute values based on this initial guess.
+
+```{r impute}
+Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)
+
+mylegend <- function() {
+  legend(
+    "topleft",
+    legend = c("True missing", "Present", "Imputed"),
+    col = c("grey", "black", "red"),
+    pch = 19,
+	bg = "white"
+  )
+}
+plot(Y_all, pch = 19, col = "grey")
+points(Ystar, pch = 19, col = "red")
+points(Y, pch = 19, col = "black")
+mylegend()
+```
+
+Clearly, these initial values are a bad fit to the data.
+Nevertheless, draw $\mu$ and $\Sigma$ conditioned on this set of imputed values.
+
+```{r draw}
+mu_star <- draw_mu(Ystar, Sigma_star, mu_0, Sigma_0)
+mu_star
+Sigma_star <- draw_Sigma(Ystar, mu_star, v0, S0)
+Sigma_star
+```
+
+Draw a new set of imputed values, conditioned on the current $\mu^*$ and $\Sigma^*$.
+
+```{r impute_2}
+Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)
+plot(Y_all, pch = 19, col = "grey")
+points(Ystar, pch = 19, col = "red")
+points(Y, pch = 19, col = "black")
+mylegend()
+```
+
+This is still a bad fit to the data, but, though it is hard to tell, it has improved slightly.
+Doing this repeatedly (in a loop) gradually improves the estimates of the covariance.
+
+```{r loop}
+par(mfrow = c(5, 4), mar = c(3, 3, 2, 0.1), cex = 0.4)
+for (i in 1:20) {
+  mu_star <- draw_mu(Ystar, Sigma_star, mu_0, Sigma_0)[1,]
+  Sigma_star <- draw_Sigma(Ystar, mu_star, v0, S0)
+  Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)
+  title <- sprintf("iteration: %d, covariance: %.2f", i, Sigma_star[2, 1])
+  plot(Y_all, pch = 19, col = "grey", xlab = "", ylab = "", main = title)
+  points(Ystar, pch = 19, col = "red")
+  points(Y, pch = 19, col = "black")
+}
+```
+
+After a lot of MCMC samples, we can generate a distribution of covariance values, which provides an uncertainty estimate on the covariance.
+
+```{r cov_estimate}
+cov_ab <- numeric(1000)
+for (i in seq_along(cov_ab)) {
+  mu_star <- draw_mu(Ystar, Sigma_star, mu_0, Sigma_0)[1,]
+  Sigma_star <- draw_Sigma(Ystar, mu_star, v0, S0)
+  Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)
+  cov_ab[i] <- Sigma_star[2, 1]
+}
+hist(cov_ab, xlab = "Covariance estimate", ylab = "Count", main = "Covariance estimate")
+abline(v = true_cov[2, 1], col = "red", lwd = 2)
+legend("topright", legend = "True cov.", lty = "solid", lwd = 2, col = "red")
+```

### 2019-08-19 supplementary_information.Rmd
Fix SI authorship

@@
     Steven Jansen,
-    Nathan Kraft,
     Koen Kramer,

### 2019-08-19 scripts/drake.R
Minor drake fixes

@@ -1,11 +1,11 @@
 library(shiklomanov2017np)
-library(drake)
+library(drake, exclude = c("expand", "gather"))
 library(knitr)
-library(kableExtra)
+library(kableExtra, exclude = c("group_rows"))
 
 import::from(here, inhere = here)
 
-pkgconfig::set_config("drake::strings_in_dots" = "literals")
+## pkgconfig::set_config("drake::strings_in_dots" = "literals")
 
 get_sample_size_long <- function(try_dat) {
   count_missing <- . %>%
@@ -42,7 +42,7 @@ base_plan <- drake_plan(
     dplyr::mutate(display = sprintf("%d (%d)", present, missing)) %>%
     dplyr::select(pft, trait, display) %>%
     tidyr::spread(trait, display),
-  rmarkdown::render(
+  paper_pdf = rmarkdown::render(
     knitr_in("supplementary_information.Rmd"),
     output_file = file_out("supplementary_information.pdf")
   )
@@ -175,11 +175,6 @@ corr_grid <- function(data) {
     ## scale_size_manual(values = c("yes" = 3, "no" = 1)) +
     scale_color_manual(name = "PFT", values = c("black", pft_colors)) +
     theme_bw()
-}
-
-if (FALSE) {
-
-
 }
 
 text_plot <- function(text) {

### 2019-08-19 scripts/run_model.R
Run models sequentially...
...because in parallel currently doesn't work.

@@ -82,7 +82,7 @@ message('Data contains ', nrow(data_mat), ' rows and ', ncol(data_mat), ' column
 
 niter <- 2500
 nchains <- 4
-parallel <- TRUE
+parallel <- FALSE
 autofit <- TRUE
 max_attempts <- 50
 keep_samples <- 20000

### 2019-08-19 scripts/process_references.R
Fix path in process_references.R

@@ -5,7 +5,7 @@ library(stringi)
 try_refs <- src_sqlite("../preprocess-try/try.sqlite") %>%
   tbl("orig_citations") %>%
   collect() %>%
-  right_join(read_csv("../preprocess-try/traits/references.csv"))
+  right_join(read_csv("../preprocess-try/processed/traits/references.csv"))
 
 try_refs_proc <- try_refs %>%
   mutate(
@@ -15,12 +15,14 @@ try_refs_proc <- try_refs %>%
   filter(!grepl("unpub.|Unpublished literature compilation",
                 reference_proc, ignore.case = TRUE))
 
+## write_csv(try_refs_proc, "~/Projects/phd_work/multivariate_traits/try_references.csv")
+
 crossref_results <- list()
 for (i in seq_len(nrow(try_refs_proc))) {
   message(i, " of ", nrow(try_refs_proc))
   qry <- try_refs_proc$reference_proc[i]
   result <- cr_works(query = try_refs_proc$reference_proc[i], limit = 1)
-  print(result$data$DOI)
+  print(result$data$doi)
   crossref_results[[i]] <- result
 }
 

### 2019-08-20 scripts/drake.R
Clean up SI drake file

@@ -1,11 +1,10 @@
+#!/usr/bin/env Rscript
+
 library(shiklomanov2017np)
 library(drake, exclude = c("expand", "gather"))
 library(knitr)
 library(kableExtra, exclude = c("group_rows"))
-
-import::from(here, inhere = here)
-
-## pkgconfig::set_config("drake::strings_in_dots" = "literals")
+library(here)
 
 get_sample_size_long <- function(try_dat) {
   count_missing <- . %>%
@@ -31,7 +30,7 @@ get_sample_size_long <- function(try_dat) {
 }
 
 base_plan <- drake_plan(
-  try_data_all = readRDS(inhere("extdata/traits_analysis.rds")),
+  try_data_all = readRDS(here("extdata", "traits_analysis.rds")),
   try_dat = try_data_all %>%
     dplyr::select(pft = clm45, !!both_params) %>%
     dplyr::mutate(
@@ -42,20 +41,25 @@ base_plan <- drake_plan(
     dplyr::mutate(display = sprintf("%d (%d)", present, missing)) %>%
     dplyr::select(pft, trait, display) %>%
     tidyr::spread(trait, display),
-  paper_pdf = rmarkdown::render(
+  supplement_pdf = rmarkdown::render(
     knitr_in("supplementary_information.Rmd"),
     output_file = file_out("supplementary_information.pdf")
   )
 )
 
+hier_mass_file <- here("output", "hier.mass.clm45.NA.2018-02-12-1154.rds")
+hier_area_file <- here("output", "hier.area.clm45.NA.2018-02-12-1227.rds")
+multi_mass_file <- here("output", "multi.mass.clm45.NA.2018-02-12-1127.rds")
+multi_area_file <- here("output", "multi.area.clm45.NA.2018-02-12-1146.rds")
+
 corr_data <- drake_plan(
-  hier_summary_mass = readRDS(file_in("output/hier.mass.clm45.NA.2018-02-12-1154.rds"))[["summary_table"]] %>%
+  hier_summary_mass = readRDS(file_in(!!hier_mass_file))[["summary_table"]] %>%
     mutate(mass_area = "mass", modeltype = "hier"),
-  hier_summary_area = readRDS(file_in("output/hier.area.clm45.NA.2018-02-12-1227.rds"))[["summary_table"]] %>%
[…]

### 2019-08-20 supplementary_information.Rmd
Fix SLA units in SI

@@
       "Leaf lifespan\n(months)",
-      "SLA\n(kg m$^{-2}$)",
+      "SLA\n(m$^2$ kg$^{-1}$)",
       "$N_\\textrm{mass}$\n(mg g$^{-1}$)",
@@
       "Leaf lifespan\n(months)",
-      "SLA\n(kg m$^{-2}$)",
+      "SLA\n(m$^2$ kg$^{-1}$)",
       "$N_\\textrm{area}$\n(g m$^{-2}$)",

### 2019-08-23 scripts/manuscript-analyses/alt-figures.R
Draft some more manuscript analyses

@@ -1,0 +1,30 @@
+library(shiklomanov2017np)
+
+manuscript_fig_dir <- here("figures", "manuscript")
+dir.create(manuscript_fig_dir, showWarnings = FALSE)
+
+resultsfile <- here("results", "mvtraits_results.rds")
+results_all <- readRDS(resultsfile)
+
+results_hier <- results_all %>%
+  filter(model_type == "hier")
+
+rh_list <- map(results_hier[["data"]], "result")
+
+xx <- rh_list %>%
+  pluck(2, "Sigma_global", list("Mean", )) %>%
+  .["SLA", "Nmass"]
+
+x <- rh_list %>%
+  pluck(2, "Sigma_group") %>%
+  map("Mean") %>%
+  map(~.["SLA", "Nmass"]) %>%
+  bind_rows() %>%
+  gather("pft", "sigma")
+
+ggplot(x) +
+  aes(x = pft, y = sigma) +
+  geom_col() +
+  theme(axis.text.x = element_text(angle = 90))
+
+rh_list

### 2019-08-23 scripts/manuscript-analyses/correlation-power.R

@@ -1,0 +1,12 @@
+library(pwr)
+out <- pwr.r.test(14, power = 0.95, sig.level = 0.05, alternative = "greater")
+
+power_pft <- function(npft) {
+  out <- numeric(length(npft))
+  for (i in seq_along(npft)) {
+    out[i] <- pwr.r.test(npft[[i]], power = 0.95, sig.level = 0.05, alternative = "greater")$r
+  }
+  out
+}
+
+curve(power_pft(x), 10, 100)

### 2019-08-23 scripts/manuscript-analyses/covariance-boxplot.R

@@ -1,0 +1,204 @@
+library(shiklomanov2017np)
+
+outdir <- here::here("output")
+hier_mass_file <- tail(dir(outdir, "hier.mass.clm45", full.names = TRUE), 1)
+hier_area_file <- tail(dir(outdir, "hier.area.clm45", full.names = TRUE), 1)
+multi_mass_file <- tail(dir(outdir, "multi.mass.clm45.NA", full.names = TRUE), 1)
+multi_area_file <- tail(dir(outdir, "multi.area.clm45.NA", full.names = TRUE), 1)
+
+hier_mass <- readRDS(hier_mass_file)
+samples <- hier_mass$samples[[1]]
+
+rma_slope <- function(m, pft, yvar, xvar) {
+  sdy_col <- paste("Sigma", pft, yvar, yvar, sep = "..")
+  sdx_col <- paste("Sigma", pft, xvar, xvar, sep = "..")
+  mcol <- paste("Sigma", pft, xvar, yvar, sep = "..")
+  sign(m[, mcol]) * sqrt(m[, sdy_col] / m[, sdx_col])
+}
+
+all_rma_slopes <- function(samples) {
+  sigma_cols <- grep("^Sigma..", colnames(samples), value = TRUE)
+  chars <- do.call(rbind, strsplit(sigma_cols, "..", fixed = TRUE))
+  pfts <- unique(chars[, 2])
+  traits <- unique(chars[, 3])
+  xyvars <- combn(traits, 2)
+  yvars <- xyvars[1, ]
+  xvars <- xyvars[2, ]
+  l <- list()
+  for (p in pfts) {
+    l <- c(l, purrr::pmap(list(yvar = yvars, xvar = xvars),
+                          rma_slope,
+                          pft = p,
+                          m = samples))
+  }
+  rma_cols <- gsub("Sigma", "RMA", sigma_cols)[chars[, 3] != chars[, 4]]
+  out <- do.call(cbind, l)
+  colnames(out) <- rma_cols
+  out
+}
+
+rma_result <- lapply(hier_mass$samples, all_rma_slopes) %>%
+  do.call(what = rbind)
+
+rma_means <- colMeans(rma_result)
+rma_q <- t(apply(rma_result, 2, quantile, c(0.025, 0.975))) %>%
+  tibble::as_tibble(rownames = "colname") %>%
+  dplyr::rename(lo = `2.5%`, hi = `97.5%`)
[…]

### 2019-08-23 scripts/manuscript-analyses/rma-boxplot.R

@@ -1,0 +1,212 @@
+library(shiklomanov2017np)
+
+outdir <- here::here("output")
+hier_mass_file <- tail(dir(outdir, "hier.mass.clm45", full.names = TRUE), 1)
+hier_area_file <- tail(dir(outdir, "hier.area.clm45", full.names = TRUE), 1)
+multi_mass_file <- tail(dir(outdir, "multi.mass.clm45.NA", full.names = TRUE), 1)
+multi_area_file <- tail(dir(outdir, "multi.area.clm45.NA", full.names = TRUE), 1)
+
+hier_mass <- readRDS(hier_mass_file)
+samples <- hier_mass$samples[[1]]
+
+rma_slope <- function(m, pft, yvar, xvar) {
+  sdy_col <- paste("Sigma", pft, yvar, yvar, sep = "..")
+  sdx_col <- paste("Sigma", pft, xvar, xvar, sep = "..")
+  mcol <- paste("Sigma", pft, xvar, yvar, sep = "..")
+  sign(m[, mcol]) * sqrt(m[, sdy_col] / m[, sdx_col])
+}
+
+all_rma_slopes <- function(samples) {
+  sigma_cols <- grep("^Sigma..", colnames(samples), value = TRUE)
+  chars <- do.call(rbind, strsplit(sigma_cols, "..", fixed = TRUE))
+  pfts <- unique(chars[, 2])
+  traits <- unique(chars[, 3])
+  xyvars <- combn(traits, 2)
+  yvars <- xyvars[1, ]
+  xvars <- xyvars[2, ]
+  l <- list()
+  for (p in pfts) {
+    l <- c(l, purrr::pmap(list(yvar = yvars, xvar = xvars),
+                          rma_slope,
+                          pft = p,
+                          m = samples))
+  }
+  rma_cols <- gsub("Sigma", "RMA", sigma_cols)[chars[, 3] != chars[, 4]]
+  out <- do.call(cbind, l)
+  colnames(out) <- rma_cols
+  out
+}
+
+tidy_rma <- function(output) {
+  raw_result <- lapply(hier_mass$samples, all_rma_slopes) %>%
+    do.call(what = rbind)
+
+  raw_means <- colMeans(raw_result)
+}
+
+rma_means <- colMeans(rma_result)
+rma_q <- t(apply(rma_result, 2, quantile, c(0.025, 0.975))) %>%
+  tibble::as_tibble(rownames = "colname") %>%
[…]

### 2019-08-23 manuscript/shiklomanov_newphyt.resubmit_2.Rmd
Add resubmit-2 file

---
title: "Does the leaf economic spectrum hold within plant functional types? A Bayesian multivariate trait meta-analysis."
author:
  - >
    Alexey N. Shiklomanov$^1$
    Elizabeth M. Cowdery$^1$,
    Michael Bahn$^2$,
    Chaeho Byun$^3$,
    Joseph Craine$^4$,
    Andrés Gonzalez-Melo$^5$,
    Steven Jansen$^6$,
    Nathan Kraft$^7$,
    Koen Kramer$^{8}$,
    Vanessa Minden$^{9,10}$,
    Ülo Niinemets$^{11}$,
    Yusuke Onoda$^{12}$,
    Enio Egon Sosinski$^{13}$,
    Nadejda A. Soudzilovskaia$^{14}$,
    Michael C. Dietze$^1$
bibliography:
  - library.bib
  - extra.bib
  - tryrefs.bib
csl: new-phytologist.csl
output:
  bookdown::word_document2:
    toc: no
    reference_docx: ~/Projects/misc/pandoc-docx-template/custom.docx
  bookdown::pdf_document2:
    latex_engine: xelatex
    keep_tex: yes
    toc: no
    includes:
      in_header: preamble.tex
---
<div custom-style="Affiliation">
$^1$Department of Earth & Environment, Boston University, Boston, MA 02215, USA;
$^2$Institute of Ecology, University of Innsbruck, 6020 Innsbruck, Austria;
$^3$School of Biological Sciences, Seoul National University, Gwanak-ro 1, 08826, Seoul, Seoul, South Korea;
$^4$Jonah Ventures, Manhattan, KS 66502, USA;
$^5$Universidad del Rosario. Facultad de Ciencias Naturales y Matemáticas. Carrera 26 #63B-48. Bogotá, Colombia;
$^6$Institute of Systematic Botany and Ecology, Ulm University, Ulm, Germany;
$^7$Department of Ecology and Evolutionary Biology, University of California Los Angeles, Los Angeles, CA 90095, USA;
$^{8}$Vegetation, Forest, and Landscape Ecology, Wageningen Environmental Research and Wageningen University, Wageningen, Gelderland, The Netherlands;
$^{9}$Institute for Biology and Environmental Sciences, Carl von Ossietzky-University of Oldenburg, Carl von Ossietzky Str. 9-11, 26129 Oldenburg, Germany;
$^{10}$Department of Biology, Ecology and Evolution, Vrije Universiteit Brussel, Pleinlaan 2, 1050 Brussels;
$^{11}$Institute of Agricultural and Environmental Sciences, Estonian University of Life Sciences, Kreutzwaldi 1, 51014 Tartu, Estonia;
$^{12}$Graduate School of Agriculture, Kyoto University, Kyoto, 606-8502, Japan;
$^{13}$Embrapa Temperate Agriculture, Rodovia BR 392, Km 78, CEP 96010-971 Pelotas, RS, Brazil;
$^{14}$Conservation Biology Department, Institute of Environmental Sciences, Leiden University, Leiden, The Netherlands
</div>

<div custom-style="Single space">
Corresponding author:

Alexey Shiklomanov,

telephone: +1 (302) 897-3348,

email: ashiklom\@bu.edu

Key words: Functional trade-off; hierarchical modeling; trait variation; ecological modeling; leaf morphology; leaf biochemistry
</div>

Section                   Word count
--------                  --------------      ------------
Introduction              1060                Figures: 5 (all colored)
Materials and Methods     1695                Tables: 2
Results                   1108                Supplementary figures: 2
Discussion                1907                Supplementary tables: 4
Acknowledgements          88                  Supplementary methods: 1
Total                     5858

<div custom-style="Heading, no number">Summary</div>

* We investigated whether global leaf economic relationships across plant functional types (PFTs) are also present within-PFTs, and the extent to which this hierarchical structure can be used to constrain trait estimates.

* We developed a hierarchical multivariate Bayesian model that assumes separate means and covariance structures within and across PFTs and fit this model to seven leaf traits from the TRY database: longevity, specific leaf area, nitrogen content, phosphorus content, dark respiration rate, maximum photosynthetic carboxylation rate ($V_{c,max}$), and maximum photosynthetic electron transport rate ($J_{max}$). 
We evaluated the ability of this model to estimate trait means and reproduce leaf economic trade-offs.

* Trait correlations were generally consistent in direction within and across PFTs, and consistent with predictions of the leaf economic spectrum.
However, correlation strength varied substantially across PFTs indicating that leaf economic relationships within PFTs are often confounded by the unique physiology of certain plant types or environmental conditions in certain biomes.

* Leveraging covariance in multivariate models reduced uncertainties in mean trait estimates, particularly for undersampled trait-PFT combinations.
However, additional constraint from the across-PFT hierarchy was limited compared to the simple multivariate models.

<!--chapter:end:index.Rmd-->

# Introduction

The diversity and dimensionality of the terrestrial biosphere is vast and complex, and therefore there has been a recurring debate in ecology about the utility of reductionist approaches for capturing this variability.
In particular, the use of functional groups with common characteristics has been widely applied in biodiversity studies [@naeem_disentangling_2003] and is essential to the structure of many ecosystem models [@lavorel_plant_1997; @wullschleger_plant_2014].
However, ecologists have long recognized the importance of individual variability and stochasticity in shaping ecosystems
[@gleason_individualistic_1926; @bolnick_why_2011; @rosindell_unified_2011; @clark_why_2016],
and the benefits of more finely-resolved representation of functional diversity for predictive ecology are supported by an increasing body of trait ecology literature
[@mayfield_diversity_2006; @mcmahon_improving_2011; @van_bodegom_going_2012; @reichstein_linking_2014; @violle_emergence_2014; @medlyn_using_2015; @moran_intraspecific_2016].

Plant functional traits can be used to link directly measurable features of individuals to their fitness within an ecosystem and, by extension, ecosystem performance as a whole [@violle_let_2007].
Recent syntheses of global trait databases have revealed that although the functional diversity across plant species is immense, this diversity is constrained by allometries and trade-offs between plant strategies [@wright_worldwide_2004; @kattge_try_2011; @kleyer_why_2015; @diaz_global_2016].
One axis of trait covariation currently receiving attention is the ‘leaf economic spectrum’, which defines a trade-off between plant investment in productive but short-lived leaves versus less productive but sturdy and long-lived leaves [@wright_worldwide_2004; @shipley_fundamental_2006; @reich_world-wide_2014; @diaz_global_2016]
Leaf economic traits are well-correlated with
individual plant productivity [@shipley_functional_2005; @niinemets_within-canopy_2016; @wu_convergence_2016],
litter decomposition rates [@bakker_leaf_2011; @hobbie_plant_2015],
community composition [@burns_patterns_2004; @cavender-bares_multiple_2004],
and ecosystem function [@diaz_plant_2004; @musavi_imprint_2015].
The relative position of plant species along the leaf economic spectrum has been shown to be influenced by climate and soil conditions
[@wright_worldwide_2004; @wright_modulation_2005; @cornwell_community_2009; @ordonez_global_2009; @wigley_leaf_2016].
As a result, relationships between leaf economic traits and climate have been incorporated into ecosystem models to allow for continuous variation in plant function and environmental responses
[@sakschewski_leaf_2015; @verheijen_inclusion_2015].

However, the use of among-trait and trait-environment correlations at the global scale, for both ecological inference and land surface modeling, has several important caveats.
First, observed correlations at the global scale do not always hold at smaller scales (such as sites, species, and individuals).
For example, some studies suggested consistent correlations across scales [@wright_worldwide_2004; @albert_multi-trait_2010; @asner_amazonian_2014]
whereas others showed no or even opposite correlations [@albert_intraspecific_2010; @messier_how_2010; @wright_does_2012; @feng_scale_2013; @grubb_relationships_2015; @wigley_leaf_2016; @messier_traitnetwork; @kichenin_2013_contrasting].
Many mechanisms have been suggested for deviation from global trait relationships at smaller scales.
Trade-offs between strategies may only be applicable when multiple competing strategies co-occur, and in cases where strong environmental filters allow only a narrow range of strategies, alternative processes can drive community assembly [@rosado_trait_dominance; @pierce_csr_cooccur; @grim_pierce_book].
Different selective pressures dominate at different scales, particularly within versus across species [@albert_intraspecific_2010; @messier_how_2010; @kichenin_2013_contrasting],
and the large heterogeneity in the variance structures of traits suggests that different traits have contrasting sensitivity to these different pressures [@messier_trait_2016].
Experimental evidence shows that species can alter different aspects of their leaf economy in a relatively uncoordinated fashion, even when the direction of univariate trait responses to environmental change is consistent [@wright_does_2012].
Meanwhile, across different plant functional types, resource allocation patterns, for instance of nutrients to photosynthesis versus structure and defense, differ substantially, suggesting different investment strategies and varying relationships among traits [@ghimire_2017_Nallocation].
Second, among-trait correlations at any scale do not provide causal evidence for functional trade-offs or even similarity in response to external stimuli [@messier_trait_2016].
Therefore, ascribing too much leverage to trait correlations can lead to an underestimation of plant functional diversity [@grubb_trade-offs_2015].
Third, plants maintain their fitness in a given environment through multiple independent strategies (corresponding to multiple mutually orthogonal axes of trait variability).
As a consequence, changes in key leaf economic traits such as leaf nitrogen content and specific leaf area area may not affect other aspects of plant function, such as
hydraulics [@li_leaf_2015],
overall plant carbon budget [@edwards_leaf_2014],
and dispersal [@westoby_plant_2002].
Finally, modeling ecosystem function based on trait correlations is sampling from the hypothetical space of potential species and communities that could have evolved, rather than constraining models to forecast the actual vegetation we have today as the result of spatial separation and constraints on convergent evolution.
Among other problems, this approach fails to account for the timescales required for adaptation as well as actual limitations of the physiology of different species and community assembly.

An alternative approach is to preserve existing PFT classifications
[though potentially with finer taxonomic, functional, or spatial resolution, e.g. @boulangeat_improving_2012]
while using statistical analyses to account for uncertainty and variability in the aggregated trait values.
For example, the Predictive Ecosystem Analyzer (PEcAn, pecanproject.org), an ecosystem model-data informatics system, parameterizes PFTs using trait probability distributions from a Bayesian meta-analysis of plant trait data across many studies
[@dietze_improving_2013; @lebauer_facilitating_2013].
This approach explicitly separates the processes driving PFT-level differentiation from processes that drive finer-scale functional variability,
and is useful for guiding future data collection and model refinement [@dietze_quantitative_2014].
However, a univariate meta-analysis, like the one currently in PEcAn, is limited by its failure to account for trait correlations, therefore neglecting useful knowledge about relationships across PFTs and between traits.
At the other extreme, existing regional and global scale analyses [e.g. @van_bodegom_going_2012; @sakschewski_leaf_2015] ignore variability within PFTs, often resulting in macroecological, evolutionary, and competitive trade-offs across PFTs being used to drive both acclimation and instantaneous responses within PFTs.

While the leaf economic spectrum has been investigated at the global scale, where it is robust, and at the site or plot scale, where deviations from it are common, it has received less attention at the intermediate scale of PFTs.
Thus, this paper seeks to answer the following questions:
First, to what extent does the leaf economic spectrum hold within vs. across plant functional types?
Second, to what extent can the leaf economic spectrum and other patterns of trait covariance be leveraged to constrain trait estimates, particularly under data limitation?
The answer to these question has implications for both functional ecology and ecosystem modelling.
To evaluate these questions, we develop a hierarchical multivariate Bayesian model that explicitly accounts for across- and within-PFT variability in trait correlations.
We then fit this model to a global database of foliar traits to estimate mean trait values and variance-covariance matrices for PFTs as defined in a major earth system model [Community Land Model, CLM, @clm45_note].
We evaluate the ability of this model to reduce uncertainties in trait estimates and reproduce observed patterns of global trait variation compared to non-hierarchical multivariate and univariate models.
Finally, we assess the generality and scale dependence of trait trade-offs by comparing covariance estimates globally and within each PFT.

<!--chapter:end:01_introduction.md-->

# Materials and Methods

## Trait data

Foliar trait data for this analysis comes from the TRY global traits database (@kattge_try_2011; Tab.\ \@ref(tab:trydata)).
We focused our research on seven foliar traits:
Leaf longevity (months),
specific leaf area (SLA, m$^2$ kg$^{-1}$),
leaf nitrogen content ($N_{mass}$, mg N g$^{-1}$ or $N_{area}$, g m$^{-2}$),
leaf phosphorus content ($P_{mass}$, mg P g$^{-1}$ or $P_{area}$, g m$^{-2}$),
leaf dark respiration at 25°C ($R_{d,mass}$, µmol g$^{-1}$ s$^{-1}$, or $R_{d,area}$, µmol m$^{-2}$ s$^{-1}$),
maximum Rubisco carboxylation rate at 25°C ($V_{c,max,mass}$, µmol g$^{-1}$ s$^{-1}$, or $V_{c,max,area}$, µmol m$^{-2}$ s$^{-1}$),
and maximum electron transport rate at 25°C ($J_{max,mass}$, µmol g$^{-1}$ s$^{-1}$, or $J_{max,area}$, µmol m$^{-2}$ s$^{-1}$).
For $V_{c,max}$, we only used values already reported in TRY as being at 25°C.
For $R_{d}$, we normalized the values to 25°C based on reported leaf temperature values following the same methods as @atkin_global_2015.
For $J_{max}$, we normalized the values to 25°C based on reported leaf temperature values using the temperature response function described in @kattge_2007_temperature (Equation 1 therein).
To avoid potential artifacts caused by different trait normalization, we performed analyses separately for both mass- and area-normalized traits [@osnas_global_2013; @lloyd_les].
We restricted our analysis to TRY data that have been quality-controlled, where at least one of the above traits was observed, and for which adequate species information was provided for functional type classification [see @kattge_try_2011].

Although the light- and CO2-saturated photosynthetic rate ($A_{max}$) was an important trait in previous studies, we did not include it in our study for two reasons.
First of all, data on raw photosynthetic rates are highly sensitive to measurement methodology and environmental conditions, which were generally inconsistent or unavailable in TRY.
Second, $A_{max}$ is not a good measure of photosynthetic capacity because it integrates over variability in many physiologically independent traits such as $V_{c,max}$, $J_{max}$, and stomatal conductance, and is therefore not used in vegetation models as a photosynthetic parameter [@Ali_2015].

Following past studies [e.g. @wright_worldwide_2004; @wright_assessing_2005; @onoda_2011_global; @diaz_global_2016], we log-transformed all trait values to correct for their strong right-skewness.

## Plant functional types

We assigned each species to a unique plant functional type (PFT) following the scheme in the Community Land Model [CLM4.5, @clm45_note; Tab.\ \@ref(tab:pfts), Fig.\ 1].
We obtained categorical data on growth form, leaf type, phenology, and photosynthetic pathway from the TRY database.
Where species attributes disagreed between datasets (e.g., categorized as a shrub in one dataset but a tree in another), we assigned the attribute that was observed most frequently between the datasets (e.g., if five datasets say "shrub" but only one says "tree", we would classify it as a shrub).
Where species lacked data on certain attributes, we assigned those attributes based on higher order phylogeny where appropriate (e.g., *Poaceae* family are usually grasses, *Larix spp.* are deciduous needleleaved trees) or otherwise omitted the species from our analyses.
For biome specification, we collected all latitude and longitude data for each species,
matched these data to 30 second ($\sim1$ km$^{2}$) mean annual temperature ($AMT$, averaged 1970-2000) data from WorldClim-2 [@worldclim],
calculated the mean AMT for all sites at which that species was observed,
and then binned these species based on the following cutoffs: boreal/arctic ($AMT \leq 5^\circ C$), temperate ($AMT \leq 20^\circ C$), and tropical ($AMT > 20^\circ C$).

## Multivariate analysis

### Basic model description

In this study, we compared three different models representing different levels of complexity.

The simplest model was the ‘univariate’ model, in which each trait was modeled independently.
For an observation $x_{i,t}$ of trait $t$ and sample $i$:

$$x_{i,t} \sim N(\mu_t, \sigma_t)$$

where $N$ is the univariate normal (Gaussian) distribution with mean $\mu_t$ and standard deviation $\sigma_t$ for trait $t$.

The second-simplest model was the ‘multivariate’ model, in which traits were modeled as samples from a multivariate distribution with a single mean vector and covariance matrix.
For the observed vector of traits ${\mathbf{x_i}}$ for sample $i$:

$$\mathbf{x_i} \sim mvN(\mathbf{\mu}, \mathbf{\Sigma})$$

where $mvN$ is the multivariate normal (Gaussian) distribution with mean vector $\mathbf{\mu}$ and variance-covariance matrix $\mathbf{\Sigma}$.
We ran both of these models independently for each PFT as well as for the entire dataset (as if every observation belonged to the same PFT).

The most complex model was the ‘hierarchical’ model, in which observed trait values were drawn from a PFT-specific multivariate normal distribution describing within-PFT variation and whose parameters were themselves sampled from a global multivariate distribution describing the variation across PFTs.
For the observed vector of traits $\mathbf{x}_{i,p}$ for sample $i$ belonging to PFT $p$:

$$\mathbf{x}_{i,p} \sim mvN(\mathbf{\mu}_p, \mathbf{\Sigma}_p)$$
$$\mathbf{\mu}_p \sim mvN(\mathbf{\mu}_g, \mathbf{\Sigma}_g)$$

where $\mathbf{\mu}_p$ and $\mathbf{\Sigma}_p$ are the mean vector and variance-covariance matrix describing variation within PFT $p$, and $\mathbf{\mu}_g$ and $\mathbf{\Sigma}_g$ are the mean vector and variance-covariance matrix describing across-PFT (global) variation.

### Model implementation

We fit the above models using a Gibbs sampling algorithm that leveraged known conjugate prior relationships for efficient exploration of the sampling space.
For priors on all multivariate mean vectors ($\mathbf{\mu}$), we used normal distributions:

$$P(\mathbf{\mu}) \sim mvN(\mathbf{\mu}_0, {\mathbf{\Sigma}}_0)$$

This gives rise to the following expression for the posterior:

$$P(\mathbf{\mu} \mid 
    \mathbf{x}, \mathbf{\Sigma}, 
    \mathbf{\mu}_0, \mathbf{\Sigma}_0)
  \sim
  mvN(\mathbf{\mu^*}, \mathbf{\Sigma^*})$$

$$\mathbf{\Sigma^*} = (\mathbf{\Sigma}_0^{-1} + n \mathbf{\Sigma}^{-1})^{-1}$$
$$\mathbf{\mu^*} = \mathbf{\mu}_0 \mathbf{\Sigma}_0^{-1} + \bar{\mathbf{x}} n \mathbf{\Sigma}^{-1}$$

where ${\bar{{\mathbf{x}}}}$ are the sample means of the data and $n$ is the number of rows in the data.

For priors on all multivariate variance-covariance matrices, we used the Wishart distribution ($W$):

$$P(\mathbf{\Sigma}) \sim W(\nu_0, \mathbf{S}_0)$$

This gives rise to the following expression for the posterior:

$$P(\mathbf{\Sigma} \mid
  \mathbf{x}, \mathbf{\mu},
  \nu_0, \mathbf{\Sigma}_0)
  \sim
  (W(\nu^*, S^*))^{-1}$$

$$\nu^* = 1 + \nu_0 + n + m$$
$$\mathbf{x^*} = \mathbf{x} - \bar{\mu}$$
$$\mathbf{SS} = \mathbf{x^*}^{T} \mathbf{x^*}$$
$$\mathbf{S^*} = (\mathbf{S}_0 + \mathbf{SS})^{-1}$$

where $n$ is the number of rows and $m$ is the number of columns in data matrix $x$. For further details on this derivation, see @gelman_bayesian.

The fundamentally multivariate nature of the sampling procedure described above makes it incapable of accommodating partially missing observations.
Therefore, our algorithm also included multiple imputation of partially missing data, which proceeded as follows:
For a block of data $\mathbf{x'}$ containing missing observations in columns $\mathbf{m}$ and present observations in columns $\mathbf{p}$,
the missing values $\mathbf{x'}[m]$ are drawn randomly from a conditional multivariate normal distribution at each iteration of the sampling algorithm:

$$\mathbf{x'}[m|p] \sim mvN(\mathbf{\mu}', \mathbf{\Sigma}')$$

$$\mathbf{\mu'} = 
  (\mathbf{x'}[p] - \mathbf{\mu'}[p]) 
  (\mathbf{\Sigma}[p,p]^{-1} \mathbf{\Sigma}[p,m])$$
$$\mathbf{\Sigma'} = \mathbf{\Sigma}[m,m] - 
  \mathbf{\Sigma}[m,p]
  (\mathbf{\Sigma}[p,p]^{-1} \mathbf{\Sigma}[p,m])$$

We emphasize that the imputation of missing data is performed iteratively as parameters are being estimated, such that each set of imputed values is conditioned on the current sampled state of the parameters.
This approach is therefore distinct from single imputation, where data are imputed first in a separate step prior to parameter estimation.
A demonstration of this multiple imputation approach and how it is used to estimate trait covariance is provided Supporting Information Method S1.

For each model fit, we ran five parallel MCMC chains, continuing the sampling until the final result achieved convergence as determined by a Gelman-Rubin potential scale reduction statistic less than 1.1 [@gelman_1992_inference].
We implemented this sampling algorithm in an open source, publicly available R [version 3.4.3; @rstats] package (<http://github.com/ashiklom/mvtraits>).

### Analysis of results

To assess the impact of multivariate and hierarchical constraint on trait estimates,
we compared the mean and 95% confidence intervals of trait estimates for each PFT from each model (Fig.\ 2, Tab.\ S1 and S2).
For reference, we also added the default parameter values of CLM 4.5 [Table 8.1 in @clm45_note] for SLA, $N_{mass}$, $N_{area}$, $V_{c,max,mass}$, and $V_{c,max,area}$ to Fig.\ 2.
To convert CLM's reported C:N ratio to $N_{mass}$, we assumed a uniform leaf C fraction of 0.46.
We then divided this calculated $N_{mass}$ by the reported SLA to obtain $N_{area}$.
We calculated $V_{c,max,mass}$ by multiplying the reported $V_{c,max,area}$ by the reported SLA.

To test the hypothesis that the multivariate and hierarchical models offer more value in terms of uncertainty constraint at smaller sample sizes, we calculated the relative uncertainty ($\alpha$) as a function of the mean ($\mu$) and upper ($q_{0.975}$) and lower ($q_{0.025}$) confidence limits of trait estimates.

$$ \alpha = \frac{q_{0.975} - q_{0.025}}{\mu} $$

We then fit a generalized linear model relating relative uncertainty to sample size ($n$) for each of the model types (univariate, multivariate, and hierarchical; Fig. 3).

$$ \log{\alpha} = b_0 + b_1 \log{n} $$

If all three models performed equally well at all sample sizes, their respective slope and intercept coefficients would be statistically indistinguishable.
On the other hand, models that perform better should have
lower intercept ($b_0$) coefficients, indicating generally lower uncertainty,
and
lower slope ($b_1$) coefficients, indicating a reduced sensitivity of uncertainty ($\alpha$) to sample size ($n$).

To assess the consistency of within- and across-PFT trait trade-offs, we looked at covariance estimates for each trait pair and, where these values were significantly different from zero ($p < 0.05$),
we calculated the eigenvalues from the pairwise variance-covariance matrix for that trait pair and plotted the corresponding dominant eigenvectors centered on the mean estimates (Fig.\ 4).
This figure provides a visual representation of relative positions of PFTs in trait space and both the direction and extent of within-PFT trait covariance, and is directly analogous to conceptual figures describing hierarchical trait variability across environmental gradients as presented in, for instance, @cornwell_community_2009 and @albert_intraspecific_2010.
Due to the small number of points used to estimate across-PFT covariance in the hierarchical model, none of the across-PFT covariances estimated in the hierarchical model were significantly different from zero ($p < 0.05$).
For this reason, we compared within-PFT covariances as estimated by the hierarchical model with the covariances estimated by fitting a multivariate model to all of the data.

Besides the consistency in the direction of trait covariance globally and between different PFTs, we also investigated the strength and predictive power of these covariances, which is represented by correlation coefficients (i.e. the pairwise covariance normalized to the variances of the component variables).
To do this, we plotted the mean and 95% confidence interval of the pairwise trait correlation coefficients for the global estimate from the pooled multivariate model and PFT-level estimates from the hierarchical model (Fig.\ 5).

The magnitude of correlation coefficients can be strongly influenced by data sampling, particularly sample size and range (correlations are likely to be stronger when data have more samples and a larger range).
To evaluate the contribution of uneven data sampling to our correlation predictions, we plotted the square of each pairwise correlation coefficient against pairwise sample size and normalized data range (Fig.\ S1 and S2).

The R code and ancillary data for running these analyses is publicly available online via the Open Science Framework at <https://osf.io/w8y73/>.
The TRY data request used for this analysis has been archived at <http://try-db.org>,
and can be retrieved by providing the TRY data request ID (#1584).
Alternatively, the exact pre-formatted data used in this analysis are available on request to Alexey Shiklomanov (ashiklom@bu.edu).

<!--chapter:end:02_methods.Rmd-->

# Results

## Estimates of PFT-level means

In general, leaf trait estimates from the univariate, multivariate, and hierarchical models were similar (Fig.\ 2, Tab.\ S1 and S2).
Where we observed differences between models, the largest were between the univariate and multivariate models, while the additional constraint from the hierarchical model tended to have a minimal effect on trait estimates.
Significant differences in trait estimates between univariate and multivariate models occurred even for traits with relatively large sample sizes, such as leaf nitrogen content.

Evergreen PFTs had by far the largest leaf lifespan, with the longest lifespan observed for temperate and boreal needleleaf evergreen species.
Meanwhile, all of deciduous species had lifespans shorter than 7 months.
Among deciduous species, lifespan was generally longer in warmer biomes than colder ones.

Across-PFT patterns in SLA and $N_{mass}$, $P_{mass}$, and $R_{d,mass}$ were similar.
Temperate broadleaved deciduous trees and shrubs generally had among the highest values of these traits, while temperate evergreen trees and shrubs had generally among the lowest.
However, none of these patterns were universal to all four traits.
For example, tropical evergreen trees had relatively high $N_{mass}$ and mean SLA and $R_{d,mass}$, but among the lowest $P_{mass}$.
Similarly, temperate and boreal shrubs had higher $N_{mass}$ and $P_{mass}$ than any of the grasses, but comparable SLA.

Across-PFT patterns in $N_{area}$, $P_{area}$, and $R_{d,area}$ were different from their mass-normalized counterparts.
For example, tropical broadleaved evergreen and needleleaf evergreen trees had among the lowest $N_{mass}$ and $P_{mass}$ basis but among the highest $N_{area}$ and $P_{area}$, while the opposite was true of deciduous temperate trees and shrubs.
Species with N contents near the middle of the observed range did not shift as dramatically depending on type of normalization.

C3 grasses had both the highest $V_{c,max,mass}$ and $V_{c,max,area}$.
Compared to broadleaved trees, temperate needleleaved evergreen trees had lower $V_{c,max,mass}$ but higher $V_{c,max,area}$.
Among broadleaved trees, deciduous trees had higher $V_{c,max,mass}$ and slightly higher $V_{c,max,area}$ than evergreen trees.
Between the deciduous and evergreen tree PFTs, we observed no significant trend by climate zone.

C3 grasses and temperate needleleaved evergreen trees had the highest $J_{max,area}$, but temperate broadleaved deciduous trees had the highest $J_{max,mass}$.
All of the shrub PFTs had the lowest $J_{max,mass}$ but average or above-average $J_{max,area}$, while the opposite was true of broadleaved tropical PFTs.
Of the tree PFTs, needleleaved evergreen trees had the highest $J_{max,area}$ but the lowest $J_{max,mass}$.

A key application of this study was to provide data-driven parameter estimates for Earth System Models.
To this end, we compared our mean parameter estimates with corresponding default parameters in CLM 4.5 [@clm45_note] (Fig.\ 2).
Our mean estimates of SLA agreed with CLM's defaults [@clm45_note, Table 8.1] only for tropical broadleaved evergreen trees, and for all other PFTs, our estimates are significantly lower.
For $N_{mass}$, our estimates agreed reasonably well with CLM for evergreen temperate trees, needleleaved trees, and C3 arctic grasses, and were substantially different for all other PFTs.
Our $N_{mass}$ estimates also varied much more across PFTs than CLM's parameters.
For $N_{area}$, our estimates were significantly higher than CLM's for all PFTs, likely due to CLM's overestimates of SLA.
Our estimates of $V_{c,max_mass}$ were lower across all PFTs, with particularly large differences for tropical and temeprate broadleaf deciduous trees and evergreen shrubs, and temperate C3 grasses.
Our estimates of $V_{c,max,area}$ showed better agreement, though our values were still significantly lower for many PFTs.
Like us, @kattge_2009_vcmax also found that $V_{c,max,area}$ was overestimated by Earth System models,
but their estimates of $V_{c,max,area}$ and $N_{area}$ are generally slightly higher than ours.

We observed clear differences in the relative uncertainties of mean estimates with respect to sample size.
All of the high-latitude PFTs consistently had among the largest error bars around their mean estimates relative to other PFTs, while the traits with the largest uncertainties were dark respiration, $V_{c,max}$, and $J_{max}$.
For many of these trait-PFT combinations, the additional constraint from trait covariance provided by the multivariate and hierarchical models substantially reduced error bars, making it possible to compare estimates against those of other PFTs.
Our analysis of the relationship between model type, sample size, and estimate relative uncertainty found that this covariance-based constraint from the multivariate model both reduced uncertainty overall (lower intercept) and reduced the sensitivity of estimate uncertainty to sample size (lower slope) compared to the univariate model (Fig.\ 3).
However, this analysis revealed no consistent significant benefit from the hierarchical model.

## Trait correlation patterns across- and within-PFTs

For all traits except leaf lifespan, pairwise trait correlations were generally consistent in direction both globally and within each PFT (Fig.\ 4).
In particular, mass- and area-normalized traits were all positively correlated with each other and, respectively, positively and negatively correlated with SLA, both globally and within each PFT.
The same was generally true of correlations of mass-based traits with leaf lifespan, but correlations of leaf lifespan with area-normalized traits were more variable.
The correlation between $N_{area}$ and leaf lifespan was positive globally and for evergreen shrubs, tropical broadleaved deciduous trees, temperate needleleaved evergreen trees but negative for temperate and boreal broadleaved deciduous trees and not significant for any other PFTs.
Similarly, the correlation between $P_{area}$ and leaf lifespan was positive globally but negative for evergreen shrubs and not significant for any other PFTs.
The correlation between leaf lifespan and $R_{d,area}$ was significant and negative globally, but was not significant within any PFTs.
The only significant correlations of leaf lifespan with $V_{c,max,area}$ and $J_{max,area}$ were negative for temperate broadleaved deciduous trees.

The strength of pairwise trait correlations varied substantially depending on scale, PFT, and trait (Fig.\ 5).
In some cases, this variability was driven by low sample sizes (Fig.\ 1, S1; Tab.\ S3, S4).
For instance, needleleaved deciduous trees, the most undersampled PFT in our analysis, were often the only PFT for which a correlation was not statistically significant.
Similarly, we had no observations of dark respirations for deciduous boreal shrubs, which explains why we found no significant correlations of dark respiration with any other trait for that PFT.
That being said, the relationship between correlations strength and sample size was very inconsistent (Fig.\ S1; Tab.\ S4).
For example, temperate needleleaf evergreen trees and C3 arctic grasses showed the strongest $N_{area}$-$P_{area}$ correlations despite having relatively average sample sizes for both traits.
In fact, every trait pair we investigated had at least one case (and often several cases) where a PFT with a larger sample size showed weaker correlations than PFTs with lower sample sizes, or where correlation strength varied significantly among PFTs with similar sample sizes (Fig.\ S1).
Relationships between correlation strength and data range were even less consistent (Fig.\ S2).
Based on these results, we conclude that the variation we observed in our correlation results is not driven primarily by sampling artifacts and at least partially captures some underlying ecophysiological differences between PFTs.

<!--chapter:end:03_results.Rmd-->

# Discussion

## Scale dependence of the leaf economic spectrum

The canonical leaf economic spectrum is defined by a negative correlation of between SLA and leaf lifespan, and a positive correlation of SLA with $N_{mass}$, $P_{mass}$, and photosynthesis and respiration rates [@wright_worldwide_2004].
The first objective of this paper was to investigate the extent to which these relationships hold within and across PFTs.
Our results indicate that the leaf economic spectrum generally still holds within PFTs, at least at the functional and phylogenetic scale of the current generation of Earth System Models.
Within PFTs, correlations between SLA, $N_{mass}$, and $P_{mass}$ were consistently positive, and correlations of these traits with leaf lifespan were generally negative (though, for many PFTs, correlations were not significantly different from zero).
Although we did not include maximum photosynthesis rate ($A_{max}$), $V_{c,max,mass}$ and $J_{max,mass}$ generally exhibited the expected positive correlations with SLA and negative correlations with leaf lifespan, as did $R_{d,mass}$, though many correlations were not significant, primarily due to sample size.

While the direction of relationships we observed was largely consistent across different PFTs, the strength of these correlations was more variable.
For example, the correlations of SLA with $N_{mass}$ and $P_{mass}$ were generally weaker in needleleaved PFTs than in broadleaved PFTs.
On the other hand, relationships of SLA and $N_{area}$ were strongly negative for all PFTs (except the data-limited needleleaved deciduous trees), and temperate needleleaved species in particular had among the strongest SLA-$N_{area}$ and SLA-$P_{area}$ correlations of any PFT.
Taken together with the finding that evergreen conifers are relatively unique in their consistently proportional allocation of N to cell walls and Rubisco [@onoda_physiological_2017], our results support the idea that needleleaved species primarily adapt to environmental changes through structural changes (i.e. increasing or decreasing SLA) rather than through modulating foliar biochemistry [@robakowski_2004_needle].

We also found that the strength of correlations between leaf nutrient concentrations and traits related to photosynthetic metabolism ($V_{c,max}$ and $J_{max}$)---which are often used in ecosystem modeling to parameterize photosynthesis [@clm45_note; @rogers_roadmap_2017]---were highly variable between PFTs.
Although trait correlations are not necessarily indicative of allocation strategies, this result generally supports the findings of @ghimire_2017_Nallocation that N allocation strategies to photosynthesis vary widely by plant functional type.
In particular, the correlations of $N_{mass}$ with traits related to photosynthesis and respiration ($R_{d,mass}$, $V_{c,max,mass}$, and $J_{max,mass}$) were weaker for tropical evergreen broadleaved trees than for most other PFTs, but the correlation of these traits with $P_{mass}$ was comparable or even slightly higher.
This suggests that the productivity of tropical species is generally P-limited [@reich_global_2004; @ghimire_2017_Nallocation], that N allocation strategies are more variable under N-poor than N-rich conditions [@ghimire_2017_Nallocation], or more generally that photosynthetic parameters are affected more strongly by environmental covariates than leaf nitrogen contents [@Ali_2015].
Meanwhile, our result that the $N_{area}$ - $V_{c,max,area}$ correlation is generally weaker for needleleaved than broadleaved species was also found by @kattge_2009_vcmax, and supports the lower allocation of N to photosynthesis found by @ghimire_2017_Nallocation.
In light of this fact and considering that boreal forests, largely populated by needleleaf species, have the largest influence on global climate of any biome [@Snyder_2004; @bonan_forests_2008], we suggest that parameterization of needleleaf tree productivity based on foliar nitrogen content in Earth System Models be treated with caution.

Correlations of all traits with leaf lifespan were weaker (often to the point of becoming insignificant) within most PFTs than globally.
This suggests that leaf economic relationships related to leaf lifespan are dominated by the fundamental differences between deciduous and evergreen PFTs, while the factors driving variability in leaf lifespan within PFTs are more complex and idiosyncratic [@Reich_2014; @wu_leaf_2016].
That being said, much of this within-PFT variability is known to be driven by variations in shade responses, and a key limitation of our study is that we have no information about the relative canopy positions at which traits were collected [@lusk_why_2008; @keenan_global_2016].

Across PFTs, the interaction between growth form and biome in PFT definitions (Table 1) confounds the interpretation of our results with respect to well established ecological and biogeographic patterns.
For example, @poorter_causes_2009 found variation in SLA by growth form was, from highest to lowest, grasses, deciduous woody plants, evergreen woody plants, and succulents, while the pattern by biome was, again from highest to lowest, grassland, tundra, tropical forest, temperate forest, woodland, shrubland, and desert.
We observed as expected that arctic grasses had lower mean SLA than temperate grasses, and that evergreen trees had lower SLA than their deciduous counterparts.
However, we observed by far the highest mean SLA values in temperate deciduous broadleaf trees, rather than in grass PFTs as expected.
Similarly to @onoda_2011_global, we found no consistent patterns in SLA with temperature:
Among broadleaved evergreen PFTs, temperate species had lower SLA than tropical, but among broadleaved deciduous PFTs, temperate species had higher SLA than both tropical and boreal species.
With respect to the expected decline in foliar N:P ratios with latitude [@reich_global_2004], we found that $N_{mass}$ was generally higher in PFTs associated with colder biomes compared to warmer ones, but observed biome-related differences in $P_{mass}$ only among deciduous broadleaved and needleleaved evergreen PFTs.
Contrary to @atkin_global_2015, our results for both $R_{d,mass}$ and $R_{d,area}$ failed to show a trend with respect to biome.
However, this comparison may not be entirely fair because our study design inherently averages over the extensive climatic variability within PFTs.

Finally, there has been some debate in the trait ecology community about the use of mass- or area-normalized traits in analyses of the leaf economic spectrum.
Two studies [@osnas_global_2013; @lloyd_les] independently concluded that leaf economic relationships among mass-based traits emerge inevitably out of variation in SLA and are therefore not ecologically meaningful.
Responses to these criticisms have suggested that both mass- and area-based normalization have merit, as mass-based traits have a natural interpretation in terms of investment costs and allocation while area-based traits are important due to the fundamentally area-based nature of energy and gas fluxes through leaf surfaces [@westoby_lloyd_response; @poorter_les_response].
Our study suggests the latter, that investigation of trait correlations on both a mass- and area-basis can yield biologically and ecologically meaningful conclusions.
For one, our discussion of differences in ecological strategies between broadleaved and needleaved species fundamentally depends on comparative analysis of mass- and area-normalized nutrient contents.
Meanwhile, our discussion of tropical tree productivity with respect to foliar nutrient contents is generally supported by both mass- and area-normalized traits (i.e. the correlation between $P$ and $V_{c,max}$ is stronger than the correlation of $N$ and $V_{c,max}$ for tropical species whether we use the mass- or area-normalized versions of these traits).

## Covariance as constraint

The second objective of this paper was to investigate the ability of trait covariance to provide additional information that could be used to reduce uncertainties in trait estimates.
In agreement with our expectations, our results show that accounting for trait covariance constrained the uncertainty around PFT-level trait means, particularly for trait-PFT combinations with low sample sizes (Fig.\ 2 and 3).
Moreover, accounting for trait covariance occasionally also changed the position of trait mean estimates, even for PFT-trait combinations with relatively large sample sizes (e.g. $N_{mass}$ for temperate broadleaved deciduous trees, Fig.\ 2).
This result echoes @diaz_global_2016 in demonstrating the importance of studying the multivariate trait space rather than individual traits. 
Such shifts suggest that the sampling of these traits in the TRY database is not representative, which is evident from looking at the relative sample sizes of different traits (Fig.\ 1; see also Kattge et al. -@kattge_try_2011). 
These shifts also indicate that parameter estimates based on univariate trait data [e.g. @lebauer_facilitating_2013; @dietze_quantitative_2014; @butler_2017_mapping] may not only be overestimating uncertainty, but may also be systematically biased.
Although our results clearly show that many time- and labor-intensive traits, such as $R_{d}$, $V_{c,max}$, and $J_{max}$, still lack the observations to estimate covariance with other traits for certain PFTs,
our results also show that the effective sample size of all traits is enhanced when covariance is taken into account.
For example, field and remote sensing studies that estimate only certain traits like SLA and $N_{mass}$ but not others (such as $P_{mass}$ and $R_{d,mass}$) can leverage trait correlations to provide indirect constraint [@serbin_spectroscopic_2014; @singh_imaging_2015; @musavi_imprint_2015; @lepine_examining_2016].
We therefore suggest that future observational campaigns consider trait covariance when deciding which traits to measure.

The additional benefit of hierarchical multivariate modeling in our study was limited, largely due to the relatively low number of points used to estimate that across-PFT covariance structure.
Therefore, for parameterizing the current generation of ecosystem models using well-sampled traits, we suggest that simple multivariate models fit independently to each PFT are sufficient and that the additional conceptual challenges and computational overhead of hierarchical modeling are not required.
However, for modeling work that requires larger numbers of PFTs [@boulangeat_improving_2012], and especially for models that are parameterized on the basis of individual species [e.g. Linkages, @linkages], the benefits of hierarchical modeling may accumulate [@clark_why_2005; @dietze_capturing_2008; @cressie_accounting_2009; @webb_structured_2010].

More generally, we foresee tremendous potential of multivariate and hierarchical modeling approaches to elucidating the role of functional traits in organismal and ecosystem function. 
Besides the many important foliar traits that we did not include (e.g.\ pigment contents, leaf hydraulic traits), our approach could readily be applied to other plant traits whose relationship to the leaf economic spectrum is less clear. 
One example is hydraulic traits:
While stem and leaf hydraulic traits are correlated [@bartlett_hydraulic], a potential scaling between hydraulic and leaf economic traits is poorly understood [@reich_world-wide_2014; @li_leaf_2015]. 
Similarly, reexamining the relationships defining wood [@chave_woodeconomics; @baraloto_decoupled_2010; @fortunel_leaf_2012] and root [@kramer-walter_root_2016; @valverde-barrantes_root_2016] economic spectra, as well as their relationship to the foliar traits, would provide useful information on possible scale-dependence of plant growth and allocation strategies.
We emphasize that the relative difficulty of measuring hydraulic and other non-foliar traits [e.g @jansen_hydraulic_traits] further increases the value of any technique that can fully leverage the information they provide. 
More generally, multivariate and hierarchical modeling has the potential to reveal functional trade-offs that are mutually confounding at different scales, thereby enhancing our understanding of processes driving functional diversity.

## Conclusions

The tremendous functional diversity of plants continues to be a major challenge for functional ecology and ecosystem modeling.
Functional diversity research fundamentally depends on dimensionality reduction through a search for meaningful pattern that can be exploited to take reasonable guesses at average behavior. 
The trait trade-offs comprising the leaf economic spectrum are one such pattern. 
In this paper, we reaffirm the existence of the leaf economic spectrum both globally and, with some caveats, within plant functional types typically used in the current generation of Earth System Models.
We also highlight how the strength of leaf economic relationships can be influenced by biotic and abiotic factors specific to certain PFTs.
Finally, we show how patterns of trait covariance like the leaf economic spectrum can be leveraged to inform trait estimates, particularly at small sample sizes.

<div custom-style="Heading, no number">Acknowledgements</div>

This project was supported by NASA grant NNX14AH65G and NSF grants 1261582 and 1458021, as well as the TRY initiative on plant traits (http://www.try-db.org).
The TRY initiative and database is hosted, developed and maintained by J. Kattge and G. Boenisch (Max Planck Institute for Biogeochemistry, Jena, Germany).
TRY is currently supported by DIVERSITAS/Future Earth and the German Centre for Integrative Biodiversity Research (iDiv) Halle-Jena-Leipzig.
The authors would also like to thank Christine Rollinson, Istem Fer, and Colin Averill for their valuable feedback on early drafts of this manuscript.

<div custom-style="Heading, no number">Author contributions</div>

ANS wrote the manuscript and implemented the analysis.
ANS and EMC designed the analysis and figures.
MCD conceived the original idea for the manuscript, guided its development, and provided financial support.
MB, SJ, KK, ÜN, and NAS provided extensive feedback on multiple drafts of the manuscript, and contributed data.
All other authors contributed data.

<!--chapter:end:04_discussion.md-->

<div custom-style="Heading, no number">Tables</div>

```{r "knitr_options", echo = FALSE, message = FALSE, warning = FALSE, error = FALSE}
library(knitr)
library(here)
opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE, error = FALSE
)
options(knitr.table.format = "pandoc")
```

<!-- Tables ---------------------------->
```{r "read_chunks"}
knitr::read_chunk(here::here("scripts/manuscript-analyses/try_ref_table.R"))
knitr::read_chunk(here::here("scripts/manuscript-analyses/table_n_species.R"))
```

<!-- TRY data citations -->
```{r "trydata"}
```

<!-- Number of species in each PFT -->
```{r "pfts"}
```

<!--chapter:end:05_figures_tables.Rmd-->

### 2019-09-02 R/trait_names.R
Add drop_mass_area helper function

@@ -25,6 +25,18 @@ both_params <- c("leaf_lifespan", "SLA", "Nmass", "Narea", "Pmass", "Parea",
                  "Rdmass", "Rdarea", "Vcmax_mass", "Vcmax_area",
                  "Jmax_mass", "Jmax_area")
 
+#' Drop the mass/area designation from trait, and return a clean factor
+#'
+#' @param trait Character or factor of traits
+#' @return Trait vector as factor, with levels [param_labels]
+#' @author Alexey Shiklomanov
+#' @export
+drop_mass_area <- function(trait) {
+  char <- gsub("_?(mass|area)", "", as.character(trait))
+  char[char == "leaf_lifespan"] <- "Leaf lifespan"
+  factor(char, param_labels)
+}
+
 #' Trait labels
 #'
 #' Various vectors for providing formatted trait labels for figures and tables.

### 2019-09-02 R/try-data.R
Make sure PFTs are factors in try_data

@@ -15,6 +15,7 @@ try_data <- function(mass_area = NULL,
   }
   dat <- readr::read_csv(f, col_types = c("PFT" = "c",
                                           .default = "n")) %>%
+    dplyr::mutate(PFT = factor(PFT, abbr2pft[-1])) %>%
     dplyr::rename(pft = PFT)
   if (is.null(mass_area)) return(dat)
   rxp <- switch(mass_area,

### 2019-09-02 scripts/n-fold-area.R
Fix levels in n-fold-* scripts

@@ -98,7 +98,7 @@ cross_validate <- function(data_mat, groups, nremove) {
 
   # Multivariate model -- by PFT
   mboot_means <- matrix(NA_real_, nrow(test_data), ncol(test_data))
-  for (g in levels(data_groups)) {
+  for (g in levels(groups)) {
     ig <- groups == g
     gdata <- test_data[ig, ]
     gmat <- multi_results[[g]]
@@ -135,20 +135,3 @@ cv_results <- replicate(ncv, cross_validate(data_mat, data_groups),
 
 cv_results_df <- bind_rows(cv_results, .id = "i")
 write_csv(cv_results_df, "results/cv-results-area.csv")
-
-plt <- cv_results_df %>%
-  gather(trait, value, -i, -model) %>%
-  group_by(model, trait) %>%
-  summarize(M = mean(value), S = sd(value)) %>%
-  ungroup() %>%
-  mutate(model = factor(model, c("uni", "multi_global",
-                                 "multi_group", "hier")),
-         trait = factor(trait, shiklomanov2017np::area_params)) %>%
-  ggplot() +
-  aes(x = model, y = M) +
-  geom_col() +
-  facet_wrap(vars(trait), scales = "free_y") +
-  cowplot::theme_cowplot() +
-  theme(
-    axis.text.x = element_text(angle = )
-  )

### 2019-09-02 scripts/n-fold-mass.R

@@ -98,7 +98,7 @@ cross_validate <- function(data_mat, groups, nremove) {
 
   # Multivariate model -- by PFT
   mboot_means <- matrix(NA_real_, nrow(test_data), ncol(test_data))
-  for (g in levels(data_groups)) {
+  for (g in levels(groups)) {
     ig <- groups == g
     gdata <- test_data[ig, ]
     gmat <- multi_results[[g]]
@@ -135,20 +135,3 @@ cv_results <- replicate(ncv, cross_validate(data_mat, data_groups),
 
 cv_results_df <- bind_rows(cv_results, .id = "i")
 write_csv(cv_results_df, file.path(resultsdir, "cv-results-mass.csv"))
-
-## plt <- cv_results_df %>%
-##   gather(trait, value, -i, -model) %>%
-##   group_by(model, trait) %>%
-##   summarize(M = mean(value), S = sd(value)) %>%
-##   ungroup() %>%
-##   mutate(model = factor(model, c("uni", "multi_global",
-##                                  "multi_group", "hier")),
-##          trait = factor(trait, shiklomanov2017np::mass_params)) %>%
-##   ggplot() +
-##   aes(x = model, y = M) +
-##   geom_col() +
-##   facet_wrap(vars(trait), scales = "free_y") +
-##   cowplot::theme_cowplot() +
-##   theme(
-##     axis.text.x = element_text(angle = )
-##   )

### 2019-09-02 scripts/n-fold-analysis.R
Add plotting code for n-fold-analysis

@@ -1,3 +1,0 @@
-
-
-nfold_

### 2019-09-02 scripts/n-fold-plot.R

@@ -1,0 +1,30 @@
+library(tidyverse)
+library(here)
+
+model_lvl <- c("uni", "multi_global", "multi_group", "hier")
+cols <- cols(i = col_number(),
+             model = col_factor(model_lvl),
+             .default = col_number())
+cv_results_mass <- here("results", "cv-results-mass.csv") %>%
+  read_csv(col_types = cols)
+cv_results_area <- here("results", "cv-results-area.csv") %>%
+  read_csv(col_types = cols)
+cv_results <- bind_rows("mass" = cv_results_mass, "area" = cv_results_area,
+                        .id = "mass_area")
+
+plt <- cv_results %>%
+  gather(trait, value, -i, -model, -mass_area) %>%
+  group_by(model, mass_area, trait) %>%
+  summarize(M = mean(value), S = sd(value)) %>%
+  ungroup() %>%
+  mutate(trait = drop_mass_area(trait)) %>%
+  ggplot() +
+  aes(x = model, y = M) +
+  geom_col() +
+  facet_grid(vars(trait), vars(mass_area), scales = "free_y") +
+  labs(y = "Mean RMSE", x = "Model") +
+  cowplot::theme_cowplot() +
+  theme(
+    axis.text.x = element_text(angle = )
+  )
+if (interactive()) plt

### 2019-09-02 scripts/n-fold-analysis.R
Unify n-fold-* scripts and create CLI

@@ -1,11 +1,25 @@
+#!/usr/bin/env Rscript
+# This script takes two command line arguments.
+#
+# The first determines if the cross validation is on "mass" or "area" normalized
+# traits. It is required.
+#
+# The second specifies the number of cross validations to do. If omitted, the
+# default is 20.
 library(shiklomanov2017np)
 library(tidyverse)
+
+argv <- commandArgs(trailingOnly = TRUE)
+stopifnot(length(argv) %in% c(1, 2))
+mass_area <- match.arg(argv[1], c("mass", "area"))
+ncv <- argv[2]
+if (is.na(ncv)) ncv <- 20
 
 outdir <- here::here("output")
 resultsdir <- here::here("results")
 dir.create(resultsdir, showWarnings = FALSE, recursive = TRUE)
 
-dat_df_all <- try_data("mass")
+dat_df_all <- try_data(mass_area)
 
 data_groups <- dat_df_all[["pft"]]
 data_mat <- dat_df_all %>%
@@ -13,7 +27,8 @@ data_mat <- dat_df_all %>%
   as.matrix() %>%
   log10()
 
-r_hm_file <- tail(list.files("output", "hier.mass.clm45", full.names = TRUE), 1)
+hier_pattern <- paste0("hier\\.", mass_area, "\\.clm45")
+r_hm_file <- tail(list.files("output", hier_pattern, full.names = TRUE), 1)
 r_hm <- readRDS(r_hm_file)
 # Remove unnecessary stuff from the hierarchical output so that
 r_hm_mat <- coda:::as.matrix.mcmc.list(
@@ -23,7 +38,8 @@ r_hm_mat <- r_hm_mat[, !grepl("^Corr", colnames(r_hm_mat))]
 r_hm_mat <- r_hm_mat[, !grepl("global", colnames(r_hm_mat))]
 
 # Multivariate results, globally
-gmulti_file <- tail(list.files(outdir, "multi.area.clm45.NA",
+gmulti_pattern <- paste0("multi\\.", mass_area, "\\.clm45\\.NA")
+gmulti_file <- tail(list.files("output", gmulti_pattern,
                                full.names = TRUE), 1)
 gmulti <- readRDS(gmulti_file)
 gmulti_mat <- coda:::as.matrix.mcmc.list(gmulti$samples)
@@ -33,7 +49,7 @@ gmulti_mat <- gmulti_mat[, !grepl("^Corr", colnames(gmulti_mat))]
 multi_results <- lapply(
[…]

### 2019-09-02 scripts/n-fold-area.R

@@ -1,137 +1,0 @@
-library(shiklomanov2017np)
-library(tidyverse)
-
-outdir <- here::here("output")
-resultsdir <- here::here("results")
-dir.create(resultsdir, showWarnings = FALSE, recursive = TRUE)
-
-dat_df_all <- try_data("area")
-
-data_groups <- dat_df_all[["pft"]]
-data_mat <- dat_df_all %>%
-  select(-pft) %>%
-  as.matrix() %>%
-  log10()
-
-r_hm_file <- tail(list.files("output", "hier.area.clm45", full.names = TRUE), 1)
-r_hm <- readRDS(r_hm_file)
-# Remove unnecessary stuff from the hierarchical output so that
-r_hm_mat <- coda:::as.matrix.mcmc.list(
-  window(r_hm$samples, start = 10000)
-)
-r_hm_mat <- r_hm_mat[, !grepl("^Corr", colnames(r_hm_mat))]
-r_hm_mat <- r_hm_mat[, !grepl("global", colnames(r_hm_mat))]
-
-# Multivariate results, globally
-gmulti_file <- tail(list.files("output", "multi.area.clm45.NA",
-                               full.names = TRUE), 1)
-gmulti <- readRDS(gmulti_file)
-gmulti_mat <- coda:::as.matrix.mcmc.list(gmulti$samples)
-gmulti_mat <- gmulti_mat[, !grepl("^Corr", colnames(gmulti_mat))]
-
-# Load all the multivariate results by PFT
-multi_results <- lapply(
-  levels(data_groups),
-  function(g) {
-    pattern <- paste0("multi\\.area\\.clm45\\.", g, ".2019")
-    f <- tail(list.files("output", pattern, full.names = TRUE), 1)
-    d <- readRDS(f)
-    dmat <- coda:::as.matrix.mcmc.list(d$samples)
-    dmat <- dmat[, !grepl("^Corr", colnames(dmat))]
-    dmat
-  }
-)
-names(multi_results) <- levels(data_groups)
-
-# Load univariate results
-uni_results <- lapply(
-  levels(data_groups),
-  function(g) {
-    f <- paste0("output/uni.area.clm45.", g, ".rds")
-    d <- readRDS(f)
-    d$stats$mu$Mean
-  }
-)
-names(uni_results) <- levels(data_groups)
-
-##################################################
-# Begin cross-validation
-##################################################
-
[…]

### 2019-09-02 scripts/n-fold-analysis.R
Bugfixes to n-fold analysis

@@ -9,10 +9,16 @@
 library(shiklomanov2017np)
 library(tidyverse)
 
-argv <- commandArgs(trailingOnly = TRUE)
+if (FALSE) {
+  # For interactive testing
+  argv <- c("area", 3)
+  argv <- c("mass", 3)
+}
+
+if (!exists("argv")) argv <- commandArgs(trailingOnly = TRUE)
 stopifnot(length(argv) %in% c(1, 2))
 mass_area <- match.arg(argv[1], c("mass", "area"))
-ncv <- argv[2]
+ncv <- as.numeric(argv[2])
 if (is.na(ncv)) ncv <- 20
 
 outdir <- here::here("output")
@@ -74,14 +80,7 @@ names(uni_results) <- levels(data_groups)
 # Begin cross-validation
 ##################################################
 
-# Create a matrix of indices (which I will eventually sample) where at least 2
-# traits are present
-present <- !is.na(data_mat)
-gt2 <- rowSums(present) >= 2
-data_mat_sub <- data_mat[gt2, ]
-groups_sub <- data_groups[gt2]
-
-cross_validate <- function(data_mat, groups, nremove) {
+cross_validate <- function(data_mat, groups, nremove = 1000) {
 
   message("Beginning cross validation")
   if (exists("incv")) {
@@ -157,8 +156,13 @@ cross_validate <- function(data_mat, groups, nremove) {
   rmse_all
 }
 
-incv <- 1
-cv_results <- replicate(ncv, cross_validate(data_mat, data_groups),
+# Create a matrix of indices (which I will eventually sample) where at least 2
+# traits are present
+present <- !is.na(data_mat)
+gtx <- rowSums(present) >= 1
+
+incv <- 0
+cv_results <- replicate(ncv, cross_validate(data_mat[gtx, ], data_groups[gtx]),
                         simplify = FALSE)
 
 cv_results_df <- bind_rows(cv_results, .id = "i")

### 2019-09-02 scripts/n-fold-plot.R

@@ -1,3 +1,4 @@
+library(shiklomanov2017np)
 library(tidyverse)
 library(here)
 

### 2019-09-02 scripts/n-fold-plot.R
Publication-ready version of n-fold plot

@@ -1,31 +1,61 @@
 library(shiklomanov2017np)
 library(tidyverse)
 library(here)
+library(cowplot)
 
 model_lvl <- c("uni", "multi_global", "multi_group", "hier")
 cols <- cols(i = col_number(),
              model = col_factor(model_lvl),
              .default = col_number())
 cv_results_mass <- here("results", "cv-results-mass.csv") %>%
-  read_csv(col_types = cols)
+  read_csv(col_types = cols) %>%
+  rename_all(~gsub("_?mass", "", .))
 cv_results_area <- here("results", "cv-results-area.csv") %>%
-  read_csv(col_types = cols)
+  read_csv(col_types = cols) %>%
+  rename_all(~gsub("_?area", "", .))
 cv_results <- bind_rows("mass" = cv_results_mass, "area" = cv_results_area,
                         .id = "mass_area")
 
-plt <- cv_results %>%
+cv_long <- cv_results %>%
   gather(trait, value, -i, -model, -mass_area) %>%
   group_by(model, mass_area, trait) %>%
   summarize(M = mean(value), S = sd(value)) %>%
   ungroup() %>%
-  mutate(trait = drop_mass_area(trait)) %>%
+  mutate(trait = drop_mass_area(trait),
+         mass_area = factor(mass_area, c("mass", "area"))) %>%
+  arrange(mass_area, trait)
+
+plt <- cv_long %>%
+  group_by(mass_area, trait) %>%
+  mutate(Mnorm = M / max(M)) %>%
   ggplot() +
-  aes(x = model, y = M) +
-  geom_col() +
+  aes(x = model, y = Mnorm, group = 1) +
+  geom_point() +
+  geom_line() +
   facet_grid(vars(trait), vars(mass_area), scales = "free_y") +
-  labs(y = "Mean RMSE", x = "Model") +
-  cowplot::theme_cowplot() +
-  theme(
-    axis.text.x = element_text(angle = )
-  )
+  labs(y = "Normalized mean RMSE", x = "Model") +
+  theme_cowplot()
 if (interactive()) plt
+figdir <- here("manuscript", "figures")
+dir.create(figdir, showWarnings = FALSE, recursive = TRUE)
+ggsave(file.path(figdir, "n-fold-rmse.png"), plt,
+       width = 8.78, height = 8.54)
+
+## # GGmatrix
+## plt_l <- cv_long %>%
[…]

### 2019-09-02 scripts/manuscript-analyses/many_results.R
Fix data loading in many_results

@@ -3,13 +3,11 @@
 
 ## ----options, echo=FALSE, message=FALSE----------------------------------
 library(shiklomanov2017np)
-manuscript_fig_dir <- here("figures", "manuscript")
+
+manuscript_fig_dir <- here("manuscript", "figures")
 dir.create(manuscript_fig_dir, showWarnings = FALSE)
 
-try_data <- readRDS(here("extdata/traits_analysis.rds"))
-
-try_sub <- try_data %>%
-  select(pft = clm45, !!both_params)
+try_sub <- try_data()
 
 sample_size <- try_sub %>%
   select(pft, one_of(both_params)) %>%

### 2019-09-02 scripts/manuscript-analyses/many_results.R
Remove sample size plot from many_results

@@ -19,29 +19,6 @@ sample_size <- try_sub %>%
   gather("param", "sample_size", -pft) %>%
   mutate(param = factor(param, both_params))
 
-## ----"samplesize", fig.cap='(ref:ssfigcap)', fig.width=7, fig.height=7----
-ss_plot <- sample_size %>%
-  mutate(param = forcats::lvls_revalue(param, param_simple[both_params])) %>%
-  ggplot() +
-  aes(x = pft, y = sample_size, fill = pft) +
-  geom_col() +
-  scale_y_log10(breaks = c(10, 100, 1000, 10000)) +
-  facet_wrap(~param, ncol = 2, labeller = label_parsed) +
-  scale_fill_manual(values = pft_colors) +
-  xlab("Plant functional type") +
-  ylab("Sample size") +
-  guides(fill = guide_legend(title = "PFT")) +
-  theme_bw() +
-  theme(
-    axis.text.x = element_blank(),
-    axis.ticks.x = element_blank()
-  )
-if (interactive()) {
-  ss_plot
-}
-ggsave(file.path(manuscript_fig_dir, "sample_size.pdf"), ss_plot,
-        width = 7, height = 7)
-
 ## ----prepresults---------------------------------------------------------
 resultsfile <- here("results/mvtraits_results.rds")
 results_all <- readRDS(resultsfile)

### 2019-09-02 scripts/manuscript-analyses/many_results.R
Revise means figure

@@ -20,7 +20,7 @@ sample_size <- try_sub %>%
   mutate(param = factor(param, both_params))
 
 ## ----prepresults---------------------------------------------------------
-resultsfile <- here("results/mvtraits_results.rds")
+resultsfile <- here("results", "mvtraits_results.rds")
 results_all <- readRDS(resultsfile)
 
 max_ss <- max(sample_size$sample_size)
@@ -123,9 +123,9 @@ table_dat <- means_dat %>%
     hi_f = formatC(mu_hi, digits = 3),
     value = sprintf("%s (%s,%s)", mid_f, lo_f, hi_f),
     value = censor("Vcmax|Jmax", "C4G", pft, param, value),
-    value = censor("Rdmass", "ShDBo", pft, param, value),
-    value = censor("Jmax_mass", c("NlEBo", "NlD"), pft, param, value),
-    value = censor("Jmax_area", c("NlD"), pft, param, value)
+    ## value = censor("Rdmass", "ShDBo", pft, param, value),
+    ## value = censor("Jmax_mass", c("NlEBo", "NlD"), pft, param, value),
+    ## value = censor("Jmax_area", c("NlD"), pft, param, value)
   ) %>%
   select(pft, param, value) %>%
   mutate(
@@ -171,8 +171,24 @@ clm_dat <- clm45_table81 %>%
     mu_lo = mu_mean
   )
 
+param_fancy_2 <- c(
+  leaf_lifespan = "atop(Leaf~lifespan, (months))",
+  Nmass = "atop(N[mass], (mg ~ g ^ -1))",
+  Pmass = "atop(P[mass], (mg ~ g ^ -1))",
+  SLA = "atop(SLA, (m ^ 2 ~ kg ^ -1))",
+  Rdmass = "atop(R[list(d,mass)], (mu * mol ~ g ^ -1 ~ s ^ -1))",
+  Vcmax_mass = "atop(V[list(c, max, mass)], (mu * mol ~ g ^ -1 ~ s ^ -1))",
+  Jmax_mass = "atop(J[list(max, mass)], (mu * mol ~ g ^ -1 ~ s ^ -1))",
+  # TODO: Fix these
+  Narea = "atop(N[area], (g ~ m ^ -2))",
+  Parea = "atop(P[area], (g ~ m ^ -2))",
[…]

### 2019-09-02 scripts/manuscript-analyses/correlation-power.R
Fix manuscript figure path

@@ -94,6 +94,6 @@ plt_color <- ggplot() +
   cowplot::theme_cowplot()
 plt_color
 
-figdir <- here::here("manuscript", "figures")
+figdir <- here::here("figures", "manuscript")
 dir.create(figdir, showWarnings = FALSE, recursive = TRUE)
 ggsave(file.path(figdir, "correlation-"))

### 2019-09-02 scripts/manuscript-analyses/many_results.R

@@ -4,7 +4,7 @@
 ## ----options, echo=FALSE, message=FALSE----------------------------------
 library(shiklomanov2017np)
 
-manuscript_fig_dir <- here("manuscript", "figures")
+manuscript_fig_dir <- here("figures", "manuscript")
 dir.create(manuscript_fig_dir, showWarnings = FALSE)
 
 try_sub <- try_data()

### 2019-09-02 scripts/manuscript-analyses/rma-plot.R

@@ -3,8 +3,6 @@ library(shiklomanov2017np)
 outdir <- here::here("output")
 hier_mass_file <- tail(dir(outdir, "hier.mass.clm45", full.names = TRUE), 1)
 hier_area_file <- tail(dir(outdir, "hier.area.clm45", full.names = TRUE), 1)
-multi_mass_file <- tail(dir(outdir, "multi.mass.clm45.NA", full.names = TRUE), 1)
-multi_area_file <- tail(dir(outdir, "multi.area.clm45.NA", full.names = TRUE), 1)
 
 rma_slope <- function(m, pft, yvar, xvar) {
   sdy_col <- paste("Sigma", pft, yvar, yvar, sep = "..")

### 2019-09-02 scripts/n-fold-plot.R

@@ -36,7 +36,7 @@ plt <- cv_long %>%
   labs(y = "Normalized mean RMSE", x = "Model") +
   theme_cowplot()
 if (interactive()) plt
-figdir <- here("manuscript", "figures")
+figdir <- here("figures", "manuscript")
 dir.create(figdir, showWarnings = FALSE, recursive = TRUE)
 ggsave(file.path(figdir, "n-fold-rmse.png"), plt,
        width = 8.78, height = 8.54)

### 2019-09-03 scripts/manuscript-analyses/rma-plot.R
Save plots as PDF

@@ -90,5 +90,5 @@ plt <- ggplot(rma_both) +
     strip.text.y = element_text(angle = 180)
   )
 
-ggsave("figures/manuscript/rma-slope-tiles.png", plt,
+ggsave("figures/manuscript/rma-slope-tiles.pdf", plt,
        width = 13.5, height = 7.9)

### 2019-09-03 scripts/n-fold-plot.R

@@ -36,9 +36,10 @@ plt <- cv_long %>%
   labs(y = "Normalized mean RMSE", x = "Model") +
   theme_cowplot()
 if (interactive()) plt
+
 figdir <- here("figures", "manuscript")
 dir.create(figdir, showWarnings = FALSE, recursive = TRUE)
-ggsave(file.path(figdir, "n-fold-rmse.png"), plt,
+ggsave(file.path(figdir, "n-fold-rmse.pdf"), plt,
        width = 8.78, height = 8.54)
 
 ## # GGmatrix

### 2019-09-03 scripts/manuscript-analyses/rma-plot.R
Correct order of RMA plot

@@ -68,7 +68,8 @@ rma_both <- bind_rows(rma_mass, rma_area) %>%
       forcats::fct_inorder(),
     pft = forcats::fct_recode(pft, !!!abbr2pft),
     significant = sign(lo) == sign(hi),
-    sigstar = if_else(significant, "*", NA_character_)
+    sigstar = if_else(significant, "*", NA_character_),
+    mass_area = factor(mass_area, c("mass", "area"))
   )
 
 plt <- ggplot(rma_both) +

### 2019-09-03 scripts/n-fold-plot.R
Drop multi-global model from RMSE figure

@@ -14,7 +14,9 @@ cv_results_area <- here("results", "cv-results-area.csv") %>%
   read_csv(col_types = cols) %>%
   rename_all(~gsub("_?area", "", .))
 cv_results <- bind_rows("mass" = cv_results_mass, "area" = cv_results_area,
-                        .id = "mass_area")
+                        .id = "mass_area") %>%
+  filter(model != "multi_global") %>%
+  mutate(model = fct_recode(model, "multi" = "multi_group"))
 
 cv_long <- cv_results %>%
   gather(trait, value, -i, -model, -mass_area) %>%

### 2019-09-03 scripts/manuscript-analyses/many_results.R
Add mean_comparison_all figure

@@ -188,7 +188,6 @@ param_fancy_2 <- c(
 )
 plot_dat <- means_dat %>%
   mutate(model_type = factor(model_type, model_type_levels)) %>%
-  filter(model_type == "hierarchical") %>%
   bind_rows(clm_dat) %>%
   mutate(
     # Convert Nmass and Pmass to correct units
@@ -209,9 +208,8 @@ plot_dat <- means_dat %>%
     mu_hi = pclip(param, "Jmax_area", mu_hi, 300),
     param = forcats::lvls_revalue(param, param_fancy_2[both_params])
   )
-p_means <- ggplot(plot_dat) +
+p_means <- ggplot(filter(plot_dat, model_type %in% c("hierarchical", "CLM 4.5"))) +
   aes(
-    ## x = interaction(model_type, pft),
     x = pft,
     y = mu_mean,
     ymin = mu_lo,
@@ -265,6 +263,60 @@ if (interactive()) {
 }
 ggsave(file.path(manuscript_fig_dir, "mean_comparison.pdf"), p_means,
        width = 6.44, height = 8.1)
+
+p_means_all <- ggplot(plot_dat) +
+  aes(
+    x = interaction(model_type, pft),
+    y = mu_mean,
+    ymin = mu_lo,
+    ymax = mu_hi,
+    color = pft,
+    shape = model_type
+  ) +
+  geom_pointrange(size = 0.3) +
+  facet_wrap(~param, scales = "free", ncol = 2,
+             labeller = label_parsed,
+             strip.position = "left") +
+  scale_y_continuous(sec.axis = dup_axis()) +
+  scale_color_manual(values = pft_colors) +
+  scale_shape_manual(values = c(
+    "univariate" = 17,
+    "multivariate" = 15,
+    "hierarchical" = 20,
+    "CLM 4.5" = 4
+  )) +
+  guides(
+    shape = guide_legend(title = "Model type",
+                         nrow = 1,
+                         override.aes = list(linetype = 0)),
+    color = guide_legend(title = "PFT",
+                         byrow = TRUE,
+                         ncol = 8,
+                         label.position = "bottom",
+                         keywidth = 2.5)
+  ) +
+  ylab("Trait estimate mean and 95% CI") +
+  xlab("Plant functional type and model type") +
+  theme_bw() +
+  theme(
+    text = element_text(size = 12),
+    axis.title.y.left = element_blank(),
+    axis.text.y.left = element_text(),
[…]

### 2019-09-03 scripts/manuscript-analyses/rma-plot.R
Add RMA display table code

@@ -93,3 +93,46 @@ plt <- ggplot(rma_both) +
 
 ggsave("figures/manuscript/rma-slope-tiles.pdf", plt,
        width = 13.5, height = 7.9)
+
+saveRDS(rma_both, "extdata/rma-results.rds")
+
+pm_missing <- . %>%
+  group_by(pft) %>%
+  nest() %>%
+  mutate(miss = map(data, get_all_pairwise_missing)) %>%
+  unnest(miss)
+
+dat_mass <- try_data("mass")
+pm_mass <- dat_mass %>% pm_missing
+
+dat_area <- try_data("area")
+pm_area <- dat_area %>% pm_missing
+
+pm_both <- bind_rows(mass = pm_mass, area = pm_area,
+                     .id = "mass_area") %>%
+  mutate(
+    yvar = drop_mass_area(yvar),
+    xvar = drop_mass_area(xvar),
+    mass_area = factor(mass_area, c("mass", "area")),
+    pft = factor(pft, abbr2pft) %>% fct_recode(!!!abbr2pft)
+  )
+
+rma_pm <- rma_both %>%
+  select(-xvar, -yvar) %>%
+  rename(xvar = xvar2, yvar = yvar2) %>%
+  mutate(yvar = as.character(yvar) %>%
+           if_else(. == "leaf_lifespan", "Leaf lifespan", .),
+         yvar = factor(yvar, levels(pm_both$yvar)),
+         xvar = factor(xvar, levels(yvar))) %>%
+  left_join(pm_both, c("xvar", "yvar", "mass_area", "pft"))
+
+rma_display <- rma_pm %>%
+  mutate(
+    present = if_else(pft == "GLOB", 14L, present),
+    sigstar = if_else(is.na(sigstar), "", "*"),
+    RMA = sprintf("%.2f (%.2f, %.2f)%s", Mean, lo, hi, sigstar)
+  ) %>%
+  select(PFT = pft, Normalization = mass_area,
+         `Trait 1` = yvar, `Trait 2` = xvar,
+         `Present` = present, RMA)
+saveRDS(rma_display, "extdata/rma-display.rds")

### 2019-09-03 supplementary_information.Rmd
Janky but functional code for RMA table

--- 2019-08-19 supplementary_information.Rmd
@@
 library(kableExtra)
-library(drake)
-library(ggplot2)
-
-correlation_scatter_cap <- paste(
-  "Relationship between squared pairwise correlation and sample size of pairwise observations for each PFT.",
-  "Grey shading is the 95% confidence interval around a local regression (LOESS).",
-  "Upper triangle is area-normalized values, and lower triangle is mass-normalized values.",
-  "Note that the x-axis uses a $\\log_{10}$ scale."
-)
+## library(drake)
+library(dplyr, exclude = "group_rows")
+library(tidyverse)
 
@@
   "as a function of sample size for each plant functional type."
-)
-
-correlation_range_cap <- paste(
-  "Relationship between pairwise correlation strength and data range for each PFT.",
-  "In each pair of sub-panels, the left panel is relationship between pairwise correlation",
-  "and the range of the variable on the x axis",
-  "while the right panel is the relationship between pairwise correlation",
-  "and the y variable",
-  "(For instance, in the two-panel figure in row 1 column 2,",
-  "the left panel is the SLA-leaf lifespan correlation as a function of the range of SLA",
-  "and the right panel is the same correlation as a function of the range of leaf lifespan).",
-  "Data range is expressed as the difference between the 2.5% and 97.5% quantiles of each normalized trait (i.e. $\\frac{X}{E[X]}$).",
-  "Grey shading is the 95% confidence interval around a local regression (LOESS).",
-  "Upper triangle is area-normalized values, and lower triangle is mass-normalized values."
 )
@@
 )
-table_cap_area <- gsub("mass", "area", table_cap_mass)
 
-readd(table_dat) %>%
-  select(-matches("area")) %>%
+table_cap_area <- paste(
+  "Mean and 95\\% confidence interval of trait estimates",
+  "for area-normalized traits from the hierarchical model.",
+  "Note that some traits for some PFTs have effectively no constraint relative to the mostly uninformative prior (as evidenced by their very large confidence intervals),",
+  "so their values should be used with caution.",
+  "See mass-normalized table for SLA and leaf lifespan estimates."
+)
+
+with_clm <- readRDS(here::here("extdata", "tidy-means.rds"))
+
+# Mass-normalized table
+with_clm %>%
+  filter(mass_area == "mass",
+         model_type == "hierarchical") %>%
+  select(pft, param, mu_hi, mu_lo, mu_mean) %>%
+  # Fix units
+  mutate_at(
+    vars(starts_with("mu")),
+    ~if_else(param %in% c("Nmass", "Pmass"), . * 1000, .)
+  ) %>%
+  mutate(string = sprintf("%.2f (%.2f, %.2f)",
+                          mu_mean, mu_lo, mu_hi)) %>%
+  select(pft, param, string) %>%
+  spread(param, string) %>%
+  arrange(pft) %>%
   kable(
@@
 
-readd(table_dat) %>%
-  select(-matches("mass")) %>%
+with_clm %>%
+  filter(mass_area == "area",
+         model_type == "hierarchical") %>%
+  select(pft, param, mu_hi, mu_lo, mu_mean) %>%
+  mutate(string = sprintf("%.2f (%.2f, %.2f)",
+                          mu_mean, mu_lo, mu_hi)) %>%
+  select(pft, param, string) %>%
+  spread(param, string) %>%
+  arrange(pft) %>%
   kable(
@@
       "PFT",
-      "Leaf lifespan\n(months)",
-      "SLA\n(kg m$^{-2}$)",
+      ## "Leaf lifespan\n(months)",
+      ## "SLA\n(m$^2$ kg$^{-1}$)",
       "$N_\\textrm{area}$\n(g m$^{-2}$)",
@@
   kableExtra::kable_styling(latex_options = c("scale_down", "hold_position"))
+
 ```
@@
 ```{r sample_size_table, echo = FALSE, results = "asis"}
-ssw <- readd(sample_size_wide)
-kable(
-  ssw %>%
-    dplyr::select(pft, leaf_lifespan, SLA,
-                  Nmass, Narea, Pmass, Parea, Rdmass, Rdarea,
-                  Vcmax_mass, Vcmax_area, Jmax_mass, Jmax_area),
-  col.names = c(
-    "PFT", "Leaf lifespan", "SLA",
-    "$N_\\textrm{mass}$", "$N_\\textrm{area}$",
-    "$P_\\textrm{mass}$", "$P_\\textrm{area}$",
-    "$R_\\textrm{d, mass}$", "$R_\\textrm{d, area}$",
-    "$V_\\textrm{c, max, mass}$", "$V_\\textrm{c, max, area}$",
-    "$J_\\textrm{max, mass}$", "$J_\\textrm{max, area}$"
-  ),
-  format = "latex",
-  booktabs = TRUE,
-  escape = FALSE,
-  caption = "The number of non-missing (outside parentheses) and imputed (inside parentheses) observations of each trait for each plant functional type used in this analysis."
-) %>%
-  kable_styling(latex_options = c("scale_down", "hold_position")) %>%
-  group_rows("Broadleaf", 2, 6) %>%
-  group_rows("Needleleaf", 7, 9) %>%
-  group_rows("Shrub", 10, 12) %>%
-  group_rows("Grass", 13, 15) %>%
-  landscape()
+## ssw <- readd(sample_size_wide)
+## kable(
+##   ssw %>%
+##     dplyr::select(pft, leaf_lifespan, SLA,
+##                   Nmass, Narea, Pmass, Parea, Rdmass, Rdarea,
+##                   Vcmax_mass, Vcmax_area, Jmax_mass, Jmax_area),
+##   col.names = c(
+##     "PFT", "Leaf lifespan", "SLA",
+##     "$N_\\textrm{mass}$", "$N_\\textrm{area}$",
+##     "$P_\\textrm{mass}$", "$P_\\textrm{area}$",
+##     "$R_\\textrm{d, mass}$", "$R_\\textrm{d, area}$",
+##     "$V_\\textrm{c, max, mass}$", "$V_\\textrm{c, max, area}$",
+##     "$J_\\textrm{max, mass}$", "$J_\\textrm{max, area}$"
+##   ),
+##   format = "latex",
+##   booktabs = TRUE,
+##   escape = FALSE,
+##   caption = "The number of non-missing (outside parentheses) and imputed (inside parentheses) observations of each trait for each plant functional type used in this analysis."
+## ) %>%
+##   kable_styling(latex_options = c("scale_down", "hold_position")) %>%
+##   group_rows("Broadleaf", 2, 6) %>%
+##   group_rows("Needleleaf", 7, 9) %>%
+##   group_rows("Shrub", 10, 12) %>%
+##   group_rows("Grass", 13, 15) %>%
+##   landscape()
 ```
@@
 ```{r pairwise_correlation_table, echo = FALSE, results = "asis"}
-corr_results <- readd(corr_processed)
+rma_display <- readRDS(here::here("extdata", "rma-display.rds"))
 kable(
-  corr_results %>%
-    dplyr::select(pft, yvar, xvar, corr_value, present, missing),
+  rma_display,
   "latex",
-  booktabs = TRUE, longtable = TRUE, escape = TRUE,
-  col.names = c("PFT", "Trait 1", "Trait 2", "Correlation (95% CI)", "present", "missing"),
-  caption = "Pairwise trait correlation values for each plant functional type. Values with asterisks indicate correlations significantly different from zero ($p < 0.05$)."
+  booktabs = TRUE, longtable = TRUE,
+  caption = paste("Pairwise reduced major axis (RMA) slope values",
+                  "(Mean and 95pct credible interval)",
+                  "for each plant functional type.",
+                  "Values with asterisks indicate slopes",
+                  "with credible intervals that don't intersect zero.")
 )
-```
 
-```{r correlation_sample_size_pairs, echo = FALSE, results = "hide", warning = FALSE, fig.cap = correlation_scatter_cap, fig.dim = c(12, 12)}
-readd(corr_ss_plot_gg)
-```
-
-```{r correlation_range_analysis, echo = FALSE, results = "hide", warning = FALSE, fig.cap = correlation_range_cap, fig.dim = c(12, 12)}
-readd(corr_range_plot_gg)
+## corr_results <- readd(corr_processed)
+## kable(
+##   corr_results %>%
+##     dplyr::select(pft, yvar, xvar, corr_value, present, missing),
+##   "latex",
+##   booktabs = TRUE, longtable = TRUE, escape = TRUE,
+##   col.names = c("PFT", "Trait 1", "Trait 2", "Correlation (95% CI)", "present", "missing"),
+##   caption = paste(
+##     "Pairwise trait correlation values (Mean and 95% credible interval)",
+##     "for each plant functional type.",
+##     "Values with asterisks indicate correlations",
+##     "with credible intervals that do not intersect zero."
+##   )
+## )
 ```

### 2019-09-03 scripts/manuscript-analyses/correlation-power.R

@@ -34,8 +34,7 @@ dat_mass <- try_data("mass")
 pm_mass <- dat_mass %>% pm_missing
 
 dat_area <- try_data("area")
-pm_area <- dat_area %>%
-  pm_missing
+pm_area <- dat_area %>% pm_missing
 
 lvl <- c("LL", "SLA", "N", "P", "Rd", "Vcmax", "Jmax")
 pm_both <- bind_rows(mass = pm_mass, area = pm_area,

### 2019-09-03 scripts/manuscript-analyses/many_results.R

@@ -186,9 +186,13 @@ param_fancy_2 <- c(
   Vcmax_area = "atop(V[{list(c, max, area)}], (mu * mol ~ m ^ -2 ~ s ^ -1))",
   Jmax_area = "atop(J[{list(max, area)}], (mu * mol ~ m ^ -2 ~ s ^ -1))"
 )
-plot_dat <- means_dat %>%
+with_clm <- means_dat %>%
   mutate(model_type = factor(model_type, model_type_levels)) %>%
-  bind_rows(clm_dat) %>%
+  bind_rows(clm_dat)
+
+saveRDS(with_clm, here::here("extdata", "tidy-means.rds"))
+
+plot_dat <- with_clm %>%
   mutate(
     # Convert Nmass and Pmass to correct units
     mu_mean = if_else(param %in% c("Nmass", "Pmass"), mu_mean * 1000, mu_mean),

### 2019-09-03 supplementary_information.Rmd
Eliminate sample size table from SI
Should be redundant with table in main text.

@@
 
-```
-
-```{r sample_size_table, echo = FALSE, results = "asis"}
-## ssw <- readd(sample_size_wide)
-## kable(
-##   ssw %>%
-##     dplyr::select(pft, leaf_lifespan, SLA,
-##                   Nmass, Narea, Pmass, Parea, Rdmass, Rdarea,
-##                   Vcmax_mass, Vcmax_area, Jmax_mass, Jmax_area),
-##   col.names = c(
-##     "PFT", "Leaf lifespan", "SLA",
-##     "$N_\\textrm{mass}$", "$N_\\textrm{area}$",
-##     "$P_\\textrm{mass}$", "$P_\\textrm{area}$",
-##     "$R_\\textrm{d, mass}$", "$R_\\textrm{d, area}$",
-##     "$V_\\textrm{c, max, mass}$", "$V_\\textrm{c, max, area}$",
-##     "$J_\\textrm{max, mass}$", "$J_\\textrm{max, area}$"
-##   ),
-##   format = "latex",
-##   booktabs = TRUE,
-##   escape = FALSE,
-##   caption = "The number of non-missing (outside parentheses) and imputed (inside parentheses) observations of each trait for each plant functional type used in this analysis."
-## ) %>%
-##   kable_styling(latex_options = c("scale_down", "hold_position")) %>%
-##   group_rows("Broadleaf", 2, 6) %>%
-##   group_rows("Needleleaf", 7, 9) %>%
-##   group_rows("Shrub", 10, 12) %>%
-##   group_rows("Grass", 13, 15) %>%
-##   landscape()
 ```
